### 2016-08-31 paper/paper.tex
Pillage from the IJCAI paper

% vim: set spell spelllang=en tw=100 :

\documentclass[twoside,11pt]{article}
\usepackage{jair}
\usepackage{theapa}
\usepackage{rawfonts}
\usepackage{complexity}
\usepackage{cleveref}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{amsmath}
\usepackage{amssymb}

\usetikzlibrary{decorations, decorations.pathreplacing, calc, backgrounds}

\jairheading{1}{2016}{0-0}{0/0}{0/0}
\ShortHeadings{Why Real-World Subgraph Matching is Easy}
{McCreesh, Prosser \& Trimble}
\firstpageno{1}

\newcommand{\citet}[1]{\citeauthor{#1} \shortcite{#1}}
\newcommand{\citep}[1]{\cite{#1}}

% http://tex.stackexchange.com/questions/22100/the-bar-and-overline-commands
\newcommand{\shortoverline}[1]{\mkern 1.5mu\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}

\definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}
\definecolor{uofglawn}{rgb}{0.517647, 0.741176, 0}
\definecolor{uofgcobalt}{rgb}{0, 0.615686, 0.92549}
\definecolor{uofgpumpkin}{rgb}{1.0, 0.72549, 0.282353}
\definecolor{uofgthistle}{rgb}{0.584314, 0.070588, 0.447059}

\begin{document}

\title{Why Real-World Subgraph Isomorphism is Easy, \\ and What This Means for Graph Databases}

\author{\name Ciaran McCreesh \thanks{This work was supported by the Engineering and Physical Sciences
            Research Council [grant number EP/K503058/1]} \email c.mccreesh.1@research.gla.ac.uk \\
       \name Patrick Prosser \email patrick.prosser@glasgow.ac.uk \\
       \name James Trimble \\
   \addr University of Glasgow, Glasgow, Scotland}
\maketitle

\begin{abstract}
    Despte being \NP-complete, modern subgraph isomorphism algorithms can operate comfortably on
    real-world graphs with thousands of vertices. However, they cannot handle arbitrary instances of
    this size. We show how to generate ``really hard'' random instances for subgraph isomorphism
    problems, which are computationally challenging with only a couple of hundred vertices. For the
    non-induced version of the problem, these instances lie on a satisfiable / unsatisfiable phase
    transition, whose location we can predict; for the induced variant, much richer behaviour is
    observed, and constrainedness gives a better measure of difficulty than does proximity to a
    phase transition. In both cases, we can use this to understand and improve variable and value
    ordering heuristics.

    When labels are introduced, as commonly seen in graph databases, additional rich behaviour
    arises.  We show that when only slightly better algorithms are used, all commonly used benchmark
    suites for evaluating graph database systems become trivial, and that existing notions of
    difficulty are based almost entirely upon unnecessarily poor performance from a widely used
    subgraph isomorphism algorithm. Finally, we argue that popular indexing techniques used in graph
    databases are flawed, and would not give any benefits even with harder instances when used with
    any reasonable modern subgraph isomorphism algorithm.
\end{abstract}

\section{Introduction}

The \emph{non-induced subgraph isomorphism problem} is to find an injective mapping from a given
pattern graph to a given target graph which preserves adjacency---in essence, we are ``finding a
copy of'' the pattern inside the target. The \emph{induced} variant of the problem additionally
requires that the mapping preserve non-adjacency, so there are no ``extra edges'' in the copy of the
pattern that we find. We illustrate both variants in \cref{figure:sip}.
Despite these problems being \NP-complete, modern practical subgraph isomorphism algorithms can
handle problem instances with many hundreds of vertices in the pattern graph, and up to ten thousand
vertices in the target graph \cite{Cordella:2004,Solnon:2010,Audemard:2014,McCreesh:2015}, leading
to successful application in areas such as computer
vision \cite{Damiand:2011,Solnon:2015}, biochemistry \cite{Giugno:2013,Carletti:2015}, and pattern
recognition \citep{Conte:2004}.

\begin{figure}[t]
    \centering
    \begin{tikzpicture}[scale=0.5,every node/.style={font=\footnotesize}]
        \begin{scope}
            \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (Na) at (1,  0) {1};
            \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (Nb) at (1, -1.5) {2};
            \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (Nc) at (0, -3) {3};
            \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (Nd) at (2, -3) {4};

            \draw [ultra thick] (Na) -- (Nb);
            \draw [ultra thick] (Nb) -- (Nc);
            \draw [ultra thick] (Nc) -- (Nd);
            \draw [ultra thick] (Nb) -- (Nd);

            \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (N1) at (5.5,  0) {1};
            \node [draw, circle, fill=white, inner sep=1.5pt] (N2) at (7.5,  0) {2};
            \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (N3) at (5.5, -1.5) {3};
            \node [draw, circle, fill=white, inner sep=1.5pt] (N4) at (7.5, -1.5) {4};
            \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (N5) at (5.5, -3) {5};
            \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (N6) at (7.5, -3) {6};

            \draw [thick, color=uofgsandstone!50] (N1) -- (N2);
            \draw [ultra thick] (N1) -- (N3);
            \draw [thick, color=uofgsandstone!50] (N1) -- (N4);
            \draw [thick, color=uofgsandstone!50] (N2) -- (N4);
            \draw [ultra thick] (N3) -- (N5);
            \draw [ultra thick] (N3) -- (N6);
            \draw [thick, color=uofgsandstone!50] (N4) -- (N6);
            \draw [ultra thick] (N5) -- (N6);
            \draw [thick, color=uofgsandstone!50] (N2) to [in=45, out=315] (N6);
            \draw [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);

            \node [draw, circle, fill=white, inner sep=1.5pt] (M1) at (-5.5,  0) {1};
            \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (M2) at (-3.5,  0) {2};
            \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (M3) at (-5.5, -1.5) {3};
            \node [draw, circle, fill=white, inner sep=1.5pt] (M4) at (-3.5, -1.5) {4};
            \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (M5) at (-5.5, -3) {5};
            \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (M6) at (-3.5, -3) {6};

            \draw [thick, color=uofgsandstone!50] (M1) -- (M2);
            \draw [thick, color=uofgsandstone!50] (M1) -- (M3);
            \draw [thick, color=uofgsandstone!50] (M1) -- (M4);
            \draw [thick, color=uofgsandstone!50] (M2) -- (M4);
            \draw [ultra thick] (M3) -- (M5);
            \draw [ultra thick] (M3) -- (M6);
            \draw [thick, color=uofgsandstone!50] (M4) -- (M6);
            \draw [ultra thick] (M5) -- (M6);
            \draw [ultra thick] (M2) to [in=45, out=315] (M6);
            \draw [thick, color=uofgsandstone!50] (M1) to [in=135, out=225] (M5);

            \node [anchor=center, font=\Large] (A1) at (-1.5, -1.5) { $\hookleftarrow$ };
            \node [anchor=center, font=\Large] (A2) at ( 3.5, -1.5) { $\rightarrowtail$ };
        \end{scope}
    \end{tikzpicture}\\[0.5cm]
    \begin{tikzpicture}[scale=0.45,every node/.style={font=\footnotesize}]
        \begin{scope}
            \foreach \n in {1, ..., 6}{
                \node [draw, circle, fill=uofgcobalt, inner sep=1pt] (N1\n) at (\n, 0) {\n};
            }

            \draw [decorate, decoration={brace, raise=0.2cm}, very thick] (N11.north west) -- (N16.north
                east) node [midway, above=0.3cm] { $1 \mapsto$ };

            \foreach \n in {1, ..., 6}{
                \node [draw, circle, fill=uofgpumpkin, inner sep=1pt] (N2\n) at (0, -\n) {\n};
            }

            \draw [decorate, decoration={brace, raise=0.2cm, mirror}, very thick] (N21.north west) --
                (N26.south west) node [midway, left=0.3cm] { $2 \mapsto$ };

            \foreach \n in {1, ..., 6}{
                \node [draw, circle, fill=uofglawn, inner sep=1pt] (N3\n) at (7, -\n) {\n};
            }

            \draw [decorate, decoration={brace, raise=0.2cm}, very thick] (N31.north east) -- (N36.south
                east) node [midway, right=0.3cm] { $3 \mapsto$ };

            \foreach \n in {1, ..., 6}{
                \node [draw, circle, fill=uofgthistle, inner sep=1pt] (N4\n) at (\n, -7) {\n};
            }

            \draw [decorate, decoration={brace, raise=0.2cm, mirror}, very thick] (N41.south west) --
                (N46.south east) node [midway, below=0.3cm] { $4 \mapsto$ };

            \begin{scope}[on background layer]
                \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N22)!0.8!(N22.east)$);
                \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N23)!0.8!(N23.east)$);
                \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N24)!0.8!(N24.east)$);
                \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N25)!0.8!(N25.east)$);
                \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N36)!0.8!(N36.west)$);
                \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N46)!0.8!(N46.north)$);
                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N12)!0.8!(N12.south)$);
                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N32)!0.8!(N32.west)$);
                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N42)!0.8!(N42.north)$);
                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N13)!0.8!(N13.south)$);
                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N33)!0.8!(N33.west)$);
                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N43)!0.8!(N43.north)$);
                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N14)!0.8!(N14.south)$);
                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N34)!0.8!(N34.west)$);
                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N44)!0.8!(N44.north)$);
                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N15)!0.8!(N15.south)$);
                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N35)!0.8!(N35.west)$);
                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N45)!0.8!(N45.north)$);
                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N22)!0.8!(N22.east)$);
                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N42)!0.8!(N42.north)$);
                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N23)!0.8!(N23.east)$);
                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N43)!0.8!(N43.north)$);
                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N24)!0.8!(N24.east)$);
                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N44)!0.8!(N44.north)$);
                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N25)!0.8!(N25.east)$);
                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N45)!0.8!(N45.north)$);
                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N16)!0.8!(N16.south)$);
                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N22)!0.8!(N22.east)$);
                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N32)!0.8!(N32.west)$);
                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N23)!0.8!(N23.east)$);
                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N33)!0.8!(N33.west)$);
                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N24)!0.8!(N24.east)$);
                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N34)!0.8!(N34.west)$);
                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N25)!0.8!(N25.east)$);
                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N35)!0.8!(N35.west)$);
                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N16)!0.8!(N16.south)$);
                \draw [color=uofgsandstone!50] ($(N12)!0.8!(N12.south)$) -- ($(N33)!0.8!(N33.west)$);
                \draw [color=uofgsandstone!50] ($(N12)!0.8!(N12.south)$) -- ($(N24)!0.8!(N24.east)$);
                \draw [color=uofgsandstone!50] ($(N12)!0.8!(N12.south)$) -- ($(N45)!0.8!(N45.north)$);
                \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N14)!0.8!(N14.south)$);
                \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N34)!0.8!(N34.west)$);
                \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N44)!0.8!(N44.north)$);
                \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N16)!0.8!(N16.south)$);
                \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N36)!0.8!(N36.west)$);
                \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N46)!0.8!(N46.north)$);
                \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N13)!0.8!(N13.south)$);
                \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N24)!0.8!(N24.east)$);
                \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N44)!0.8!(N44.north)$);
                \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N15)!0.8!(N15.south)$);
                \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N26)!0.8!(N26.east)$);
                \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N46)!0.8!(N46.north)$);
                \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N13)!0.8!(N13.south)$);
                \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N24)!0.8!(N24.east)$);
                \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N34)!0.8!(N34.west)$);
                \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N15)!0.8!(N15.south)$);
                \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N26)!0.8!(N26.east)$);
                \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N36)!0.8!(N36.west)$);
                \draw [color=uofgsandstone!50] ($(N13)!0.8!(N13.south)$) -- ($(N34)!0.8!(N34.west)$);
                \draw [color=uofgsandstone!50] ($(N13)!0.8!(N13.south)$) -- ($(N44)!0.8!(N44.north)$);
                \draw [color=uofgsandstone!50] ($(N13)!0.8!(N13.south)$) -- ($(N25)!0.8!(N25.east)$);
                \draw [color=uofgsandstone!50] ($(N13)!0.8!(N13.south)$) -- ($(N26)!0.8!(N26.east)$);
                \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N15)!0.8!(N15.south)$);
                \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N35)!0.8!(N35.west)$);
                \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N45)!0.8!(N45.north)$);
                \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N16)!0.8!(N16.south)$);
                \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N36)!0.8!(N36.west)$);
                \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N46)!0.8!(N46.north)$);
                \draw [color=uofgsandstone!50] ($(N33)!0.8!(N33.west)$) -- ($(N14)!0.8!(N14.south)$);
                \draw [color=uofgsandstone!50] ($(N33)!0.8!(N33.west)$) -- ($(N25)!0.8!(N25.east)$);
                \draw [color=uofgsandstone!50] ($(N33)!0.8!(N33.west)$) -- ($(N45)!0.8!(N45.north)$);
                \draw [color=uofgsandstone!50] ($(N33)!0.8!(N33.west)$) -- ($(N26)!0.8!(N26.east)$);
                \draw [color=uofgsandstone!50] ($(N33)!0.8!(N33.west)$) -- ($(N46)!0.8!(N46.north)$);
                \draw [color=uofgsandstone!50] ($(N43)!0.8!(N43.north)$) -- ($(N14)!0.8!(N14.south)$);
                \draw [color=uofgsandstone!50] ($(N43)!0.8!(N43.north)$) -- ($(N25)!0.8!(N25.east)$);
                \draw [color=uofgsandstone!50] ($(N43)!0.8!(N43.north)$) -- ($(N36)!0.8!(N36.west)$);
                \draw [color=uofgsandstone!50] ($(N14)!0.8!(N14.south)$) -- ($(N35)!0.8!(N35.west)$);
                \draw [color=uofgsandstone!50] ($(N14)!0.8!(N14.south)$) -- ($(N45)!0.8!(N45.north)$);
                \draw [color=uofgsandstone!50] ($(N14)!0.8!(N14.south)$) -- ($(N26)!0.8!(N26.east)$);
                \draw [color=uofgsandstone!50] ($(N24)!0.8!(N24.east)$) -- ($(N16)!0.8!(N16.south)$);
                \draw [color=uofgsandstone!50] ($(N24)!0.8!(N24.east)$) -- ($(N36)!0.8!(N36.west)$);
                \draw [color=uofgsandstone!50] ($(N24)!0.8!(N24.east)$) -- ($(N46)!0.8!(N46.north)$);
                \draw [color=uofgsandstone!50] ($(N34)!0.8!(N34.west)$) -- ($(N15)!0.8!(N15.south)$);
                \draw [color=uofgsandstone!50] ($(N34)!0.8!(N34.west)$) -- ($(N26)!0.8!(N26.east)$);
                \draw [color=uofgsandstone!50] ($(N34)!0.8!(N34.west)$) -- ($(N46)!0.8!(N46.north)$);
                \draw [color=uofgsandstone!50] ($(N44)!0.8!(N44.north)$) -- ($(N15)!0.8!(N15.south)$);
                \draw [color=uofgsandstone!50] ($(N44)!0.8!(N44.north)$) -- ($(N26)!0.8!(N26.east)$);
                \draw [color=uofgsandstone!50] ($(N44)!0.8!(N44.north)$) -- ($(N36)!0.8!(N36.west)$);
                \draw [color=uofgsandstone!50] ($(N15)!0.8!(N15.south)$) -- ($(N26)!0.8!(N26.east)$);
                \draw [color=uofgsandstone!50] ($(N25)!0.8!(N25.east)$) -- ($(N16)!0.8!(N16.south)$);
                \draw [color=uofgsandstone!50] ($(N25)!0.8!(N25.east)$) -- ($(N36)!0.8!(N36.west)$);
                \draw [color=uofgsandstone!50] ($(N25)!0.8!(N25.east)$) -- ($(N46)!0.8!(N46.north)$);
                \draw [color=uofgsandstone!50] ($(N35)!0.8!(N35.west)$) -- ($(N46)!0.8!(N46.north)$);
                \draw [color=uofgsandstone!50] ($(N45)!0.8!(N45.north)$) -- ($(N26)!0.8!(N26.east)$);
                \draw [color=uofgsandstone!50] ($(N45)!0.8!(N45.north)$) -- ($(N36)!0.8!(N36.west)$);
                \draw [ultra thick] ($(N12)!0.8!(N12.south)$) -- ($(N43)!0.8!(N43.north)$);
                \draw [ultra thick] ($(N12)!0.8!(N12.south)$) -- ($(N35)!0.8!(N35.west)$);
                \draw [ultra thick] ($(N12)!0.8!(N12.south)$) -- ($(N26)!0.8!(N26.east)$);
                \draw [ultra thick] ($(N43)!0.8!(N43.north)$) -- ($(N35)!0.8!(N35.west)$);
                \draw [ultra thick] ($(N43)!0.8!(N43.north)$) -- ($(N26)!0.8!(N26.east)$);
                \draw [ultra thick] ($(N35)!0.8!(N35.west)$) -- ($(N26)!0.8!(N26.east)$);
            \end{scope}
        \end{scope}
    \end{tikzpicture}

    \caption{On top, there is a non-induced isomorphism from the pattern graph, in the center, to the target
    on the right, mapping vertex 1 to 1, 2 to 3, 3 to 5 and 4 to 6. This is not an induced
    isomorphism, since there is an edge between 1 and 5 in the target but not between 1 and 3 in the
    pattern. The mapping from the pattern to the (same) target on the left, sending 1 to 2, 2 to 6, 3 to
    5 and 4 to 3, is both non-induced and induced.
    Below, the association graph encoding for the induced version: the highlighted clique
    corresponds to the same mapping.}
    \label{figure:sip}
\end{figure}

However, these algorithms cannot handle \emph{arbitrary} instances of this size. The experimental
evaluations of these algorithms were performed using a mix of real-world graphs, graphs that encode
biochemistry and computer vision problems, and randomly generated graph pairs. Using random
instances to evaluate algorithm behaviour can be beneficial, because it provides a way of generating
many instances cheaply, and reduces the risk of over-fitting when tuning design parameters. The
random instances used in each case came from common datasets \citep{DeSanto:2003,Zampelli:2010},
which were generated by taking a random subgraph of a random (Erd\H{o}s-R\'enyi, scale-free, bounded
degree, or mesh) graph and permuting the vertices. Such instances are guaranteed to be
satisfiable---\citet{Anton:2009} exploited this property to create large sets of random satisfiable
boolean satisfiability instances.  However, since this has been the only approach used to generate
random subgraph isomorphism instances, existing benchmark suites contain relatively few non-trivial
unsatisfiable instances, and the satisfiable instances tend to be computationally fairly easy, with
most of the difficulty being in dealing with the size of the model. This has lead to bias in
algorithm design, to the extent that some proposed techniques will \emph{only} work on satisfiable
instances \citep{Battiti:2007}.

The lack of unsatisfiable instances cannot be addressed by using a pattern graph from one of the
random suites with the ``wrong'' target graph: this tends to give either a trivially unsatisfiable
instance, or a satisfiable instance. (In particular, it is \emph{not} the case that a relatively
small random graph is unlikely to appear in a larger random graph.)

Here we present and evaluate a new method for creating random pattern/target pairs. This method
generates both satisfiable and unsatisfiable instances, and can produce computationally challenging
instances with only a few tens of vertices in the pattern, and 150 vertices in the target. Our work
builds upon the phase transition phenomena observed in satisfiability and graph colouring problems
first described by \citet{DBLP:conf/ijcai/CheesemanKT91} and \citet{Mitchell:1992}.  For subgraph isomorphism we
identify three relevant control parameters: we can independently alter the edge probability of the
pattern graph, the edge probability of the target graph, and the relative orders (number of
vertices) of the pattern and target graphs.  For non-induced isomorphisms, with the correct choice
of parameters we see results very similar to those observed with boolean satisfiability problems:
there is a phase transition (whose location we can predict) from satisfiable to unsatisfiable,
we see a solver-independent complexity peak occur near this phase transition, and understanding this
behaviour helps us to select variable and value ordering heuristics.

For certain choices of parameters for induced isomorphisms, there are two phase transitions, going
from satisfiable to unsatisfiable, and then from unsatisfiable back to satisfiable. Again, when
going from satisfiable to unsatisfiable (from either direction), instances go from being trivial to
really hard to solve. However, each of the three solvers we tested also finds the central
unsatisfiable region to be hard, despite it not being near a phase transition. To show that this is
not a simple weakness of current subgraph isomorphism algorithms, we verify that this region is also
hard when using a pseudo-boolean encoding, and under reduction to the clique problem. Interestingly,
the constrainedness measure proposed by \citet{Gent:1996:Kappa} \emph{does} predict this difficult
region---these instances provide evidence in favour of constrainedness, rather than proximity to a
phase transition, being an accurate predictor of difficulty, and show that constrainedness is not
simply a refinement of a phase transition prediction.

\subsection{Definitions}

Throughout, our graphs are unlabelled, undirected, and do not have any loops.  The \emph{order} of a
graph is the cardinality of its vertex set. We write $\operatorname{V}(G)$ for the vertex set of a
graph $G$. The \emph{complement} of a graph $G$, denoted $\shortoverline{G}$, is the graph with the
same vertex set as $G$, and with an edge between distinct vertices $v$ and $w$ if and only if $v$
and $w$ are not adjacent in $G$. We write $G(n, p)$ for an Erd\H{o}s-R\'enyi random graph with $n$
vertices, and an edge between each distinct pair of vertices with independent probability $p$.

A \emph{non-induced subgraph isomorphism} from a graph $P$ (called the \emph{pattern}) to a graph
$T$ (called the \emph{target}) is an injective mapping from $\operatorname{V}(P)$ to
$\operatorname{V}(T)$ which preserves adjacency---that is, for every adjacent $v$ and $w$ in
$\operatorname{V}(P)$, the vertices $i(v)$ and $i(w)$ are adjacent in $T$. An \emph{induced subgraph
isomorphism} additionally preserves non-adjacency---that is, if $v$ and $w$ are not adjacent in $P$,
then $i(v)$ and $i(w)$ must not be adjacent in $T$. We use the notation $i : P \rightarrowtail T$
for a non-induced isomorphism, and $i : P \hookrightarrow T$ for an induced isomorphism. Observe
that an induced isomorphism $i : P \hookrightarrow T$ is a non-induced isomorphism $i : P
\rightarrowtail T$ which is also a non-induced isomorphism $\shortoverline{i} : \shortoverline{P}
\rightarrowtail \shortoverline{T}$.

\subsection{Experimental Setup}

Our experiments were performed on systems with Intel Xeon E5-4650 v2 CPUs, running Scientific Linux
release 6.7. We selected three subgraph isomorphism solvers: the Glasgow solver
\citep{McCreesh:2015}, LAD \citep{Solnon:2010}, and VF2 \citep{Cordella:2004}; each was compiled
using GCC 4.9.

The Glasgow and LAD solvers use backtracking search to build up an assignment of target vertices
(values) to pattern vertices (variables), but differ in terms of inference and ordering heuristics.
The approach used by VF2 is similar, although the domains of variables are not stored (in the style
of conventional backtracking, rather than forward-checking), and so domain wipeouts are not detected
until an assignment is made.

In each case we measure the number of recursive calls (guessed assignments) made, not runtimes. We
are not aiming to compare absolute performance between solvers; rather, we are looking for
solver-independent patterns of difficulty. We used a timeout of 1,000 seconds, which was enough for
the Glasgow solver to solve nearly all our instances (whose orders were selected with this timeout
in mind), although we may slightly overestimate the proportion of unsatisfiable instances for
extremely sparse or dense pattern graphs. The LAD and VF2 solvers experienced many more failures
with this timeout, so our picture of just how hard the hardest instances are with these solvers is
less detailed.

\section{Non-Induced Subgraph Isomorphisms}

\begin{figure}[tb]
    %% \includegraphics{gen-graph-phase-transition.pdf}
    \caption{With a fixed pattern graph order of 20, a target graph order of 150, a target edge
        probability of 0.40, and varying pattern edge probability, we observe a phase transition and
        complexity peak with the Glasgow solver in the non-induced variant. Each point represents
        one instance. The lines show mean search effort and mean proportion satisfiable.}
    \label{figure:phase-transition}
\end{figure}

Suppose we arbitrarily decide upon a pattern graph order of 20, a target graph order of 150, and a
fixed target edge probability of 0.40. As we vary the pattern edge probability from 0 to 1, we would
expect to see a shift from entirely satisfiable instances (with no edges in the pattern, we can
always find a match) to entirely unsatisfiable instances (a maximum clique in this order and edge
probability of target graph will usually have between 9 and 12 vertices). The dark line in
\cref{figure:phase-transition} shows that this is the case. For densities of 0.67 or greater, no
instance is satisfiable; with densities of 0.44 or less, every instance is satisfiable; and with a
density of 0.55, roughly half the instances are satisfiable.

The light line plots mean search effort using the Glasgow solver: for sparse patterns, the problem
is trivial, for dense patterns proving unsatisfiability is not particularly difficult, and we see a
complexity peak around the point where half the instances are satisfiable.  We also plot the search
cost of individual instances, as points. The behaviour we observe looks remarkably similar to random
3SAT problems---compare, for example, Figure 1 of \citet{LeytonBrown:2014}. In particular,
satisfiable instances tend to be easier, but show greater variation than unsatisfiable instances, and
there are exceptionally hard satisfiable instances \citep{Smith:1997}. (The Glasgow solver supports
parallel search with a work-stealing strategy explicitly designed to eliminate these. We have not
enabled this option to avoid dealing with the complexity of search tree measurements under
parallelism.)

\begin{figure}[tb]
    %% \includegraphics{gen-graph-non-induced.pdf}
    \caption{Behaviour of algorithms on the non-induced variant. For each plot, the x-axis is the
        pattern edge probability and the y-axis is the target edge probability, both from 0 to 1.
        Along the top row, we show the proportion of instances which are satisfiable; the white
        bands shows the phase transitions, and the black lines are our predictions of where the
        phase transition will occur. On the final three rows, we show the number of search nodes used by the
        Glasgow, LAD and VF2 solvers; the dark regions indicate ``really hard'' instances.}
    \label{figure:non-induced}
\end{figure}

What if we alter the edge probabilities for both the pattern graph and the target graph?  In the top
row of \cref{figure:non-induced} we show the satisfiability phase transition for the non-induced
variant, for patterns of order 10, 20 and 30, targets of order 150, and varying pattern (x-axis) and
target (y-axis) edge probabilities. Each axis runs over 101 edge probabilities, from 0 to 1 in steps
of 0.01, except for the VF2 row which uses steps of 0.02. For each of these points, we generate ten
random instances. The colour denotes the proportion of these instances which were found to be
satisfiable.  Inside the orange region, at the bottom right of each plot, every instance is
unsatisfiable---here we are trying to find a dense pattern in a sparse target. In the purple region,
at the top left, every instance is satisfiable---we are looking for a sparse pattern in a dense
target (which is easy, since we only have to preserve adjacency, not non-adjacency). The white band
between the regions shows the location of the phase transition: here, roughly half the instances are
satisfiable. (We discuss the black line below.)

On subsequent rows, we show the average number of search nodes used by the different algorithms. In
general, satisfiable instances are easy, until very close to the phase transition. As we hit the
phase transition and move into the unsatisfiable region, we see complexity increase. Finally, as
we pass through the phase transition and move deeper into the unsatisfiable region, instances become
easier again. This behaviour is largely solver-independent, although VF2 has a larger hard region
than Glasgow or LAD. Thus, although we have moved away from a single control parameter, we still
observe the easy-hard-easy pattern seen in many other \NP-complete problems.

\subsection{Locating the Phase Transition}

We can approximately predict the location of the phase transition by calculating (with
simplifications regarding rounding and independence) the expected number of solutions for given
parameters. Since we are trying to find an \emph{injective} mapping from a pattern $P = G(p, d_p)$
to a target $T = G(t, d_t)$, there are \[ t^{\underline{p}} = t \cdot (t - 1) \cdot \ldots \cdot (t -
p + 1) \] possible assignments of target vertices to pattern vertices.  We expect the pattern to have
$d_p \cdot \binom{p}{2}$ edges, so we obtain the probability of all of these edges being mapped to
edges in the target by raising $d_t$ to this power, giving an expected number of solutions of \[
\langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \textnormal{.} \] This
formula predicts a very sharp phase transition from $\langle Sol \rangle \ll 1$ to $\langle Sol
\rangle \gg 1$, which may easily be located numerically. We plot where this occurs using black lines
in the first row of \cref{figure:non-induced}.

This prediction is generally reasonably accurate, except that for very low and very high pattern
densities, we overestimate the satisfiable region. This is due to variance: although an expected
number of solutions much below one implies a high likelihood of unsatisfiability, it is not true
that a high expected number of solutions implies that any particular instance is likely to be
satisfiable. (Consider, for example, a sparse graph which has several isolated vertices. If one
solution exists, other symmetric solutions can be obtained by permuting the isolated vertices.
Thus although the expected number of solutions may be one, there cannot be exactly one solution.) A
similar behaviour is seen with random constraint satisfaction problems
\citep{Smith:1996}.

\subsection{Variable and Value Ordering Heuristics}

Various general principles have been considered when designing variable and value ordering
heuristics for backtracking search algorithms---one of these is to try to maximise the expected
number of solutions inside any subproblem considered during search \citep{Gent:1996:EN}.  This is
usually done by cheaper surrogates, rather than direct calculation. When branching, both LAD and
Glasgow pick a variable with fewest remaining values in its domain: doing this will generally reduce
the first part of the $\langle Sol \rangle$ equation by as little as possible. When two or more
domains are of equal size, LAD simply breaks ties lexicographically, whereas Glasgow will pick a
variable corresponding to a pattern vertex of highest degree. This strategy was determined
empirically, but could have been derived from the $\langle Sol \rangle$ formula: picking a pattern
vertex of high degree will make the remaining pattern subgraph sparser, which will decrease the
exponent in the second half of the formula, maximising the overall value. LAD does not apply a value
ordering heuristic, but Glasgow does: it prefers target vertices of lowest degree.  Again, this was
determined empirically, but it has the effect of increasing $\langle Sol \rangle$ by increasing the
remaining target density. The VF2 heuristics, in contrast, are based around preserving connectivity,
which gives very little discrimination except on the sparsest of inputs.

\section{Induced Subgraph Isomorphisms}

\begin{figure*}[p]
    %% \includegraphics{gen-graph-induced.pdf}
    \caption{Behaviour of algorithms on the induced variant, shown in the style
    of \cref{figure:non-induced}. The second, third and fourth rows show the number of search nodes used by the
    Glasgow, LAD and VF2 algorithms. The fifth row plots constrainedness: the darkest region is where
    $\kappa = 1$, and the lighter regions show where the problem is either over-
    or under-constrained. The final row shows when the Glasgow algorithm performs better when given
    the complements of the pattern and target graphs as inputs---the solid lines show the location
of the phase transition, and the dotted lines are $t_d=0.5$ and the $p_d=t_d$ diagonal.}\label{figure:induced}
\end{figure*}

\begin{figure*}[t]
    %% \includegraphics{gen-graph-sat.pdf}
    \caption{Behaviour of other solvers on the induced variant on smaller graphs, shown in the style of
        \cref{figure:non-induced}. The second row shows the number of search nodes used by the
    Glasgow algorithm, the third and fourth rows show the number of decisions made by the pseudo-boolean
and SAT solvers, and the final shows the number of search nodes used on the clique
encoding.}\label{figure:alt}
\end{figure*}

In the first four rows of \cref{figure:induced} we repeat our experiments, finding induced
isomorphisms. With a pattern of order 10, we get two independent phase transitions: the bottom right
half of the plots resemble the non-induced results, and the top left half is close to a mirror
image. The central satisfiable region, which is away from either phase transition, is
computationally easy, but instances near the phase transition are hard.

For larger patterns of order 20 and 30, we have a large unsatisfiable region in the middle. Despite
not being near either phase transition, instances in the centre remain computationally challenging.
We also plot patterns of orders 14, 15 and 16, to show the transition between the two behaviours.

\subsection{Predictions and Heuristics}

To predict the location of the induced phase transition, we repeat the argument for locating the
non-induced phase transition and additionally considering non-edges, to get an expected number of
solutions of \[ \langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot
{(1 - d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}} \textnormal{.} \] We plot this using black lines on
the top row of \cref{figure:induced}---again, our prediction is accurate except for very sparse or
very dense patterns.

We might guess that degree-based heuristics would just not work for the induced problem: for any
claim about the degree, the opposite will hold for the complement constraints. However, empirically,
this is not the case: on the final row of \cref{figure:induced}, we show whether it is better to use
the original pattern and target as the input to the Glasgow algorithm, or to take the complements.
(The only steps performed by the Glagsow algorithm which differ under taking the complements are
the degree-based heuristics.  LAD and VF2 are not symmetric in this way: LAD performs a filtering
step using degree information, but does not consider the complement degree, and VF2 uses
connectivity in the pattern graph.)

For patterns of order 10, it is always better to try to move towards the satisfiable region: if we
are in the bottom right diagonal half, we are best retaining the original heuristics (which move us
towards the top left), and if we are in the top left we should use the complement instead. This
goes against a suggestion by \citet{Walsh:1998} that switching heuristics based upon an estimate of
the solubility of the problem may offer good performance.

For larger patterns, more complex behaviour emerges. If we are in the intersection of the bottom half
and the bottom right diagonal of the search space, we should always retain the original heuristic,
and if we are in the intersection of the top half and the top left diagonal, we should always use
the complements. This behaviour can be predicted by taking the partial derivatives of $\langle Sol
\rangle$ in the $-p_d$ and $t_d$ directions.  However, when inside the remaining two eighths of the
parameter space, the partial derivatives of $\langle Sol \rangle$ disagree on which heuristic to
use, and using directional derivatives is not enough to resolve the problem. A close observation of
the data suggests that the actual location of the phase transition may be involved (and perhaps
\citeauthor{Walsh:1998}'s suggestion applies only in these conditions). In any case, $\langle Sol
\rangle$ is insufficient to explain the observed behaviour in these two eighths of the parameter space.

In practice, this is unlikely to be a problem: most real-world instances are extremely sparse and
are usually easy, which perhaps explains the continuing popularity of VF2's connectivity-based
heuristics \citep{Carletti:2015}. In this situation, these experiments justify reusing the
non-induced heuristics on induced problems.

\subsection{Is the Central Region Genuinely Hard?}

The region in the parameter space where both pattern and target have medium density is far from a
phase transition, but nevertheless contains instances that are hard for all three solvers. We would
like to know whether this is due to a weakness in current solvers (perhaps our solvers cannot reason
about adjacency and non-adjacency simultaneously?), or whether instances in this region are
inherently difficult to solve.  Thus we repeat the induced experiments on smaller pattern and target
graphs, using different solving techniques.  Although these techniques are not competitive in
absolute terms, we wish to see if the same pattern of behaviour occurs. The results are plotted in
\cref{figure:alt}.

The pseudo-boolean (PB) encoding is as follows. For each pattern vertex $v$ and each target vertex
$w$, we have a binary variable which takes the value 1 if and only if $v$ is mapped to $w$.
Constraints are added to ensure that each pattern vertex maps to exactly one target vertex, that
each target vertex is mapped to by at most one pattern vertex, that adjacent vertices are mapped to
adjacent vertices, and that non-adjacent vertices are mapped to non-adjacent vertices. We used the
Clasp solver~\citep{gekakaosscsc11a} version 3.1.3 to solve the pseudo-boolean instances.  The
instances that are hard for the Glasgow solver remain hard for the PB solver, including instances
inside the central region, and the easy satisfiable instances remain easy. Similar results were
seen with the Glucose SAT solver \cite{glucose} using a direct encoding of the cardinality constraints. (We also
implemented an integer program encoding; the Gurobi solver was only able to solve some of the
trivial satisfiable instances, and was almost never able to prove unsatisfiability within the time
limit.)

The \emph{association graph encoding} of a subgraph isomorphism problem (illustrated in
\cref{figure:sip}) is constructed by creating a new graph with a vertex for each pair $(p, t)$ of
vertices from the pattern and target graphs respectively. There is an edge between vertex $(p_1,
t_1)$ and vertex $(p_2, t_2)$ if mapping $p_1$ to $t_1$ and $p_2$ to $t_2$ simultaneously is
permitted, i.e.\ $p_1$ is adjacent to $p_2$ if and only if $t_1$ is adjacent to $t_2$. A clique of
size equal to the order of the pattern graph exists in the association graph if and only if the
problem is satisfiable \citep{Levi:1973}. We used this encoding with the BBMC clique algorithm
\citep{SanSegundo:2011}, which we implemented in C++. Usually BBMC solves the optimisation version
of the clique problem; we adapted it to solve the decision problem by initialising the incumbent to
be one less than the decision value, and allowing it to exit as soon as a clique with size equal to
the decision value is encountered.  Again, our results show that the instances in the central region
remain hard, and additionally, some of the easy unsatisfiable instances become hard.

Together, these experiments suggest that the central region may be genuinely hard, despite not being
near a phase transition. The clique results in particular rule out the hypothesis that subgraph
isomorphism solvers only find this region hard due to not reasoning simultaneously about adjacency
and non-adjacency, since the association graph encoding constraints consider compatibility rather
than adjacency and non-adjacency.

\subsection{Constrainedness}

Constrainedness, denoted $\kappa$, is an alternative measure of difficulty designed to refine the
phase transition concept, and to generalise hardness parameters across different combinatorial
problems \citep{Gent:1996:Kappa}. A problem with $\kappa < 1$ is said to be underconstrained, and is
likely to be satisfiable; a problem with $\kappa > 1$ is overconstrained, and is likely to be
unsatisfiable. Empirically, problems with $\kappa$ close to 1 are hard, and problems where $\kappa$
is very small or very large are usually easy. By handling injectivity as a restriction on the size
of the state space rather than as a constraint, we derive \[ \kappa = 1 - \frac{\log \left(
t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 - d_{t})}^{(1 - d_{p}) \cdot
\binom{p}{2}} \right)}{\log t^{\underline{p}}}\] for induced isomorphisms, which we plot on the
fifth row of \cref{figure:induced}. We see that constrainedness predicts that the central region
will still be relatively difficult for larger pattern graphs: although the problem is
overconstrained, it is less overconstrained than in the regions the Glasgow and LAD solvers found
easy.  Thus it seems that rather than just being a unification of existing generalised heuristic
techniques, constrainedness also gives a better predictor of difficulty than proximity to a phase
transition---our method generates instances where constrainedness and ``close to a phase
transition'' give very different predictions, and constrainedness gives the better prediction.

Unfortunately, constrainedness does not help us with heuristics: minimising constrainedness gives
the same predictions as maximising the expected number of solutions.

\section{Labelled Graphs}

\section{Other Models of Randomness}

\section{Techniques Used in Graph Databases}

\section{Conclusion}

We have shown how to generate small but hard instances for the non-induced and induced subgraph
isomorphism problems, which will help offset the bias in existing datasets. For non-induced
isomorphisms, behaviour was as in many other hard problems, but for induced isomorphisms we
uncovered several interesting phenomena: there are hard instances far from a phase transition,
constrainedness predicts this, and existing general techniques for designing heuristics do not work
in certain portions of the parameter space.

The model we have proposed may be extended to graphs with labels on the vertices or edges, which
arise naturally in many applications. Broadly speaking, randomly allocated labels make the problem
easier by restricting the search space, although there are pathological cases where increasing the
number of labels shifts an instance from the easy satisfiable region to being close to the phase
transition. The model may also be extended to the (decision version of the) maximum common subgraph
problem. The maximum clique approach is competitive for the maximum common subgraph problem, so it
is interesting to observe that some of the easy portions of the search space become hard under this
reduction---this suggests a possible unnecessary weakness of maximum clique algorithms on certain
kinds of input. (Maximum clique algorithms use a greedy colouring as a bound, and a preliminary
investigation suggests that the association graphs produced by this reduction can be very bad for
greedy colourings in the same way that crown graphs are.)

In contrast, it is worth noting that this technique does \emph{not} give a way of generating hard
instances for graph isomorphism problems: the pattern graph must be substantially smaller than the
target graph for independent pairs of randomly generated instances to give interesting behaviour.

In future work, we intend to repeat the experiments using other random models, including bounded,
regular degree, and scale-free; regular degree graphs in particular foil existing degree-based
heuristics. Similarly, we will look at alternatives to simple randomness for labelling
strategies---for example, in chemical datasets, label sets tend to differ based upon the degrees of
their corresponding vertices.  We will also look at dynamic heuristics, and switching pattern and
target heuristics independently.  Finally, we intend to investigate whether variance can be
calculated efficiently enough to give better predictions for very sparse or dense pattern graphs.

\section*{Acknowledgements}

The authors wish to thank Kitty Meeks and Craig Reilly for their comments.

\vskip 0.2in
\bibliography{paper}
\bibliographystyle{theapa}

\end{document}

### 2016-09-21 paper/paper.tex
Make a start on this

@@
 \usepackage{amssymb}
+\usepackage{nicefrac}
+\usepackage{csquotes}
+
+% \usepackage{showframe}
 
@@
 
-\newcommand{\citet}[1]{\citeauthor{#1} \shortcite{#1}}
+\newcommand{\citet}[1]{\citeA{#1}}
 \newcommand{\citep}[1]{\cite{#1}}
@@
 
-\title{Why Real-World Subgraph Isomorphism is Easy, \\ and What This Means for Graph Databases}
+\title{Why Real-World Subgraph Matching is Easy, and \\ What This Means for Filter / Verify in
+Graph Databases \thanks{Part of this paper appeared at IJCAI 2016
+\citep{DBLP:conf/ijcai/McCreeshPT16}; \cref{section:introduction} has been expanded and
+\cref{section:models,section:labelled,section:filterverify} are new material.}}
 
@@
     difficulty are based almost entirely upon unnecessarily poor performance from a widely used
-    subgraph isomorphism algorithm. Finally, we argue that popular indexing techniques used in graph
-    databases are flawed, and would not give any benefits even with harder instances when used with
+    subgraph isomorphism algorithm. With this in mind, we argue that widely researched ``filter /
+    verify'' indexing technique used in graph databases is a flawed design based upon a widely
+    repeated misunderstanding of what makes subgraph isomorphism hard. We explain why this technique
+    would not be beneficial, even with harder instances than those currently used, when paired with
     any reasonable modern subgraph isomorphism algorithm.
@@
 
-\section{Introduction}
+\section{Introduction}\label{section:introduction}
 
@@
 requires that the mapping preserve non-adjacency, so there are no ``extra edges'' in the copy of the
-pattern that we find. We illustrate both variants in \cref{figure:sip}.
-Despite these problems being \NP-complete, modern practical subgraph isomorphism algorithms can
-handle problem instances with many hundreds of vertices in the pattern graph, and up to ten thousand
-vertices in the target graph \cite{Cordella:2004,Solnon:2010,Audemard:2014,McCreesh:2015}, leading
-to successful application in areas such as computer
-vision \cite{Damiand:2011,Solnon:2015}, biochemistry \cite{Giugno:2013,Carletti:2015}, and pattern
-recognition \citep{Conte:2004}.
+pattern that we find. We illustrate both variants in \cref{figure:sip}.  Although these problems
+are \NP-complete, modern practical subgraph isomorphism algorithms can handle problem instances
+with many hundreds of vertices in the pattern graph, and up to ten thousand vertices in the target
+graph \cite{Cordella:2004,Solnon:2010,Audemard:2014,McCreesh:2015}, leading to successful
+application in areas such as computer vision \cite{Damiand:2011,Solnon:2015}, biochemistry
+\cite{Giugno:2013,Carletti:2015}, and pattern recognition \citep{Conte:2004}.  In particular,
+subgraph isomorphism is one of the key components in supporting complex queries in graph
+databases---typically, these systems store a fixed set of target graphs, and for a sequence of
+pattern queries, they must return every target graph which contains that pattern.  One common
+approach to this problem is to combine a subgraph isomorphism algorithm with an indexing system in a
+so-called ``filter / verify'' model, where invariants are used to attempt to pre-exclude
+unsatisfiable instances to avoid the cost of a subgraph isomorphism call. In this context, the terms
+\emph{matching} and \emph{verification} are often used for the subgraph isomorphism step (or a
+slightly broader problem, for example permitting wildcards).
 
@@
     \begin{tikzpicture}[scale=0.5,every node/.style={font=\footnotesize}]
-        \begin{scope}
+        \begin{scope}[yshift=-2cm]
             \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (Na) at (1,  0) {1};
@@
         \end{scope}
-    \end{tikzpicture}\\[0.5cm]
-    \begin{tikzpicture}[scale=0.45,every node/.style={font=\footnotesize}]
-        \begin{scope}
+        \begin{scope}[xshift=13cm]
             \foreach \n in {1, ..., 6}{
@@
 
-    \caption{On top, there is a non-induced isomorphism from the pattern graph, in the center, to the target
+    \caption{On the left, there is a non-induced isomorphism from the pattern graph, in the center, to the target
     on the right, mapping vertex 1 to 1, 2 to 3, 3 to 5 and 4 to 6. This is not an induced
@@
     5 and 4 to 3, is both non-induced and induced.
-    Below, the association graph encoding for the induced version: the highlighted clique
+    To the right, the association graph encoding for the induced version: the highlighted clique
     corresponds to the same mapping.}
@@
 
-However, these algorithms cannot handle \emph{arbitrary} instances of this size. The experimental
-evaluations of these algorithms were performed using a mix of real-world graphs, graphs that encode
-biochemistry and computer vision problems, and randomly generated graph pairs. Using random
-instances to evaluate algorithm behaviour can be beneficial, because it provides a way of generating
-many instances cheaply, and reduces the risk of over-fitting when tuning design parameters. The
-random instances used in each case came from common datasets \citep{DeSanto:2003,Zampelli:2010},
-which were generated by taking a random subgraph of a random (Erd\H{o}s-R\'enyi, scale-free, bounded
-degree, or mesh) graph and permuting the vertices. Such instances are guaranteed to be
+Despite these successes, subgraph isomorphism algorithms cannot handle \emph{arbitrary} instances
+involving hundreds or thousands of vertices. Experimental evaluations of subgraph isomorphism
+algorithms are usually performed using a mix of real-world graphs, graphs that encode biochemistry
+and computer vision problems, and randomly generated graph pairs. Using random instances to evaluate
+algorithm behaviour can be beneficial, because it provides a way of generating many instances
+cheaply, and reduces the risk of over-fitting when tuning design parameters. The random instances
+used in each case came from common datasets \citep{DeSanto:2003,Zampelli:2010}, which were generated
+by taking a random subgraph of a random (Erd\H{o}s-R\'enyi, scale-free, bounded degree, or mesh)
+graph and permuting the vertices. Such instances are guaranteed to be
 satisfiable---\citet{Anton:2009} exploited this property to create large sets of random satisfiable
-boolean satisfiability instances.  However, since this has been the only approach used to generate
-random subgraph isomorphism instances, existing benchmark suites contain relatively few non-trivial
-unsatisfiable instances, and the satisfiable instances tend to be computationally fairly easy, with
-most of the difficulty being in dealing with the size of the model. This has lead to bias in
-algorithm design, to the extent that some proposed techniques will \emph{only} work on satisfiable
-instances \citep{Battiti:2007}.
+boolean satisfiability instances.  This has been by far the most common approach used so far to
+generate random subgraph isomorphism instances so far, meaning existing benchmark suites contain
+relatively few non-trivial unsatisfiable instances.  Also, the satisfiable instances tend to be
+computationally fairly easy, with most of the difficulty being in dealing with the size of the
+model. This has lead to bias in algorithm design, to the extent that some proposed techniques will
+\emph{only} work on satisfiable instances \citep{Battiti:2007}.
 
-The lack of unsatisfiable instances cannot be addressed by using a pattern graph from one of the
-random suites with the ``wrong'' target graph: this tends to give either a trivially unsatisfiable
-instance, or a satisfiable instance. (In particular, it is \emph{not} the case that a relatively
-small random graph is unlikely to appear in a larger random graph.)
+Our first contribution is to present and evaluate a new method for creating random pattern/target
+pairs. This method generates both satisfiable and unsatisfiable instances, and can produce
+computationally challenging instances with only a few tens of vertices in the pattern, and 150
+vertices in the target. This is not entirely straightforward---the lack of unsatisfiable instances
+for testing purposes cannot be addressed simply by using taking a pattern graph from one of the
+existing random suites with the ``wrong'' target graph, as this tends to give either a trivially
+unsatisfiable instance, or a satisfiable instance. (In particular, it is \emph{not} the case that a
+relatively small random graph is unlikely to appear in a larger random graph.)
 
-Here we present and evaluate a new method for creating random pattern/target pairs. This method
-generates both satisfiable and unsatisfiable instances, and can produce computationally challenging
-instances with only a few tens of vertices in the pattern, and 150 vertices in the target. Our work
-builds upon the phase transition phenomena observed in satisfiability and graph colouring problems
-first described by \citet{DBLP:conf/ijcai/CheesemanKT91} and \citet{Mitchell:1992}.  For subgraph isomorphism we
-identify three relevant control parameters: we can independently alter the edge probability of the
-pattern graph, the edge probability of the target graph, and the relative orders (number of
-vertices) of the pattern and target graphs.  For non-induced isomorphisms, with the correct choice
-of parameters we see results very similar to those observed with boolean satisfiability problems:
-there is a phase transition (whose location we can predict) from satisfiable to unsatisfiable,
-we see a solver-independent complexity peak occur near this phase transition, and understanding this
-behaviour helps us to select variable and value ordering heuristics.
+This work builds upon the phase transition phenomena observed in satisfiability and graph colouring
+problems first described by \citet{DBLP:conf/ijcai/CheesemanKT91} and \citet{Mitchell:1992}.  For
+subgraph isomorphism we identify three relevant control parameters: we can independently alter the
+edge probability of the pattern graph, the edge probability of the target graph, and the relative
+orders (number of vertices) of the pattern and target graphs.  For non-induced isomorphisms, with
+the correct choice of parameters we see results very similar to those observed with boolean
+satisfiability problems: there is a phase transition (whose location we can predict) from
+satisfiable to unsatisfiable, we see a solver-independent complexity peak occur near this phase
+transition, and understanding this behaviour helps us to select variable and value ordering
+heuristics.
 
@@
 
+When labels on vertices are introduced, as is commonly seen in graph database systems, richer
+behaviour emerges. We look at some of the datasets commonly used to test graph database systems,
+which the literature suggests are extremely hard to solve. Our theory predicts that these instances
+should be trivial, and indeed we see that the entire perceived difficulty of every dataset we
+investigate comes down to the widespread use of a particular subgraph isomorphsim algorithm which
+does not use a theoretically-strong variable ordering heuristic. Our second contribution is to show
+that the use of the right variable ordering heuristic, even without the more sophisticated recent
+advances in subgraph algorithms (which have been largely neglected outside of the artificial
+intelligence community), makes the entire graph database filter / verify paradigm unnecessary.
+Finally, we explain why filter / verify cannot be beneficial even on more challenging instances
+except if an extremely poor choice of subgraph isomorphism algorithm is made.
+
 \subsection{Definitions}
 
-Throughout, our graphs are unlabelled, undirected, and do not have any loops.  The \emph{order} of a
-graph is the cardinality of its vertex set. We write $\operatorname{V}(G)$ for the vertex set of a
-graph $G$. The \emph{complement} of a graph $G$, denoted $\shortoverline{G}$, is the graph with the
-same vertex set as $G$, and with an edge between distinct vertices $v$ and $w$ if and only if $v$
-and $w$ are not adjacent in $G$. We write $G(n, p)$ for an Erd\H{o}s-R\'enyi random graph with $n$
-vertices, and an edge between each distinct pair of vertices with independent probability $p$.
+Throughout, our graphs are undirected, and do not have any loops.  The \emph{order} of a graph is
+the cardinality of its vertex set. We write $\operatorname{V}(G)$ for the vertex set of a graph $G$.
+The \emph{complement} of a graph $G$, denoted $\shortoverline{G}$, is the graph with the same vertex
+set as $G$, and with an edge between distinct vertices $v$ and $w$ if and only if $v$ and $w$ are
+not adjacent in $G$. We write $G(n, p)$ for an Erd\H{o}s-R\'enyi random graph with $n$ vertices, and
+an edge between each distinct pair of vertices with independent probability $p$.
+
+A graph may have labels on its vertices, in which case we write $\ell(v)$ for the label of a given
+vertex $v$. We write $G(n, p, l)$ for the graph $G(n, p)$ with vertex labels selected uniformly at
+random from a set of $l$ different labels.
 
@@
 $\operatorname{V}(T)$ which preserves adjacency---that is, for every adjacent $v$ and $w$ in
-$\operatorname{V}(P)$, the vertices $i(v)$ and $i(w)$ are adjacent in $T$. An \emph{induced subgraph
-isomorphism} additionally preserves non-adjacency---that is, if $v$ and $w$ are not adjacent in $P$,
-then $i(v)$ and $i(w)$ must not be adjacent in $T$. We use the notation $i : P \rightarrowtail T$
-for a non-induced isomorphism, and $i : P \hookrightarrow T$ for an induced isomorphism. Observe
-that an induced isomorphism $i : P \hookrightarrow T$ is a non-induced isomorphism $i : P
-\rightarrowtail T$ which is also a non-induced isomorphism $\shortoverline{i} : \shortoverline{P}
-\rightarrowtail \shortoverline{T}$.
+$\operatorname{V}(P)$, the vertices $i(v)$ and $i(w)$ are adjacent in $T$. If vertices are labelled,
+the mapping must preserve labels, so $\ell(v) = \ell(i(v))$. An \emph{induced subgraph isomorphism}
+additionally preserves non-adjacency---that is, if $v$ and $w$ are not adjacent in $P$, then $i(v)$
+and $i(w)$ must not be adjacent in $T$. We use the notation $i : P \rightarrowtail T$ for a
+non-induced isomorphism, and $i : P \hookrightarrow T$ for an induced isomorphism. Observe that an
+induced isomorphism $i : P \hookrightarrow T$ is a non-induced isomorphism $i : P \rightarrowtail T$
+which is also a non-induced isomorphism $\shortoverline{i} : \shortoverline{P} \rightarrowtail
+\shortoverline{T}$.
 
@@
 \begin{figure}[tb]
-    %% \includegraphics{gen-graph-phase-transition.pdf}
+    \centering\includegraphics{gen-graph-phase-transition.pdf}
     \caption{With a fixed pattern graph order of 20, a target graph order of 150, a target edge
@@
 
-Suppose we arbitrarily decide upon a pattern graph order of 20, a target graph order of 150, and a
-fixed target edge probability of 0.40. As we vary the pattern edge probability from 0 to 1, we would
-expect to see a shift from entirely satisfiable instances (with no edges in the pattern, we can
-always find a match) to entirely unsatisfiable instances (a maximum clique in this order and edge
-probability of target graph will usually have between 9 and 12 vertices). The dark line in
-\cref{figure:phase-transition} shows that this is the case. For densities of 0.67 or greater, no
-instance is satisfiable; with densities of 0.44 or less, every instance is satisfiable; and with a
-density of 0.55, roughly half the instances are satisfiable.
+Suppose we arbitrarily decide upon a pattern graph order of 20, a target graph order of 150, a fixed
+target edge probability of 0.40, and no vertex labels. As we vary the pattern edge probability from
+0 to 1, we would expect to see a shift from entirely satisfiable instances (with no edges in the
+pattern, we can always find a match) to entirely unsatisfiable instances (a maximum clique in this
+order and edge probability of target graph will usually have between 9 and 12 vertices). The dark
+line in \cref{figure:phase-transition} shows that this is the case. For densities of 0.67 or
+greater, no instance is satisfiable; with densities of 0.44 or less, every instance is satisfiable;
+and with a density of 0.55, roughly half the instances are satisfiable.
 
@@
 \begin{figure}[tb]
-    %% \includegraphics{gen-graph-non-induced.pdf}
+    \centering\includegraphics{gen-graph-non-induced.pdf}
     \caption{Behaviour of algorithms on the non-induced variant. For each plot, the x-axis is the
@@
 
-\begin{figure*}[p]
-    %% \includegraphics{gen-graph-induced.pdf}
+\begin{figure}[p]
+    \centering\includegraphics{gen-graph-induced.pdf}
     \caption{Behaviour of algorithms on the induced variant, shown in the style
-    of \cref{figure:non-induced}. The second, third and fourth rows show the number of search nodes used by the
-    Glasgow, LAD and VF2 algorithms. The fifth row plots constrainedness: the darkest region is where
+    of \cref{figure:non-induced}. The fifth row plots constrainedness: the darkest region is where
     $\kappa = 1$, and the lighter regions show where the problem is either over-
@@
 of the phase transition, and the dotted lines are $t_d=0.5$ and the $p_d=t_d$ diagonal.}\label{figure:induced}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}[t]
-    %% \includegraphics{gen-graph-sat.pdf}
+\begin{figure}[tb]
+    \centering\includegraphics{gen-graph-sat.pdf}
     \caption{Behaviour of other solvers on the induced variant on smaller graphs, shown in the style of
@@
 encoding.}\label{figure:alt}
-\end{figure*}
+\end{figure}
 
@@
 the data suggests that the actual location of the phase transition may be involved (and perhaps
-\citeauthor{Walsh:1998}'s suggestion applies only in these conditions). In any case, $\langle Sol
+\citeS{Walsh:1998} suggestion applies only in these conditions). In any case, $\langle Sol
 \rangle$ is insufficient to explain the observed behaviour in these two eighths of the parameter space.
@@
 
-\section{Labelled Graphs}
+\section{Other Models of Randomness}\label{section:models}
 
-\section{Other Models of Randomness}
+\section{Labelled Graphs}\label{section:labelled}
 
-\section{Techniques Used in Graph Databases}
+\begin{figure}[t]
+    \centering\includegraphics{gen-graph-labels.pdf}
+    \caption{On the top row, predicted and actual location of the phase transition for labelled
+    non-induced random subgraph isomorphism, with a pattern order of 20, a target order of 100,
+    varying pattern (x-axis) and target (y-axis) density, and varying numbers of labels. On the second
+    row, the average number of search nodes needed to solve an instance.}\label{figure:labels}
+\end{figure}
+
+So far, we have looked at unlabelled graphs. What happens when labels on vertices are introduced?
+This is common in real-world applications---for example, when working with graphs representing
+chemical molecules, mappings are typically expected only to match carbon atoms with carbon atoms,
+hydrogen atoms with hydrogen atoms, and so on.  We will look at the non-induced variant, as this
+seems to be more common in the literature, and return to the Erd\H{o}s-R\'enyi model.
+
+Suppose our labels are drawn randomly from a set $L = \{ 1, \ldots, k \}$, where $k$ is reasonably
+small compared to $p$. We can partition the pattern vertices by label into disjoint sets \[\{
+    \operatorname{V}(P)\vert_1 = \{ v \in \operatorname{V}(P) : \ell(v) = 1 \}, \ldots,
+    \operatorname{V}(P)\vert_k \}\textnormal{\,,}\] each of which is expected to contain
+$\nicefrac{p}{k}$ vertices. Similarly, we may partition the target vertices into disjoint sets $\{
+    \operatorname{V}(T)\vert_1, \ldots, \operatorname{V}(T)\vert_k \}$.
+
+Without labels, there are $t^{\underline{p}} = t \cdot (t - 1) \cdot \ldots \cdot (t - p + 1)$
+possible injective assignments of target vertices to pattern vertices.  With labels, observe that
+for any label $x$, vertices in $\operatorname{V}(P)|_x$ may only be mapped to vertices in
+$\operatorname{V}(T)|_x$.  Thus for each label $x$, we have an expected $\nicefrac{p}{k}$ variables,
+each of whose domains contain $\nicefrac{t}{k}$ values. We would like to say that the size of the
+state space is now \[ |S| = \left((\nicefrac{t}{k})^{\underline{\nicefrac{p}{k}}}\right)^{k}
+\textnormal{,} \] but to do this we must state what $a^{\underline{b}}$ means when $b$ is fractional.
+A reasonable continuous extension may be obtained by taking \[ |S| = \left(\frac{\Gamma\left(\nicefrac{t}{k}
++ 1\right)}{\Gamma\left(\nicefrac{t}{k} - \nicefrac{p}{k} + 1\right)}\right)^{k} \textnormal{\,.} \]
+
+As before, we expect the pattern to have $d_p \cdot \binom{p}{2}$ edges, so we obtain the
+probability of all of these edges being mapped to edges in the target by raising $d_t$ to this
+power, giving an expected number of solutions of \[ \langle Sol \rangle = \left(
+    \frac{\Gamma\left(\nicefrac{t}{k} + 1\right)}{\Gamma\left(\nicefrac{t}{k} - \nicefrac{p}{k} +
+1\right)} \right)^{k}  \cdot
+{d_t}^{d_p \cdot \binom{p}{2}} \textnormal{\,.} \]
+
+So how good are these predictions? We draw some heatmaps in \cref{figure:labels}. For small numbers
+of labels, our predictions are slightly better than in the unlabelled case: there seems to be less
+of a variance problem for very dense patterns. Even as the numbers of labels becomes relatively
+large, the prediction of the phase transition still occurs in the right place, but we start to see
+sporadic unsatisfiable instances deep inside the satisfiable reason. With sufficiently many labels,
+we get a kind of phase transition from ``all unsatisfiable'' to ``mixed satisfiable and
+unsatisfiable''.
+
+\subsection{Flawed Instances}
+
+What goes on here?
+
+Min unsat core.
+
+Insufficient vertices.
+
+Two adjacent differently-coloured vertices.
+
+VF2 does not detect the case that there are insufficient, or even no, labels.
+
+\section{Querying Graph Databases}\label{section:filterverify}
+
+The general problem is, for a set of target graphs, to process a pattern query and return every
+target graph which is subgraph-isomorphic to that pattern. The set of target graphs is usually seen
+as fixed, or at least rarely-changing, whilst the patterns arrive dynamically. This has lead to the
+development of systems which perform extensive computations on the target graphs beforehand, in the
+hopes of reducing the response times for individual pattern queries. The most popular of these
+strategies is a form of indexing which is often named \emph{filter / verify}.
+
+\subsection{The Filter / Verify Paradigm}
+
+The filter / verify approach has an interesting history, of which we now give a very selective and
+incomplete overview. Our description is biased by a general modern understanding of \NP-completeness
+which was not widely known at least at the time of the earlier papers we discuss. The common theme
+of all of the following papers is that pre-computed information is used to eliminate certain
+unsatisfiable instances from consideration, without performing a subgraph isomorphism test.
+
+An early graph database system by \citet{DBLP:conf/pods/ShashaWG02} uses a filtering heuristic to
+eliminate unsatisfiable instances based upon simple structural elements. It is not clear whether the
+aim is to reduce I/O costs, or to reduce the number of queries which must be tested, and the
+experiments do not answer whether the indexing is effective. (?? This isn't the first paper on this,
+but before this seem to be commercial applications, in particular a system called Daylight. Maybe
+more digging is required.)
+
+Subsequently, in a widely cited paper, \citet{DBLP:conf/sigmod/YanYH04} introduce an indexing system
+called gIndex. Again, this system handles queries by first producing a set of candidates by
+eliminating certain unsatisfiable instances, this time by using substructures. They argue that the
+query response time, which is to be minimised, is governed by the equation \[ T =
+T_{\mathit{search}} + \left|C_q\right| \cdot T_{\mathit{iso\_test}} \textnormal{,}\] where
+$T_{\mathit{search}}$ is the time taken to search for a candidate set of potential solutions of size
+$\left|C_q\right|$, and $T_{\mathit{iso\_test}}$ is the cost of a subgraph isomorphism test. They
+reason that since isomorphism testing is \NP-complete, by making $\left|C_q\right|$ as small as
+possible, the query response time will be reduced.  (Importantly, it is \emph{not} time taken to
+load graphs from disk which contributes to the per-candidate cost, but rather the time to perform a
+subgraph isomorphism call.) They conclude that ``graph indexing plays a critical role at efficient
+query processing in graph databases''.
+
+This equation is repeated and expanded upon by \citet{DBLP:journals/tods/YanYH05} to explicitly
+separate I/O and isomorphism testing costs, obtaining a query response time of \[
+    T_{\mathit{search}} + \left|C_q\right| \cdot (T_{\mathit{io}} + T_{\mathit{iso\_test}})
+    \textnormal{.} \] The authors explicitly state that ``the value of $T_{\mathit{iso\_test}}$ does
+not change much for a given query''. The argument presented is that
+\begin{displayquote}``Sequential scan is very costly
+because one has to not only access the whole graph database but also check subgraph isomorphism. It
+is known that subgraph isomorphism is an \NP-complete problem. Clearly, it is necessary to build
+graph indices in order to help processing graph queries.''\end{displayquote}
+
+With what we now know about the behaviour of modern subgraph isomorphism algorithms, and the nature
+of solving \NP-complete problems in general, we should immediately be suspicious of these claims. We
+do not expect $T_{\mathit{iso\_test}}$ to be anything like a constant, even if the orders of the
+input graphs are similar. In particular, any instance which can be excluded based upon filtering
+must have a very small proof of unsatisfiability. These instances \emph{should} be trivial with any
+decent matching algorithm. Thus, all filtering \emph{should} be doing is eliminating the startup
+costs of a trivial subgraph isomorphism call. The fact that filtering was successful empirically
+should make us wonder whether the subgraph isomorphism algorithm being used was excessively
+primitive. Indeed, the matching algorithm used is described only as ``the simplest approach'' in the
+paper.  Additionally, the experiments focus on reducing the size of the candidate set, without
+considering the time taken to verify different candidate set instances.  The claim that
+$T_{\mathit{iso\_test}}$ does not change much is not justified experimentally, and no consideration
+is given as to whether this would hold true for other subgraph matching algorithms.
+
+Moving forwards, the (simpler form of the) query response time equation is repeated by
+\citet{DBLP:conf/vldb/ZhaoYY07}. Again, the work has a focus on reducing the candidate set size
+through indexing. The authors appear to believe that the cost of the isomorphism test is not a major
+factor in influencing the result, and focus on the remaining terms in the equation. They use the
+``average cost'' of a subgraph isomorphism test as a constant, without considering that the average
+cost could be influenced by the character of the candidate set.
+
+The equation is also used by \citet{DBLP:conf/icde/JiangWYZ07}, who claim that ``usually the
+verification time dominates the Query Response Time [s]ince the computational complexity of
+$T_{\mathit{iso\_test}}$ is NP-Complete''. They note that \begin{displayquote}``Approximately, the
+    value of $T_{\mathit{iso\_test}}$ does not change too much with the difference of query. Thus,
+the key to reducing query response time is to minimize the size of the candidate answer
+set''.\end{displayquote} This claim becomes understandable when one examines the subgraph matching
+algorithm chosen for the verification step \citep{DBLP:journals/jacm/Ullmann76}: as the algorithm
+predates techniques like arc-consistent all-different propagation \citep{DBLP:conf/aaai/Regin94}, it
+cannot immediately detect unsatisfiability in simple cases like there being two red vertices in the
+pattern but only one in the target.
+
+Without using the equation, \citet{DBLP:conf/sigmod/ChengKNL07} state that since subgraph
+isomorphism is NP-complete, processing by a sequential scan is infeasible. They introduce new
+filtering techniques to try to avoid the subgraph isomorphism step. A similar claim is made by
+\citet{DBLP:conf/icde/ZhangHY07} in a introduction of another indexing technique: ``obviously it is
+inefficient to perform a sequential scan on every graph in the database, because the subgraph
+isomorphism test is expensive''.
+
+Muddying the waters slightly, a survey by \citet{DBLP:journals/datamine/HanCXY07} states that
+``large volumes of data'' (not \NP-completeness) is the reason for using indexing in these systems.
+However, in a description of a system tailored to biological networks,
+\citet{DBLP:conf/edbt/ZhangLY09} state that \begin{displayquote}``Since the size of the raw database
+    graph is small, it can be easily fit in the main memory.  However, the query (matching) time
+will be very long due to the NP-hard complexity.''\end{displayquote} They suggest that indexing is
+a way of avoiding this.
+
+\citet{DBLP:journals/tkde/WangWYY12} look at indexing large sparse graphs. They state that ``because
+subgraph isomorphism is an \NP-complete problem, a filter-and-verification method is usually
+employed to speed up the search efficiency of graph similarity matching over a graph set''.
+Similarly, after reviewing the literature, \citet{DBLP:journals/vldb/YuanM13} argue that subgraph
+querying is costly because it is \NP-complete, and that indices can improve the performance of graph
+database queries. Again, new indexing techniques are introduced.  Subsequently,
+\citet{DBLP:journals/pvldb/KatsarouNT15} perform a comprehensive comparison of graph database
+filtering techniques. They state that performing a query against each graph in the dataset
+``obviously does not scale, as subgraph isomorphism is \NP-complete''.
+
+The filter / verify paradigm also influences other research. For example, recently
+\citet{DBLP:journals/pvldb/YuanMG13} continue a line of supergraph search work, again using a
+filtering step to avoid subgraph isomorphism calls. There is also research into maintaining indices
+when the set of target graph changes: for example \citet{DBLP:conf/sigmod/YuanMYG15} look at
+algorithms for updating graph indices. And in describing a system for reusing results of queries
+which are sub- or super-graphs of previous queries, \citet{DBLP:conf/edbt/WangNT16} state that
+querying is a ``very costly operation as it entails the \NP-complete problem of subgraph
+isomorphism'', and place ``an emphasis on the number of unnecessary subgraph isomorphism tests''.
+
+After some early ambiguity, it becomes clear that the intent behind filter / verify systems is to
+reduce the number of subgraph isomorphism calls, and that the cost of loading graphs from disk is
+not considered to be problematic. It is worth noting that the entire test datasets from most of
+these papers will comfortably fit in RAM on a modern desktop machine, even when an adjacency matrix
+representation is used.
+
+Thus we can see that there are two critical beliefs underlying all of this work---firstly, that
+subgraph isomorphism is necessarily hard because it is \NP-complete, and secondly, that there are
+ways of identifying unsatisfiable instances using short proofs that a subgraph isomorphism algorithm
+will not detect, but that an indexing system can. Throughout, the cost models used assume that the
+time for a subgraph isomorphism query does not particularly depend upon the instance, and nowhere is
+it considered that a good subgraph isomorphism algorithm should be able to eliminate
+obviously-unsatisfiable instances with a similar time requirement to an indexing system.
+
+\subsection{An Algorithm which Requires No Filtering}
+
+These beliefs are not entirely unfounded: none of the matching algorithms considered in these papers
+will immediately detect if a pattern contains two red vertices, whilst the target graph contains
+only one. This kind of flaw \emph{should} be picked up at the top of search by an all-different
+propagator, but such techniques are not used in graph databases systems, let alone the more powerful
+degree- and difference-based propagation rules introduced LAD \cite{Solnon:2010} or the
+distance-based reasoning in more recent work \cite{Audemard:2014,McCreesh:2015}.
+
+Indeed, as \citet{DBLP:journals/pvldb/KatsarouNT15} note, some variation on VF2
+\citep{Cordella:2004} is the usual matching algorithm of choice, although
+\citeS{DBLP:journals/jacm/Ullmann76}'s algorithm is sometimes chosen. Other approaches have been
+considered: for example, \citet{DBLP:journals/pvldb/ShangZLY08} propose an algorithm which makes use
+of the frequency of various features to guide search.  \citet{DBLP:journals/pvldb/LeeHKL12}
+determine experimentally that this technique tends to be very effective, even on families of graphs
+for which it was not designed.
+
+Thus we believe it would be unlikely to cause too much astonishment if we suggested that a better
+matching algorithm could be dropped in as a black box replacement in graph databases systems to
+improve their performance. However, it appears not to have occurred to the graph databases community
+that a better subgraph isomorphism algorithm could invalidate the premise underlying the entire
+filter / verify approach. Eliminating this filtering step would not merely be a matter of
+simplicity: indices take ??  hours to build.
+
+We build up a simple algorithm. Start with VF2, adapt to track domains, smallest domain first, all
+different. Show its effectiveness at eliminating unsatisfiable instances, and give its total
+performance.  None of the steps in this algorithm are new. Our aim in this section is to show which
+aspects of modern algorithm design are critical, and that an indexless sequential scan is entirely
+plausible with a decent algorithm.
+
+\subsection{What Should the Future of Graph Matching in Databases Be?}
+
+We do not present the ultimate big data subgraph matching algorithm.
+
+Read outside the field. AI has a lot to offer.
+
+Much more focus on the matching algorithm.
+
+Optimisations for trivial cases. Lazy forward checking, cheap alternatives to heuristics and
+all-different. Presolving.
+
+Invariants on labels used to inform algorithm design and to provide filtering during search.
+
+Indexing can be used to enhance propagation.
+
+On the other hand, our key message is not just saying use a better algorithm. Need a full
+understanding of every aspect of the system, and cannot treat subgraph matching as a black box.
 
@@
 
-The model we have proposed may be extended to graphs with labels on the vertices or edges, which
-arise naturally in many applications. Broadly speaking, randomly allocated labels make the problem
-easier by restricting the search space, although there are pathological cases where increasing the
-number of labels shifts an instance from the easy satisfiable region to being close to the phase
-transition. The model may also be extended to the (decision version of the) maximum common subgraph
-problem. The maximum clique approach is competitive for the maximum common subgraph problem, so it
-is interesting to observe that some of the easy portions of the search space become hard under this
-reduction---this suggests a possible unnecessary weakness of maximum clique algorithms on certain
-kinds of input. (Maximum clique algorithms use a greedy colouring as a bound, and a preliminary
-investigation suggests that the association graphs produced by this reduction can be very bad for
-greedy colourings in the same way that crown graphs are.)
+?? Stuff about graph databases.
 
-In contrast, it is worth noting that this technique does \emph{not} give a way of generating hard
-instances for graph isomorphism problems: the pattern graph must be substantially smaller than the
-target graph for independent pairs of randomly generated instances to give interesting behaviour.
-
-In future work, we intend to repeat the experiments using other random models, including bounded,
-regular degree, and scale-free; regular degree graphs in particular foil existing degree-based
-heuristics. Similarly, we will look at alternatives to simple randomness for labelling
-strategies---for example, in chemical datasets, label sets tend to differ based upon the degrees of
-their corresponding vertices.  We will also look at dynamic heuristics, and switching pattern and
-target heuristics independently.  Finally, we intend to investigate whether variance can be
-calculated efficiently enough to give better predictions for very sparse or dense pattern graphs.
+It is worth noting that this technique does \emph{not} give a way of generating hard instances for
+graph isomorphism problems: the pattern graph must be substantially smaller than the target graph
+for independent pairs of randomly generated instances to give interesting behaviour.
 
@@
 
-\vskip 0.2in
 \bibliography{paper}

### 2016-09-23 paper/paper.tex
Write a bit more

@@
     suites for evaluating graph database systems become trivial, and that existing notions of
-    difficulty are based almost entirely upon unnecessarily poor performance from a widely used
-    subgraph isomorphism algorithm. With this in mind, we argue that widely researched ``filter /
-    verify'' indexing technique used in graph databases is a flawed design based upon a widely
-    repeated misunderstanding of what makes subgraph isomorphism hard. We explain why this technique
-    would not be beneficial, even with harder instances than those currently used, when paired with
-    any reasonable modern subgraph isomorphism algorithm.
+    difficulty are based almost entirely upon unnecessarily poor performance from overly simplistic
+    subgraph isomorphism algorithms. With this in mind, we argue that the widely researched ``filter
+    / verify'' indexing technique used in graph databases is a flawed design based upon a repeated
+    misunderstanding of what makes subgraph isomorphism hard. We explain why this technique would
+    not be beneficial, even with harder instances than those currently used, when paired with any
+    reasonable modern subgraph isomorphism algorithm.
 \end{abstract}
@@
 not adjacent in $G$. We write $G(n, p)$ for an Erd\H{o}s-R\'enyi random graph with $n$ vertices, and
-an edge between each distinct pair of vertices with independent probability $p$.
-
-A graph may have labels on its vertices, in which case we write $\ell(v)$ for the label of a given
-vertex $v$. We write $G(n, p, l)$ for the graph $G(n, p)$ with vertex labels selected uniformly at
-random from a set of $l$ different labels.
+an edge between each distinct pair of vertices with independent probability $p$.  A graph may have
+labels on its vertices, in which case we write $\ell(v)$ for the label of a given vertex $v$.
 
 A \emph{non-induced subgraph isomorphism} from a graph $P$ (called the \emph{pattern}) to a graph
-$T$ (called the \emph{target}) is an injective mapping from $\operatorname{V}(P)$ to
+$T$ (called the \emph{target}) is an injective mapping $i$ from $\operatorname{V}(P)$ to
 $\operatorname{V}(T)$ which preserves adjacency---that is, for every adjacent $v$ and $w$ in
@@
 easier again. This behaviour is largely solver-independent, although VF2 has a larger hard region
-than Glasgow or LAD. Thus, although we have moved away from a single control parameter, we still
-observe the easy-hard-easy pattern seen in many other \NP-complete problems.
+than Glasgow or LAD (and the observant reader may notice that VF2 finds some instances extremely
+hard when the target is empty but the pattern is not---this turns out to be extremely important, and
+we return to it in \cref{section:labelled,section:filterverify}). Thus, although we have moved away
+from a single control parameter, we still observe the easy-hard-easy pattern seen in many other
+\NP-complete problems.
 
@@
 
-\subsection{Flawed Instances}
+\subsection{Flawed and Obviously Unsatisfiable Instances}
 
@@
 
-Two adjacent differently-coloured vertices.
-
-VF2 does not detect the case that there are insufficient, or even no, labels.
+VF2 does not detect the case that there are insufficient labels. Also empty pattern but
+not empty target.
 
@@
 cannot immediately detect unsatisfiability in simple cases like there being two red vertices in the
-pattern but only one in the target.
+pattern but only one in the target\footnote{?? expand on: Although interestingly, this algorithm
+effectively does forward-checking and decent value ordering heuristics before these concepts
+appeared in the constraints literature.}.
 
@@
 only one. This kind of flaw \emph{should} be picked up at the top of search by an all-different
-propagator, but such techniques are not used in graph databases systems, let alone the more powerful
-degree- and difference-based propagation rules introduced LAD \cite{Solnon:2010} or the
-distance-based reasoning in more recent work \cite{Audemard:2014,McCreesh:2015}.
-
-Indeed, as \citet{DBLP:journals/pvldb/KatsarouNT15} note, some variation on VF2
-\citep{Cordella:2004} is the usual matching algorithm of choice, although
-\citeS{DBLP:journals/jacm/Ullmann76}'s algorithm is sometimes chosen. Other approaches have been
-considered: for example, \citet{DBLP:journals/pvldb/ShangZLY08} propose an algorithm which makes use
-of the frequency of various features to guide search.  \citet{DBLP:journals/pvldb/LeeHKL12}
-determine experimentally that this technique tends to be very effective, even on families of graphs
-for which it was not designed.
+propagator \citep{DBLP:conf/aaai/Regin94}.  However, as \citet{DBLP:journals/pvldb/KatsarouNT15}
+note, some variation on VF2 \citep{Cordella:2004} is the usual matching algorithm of choice for
+graph database systems, although \citeS{DBLP:journals/jacm/Ullmann76}'s algorithm is sometimes
+chosen. Other approaches have been considered: for example, \citet{DBLP:journals/pvldb/ShangZLY08}
+propose an algorithm which makes use of the frequency of various features to guide search.
+\citet{DBLP:journals/pvldb/LeeHKL12} determine experimentally that this technique tends to be very
+effective, even on families of graphs for which it was not designed.
 
@@
 
-We do not present the ultimate big data subgraph matching algorithm.
+?? The point of this paper is not to present the ultimate big data subgraph matching algorithm.
 
-Read outside the field. AI has a lot to offer.
+Read outside the field. AI has a lot to offer. Much more focus on the matching algorithm.
 
-Much more focus on the matching algorithm.
-
-Optimisations for trivial cases. Lazy forward checking, cheap alternatives to heuristics and
+?? Optimisations for trivial cases. Lazy forward checking, cheap alternatives to heuristics and
 all-different. Presolving.
 
-Invariants on labels used to inform algorithm design and to provide filtering during search.
+?? Invariant which are effective on labels used to inform algorithm design and to provide filtering
+during search. Minimum unsat core. Indexing can be used to enhance propagation.
 
-Indexing can be used to enhance propagation.
-
-On the other hand, our key message is not just saying use a better algorithm. Need a full
+?? On the other hand, our key message is not just saying use a better algorithm. Need a full
 understanding of every aspect of the system, and cannot treat subgraph matching as a black box.
@@
 
-?? Stuff about graph databases.
-
 It is worth noting that this technique does \emph{not} give a way of generating hard instances for
@@
 for independent pairs of randomly generated instances to give interesting behaviour.
+
+?? This is not purely of theoretical interest. Stuff about graph databases.
 

### 2016-09-23 code/dimacs.cc
Some code

@@ -1,0 +1,65 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "dimacs.hh"
+#include "graph.hh"
+#include <boost/regex.hpp>
+#include <fstream>
+
+auto read_dimacs(const std::string & filename) -> Graph
+{
+    Graph result;
+
+    std::ifstream infile{ filename };
+    if (! infile)
+        throw GraphFileError{ filename, "unable to open file" };
+
+    std::string line;
+    while (std::getline(infile, line)) {
+        if (line.empty())
+            continue;
+
+        /* Lines are comments, a problem description (contains the number of
+         * vertices), or an edge. */
+        static const boost::regex
+            comment{ R"(c(\s.*)?)" },
+            problem{ R"(p\s+(edge|col)\s+(\d+)\s+(\d+)?\s*)" },
+            label{ R"(l\s+(\d+)\s+(\d+)\s*)" },
+            edge{ R"(e\s+(\d+)\s+(\d+)\s*)" };
+
+        boost::smatch match;
+        if (regex_match(line, match, comment)) {
+            /* Comment, ignore */
+        }
+        else if (regex_match(line, match, problem)) {
+            /* Problem. Specifies the size of the graph. Must happen exactly
+             * once. */
+            if (0 != result.size())
+                throw GraphFileError{ filename, "multiple 'p' lines encountered" };
+            result.resize(std::stoi(match.str(2)));
+        }
+        else if (regex_match(line, match, edge)) {
+            /* An edge. DIMACS files are 1-indexed. We assume we've already had
+             * a problem line (if not our size will be 0, so we'll throw). */
+            int a{ std::stoi(match.str(1)) }, b{ std::stoi(match.str(2)) };
+            if (0 == a || 0 == b || a > int(result.size()) || b > int(result.size()))
+                throw GraphFileError{ filename, "line '" + line + "' edge index out of bounds" };
+            else if (a == b)
+                throw GraphFileError{ filename, "line '" + line + "' contains a loop" };
+            result.add_edge(a - 1, b - 1);
+        }
+        else if (regex_match(line, match, label)) {
+            int a{ std::stoi(match.str(1)) }, l{ std::stoi(match.str(2)) };
+            if (0 == a || a > int(result.size()))
+                throw GraphFileError{ filename, "line '" + line + "' label index out of bounds" };
+            result.set_label(a - 1, l);
+        }
+        else
[…]

### 2016-09-23 code/graph.cc

@@ -1,0 +1,67 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "graph.hh"
+#include <algorithm>
+
+GraphFileError::GraphFileError(const std::string & filename, const std::string & message) throw () :
+    _what("Error reading graph file '" + filename + "': " + message)
+{
+}
+
+auto GraphFileError::what() const throw () -> const char *
+{
+    return _what.c_str();
+}
+
+Graph::Graph(unsigned size)
+{
+    if (0 != size)
+        resize(size);
+}
+
+auto Graph::_position(unsigned a, unsigned b) const -> AdjacencyMatrix::size_type
+{
+    return (a * _size) + b;
+}
+
+auto Graph::resize(unsigned size) -> void
+{
+    _size = size;
+    _adjacency.resize(size * size);
+    _labels.resize(size);
+}
+
+auto Graph::add_edge(unsigned a, unsigned b) -> void
+{
+    _adjacency[_position(a, b)] = 1;
+    _adjacency[_position(b, a)] = 1;
+}
+
+auto Graph::adjacent(unsigned a, unsigned b) const -> bool
+{
+    return _adjacency[_position(a, b)];
+}
+
+auto Graph::size() const -> unsigned
+{
+    return _size;
+}
+
+auto Graph::degree(unsigned a) const -> unsigned
+{
+    return std::count_if(
+            _adjacency.begin() + a * _size,
+            _adjacency.begin() + (a + 1) * _size,
+            [] (unsigned x) { return !!x; });
+}
+
+auto Graph::set_label(unsigned a, unsigned l) -> void
+{
+    _labels[a] = l;
+}
+
+auto Graph::label(unsigned a) const -> unsigned
+{
+    return _labels[a];
+}
+

### 2016-09-23 code/sequential.cc

@@ -1,0 +1,744 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "sequential.hh"
+
+#include <algorithm>
+#include <chrono>
+#include <functional>
+#include <limits>
+#include <list>
+#include <map>
+#include <numeric>
+#include <tuple>
+#include <utility>
+#include <vector>
+
+#include <iostream>
+
+#include <boost/dynamic_bitset.hpp>
+
+using std::find_if;
+using std::get;
+using std::greater;
+using std::list;
+using std::make_pair;
+using std::map;
+using std::numeric_limits;
+using std::pair;
+using std::to_string;
+using std::tuple;
+using std::unique;
+using std::vector;
+
+using std::chrono::duration_cast;
+using std::chrono::milliseconds;
+using std::chrono::steady_clock;
+
+using std::cerr;
+using std::endl;
+
+namespace
+{
+    /// We'll use an array of unsigned long longs to represent our bits.
+    using BitWord = unsigned long long;
+
+    /// Number of bits per word.
+    static const constexpr int bits_per_word = sizeof(BitWord) * 8;
+
+    /**
+     * A bitset with a fixed maximum size. This only provides the operations
+     * we actually use in the bitset algorithms: it's more readable this way
+     * than doing all the bit voodoo inline.
+     *
+     * Indices start at 0.
+     */
+    template <unsigned words_>
+    class FixedBitSet
+    {
+        private:
+            using Bits = std::array<BitWord, words_>;
+
+            Bits _bits = {{ }};
+
+        public:
+            FixedBitSet() = default;
+
+            FixedBitSet(unsigned size)
+            {
+                assert(size <= words_ * bits_per_word);
+            }
+
+            FixedBitSet(const FixedBitSet &) = default;
+
+            FixedBitSet & operator= (const FixedBitSet &) = default;
+
+            /**
+             * Set a given bit 'on'.
+             */
+            auto set(int a) -> void
+            {
+                // The 1 does have to be of type BitWord. If we just specify a
+                // literal, it ends up being an int, and it isn't converted
+                // upwards until after the shift is done.
+                _bits[a / bits_per_word] |= (BitWord{ 1 } << (a % bits_per_word));
+            }
+
+            /**
+             * Set a given bit 'off'.
+             */
+            auto reset(int a) -> void
+            {
+                _bits[a / bits_per_word] &= ~(BitWord{ 1 } << (a % bits_per_word));
+            }
+
[…]

### 2016-09-23 code/solve_subgraph_isomorphism.cc

@@ -1,0 +1,276 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "lad.hh"
+#include "vf.hh"
+#include "dimacs.hh"
+#include "sequential.hh"
+
+#include <boost/program_options.hpp>
+
+#include <iostream>
+#include <exception>
+#include <cstdlib>
+#include <chrono>
+#include <thread>
+#include <mutex>
+#include <condition_variable>
+
+namespace po = boost::program_options;
+
+using std::chrono::steady_clock;
+using std::chrono::duration_cast;
+using std::chrono::milliseconds;
+
+/* Helper: return a function that runs the specified algorithm, dealing
+ * with timing information and timeouts. */
+template <typename Result_, typename Params_, typename Data_>
+auto run_this_wrapped(const std::function<Result_ (const Data_ &, const Params_ &)> & func)
+    -> std::function<Result_ (const Data_ &, Params_ &, bool &, int)>
+{
+    return [func] (const Data_ & data, Params_ & params, bool & aborted, int timeout) -> Result_ {
+        /* For a timeout, we use a thread and a timed CV. We also wake the
+         * CV up if we're done, so the timeout thread can terminate. */
+        std::thread timeout_thread;
+        std::mutex timeout_mutex;
+        std::condition_variable timeout_cv;
+        std::atomic<bool> abort;
+        abort.store(false);
+        params.abort = &abort;
+        if (0 != timeout) {
+            timeout_thread = std::thread([&] {
+                    auto abort_time = std::chrono::steady_clock::now() + std::chrono::seconds(timeout);
+                    {
+                        /* Sleep until either we've reached the time limit,
+                         * or we've finished all the work. */
+                        std::unique_lock<std::mutex> guard(timeout_mutex);
+                        while (! abort.load()) {
+                            if (std::cv_status::timeout == timeout_cv.wait_until(guard, abort_time)) {
+                                /* We've woken up, and it's due to a timeout. */
+                                aborted = true;
+                                break;
+                            }
+                        }
+                    }
+                    abort.store(true);
+                    });
+        }
+
+        /* Start the clock */
+        params.start_time = std::chrono::steady_clock::now();
+
+        try {
+            auto result = func(data, params);
+
+            /* Clean up the timeout thread */
+            if (timeout_thread.joinable()) {
+                {
+                    std::unique_lock<std::mutex> guard(timeout_mutex);
+                    abort.store(true);
+                    timeout_cv.notify_all();
+                }
[…]

### 2016-09-23 vflib/solve_vf.cc
Some VFLib code

@@ -1,0 +1,215 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "argraph.h"
+#include "argedit.h"
+#include "argloader.h"
+#include "match.h"
+#include "vf2_mono_state.h"
+
+#include <fstream>
+#include <iostream>
+#include <thread>
+#include <condition_variable>
+#include <mutex>
+#include <atomic>
+#include <cstdlib>
+#include <chrono>
+#include <vector>
+
+#include <boost/regex.hpp>
+
+using std::chrono::milliseconds;
+using std::chrono::steady_clock;
+using std::chrono::duration_cast;
+
+class GraphFileError :
+    public std::exception
+{
+    private:
+        std::string _what;
+
+    public:
+        GraphFileError(const std::string & filename, const std::string & message) throw ();
+
+        auto what() const throw () -> const char *;
+};
+
+GraphFileError::GraphFileError(const std::string & filename, const std::string & message) throw () :
+    _what("Error reading graph file '" + filename + "': " + message)
+{
+}
+
+auto GraphFileError::what() const throw () -> const char *
+{
+    return _what.c_str();
+}
+
+namespace
+{
+    unsigned long long hacky_global_nodes = 0;
+
+    void load(const std::string & filename, ARGEdit & ed, std::vector<int> & attrs)
+    {
+        std::ifstream infile{ filename };
+        if (! infile)
+            throw GraphFileError{ filename, "unable to open file" };
+
+        std::string line;
+        while (std::getline(infile, line)) {
+            if (line.empty())
+                continue;
+
+            /* Lines are comments, a problem description (contains the number of
+             * vertices), or an edge. */
+            static const boost::regex
+                comment{ R"(c(\s.*)?)" },
+                problem{ R"(p\s+(edge|col)\s+(\d+)\s+(\d+)?\s*)" },
+                label{ R"(l\s+(\d+)\s+(\d+)\s*)" },
+                edge{ R"(e\s+(\d+)\s+(\d+)\s*)" };
+
+            boost::smatch match;
+            if (regex_match(line, match, comment)) {
+                /* Comment, ignore */
+            }
+            else if (regex_match(line, match, problem)) {
+                /* Problem. Specifies the size of the graph. Must happen exactly
+                 * once. */
+                auto nv = std::stoi(match.str(2));
+                attrs.resize(nv);
+                for (int i = 0 ; i < nv ; ++i)
[…]

### 2016-09-24 paper/paper.tex
More ideas

@@
 aim is to reduce I/O costs, or to reduce the number of queries which must be tested, and the
-experiments do not answer whether the indexing is effective. (?? This isn't the first paper on this,
-but before this seem to be commercial applications, in particular a system called Daylight. Maybe
-more digging is required.)
+experiments do not answer whether the indexing is effective. However, the work was influenced by a commercial
+graph database system, whose documentation \cite[section 7.1]{Daylight} states that indexing is
+used to minimise disk accesses.
 
@@
 
-Read outside the field. AI has a lot to offer. Much more focus on the matching algorithm.
+?? They need to read outside the field. AI has a lot to offer. There needs to be much more focus on
+the matching algorithm, and how it interacts with the system as a whole.
 
-?? Optimisations for trivial cases. Lazy forward checking, cheap alternatives to heuristics and
-all-different. Presolving.
+?? We need optimisations for trivial cases. Lazy forward checking, cheap alternatives to heuristics and
+all-different. Presolving. Huge domains.
+
+?? Degrade to clique.
 
 ?? Invariant which are effective on labels used to inform algorithm design and to provide filtering
-during search. Minimum unsat core. Indexing can be used to enhance propagation.
+during search. Look at minimum unsat core. If we are prepared to spend hours, indexing-like system
+can be used to enhance propagation, for example by pre-computing degree sequences and paths on
+target graphs. This is stronger, because, say, counting red vertices and counting degrees are
+automatically combined.
 

### 2016-09-25 line/noninduced-line.rb
Code to draw a line

@@ -1,0 +1,43 @@
+#!/usr/bin/ruby
+# vim: set sw=4 sts=4 et tw=80 :
+
+require 'bigdecimal'
+
+ps = ARGV[0].to_i
+ts = ARGV[1].to_i
+vl = ARGV[2].to_i
+
+0.upto(500) do | x |
+    pd = BigDecimal.new(x) / 500;
+
+    tdupper = 500
+    tdlower = 0
+    expsols = 0
+
+    0.upto(100) do | y |
+        td = BigDecimal.new(tdupper + tdlower) / 1000;
+
+        cliqueedges = BigDecimal.new(ps * (ps - 1) / 2);
+        nedges = pd * cliqueedges;
+        nnonedges = (1 - pd) * cliqueedges;
+
+        edgesokprob = td ** nedges;
+
+        solprob = edgesokprob;
+
+        nstates = Math.gamma((ts / vl) + 1) / Math.gamma((ts / vl - ps / vl + 1))
+        nstates = nstates ** vl;
+
+        expsols = nstates * solprob;
+
+        if (expsols < 1)
+            tdlower = (tdupper + tdlower) / 2
+        elsif (expsols > 1)
+            tdupper = (tdupper + tdlower) / 2
+        else
+            break
+        end
+    end
+
+    puts(pd.to_f.to_s + " " + ((tdupper + tdlower) / 1000.0).to_s + " " + expsols.to_s)
+end

### 2016-09-25 paper/paper.tex
Some more things

@@
 target (y-axis) edge probabilities. Each axis runs over 101 edge probabilities, from 0 to 1 in steps
-of 0.01, except for the VF2 row which uses steps of 0.02. For each of these points, we generate ten
-random instances. The colour denotes the proportion of these instances which were found to be
-satisfiable.  Inside the orange region, at the bottom right of each plot, every instance is
-unsatisfiable---here we are trying to find a dense pattern in a sparse target. In the purple region,
-at the top left, every instance is satisfiable---we are looking for a sparse pattern in a dense
-target (which is easy, since we only have to preserve adjacency, not non-adjacency). The white band
-between the regions shows the location of the phase transition: here, roughly half the instances are
-satisfiable. (We discuss the black line below.)
+of 0.01. For each of these points, we generate ten random instances. The colour denotes the
+proportion of these instances which were found to be satisfiable.  Inside the orange region, at the
+bottom right of each plot, every instance is unsatisfiable---here we are trying to find a dense
+pattern in a sparse target. In the purple region, at the top left, every instance is
+satisfiable---we are looking for a sparse pattern in a dense target (which is easy, since we only
+have to preserve adjacency, not non-adjacency). The white band between the regions shows the
+location of the phase transition: here, roughly half the instances are satisfiable. (We discuss the
+black line below.)
 
@@
     \caption{On the top row, predicted and actual location of the phase transition for labelled
-    non-induced random subgraph isomorphism, with a pattern order of 20, a target order of 100,
-    varying pattern (x-axis) and target (y-axis) density, and varying numbers of labels. On the second
-    row, the average number of search nodes needed to solve an instance.}\label{figure:labels}
+    non-induced random subgraph isomorphism, with a pattern order of 20, a target order of 150,
+    varying pattern (x-axis) and target (y-axis) density, and varying numbers of labels. On
+    subsequent rows, the average number of search nodes needed to solve an instance, for three
+    different solvers.}\label{figure:labels}
 \end{figure}
@@
 
-\subsection{Flawed and Obviously Unsatisfiable Instances}
+\subsection{What's Wrong with VF2?}
 
@@
 
-\subsection{An Algorithm which Requires No Filtering}
-
 These beliefs are not entirely unfounded: none of the matching algorithms considered in these papers
@@
 graph database systems, although \citeS{DBLP:journals/jacm/Ullmann76}'s algorithm is sometimes
-chosen. Other approaches have been considered: for example, \citet{DBLP:journals/pvldb/ShangZLY08}
+chosen. (Other approaches have been considered: for example, \citet{DBLP:journals/pvldb/ShangZLY08}
 propose an algorithm which makes use of the frequency of various features to guide search.
 \citet{DBLP:journals/pvldb/LeeHKL12} determine experimentally that this technique tends to be very
-effective, even on families of graphs for which it was not designed.
+effective, even on families of graphs for which it was not designed.) We believe it would be
+unlikely to cause too much astonishment if we suggested that a better matching algorithm could be
+dropped in as a black box replacement in graph databases systems to improve their performance.
+However, it appears not to have occurred to the graph databases community that a better subgraph
+isomorphism algorithm could invalidate the premise underlying the entire filter / verify approach.
+Eliminating this filtering step would not merely be a matter of simplicity: indices take ??  hours
+to build.
 
-Thus we believe it would be unlikely to cause too much astonishment if we suggested that a better
-matching algorithm could be dropped in as a black box replacement in graph databases systems to
-improve their performance. However, it appears not to have occurred to the graph databases community
-that a better subgraph isomorphism algorithm could invalidate the premise underlying the entire
-filter / verify approach. Eliminating this filtering step would not merely be a matter of
-simplicity: indices take ??  hours to build.
+\subsection{An Algorithm which Requires No Filtering}
 
@@
 
-?? The point of this paper is not to present the ultimate big data subgraph matching algorithm.
+?? Presenting the ultimate big data subgraph matching algorithm is outwith the scope of this paper.
 
@@
 ?? We need optimisations for trivial cases. Lazy forward checking, cheap alternatives to heuristics and
-all-different. Presolving. Huge domains.
+all-different. Presolving. Huge domains. Ongoing use of VF2 and micro vs millisecond benchmarks.
 
@@
 understanding of every aspect of the system, and cannot treat subgraph matching as a black box.
+
+?? Use of VF2 in other areas.
 
@@
 
+\clearpage
+\appendix
+
+\section{Reproducing These Experiments}
+
 \end{document}

### 2016-09-25 lad/lad_random_subgraph_isomorphism.cc
Code for LAD

@@ -1,0 +1,132 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include <boost/program_options.hpp>
+
+#include <iostream>
+#include <exception>
+#include <cstdlib>
+#include <vector>
+#include <random>
+
+namespace po = boost::program_options;
+
+auto create_random_graph(int size, double density, int seed, bool complement, int labels) -> std::pair<std::vector<std::vector<uint8_t>>, std::vector<unsigned> >
+{
+    std::mt19937 rand;
+    rand.seed(seed);
+    std::uniform_int_distribution<int> label_dist(0, labels - 1);
+    std::uniform_real_distribution<double> dist(0.0, 1.0);
+
+    std::vector<unsigned> result_labels(size, 0);
+
+    if (labels != 0)
+        for (int e = 0 ; e < size ; ++e)
+            result_labels[e] = label_dist(rand);
+
+    std::vector<std::vector<uint8_t>> result(size, std::vector<uint8_t>(size, 0));
+
+    for (int e = 0 ; e < size ; ++e) {
+        for (int f = e + 1 ; f < size ; ++f) {
+            if (dist(rand) <= density) {
+                result[e][f] = complement ? 0 : 1;
+                result[f][e] = complement ? 0 : 1;
+            } else {
+                result[e][f] = complement ? 1 : 0;
+                result[f][e] = complement ? 1 : 0;
+            }
+        }
+    }
+
+    return { result, result_labels };
+}
+
+auto var(int a, int b, int target_size) -> int
+{
+    return a * target_size + b + 1;
+}
+
+auto main(int argc, char * argv[]) -> int
+{
+    try {
+        po::options_description display_options{ "Program options" };
+        display_options.add_options()
+            ("help",                                  "Display help information")
+            ;
+
+        po::options_description all_options{ "All options" };
+        all_options.add_options()
+            ("size",    po::value<int>(),     "Number of vertices")
+            ("density", po::value<double>(),  "Density")
+            ("labels",  po::value<int>(),     "Number of labels")
+            ("seed",    po::value<int>(),     "Seed")
+            ("complement",                    "Take the complement")
+            ;
+
+        all_options.add(display_options);
+
+        po::positional_options_description positional_options;
+        positional_options
+            .add("size", 1)
+            .add("density", 1)
+            .add("labels", 1)
+            .add("seed", 1)
+            ;
+
+        po::variables_map options_vars;
[…]

### 2016-09-26 paper/paper.tex
Reword to keep Craig happy

@@
 \begin{abstract}
-    Despte being \NP-complete, modern subgraph isomorphism algorithms can operate comfortably on
+    Despite subgraph isomorphism being \NP-complete, modern algorithms can operate comfortably on
     real-world graphs with thousands of vertices. However, they cannot handle arbitrary instances of

### 2016-09-27 paper/paper.tex
Works on sat too

@@
 graph database systems, although \citeS{DBLP:journals/jacm/Ullmann76}'s algorithm is sometimes
-chosen. (Other approaches have been considered: for example, \citet{DBLP:journals/pvldb/ShangZLY08}
+chosen. Other approaches have been considered: for example, \citet{DBLP:journals/pvldb/ShangZLY08}
 propose an algorithm which makes use of the frequency of various features to guide search.
@@
 isomorphism algorithm could invalidate the premise underlying the entire filter / verify approach.
-Eliminating this filtering step would not merely be a matter of simplicity: indices take ??  hours
+
+?? Eliminating this filtering step would not merely be a matter of simplicity: indices take ??  hours
 to build.
+
+?? Additionally, this stuff works on satisfiable instances too.
 

### 2016-09-27 paper/paper.tex
Another one

@@
 \citet{DBLP:journals/pvldb/LeeHKL12} determine experimentally that this technique tends to be very
-effective, even on families of graphs for which it was not designed.) We believe it would be
-unlikely to cause too much astonishment if we suggested that a better matching algorithm could be
-dropped in as a black box replacement in graph databases systems to improve their performance.
-However, it appears not to have occurred to the graph databases community that a better subgraph
-isomorphism algorithm could invalidate the premise underlying the entire filter / verify approach.
+effective, even on families of graphs for which it was not designed. However, these comparisons have
+either ignored strong propagation algorithms, or in one case misunderstood them to be part of a
+filtering step rather than matching algorithms \citep{DBLP:conf/icdm/AlmasriGF14}.
 
-?? Eliminating this filtering step would not merely be a matter of simplicity: indices take ??  hours
-to build.
-
-?? Additionally, this stuff works on satisfiable instances too.
+We believe it would be unlikely to cause too much astonishment if we suggested that a better
+matching algorithm could be dropped in as a black box replacement in graph databases systems to
+improve their performance.  However, it appears not to have occurred to the graph databases
+community that a better subgraph isomorphism algorithm could invalidate the premise underlying the
+entire filter / verify approach.  Eliminating this filtering step would not merely be a matter of
+simplicity: indices take ??  hours to build.
 

### 2016-09-28 paper/paper.tex
More data

--- 2016-09-27 paper/paper.tex
@@
 \cite{Giugno:2013,Carletti:2015}, and pattern recognition \citep{Conte:2004}.  In particular,
-subgraph isomorphism is one of the key components in supporting complex queries in graph
-databases---typically, these systems store a fixed set of target graphs, and for a sequence of
-pattern queries, they must return every target graph which contains that pattern.  One common
-approach to this problem is to combine a subgraph isomorphism algorithm with an indexing system in a
-so-called ``filter / verify'' model, where invariants are used to attempt to pre-exclude
+a labelled version of subgraph isomorphism is one of the key components in supporting complex
+queries in graph databases---typically, these systems store a fixed set of target graphs, and for a
+sequence of pattern queries, they must return every target graph which contains that pattern.  One
+common approach to this problem is to combine a subgraph isomorphism algorithm with an indexing
+system in a so-called ``filter / verify'' model, where invariants are used to attempt to pre-exclude
 unsatisfiable instances to avoid the cost of a subgraph isomorphism call. In this context, the terms
@@
 
-What goes on here?
+?? What goes on here? With labels, there are two regions where VF2 often finds instances really
+hard, but Glasgow and LAD do not. Looking at the data more closely, although the average runtime is
+high, there is a wide distribution of runtimes in these areas, and VF2 does sometimes find these
+instances trivial too.
 
-Min unsat core.
+?? Give examples here: VF2 does not detect simple things like there being two red vertices in the
+pattern but one in the target, or even no red vertices in the target. In contrast, Glasgow and LAD
+do. Explain thrashing.
 
-Insufficient vertices.
+?? Explain what's going on in the satisfiable region. This is a bit more subtle: is it to do with
+neighbours, rather than a root node property? Would be nice if we could find a representative
+example.
 
-VF2 does not detect the case that there are insufficient labels. Also empty pattern but
-not empty target.
+?? Look more at size of min unsat core in these areas?
 
@@
 unsatisfiable instances from consideration, without performing a subgraph isomorphism test.
+
+?? Give a quick example here of how an index would eliminate the two red vertices in the pattern but
+one red vertex in the target example.
 
@@
 
-?? Additionally, this stuff works on satisfiable instances too.
+?? Additionally, this stuff works on satisfiable instances too. Indexing addresses some of the
+obviously-unsatisfiable cases that VF2 struggles with, but does nothing on the satisfiable
+instances.
 
@@
 
-We build up a simple algorithm. Start with VF2, adapt to track domains, smallest domain first, all
-different. Show its effectiveness at eliminating unsatisfiable instances, and give its total
-performance.  None of the steps in this algorithm are new. Our aim in this section is to show which
-aspects of modern algorithm design are critical, and that an indexless sequential scan is entirely
-plausible with a decent algorithm.
+?? We build up a simple algorithm. Use smallest domain first, all different, NDS, possibly distance
+2 filtering. Show its effectiveness at eliminating unsatisfiable instances from graph database test
+sets, and give its total performance.  None of the steps in this algorithm are new. Our aim in this
+section is to show which aspects of modern algorithm design are critical, and that an indexless
+sequential scan is entirely plausible with a decent algorithm.
+
+?? To tie this into earlier parts of the paper, we also show what happens if we use the VF2
+heuristic in this setting, rather than smallest domains first.
+
+?? This stuff eliminates all the spuriously hard unsatisfiable instances along the bottom of the
+graph, but also helps on the satisfiable instances that an indexing system cannot touch.
 
@@
 
+?? Clearly lots of scope for cooperation between AI and graph databases. AI is not just good for
+better algorithms, but the design of entire systems.
+
 ?? Presenting the ultimate big data subgraph matching algorithm is outwith the scope of this paper.
-
-?? They need to read outside the field. AI has a lot to offer. There needs to be much more focus on
-the matching algorithm, and how it interacts with the system as a whole.
 
@@
 all-different. Presolving. Huge domains. Ongoing use of VF2 and micro vs millisecond benchmarks.
+Precalculating NDS, distance 2 etc, not for indexing. Edge labels.
 
-?? Degrade to clique.
+?? Degrade to clique. Symmetries.
 

### 2016-09-28 paper/paper.tex
Tweak slightly

@@
 
+?? We call these \emph{spuriously hard} instances, to distinguish them from \emph{really hard}
+problems. Not the same as \emph{exceptionally hard}.
+
 ?? Look more at size of min unsat core in these areas?
@@
 
-?? Presenting the ultimate big data subgraph matching algorithm is outwith the scope of this paper.
+?? We are not presenting the ultimate big data subgraph matching algorithm.
 
-?? We need optimisations for trivial cases. Lazy forward checking, cheap alternatives to heuristics and
-all-different. Presolving. Huge domains. Ongoing use of VF2 and micro vs millisecond benchmarks.
-Precalculating NDS, distance 2 etc, not for indexing. Edge labels.
+?? We need optimisations for trivial cases. Lazy forward checking, cheap alternatives to heuristics
+and all-different. Presolving. Huge domains.  Precalculating NDS, distance 2 etc, not for indexing.
+Edge labels.
 
@@
 
+\section{Spuriously Hard Instances Outwith Graph Databases}
+
 ?? Use of VF2 in other areas.
+
+?? Need for better benchmark instances, not running millisecond versus microsecond benchmarks.
 

### 2016-10-06 paper/paper.tex
Better here
A note
Experiment with a different style of plotting

@@
 \begin{figure}[tb]
-    \centering\includegraphics{gen-graph-non-induced.pdf}
+    \centering
+    \begin{tikzpicture}
+        \matrix {
+            \node [anchor=center] {}; &
+            \node [anchor=center] {\scriptsize $G(10, x){\rightarrowtail}G(150, y)$}; &
+            \node [anchor=center] {\scriptsize $G(20, x){\rightarrowtail}G(150, y)$}; &
+            \node [anchor=center] {\scriptsize $G(30, x){\rightarrowtail}G(150, y)$}; &
+            \\
+            \node [anchor=center] [rotate=90] {\scriptsize Satisfiable?}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-satisfiable-10.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-satisfiable-20.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-satisfiable-30.pdf}}; &
+            \\
+            \node [anchor=center] [rotate=90] {\scriptsize Glasgow}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-nodes-10.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-nodes-20.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-nodes-30.pdf}}; &
+            \\
+            \node [anchor=center] [rotate=90] {\scriptsize LAD}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-lad-non-induced-nodes-10.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-lad-non-induced-nodes-20.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-lad-non-induced-nodes-30.pdf}}; &
+            \\
+            \node [anchor=center] [rotate=90] {\scriptsize VF2}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-vf2-non-induced-nodes-10.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-vf2-non-induced-nodes-20.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-vf2-non-induced-nodes-30.pdf}}; \\
+        };
+    \end{tikzpicture}
+
     \caption{Behaviour of algorithms on the non-induced variant. For each plot, the x-axis is the
@@
 
+\subsection{Spuriously Hard Instances?}
+
+?? Use of VF2 in other areas.
+
+?? Need for better benchmark instances, not running millisecond versus microsecond benchmarks.
+
+?? However, by far the biggest problem is in graph databases. Not just lead to overly pessimistic
+conclusions, but has misdirected the design of systems. Note that these deal with sparse, non-random
+graphs.
+
 \section{Querying Graph Databases}\label{section:filterverify}
@@
 
-\section{Spuriously Hard Instances Outwith Graph Databases}
-
-?? Use of VF2 in other areas.
-
-?? Need for better benchmark instances, not running millisecond versus microsecond benchmarks.
-
 \section{Conclusion}

### 2016-10-07 paper/paper.tex
Plot better

@@
 
-% \usepackage{showframe}
+\usepackage{showframe}
 
@@
 \begin{figure}[p]
-    \centering\includegraphics{gen-graph-induced.pdf}
+    \centering
+    \begin{tikzpicture}
+        \matrix {
+            \node [anchor=center] {}; &
+            \node [anchor=center] {\scriptsize $G(10, x)$}; &
+            \node [anchor=center] {\scriptsize $G(14, x)$}; &
+            \node [anchor=center] {\scriptsize $G(15, x)$}; &
+            \node [anchor=center] {\scriptsize $G(16, x)$}; &
+            \node [anchor=center] {\scriptsize $G(20, x)$}; &
+            \node [anchor=center] {\scriptsize $G(30, x)$}; &
+            \\
+            \node [anchor=center] [rotate=90] {\scriptsize Satisfiable?}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-10.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-14.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-15.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-16.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-20.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-30.pdf}}; &
+            \\
+            \node [anchor=center] [rotate=90] {\scriptsize Glasgow}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-10.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-14.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-15.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-16.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-20.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-30.pdf}}; &
+            \\
+            \node [anchor=center] [rotate=90] {\scriptsize LAD}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-lad-induced-nodes-10.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-lad-induced-nodes-14.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-lad-induced-nodes-15.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-lad-induced-nodes-16.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-lad-induced-nodes-20.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-lad-induced-nodes-30.pdf}}; &
+            \\
+            \node [anchor=center] [rotate=90] {\scriptsize VF2}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-vf2-induced-nodes-10.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-vf2-induced-nodes-14.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-vf2-induced-nodes-15.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-vf2-induced-nodes-16.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-vf2-induced-nodes-20.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-vf2-induced-nodes-30.pdf}}; \\
+            \\
+            \node [anchor=center] [rotate=90] {\scriptsize Constrainedness}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-kappa-10.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-kappa-14.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-kappa-15.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-kappa-16.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-kappa-20.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-kappa-30.pdf}}; \\
+            \\
+            \node [anchor=center] [rotate=90] {\scriptsize Complement?}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-which-10.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-which-14.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-which-15.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-which-16.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-which-20.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-which-30.pdf}}; \\
+        };
+    \end{tikzpicture}
     \caption{Behaviour of algorithms on the induced variant, shown in the style

### 2016-10-08 paper/paper.tex
More replottery

--- 2016-09-28 paper/paper.tex
@@
 \begin{figure}[tb]
-    \centering\includegraphics{gen-graph-non-induced.pdf}
+    \centering
+    \begin{tikzpicture}[every node/.style={inner sep=0pt, outer sep=0pt}]
+        \matrix {
+            \node [anchor=center] {}; &
+            \node [anchor=center] {\scriptsize $G(10, x){\rightarrowtail}G(150, y)$}; &
+            \node [anchor=center] {\scriptsize $G(20, x){\rightarrowtail}G(150, y)$}; &
+            \node [anchor=center] {\scriptsize $G(30, x){\rightarrowtail}G(150, y)$}; &
+            \\
+            \node [anchor=center] [rotate=90] {\scriptsize Satisfiable?}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-satisfiable-10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-satisfiable-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-satisfiable-30-150.pdf}}; &
+            \\
+            \node [anchor=center] [rotate=90] {\scriptsize Glasgow}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-nodes-10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-nodes-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-nodes-30-150.pdf}}; &
+            \\
+            \node [anchor=center] [rotate=90] {\scriptsize LAD}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-lad-non-induced-nodes-10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-lad-non-induced-nodes-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-lad-non-induced-nodes-30-150.pdf}}; &
+            \\
+            \node [anchor=center] [rotate=90] {\scriptsize VF2}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-vf2-non-induced-nodes-10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-vf2-non-induced-nodes-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-vf2-non-induced-nodes-30-150.pdf}}; \\
+        };
+    \end{tikzpicture}
+
     \caption{Behaviour of algorithms on the non-induced variant. For each plot, the x-axis is the
@@
 \begin{figure}[p]
-    \centering\includegraphics{gen-graph-induced.pdf}
+    \centering
+    \begin{tikzpicture}[every node/.style={inner sep=0pt, outer sep=0pt}]
+        \matrix {
+            \node [anchor=center] {}; &
+            \node [anchor=center] {\scriptsize $G(10, x)$}; &
+            \node [anchor=center] {\scriptsize $G(14, x)$}; &
+            \node [anchor=center] {\scriptsize $G(15, x)$}; &
+            \node [anchor=center] {\scriptsize $G(16, x)$}; &
+            \node [anchor=center] {\scriptsize $G(20, x)$}; &
+            \node [anchor=center] {\scriptsize $G(30, x)$}; &
+            \\
+            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Satisfiable?}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-14-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-15-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-16-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-30-150.pdf}}; &
+            \\
+            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Glasgow}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-14-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-15-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-16-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-30-150.pdf}}; &
+            \\
+            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize LAD}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-lad-induced-nodes-10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-lad-induced-nodes-14-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-lad-induced-nodes-15-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-lad-induced-nodes-16-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-lad-induced-nodes-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-lad-induced-nodes-30-150.pdf}}; &
+            \\
+            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize VF2}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-vf2-induced-nodes-10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-vf2-induced-nodes-14-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-vf2-induced-nodes-15-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-vf2-induced-nodes-16-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-vf2-induced-nodes-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-vf2-induced-nodes-30-150.pdf}}; \\
+            \\
+            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Constrainedness}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-kappa-10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-kappa-14-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-kappa-15-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-kappa-16-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-kappa-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-kappa-30-150.pdf}}; \\
+            \\
+            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Complement?}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-which-10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-which-14-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-which-15-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-which-16-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-which-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-which-30-150.pdf}}; \\
+        };
+    \end{tikzpicture}
     \caption{Behaviour of algorithms on the induced variant, shown in the style
@@
 \begin{figure}[tb]
-    \centering\includegraphics{gen-graph-sat.pdf}
+    \centering
+    \begin{tikzpicture}[every node/.style={inner sep=0pt, outer sep=0pt}]
+        \matrix {
+            \node [anchor=center] {}; &
+            \node [anchor=center] {\scriptsize $G(10, x)$}; &
+            \node [anchor=center] {\scriptsize $G(12, x)$}; &
+            \node [anchor=center] {\scriptsize $G(14, x)$}; &
+            \node [anchor=center] {\scriptsize $G(16, x)$}; &
+            \node [anchor=center] {\scriptsize $G(18, x)$}; &
+            \node [anchor=center] {\scriptsize $G(25, x)$}; &
+            \\
+            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Satisfiable?}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-10-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-12-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-14-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-16-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-18-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-25-75.pdf}}; &
+            \\
+            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Glasgow}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-10-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-12-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-14-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-16-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-18-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-25-75.pdf}}; &
+            \\
+            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Clasp (PB)}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-clasp-induced-nodes-10-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-clasp-induced-nodes-12-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-clasp-induced-nodes-14-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-clasp-induced-nodes-16-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-clasp-induced-nodes-18-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-clasp-induced-nodes-25-75.pdf}}; &
+            \\
+            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Glucose (SAT)}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-glucose-induced-nodes-10-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-glucose-induced-nodes-12-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-glucose-induced-nodes-14-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-glucose-induced-nodes-16-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-glucose-induced-nodes-18-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-glucose-induced-nodes-25-75.pdf}}; &
+            \\
+            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize BBMC (Clique)}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-clique-induced-nodes-10-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-clique-induced-nodes-12-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-clique-induced-nodes-14-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-clique-induced-nodes-16-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-clique-induced-nodes-18-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-clique-induced-nodes-25-75.pdf}}; &
+            \\
+            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Gurobi (MIP)}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-gurobi-induced-nodes-10-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-gurobi-induced-nodes-12-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-gurobi-induced-nodes-14-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-gurobi-induced-nodes-16-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-gurobi-induced-nodes-18-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-gurobi-induced-nodes-25-75.pdf}}; &
+            \\
+        };
+    \end{tikzpicture}
     \caption{Behaviour of other solvers on the induced variant on smaller graphs, shown in the style of

### 2016-10-08 paper/paper.tex
Move around a bit

@@
 
-\begin{figure}[p]
+\begin{figure}[tb]
     \centering
@@
 
+In the first four rows of \cref{figure:induced} we repeat our experiments, finding induced
+isomorphisms. With a pattern of order 10, we get two independent phase transitions: the bottom right
+half of the plots resemble the non-induced results, and the top left half is close to a mirror
+image. The central satisfiable region, which is away from either phase transition, is
+computationally easy, but instances near the phase transition are hard.
+
+For larger patterns of order 20 and 30, we have a large unsatisfiable region in the middle. Despite
+not being near either phase transition, instances in the centre remain computationally challenging.
+We also plot patterns of orders 14, 15 and 16, to show the transition between the two behaviours.
+
+\subsection{Predictions and Heuristics}
+
+To predict the location of the induced phase transition, we repeat the argument for locating the
+non-induced phase transition and additionally considering non-edges, to get an expected number of
+solutions of \[ \langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot
+{(1 - d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}} \textnormal{.} \] We plot this using black lines on
+the top row of \cref{figure:induced}---again, our prediction is accurate except for very sparse or
+very dense patterns.
+
+We might guess that degree-based heuristics would just not work for the induced problem: for any
+claim about the degree, the opposite will hold for the complement constraints. However, empirically,
+this is not the case: on the final row of \cref{figure:induced}, we show whether it is better to use
+the original pattern and target as the input to the Glasgow algorithm, or to take the complements.
+(The only steps performed by the Glagsow algorithm which differ under taking the complements are
+the degree-based heuristics.  LAD and VF2 are not symmetric in this way: LAD performs a filtering
+step using degree information, but does not consider the complement degree, and VF2 uses
+connectivity in the pattern graph.)
+
+For patterns of order 10, it is always better to try to move towards the satisfiable region: if we
+are in the bottom right diagonal half, we are best retaining the original heuristics (which move us
+towards the top left), and if we are in the top left we should use the complement instead. This
+goes against a suggestion by \citet{Walsh:1998} that switching heuristics based upon an estimate of
+the solubility of the problem may offer good performance.
+
+For larger patterns, more complex behaviour emerges. If we are in the intersection of the bottom half
+and the bottom right diagonal of the search space, we should always retain the original heuristic,
+and if we are in the intersection of the top half and the top left diagonal, we should always use
+the complements. This behaviour can be predicted by taking the partial derivatives of $\langle Sol
+\rangle$ in the $-p_d$ and $t_d$ directions.  However, when inside the remaining two eighths of the
+parameter space, the partial derivatives of $\langle Sol \rangle$ disagree on which heuristic to
+use, and using directional derivatives is not enough to resolve the problem. A close observation of
+the data suggests that the actual location of the phase transition may be involved (and perhaps
+\citeS{Walsh:1998} suggestion applies only in these conditions). In any case, $\langle Sol
+\rangle$ is insufficient to explain the observed behaviour in these two eighths of the parameter space.
+
+In practice, this is unlikely to be a problem: most real-world instances are extremely sparse and
+are usually easy, which perhaps explains the continuing popularity of VF2's connectivity-based
+heuristics \citep{Carletti:2015}. In this situation, these experiments justify reusing the
+non-induced heuristics on induced problems.
+
+\subsection{Is the Central Region Genuinely Hard?}
+
 \begin{figure}[tb]
@@
     Glasgow algorithm, the third and fourth rows show the number of decisions made by the pseudo-boolean
-and SAT solvers, and the final shows the number of search nodes used on the clique
-encoding.}\label{figure:alt}
+and SAT solvers, the fifth shows the number of search nodes used on the clique encoding, and the
+final row the MIP encoding.}\label{figure:alt}
 \end{figure}
-
-In the first four rows of \cref{figure:induced} we repeat our experiments, finding induced
-isomorphisms. With a pattern of order 10, we get two independent phase transitions: the bottom right
-half of the plots resemble the non-induced results, and the top left half is close to a mirror
-image. The central satisfiable region, which is away from either phase transition, is
-computationally easy, but instances near the phase transition are hard.
-
-For larger patterns of order 20 and 30, we have a large unsatisfiable region in the middle. Despite
-not being near either phase transition, instances in the centre remain computationally challenging.
-We also plot patterns of orders 14, 15 and 16, to show the transition between the two behaviours.
-
-\subsection{Predictions and Heuristics}
-
-To predict the location of the induced phase transition, we repeat the argument for locating the
-non-induced phase transition and additionally considering non-edges, to get an expected number of
-solutions of \[ \langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot
-{(1 - d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}} \textnormal{.} \] We plot this using black lines on
-the top row of \cref{figure:induced}---again, our prediction is accurate except for very sparse or
-very dense patterns.
-
-We might guess that degree-based heuristics would just not work for the induced problem: for any
-claim about the degree, the opposite will hold for the complement constraints. However, empirically,
-this is not the case: on the final row of \cref{figure:induced}, we show whether it is better to use
-the original pattern and target as the input to the Glasgow algorithm, or to take the complements.
-(The only steps performed by the Glagsow algorithm which differ under taking the complements are
-the degree-based heuristics.  LAD and VF2 are not symmetric in this way: LAD performs a filtering
-step using degree information, but does not consider the complement degree, and VF2 uses
-connectivity in the pattern graph.)
-
-For patterns of order 10, it is always better to try to move towards the satisfiable region: if we
-are in the bottom right diagonal half, we are best retaining the original heuristics (which move us
-towards the top left), and if we are in the top left we should use the complement instead. This
-goes against a suggestion by \citet{Walsh:1998} that switching heuristics based upon an estimate of
-the solubility of the problem may offer good performance.
-
-For larger patterns, more complex behaviour emerges. If we are in the intersection of the bottom half
-and the bottom right diagonal of the search space, we should always retain the original heuristic,
-and if we are in the intersection of the top half and the top left diagonal, we should always use
-the complements. This behaviour can be predicted by taking the partial derivatives of $\langle Sol
-\rangle$ in the $-p_d$ and $t_d$ directions.  However, when inside the remaining two eighths of the
-parameter space, the partial derivatives of $\langle Sol \rangle$ disagree on which heuristic to
-use, and using directional derivatives is not enough to resolve the problem. A close observation of
-the data suggests that the actual location of the phase transition may be involved (and perhaps
-\citeS{Walsh:1998} suggestion applies only in these conditions). In any case, $\langle Sol
-\rangle$ is insufficient to explain the observed behaviour in these two eighths of the parameter space.
-
-In practice, this is unlikely to be a problem: most real-world instances are extremely sparse and
-are usually easy, which perhaps explains the continuing popularity of VF2's connectivity-based
-heuristics \citep{Carletti:2015}. In this situation, these experiments justify reusing the
-non-induced heuristics on induced problems.
-
-\subsection{Is the Central Region Genuinely Hard?}
 
@@
 instances that are hard for the Glasgow solver remain hard for the PB solver, including instances
-inside the central region, and the easy satisfiable instances remain easy. Similar results were
-seen with the Glucose SAT solver \cite{glucose} using a direct encoding of the cardinality constraints. (We also
-implemented an integer program encoding; the Gurobi solver was only able to solve some of the
-trivial satisfiable instances, and was almost never able to prove unsatisfiability within the time
-limit.)
+inside the central region, and the easy satisfiable instances remain easy. Similar results were seen
+with the Glucose SAT solver \cite{glucose} using a direct encoding of the cardinality constraints.
+We also implemented an integer program encoding: the Gurobi solver was only able to solve some of
+the trivial satisfiable instances, and was almost never able to prove unsatisfiability within the
+time limit.
 

### 2016-10-08 paper/paper.tex
More replottering

@@
 \usepackage{csquotes}
-
-\usepackage{showframe}
 
@@
 \begin{figure}[t]
-    \centering\includegraphics{gen-graph-labels.pdf}
+    \begin{tikzpicture}[every node/.style={inner sep=0pt, outer sep=0pt}]
+        \matrix {
+            \node [anchor=center] {}; &
+            \node [anchor=center] {\scriptsize No labels}; &
+            \node [anchor=center] {\scriptsize 1 label}; &
+            \node [anchor=center] {\scriptsize 2 labels}; &
+            \node [anchor=center] {\scriptsize 5 labels}; &
+            \node [anchor=center] {\scriptsize 10 labels}; &
+            \node [anchor=center] {\scriptsize 20 labels}; &
+            \\
+            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Satisfiable?}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-satisfiable-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-satisfiable-20-l2-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-satisfiable-20-l3-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-satisfiable-20-l5-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-satisfiable-20-l10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-satisfiable-20-l20-150.pdf}}; &
+            \\
+            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Glasgow}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-nodes-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-nodes-20-l2-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-nodes-20-l3-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-nodes-20-l5-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-nodes-20-l10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-nodes-20-l20-150.pdf}}; &
+            \\
+            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize LAD}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-lad-non-induced-nodes-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-lad-non-induced-nodes-20-l2-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-lad-non-induced-nodes-20-l3-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-lad-non-induced-nodes-20-l5-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-lad-non-induced-nodes-20-l10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-lad-non-induced-nodes-20-l20-150.pdf}}; &
+            \\
+            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize VF2}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-vf2-non-induced-nodes-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-vf2-non-induced-nodes-20-l2-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-vf2-non-induced-nodes-20-l3-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-vf2-non-induced-nodes-20-l5-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-vf2-non-induced-nodes-20-l10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-vf2-non-induced-nodes-20-l20-150.pdf}}; &
+            \\
+        };
+    \end{tikzpicture}
     \caption{On the top row, predicted and actual location of the phase transition for labelled

### 2016-10-11 paper/paper.tex
Number some equations

@@
 $d_p \cdot \binom{p}{2}$ edges, so we obtain the probability of all of these edges being mapped to
-edges in the target by raising $d_t$ to this power, giving an expected number of solutions of \[
-\langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \textnormal{.} \] This
+edges in the target by raising $d_t$ to this power, giving an expected number of solutions of
+\begin{equation}
+    \langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \textnormal{.}
+\end{equation} This
 formula predicts a very sharp phase transition from $\langle Sol \rangle \ll 1$ to $\langle Sol
@@
 non-induced phase transition and additionally considering non-edges, to get an expected number of
-solutions of \[ \langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot
-{(1 - d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}} \textnormal{.} \] We plot this using black lines on
+solutions of \begin{equation} \langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot
+{(1 - d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}} \textnormal{.} \end{equation} We plot this using black lines on
 the top row of \cref{figure:induced}---again, our prediction is accurate except for very sparse or
@@
 is very small or very large are usually easy. By handling injectivity as a restriction on the size
-of the state space rather than as a constraint, we derive \[ \kappa = 1 - \frac{\log \left(
+of the state space rather than as a constraint, we derive \begin{equation} \kappa = 1 - \frac{\log \left(
 t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 - d_{t})}^{(1 - d_{p}) \cdot
-\binom{p}{2}} \right)}{\log t^{\underline{p}}}\] for induced isomorphisms, which we plot on the
+\binom{p}{2}} \right)}{\log t^{\underline{p}}} \end{equation} for induced isomorphisms, which we plot on the
 fifth row of \cref{figure:induced}. We see that constrainedness predicts that the central region
@@
 probability of all of these edges being mapped to edges in the target by raising $d_t$ to this
-power, giving an expected number of solutions of \[ \langle Sol \rangle = \left(
+power, giving an expected number of solutions of \begin{equation} \langle Sol \rangle = \left(
     \frac{\Gamma\left(\nicefrac{t}{k} + 1\right)}{\Gamma\left(\nicefrac{t}{k} - \nicefrac{p}{k} +
 1\right)} \right)^{k}  \cdot
-{d_t}^{d_p \cdot \binom{p}{2}} \textnormal{\,.} \]
+{d_t}^{d_p \cdot \binom{p}{2}} \textnormal{\,.} \end{equation}
 

### 2016-10-20 paper/paper.tex
James
Feedback from Christine

@@
        \name Patrick Prosser \email patrick.prosser@glasgow.ac.uk \\
-       \name James Trimble \\
+       \name James Trimble \thanks{This work was supported by the Engineering and Physical Sciences
+           Research Council [grant number EP/M508056/1]} \email j.trimble.1@research.gla.ac.uk \\
    \addr University of Glasgow, Glasgow, Scotland}
@@
     Despite subgraph isomorphism being \NP-complete, modern algorithms can operate comfortably on
-    real-world graphs with thousands of vertices. However, they cannot handle arbitrary instances of
+    most real-world graphs with thousands of vertices. However, they cannot handle arbitrary instances of
     this size. We show how to generate ``really hard'' random instances for subgraph isomorphism
@@
 
-    When labels are introduced, as commonly seen in graph databases, additional rich behaviour
-    arises.  We show that when only slightly better algorithms are used, all commonly used benchmark
-    suites for evaluating graph database systems become trivial, and that existing notions of
-    difficulty are based almost entirely upon unnecessarily poor performance from overly simplistic
-    subgraph isomorphism algorithms. With this in mind, we argue that the widely researched ``filter
-    / verify'' indexing technique used in graph databases is a flawed design based upon a repeated
-    misunderstanding of what makes subgraph isomorphism hard. We explain why this technique would
-    not be beneficial, even with harder instances than those currently used, when paired with any
-    reasonable modern subgraph isomorphism algorithm.
+    When labels are introduced, as commonly seen in real-world applications and in graph databases,
+    additional rich behaviour arises.  We show that when only slightly better algorithms are used,
+    all commonly used benchmark suites for evaluating graph database systems become trivial, and
+    that existing notions of difficulty are based almost entirely upon unnecessarily poor
+    performance from overly simplistic subgraph isomorphism algorithms. With this in mind, we argue
+    that the widely researched ``filter / verify'' indexing technique used in graph databases is a
+    flawed design based upon a repeated misunderstanding of what makes subgraph isomorphism hard. We
+    explain why this technique would not be beneficial, even with harder instances than those
+    currently used, when paired with any reasonable modern subgraph isomorphism algorithm.
 \end{abstract}
@@
 
-The \emph{non-induced subgraph isomorphism problem} is to find an injective mapping from a given
-pattern graph to a given target graph which preserves adjacency---in essence, we are ``finding a
-copy of'' the pattern inside the target. The \emph{induced} variant of the problem additionally
-requires that the mapping preserve non-adjacency, so there are no ``extra edges'' in the copy of the
-pattern that we find. We illustrate both variants in \cref{figure:sip}.  Although these problems
-are \NP-complete, modern practical subgraph isomorphism algorithms can handle problem instances
-with many hundreds of vertices in the pattern graph, and up to ten thousand vertices in the target
-graph \cite{Cordella:2004,Solnon:2010,Audemard:2014,McCreesh:2015}, leading to successful
-application in areas such as computer vision \cite{Damiand:2011,Solnon:2015}, biochemistry
-\cite{Giugno:2013,Carletti:2015}, and pattern recognition \citep{Conte:2004}.  In particular,
-a labelled version of subgraph isomorphism is one of the key components in supporting complex
-queries in graph databases---typically, these systems store a fixed set of target graphs, and for a
-sequence of pattern queries, they must return every target graph which contains that pattern.  One
-common approach to this problem is to combine a subgraph isomorphism algorithm with an indexing
-system in a so-called ``filter / verify'' model, where invariants are used to attempt to pre-exclude
+The \emph{non-induced subgraph isomorphism problem} is to find an injective mapping from the
+vertices of a given pattern graph to the vertices of a given target graph which preserves
+adjacency---in essence, we are ``finding a copy of'' the pattern inside the target. The
+\emph{induced} variant of the problem additionally requires that the mapping preserve non-adjacency,
+so there are no ``extra edges'' in the copy of the pattern that we find. We illustrate both variants
+in \cref{figure:sip}.  Although these problems are \NP-complete, modern practical subgraph
+isomorphism algorithms can handle problem instances with many hundreds of vertices in the pattern
+graph, and up to ten thousand vertices in the target graph
+\cite{Cordella:2004,Solnon:2010,Audemard:2014,McCreesh:2015}, leading to successful applications in
+areas such as computer vision \cite{Damiand:2011,Solnon:2015}, biochemistry
+\cite{Giugno:2013,Carletti:2015}, and pattern recognition \citep{Conte:2004}.  In particular, a
+labelled version of subgraph isomorphism is one of the key components in supporting complex queries
+in graph databases---typically, these systems store a fixed set of target graphs, and for a sequence
+of pattern queries, they must return every target graph which contains that pattern.  One common
+approach to this problem is to combine a subgraph isomorphism algorithm with an indexing system in a
+so-called ``filter / verify'' model, where invariants are used to attempt to pre-exclude
 unsatisfiable instances to avoid the cost of a subgraph isomorphism call. In this context, the terms
@@
 generate random subgraph isomorphism instances so far, meaning existing benchmark suites contain
-relatively few non-trivial unsatisfiable instances.  Also, the satisfiable instances tend to be
-computationally fairly easy, with most of the difficulty being in dealing with the size of the
-model. This has lead to bias in algorithm design, to the extent that some proposed techniques will
-\emph{only} work on satisfiable instances \citep{Battiti:2007}.
+relatively few non-trivial unsatisfiable instances (although some of the patterns in the instances
+by \citeauthor{Zampelli:2010} have had extra edges added, making them unsatisfiable). Also, the
+satisfiable instances tend to be computationally fairly easy, with most of the difficulty being in
+dealing with the size of the model. This has lead to bias in algorithm design, to the extent that
+some proposed techniques will \emph{only} work on satisfiable instances \citep{Battiti:2007}.
 
@@
 relatively small random graph is unlikely to appear in a larger random graph.)
+
+?? This needs to be reworded if we include non-ER random graphs.
 
@@
 
-When labels on vertices are introduced, as is commonly seen in graph database systems, richer
-behaviour emerges. We look at some of the datasets commonly used to test graph database systems,
-which the literature suggests are extremely hard to solve. Our theory predicts that these instances
-should be trivial, and indeed we see that the entire perceived difficulty of every dataset we
-investigate comes down to the widespread use of a particular subgraph isomorphsim algorithm which
-does not use a theoretically-strong variable ordering heuristic. Our second contribution is to show
-that the use of the right variable ordering heuristic, even without the more sophisticated recent
-advances in subgraph algorithms (which have been largely neglected outside of the artificial
-intelligence community), makes the entire graph database filter / verify paradigm unnecessary.
-Finally, we explain why filter / verify cannot be beneficial even on more challenging instances
-except if an extremely poor choice of subgraph isomorphism algorithm is made.
+When labels on vertices are introduced, as is commonly seen in real-world applications and in graph
+database systems, richer behaviour emerges. We look at some of the datasets commonly used to test
+graph database systems, which the literature suggests are extremely hard to solve. Our theory
+predicts that these instances should be trivial, and indeed we see that the entire perceived
+difficulty of every dataset we investigate comes down to the widespread use of a particular subgraph
+isomorphsim algorithm which does not use a theoretically-strong variable ordering heuristic. Our
+second contribution is to show that the use of the right variable ordering heuristic, even without
+the more sophisticated recent advances in subgraph algorithms, makes the entire graph database
+filter / verify paradigm unnecessary.  Finally, we explain why filter / verify cannot be beneficial
+even on more challenging instances except if an extremely poor choice of subgraph isomorphism
+algorithm is made.
 
 \subsection{Definitions}
+
+?? Check whether any of the real-world graphs we look at have loops.
 
@@
 $T$ (called the \emph{target}) is an injective mapping $i$ from $\operatorname{V}(P)$ to
-$\operatorname{V}(T)$ which preserves adjacency---that is, for every adjacent $v$ and $w$ in
-$\operatorname{V}(P)$, the vertices $i(v)$ and $i(w)$ are adjacent in $T$. If vertices are labelled,
-the mapping must preserve labels, so $\ell(v) = \ell(i(v))$. An \emph{induced subgraph isomorphism}
-additionally preserves non-adjacency---that is, if $v$ and $w$ are not adjacent in $P$, then $i(v)$
-and $i(w)$ must not be adjacent in $T$. We use the notation $i : P \rightarrowtail T$ for a
-non-induced isomorphism, and $i : P \hookrightarrow T$ for an induced isomorphism. Observe that an
-induced isomorphism $i : P \hookrightarrow T$ is a non-induced isomorphism $i : P \rightarrowtail T$
-which is also a non-induced isomorphism $\shortoverline{i} : \shortoverline{P} \rightarrowtail
-\shortoverline{T}$.
+$\operatorname{V}(T)$ which preserves adjacency---that is, for every adjacent vertices $v$ and $w$
+in $\operatorname{V}(P)$, the vertices $i(v)$ and $i(w)$ are adjacent in $T$. If vertices are
+labelled, the mapping must preserve labels, so $\ell(v) = \ell(i(v)) \forall v \in
+\operatorname{V}(P)$. An \emph{induced subgraph isomorphism} additionally preserves
+non-adjacency---that is, if $v$ and $w$ are not adjacent in $P$, then $i(v)$ and $i(w)$ must not be
+adjacent in $T$. We use the notation $i : P \rightarrowtail T$ for a non-induced subgraph
+isomorphism, and $i : P \hookrightarrow T$ for an induced subgraph isomorphism. Observe that an
+induced subgraph isomorphism $i : P \hookrightarrow T$ is a non-induced subgraph isomorphism $i : P
+\rightarrowtail T$ which is also a non-induced subgraph isomorphism $\shortoverline{i} :
+\shortoverline{P} \rightarrowtail \shortoverline{T}$.
 
@@
         Along the top row, we show the proportion of instances which are satisfiable; the white
-        bands shows the phase transitions, and the black lines are our predictions of where the
+        bands show the phase transitions, and the black lines are our predictions of where the
         phase transition will occur. On the final three rows, we show the number of search nodes used by the
@@
 Suppose our labels are drawn randomly from a set $L = \{ 1, \ldots, k \}$, where $k$ is reasonably
-small compared to $p$. We can partition the pattern vertices by label into disjoint sets \[\{
-    \operatorname{V}(P)\vert_1 = \{ v \in \operatorname{V}(P) : \ell(v) = 1 \}, \ldots,
+small compared to $p$. By defining \[ \operatorname{V}(P)\vert_n = \{ v \in \operatorname{V}(P) :
+\ell(v) = n \} \] to be the set of vertices with label $k$, we can partition the pattern
+vertices by label into disjoint sets \[\{ \operatorname{V}(P)\vert_1, \ldots,
 \operatorname{V}(P)\vert_k \}\textnormal{\,,}\] each of which is expected to contain
-$\nicefrac{p}{k}$ vertices. Similarly, we may partition the target vertices into disjoint sets $\{
-    \operatorname{V}(T)\vert_1, \ldots, \operatorname{V}(T)\vert_k \}$.
+$\nicefrac{p}{k}$ vertices. Similarly, we may partition the target vertices into disjoint
+sets $\{ \operatorname{V}(T)\vert_1, \ldots, \operatorname{V}(T)\vert_k \}$.
 
@@
 state space is now \[ |S| = \left((\nicefrac{t}{k})^{\underline{\nicefrac{p}{k}}}\right)^{k}
-\textnormal{,} \] but to do this we must state what $a^{\underline{b}}$ means when $b$ is fractional.
-A reasonable continuous extension may be obtained by taking \[ |S| = \left(\frac{\Gamma\left(\nicefrac{t}{k}
-+ 1\right)}{\Gamma\left(\nicefrac{t}{k} - \nicefrac{p}{k} + 1\right)}\right)^{k} \textnormal{\,.} \]
+\textnormal{,} \] but to do this we must state what $a^{\underline{b}}$ means when $b$ is
+fractional. The gamma function $\Gamma(n)$ is equal to $(n - 1)!$ for integers $n \ge 1$, but is also
+defined for positive real numbers, obeying the identify $\Gamma(x + 1) = x\Gamma(x)$. By noting that
+$t^{\underline{p}} = \nicefrac{t!}{(t - p)!}$, we may obtain a reasonable
+continuous extension by taking \[ |S| = \left(\frac{\Gamma\left(\nicefrac{t}{k} +
+1\right)}{\Gamma\left(\nicefrac{t}{k} - \nicefrac{p}{k} + 1\right)}\right)^{k} \textnormal{\,,} \]
 
@@
 large, the prediction of the phase transition still occurs in the right place, but we start to see
-sporadic unsatisfiable instances deep inside the satisfiable reason. With sufficiently many labels,
+sporadic unsatisfiable instances deep inside the satisfiable region. With sufficiently many labels,
 we get a kind of phase transition from ``all unsatisfiable'' to ``mixed satisfiable and
@@
 filtering techniques to try to avoid the subgraph isomorphism step. A similar claim is made by
-\citet{DBLP:conf/icde/ZhangHY07} in a introduction of another indexing technique: ``obviously it is
+\citet{DBLP:conf/icde/ZhangHY07} in an introduction of another indexing technique: ``obviously it is
 inefficient to perform a sequential scan on every graph in the database, because the subgraph

### 2016-10-20 paper/paper.tex
Typo

@@
 vertices in the target. This is not entirely straightforward---the lack of unsatisfiable instances
-for testing purposes cannot be addressed simply by using taking a pattern graph from one of the
+for testing purposes cannot be addressed simply by taking a pattern graph from one of the
 existing random suites with the ``wrong'' target graph, as this tends to give either a trivially

### 2016-10-22 paper/paper.tex
Another one

@@
 \citet{DBLP:journals/pvldb/YuanMG13} continue a line of supergraph search work, again using a
-filtering step to avoid subgraph isomorphism calls. There is also research into maintaining indices
-when the set of target graph changes: for example \citet{DBLP:conf/sigmod/YuanMYG15} look at
-algorithms for updating graph indices. And in describing a system for reusing results of queries
-which are sub- or super-graphs of previous queries, \citet{DBLP:conf/edbt/WangNT16} state that
-querying is a ``very costly operation as it entails the \NP-complete problem of subgraph
-isomorphism'', and place ``an emphasis on the number of unnecessary subgraph isomorphism tests''.
+filtering step to avoid subgraph isomorphism calls.  \citet{DBLP:journals/tkde/HongZLY15} look at
+graph database subgraph matching with an additional set similarity constraint, and state that
+Ullmann's algorithm and VF2 are ``usually costly for large graphs'' because they ``do not utilize
+any index structure''. They propose an indexing structure, which takes over 2,000 seconds to
+construct, and uses nearly 2GBytes of space.  There is also research into maintaining indices when
+the set of target graph changes: for example \citet{DBLP:conf/sigmod/YuanMYG15} look at algorithms
+for updating graph indices. And in describing a system for reusing results of queries which are sub-
+or super-graphs of previous queries, \citet{DBLP:conf/edbt/WangNT16} state that querying is a ``very
+costly operation as it entails the \NP-complete problem of subgraph isomorphism'', and place ``an
+emphasis on the number of unnecessary subgraph isomorphism tests''.
 

### 2016-10-23 paper/paper.tex
I found another one
Yet another

@@
 
+\citet{DBLP:conf/icdcs/CaoYWRL11} propose a privacy-preserving cloud graph database system using
+filter / verify, stating that ``checking subgraph isomorphism is \NP-complete, and therefore it is
+infeasible to employ such a costly solution'' which simply checks every graph for a match.
 \citet{DBLP:journals/tkde/WangWYY12} look at indexing large sparse graphs. They state that ``because
@@
 filtering techniques. They state that performing a query against each graph in the dataset
-``obviously does not scale, as subgraph isomorphism is \NP-complete''.
+``obviously does not scale, as subgraph isomorphism is \NP-complete''. In describing a system which
+returns a special subset of graphs which match a query, \citet{DBLP:journals/tkde/ZhengLZHZ16}
+suggest that it is ``NP-hard to check the graph isomorphism'' (meaning subgraph isomorphism), and
+``in order to improve the time efficiency'' they use an indexing system to ``avoid as many costly
+subgraph isomorphism checkings as possible''. Their index takes tens of thousands of seconds to
+build, and they suggest that Ullmann's algorithm and VF2 are state of the art for verification.
 

### 2016-10-23 paper/paper.tex
They just keep on saying that

@@
 querying is costly because it is \NP-complete, and that indices can improve the performance of graph
-database queries. Again, new indexing techniques are introduced.  Subsequently,
-\citet{DBLP:journals/pvldb/KatsarouNT15} perform a comprehensive comparison of graph database
+database queries. Again, new indexing techniques are introduced. Subsequently, \citet{DBLP:journals/pvldb/KatsarouNT15} perform a comprehensive comparison of graph database
 filtering techniques. They state that performing a query against each graph in the dataset
@@
 build, and they suggest that Ullmann's algorithm and VF2 are state of the art for verification.
+And \citet{DBLP:journals/www/PengZ0LZ16} state that ``obviously, it is impossible to employ some
+subgraph isomorphism algorithm, such as Ullmann or VF2'', and argue that ``in order to speed up
+query processing'', they need to create indices.
 

### 2016-10-23 paper/paper.tex
Please Sir, can I have some more?

@@
 
-The filter / verify paradigm also influences other research. For example, recently
-\citet{DBLP:journals/pvldb/YuanMG13} continue a line of supergraph search work, again using a
-filtering step to avoid subgraph isomorphism calls.  \citet{DBLP:journals/tkde/HongZLY15} look at
-graph database subgraph matching with an additional set similarity constraint, and state that
-Ullmann's algorithm and VF2 are ``usually costly for large graphs'' because they ``do not utilize
-any index structure''. They propose an indexing structure, which takes over 2,000 seconds to
-construct, and uses nearly 2GBytes of space.  There is also research into maintaining indices when
-the set of target graph changes: for example \citet{DBLP:conf/sigmod/YuanMYG15} look at algorithms
-for updating graph indices. And in describing a system for reusing results of queries which are sub-
-or super-graphs of previous queries, \citet{DBLP:conf/edbt/WangNT16} state that querying is a ``very
-costly operation as it entails the \NP-complete problem of subgraph isomorphism'', and place ``an
-emphasis on the number of unnecessary subgraph isomorphism tests''.
+The filter / verify paradigm also influences other research. For example, it is used by
+\citet{DBLP:conf/icde/TianP08} in an approximate subgraph searching system: they state that
+Ullmann's algorithm ``is prohibitively expensive for querying against [a] database with a large
+number of graphs'', and that indices are used ``to filter out graphs that do not match the query''.
+More recently, \citet{DBLP:journals/pvldb/YuanMG13} continue a line of supergraph search work, again
+using a filtering step to avoid subgraph isomorphism calls (in the opposite direction, so queries
+are now target graphs). \citet{DBLP:journals/tkde/HongZLY15} look at graph database subgraph
+matching with an additional set similarity constraint, and state that Ullmann's algorithm and VF2
+are ``usually costly for large graphs'' because they ``do not utilize any index structure''. They
+propose an indexing structure, which takes over 2,000 seconds to construct, and uses nearly 2GBytes
+of space.  There is also research into maintaining indices when the set of target graph changes: for
+example \citet{DBLP:conf/sigmod/YuanMYG15} look at algorithms for updating graph indices. And in
+describing a system for reusing results of queries which are sub- or super-graphs of previous
+queries, \citet{DBLP:conf/edbt/WangNT16} state that querying is a ``very costly operation as it
+entails the \NP-complete problem of subgraph isomorphism'', and place ``an emphasis on the number of
+unnecessary subgraph isomorphism tests''.
 

### 2016-10-24 paper/paper.tex
Regular things

@@
 
+\begin{figure}[tb]
+    \centering
+    \begin{tikzpicture}[every node/.style={inner sep=0pt, outer sep=0pt}]
+        \matrix {
+            \node [anchor=center] {}; &
+            \node [anchor=center] {\scriptsize $\mathcal{G}(10, x){\rightarrowtail}\mathcal{G}(150, y)$}; &
+            \node [anchor=center] {\scriptsize $\mathcal{G}(20, x){\rightarrowtail}\mathcal{G}(150, y)$}; &
+            \node [anchor=center] {\scriptsize $\mathcal{G}(30, x){\rightarrowtail}\mathcal{G}(150, y)$}; &
+            \\
+            \node [anchor=center] [rotate=90] {\scriptsize Satisfiable?}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-kr-satisfiable-10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-kr-satisfiable-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-kr-satisfiable-30-150.pdf}}; &
+            \\
+            \node [anchor=center] [rotate=90] {\scriptsize Glasgow}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-kr-nodes-10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-kr-nodes-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-kr-nodes-30-150.pdf}}; &
+            \\
+        };
+    \end{tikzpicture}
+
+    \caption{Behaviour of algorithms on random $k$-regular graphs, as the pattern degree (x-axis)
+    and target degree (y-axis) are varied from 0 to $\left|\operatorname{V}\right| - 1$.}
+    \label{figure:kr}
+\end{figure}
+
+By $\mathcal{G}(n, k)$ we mean a random graph with $k$ vertices, each of which has degree $k$. In
+\cref{figure:kr} we recreate \cref{figure:non-induced}, using the NetworkX implementation
+\citep{NetworkX} of the \citet{DBLP:journals/cpc/StegerW99} algorithm to generate regular graphs.
+The $x$-axes range from degree 0 to degree $n - 1$, and the y-axes range from degree 0 to degree
+149.
+
 \section{Labelled Graphs}\label{section:labelled}

### 2016-10-24 line/noninduced-kr-line.rb
Some more data

@@ -1,0 +1,45 @@
+#!/usr/bin/ruby
+# vim: set sw=4 sts=4 et tw=80 :
+
+require 'bigdecimal'
+
+ps = ARGV[0].to_i
+ts = ARGV[1].to_i
+
+0.upto(ps - 1) do | x |
+    pd = BigDecimal.new(x) / BigDecimal.new(ps - 1);
+
+    tdupper = 500
+    tdlower = 0
+    expsols = 0
+
+    0.upto(100) do | y |
+        td = BigDecimal.new(tdupper + tdlower) / 1000;
+
+        cliqueedges = BigDecimal.new(ps * (ps - 1) / 2);
+        nedges = pd * cliqueedges;
+        nnonedges = (1 - pd) * cliqueedges;
+
+        edgesokprob = td ** nedges;
+
+        solprob = edgesokprob;
+
+        nstates = BigDecimal.new(1);
+        1.upto(ps) do | n |
+            nstates *= (ts - n + 1);
+        end
+
+        expsols = nstates * solprob;
+
+        if (expsols < 1)
+            tdlower = (tdupper + tdlower) / 2
+        elsif (expsols > 1)
+            tdupper = (tdupper + tdlower) / 2
+        else
+            break
+        end
+    end
+
+    puts(pd.to_f.to_s + " " + ((tdupper + tdlower) / 1000.0).to_s + " " + expsols.to_s)
+end
+

### 2016-11-01 paper/paper.tex
More
Start on VF2 kr

@@
             \\
+            \node [anchor=center] [rotate=90] {\scriptsize LAD}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-lad-non-induced-kr-nodes-10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-lad-non-induced-kr-nodes-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-lad-non-induced-kr-nodes-30-150.pdf}}; &
+            \\
+            \node [anchor=center] [rotate=90] {\scriptsize VF2}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-vf2-non-induced-kr-nodes-10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-vf2-non-induced-kr-nodes-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{gen-graph-vf2-non-induced-kr-nodes-30-150.pdf}}; &
+            \\
         };

### 2016-11-07 paper/paper.tex
Combine EPSRC lines

@@
 
+\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}
+
 \definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}
@@
 \author{\name Ciaran McCreesh \thanks{This work was supported by the Engineering and Physical Sciences
-           Research Council [grant number EP/K503058/1]} \email c.mccreesh.1@research.gla.ac.uk \\
+           Research Council [grant numbers EP/K503058/1 and EP/M508056/1]} \email c.mccreesh.1@research.gla.ac.uk \\
        \name Patrick Prosser \email patrick.prosser@glasgow.ac.uk \\
-       \name James Trimble \thanks{This work was supported by the Engineering and Physical Sciences
-           Research Council [grant number EP/M508056/1]} \email j.trimble.1@research.gla.ac.uk \\
+       \name James Trimble \samethanks \email j.trimble.1@research.gla.ac.uk \\
    \addr University of Glasgow, Glasgow, Scotland}

### 2016-11-07 paper/paper.tex
No need to repeat the year

@@
 \title{Why Real-World Subgraph Matching is Easy, and \\ What This Means for Filter / Verify in
-Graph Databases \thanks{Part of this paper appeared at IJCAI 2016
+Graph Databases \thanks{Part of this paper appeared at IJCAI
 \citep{DBLP:conf/ijcai/McCreeshPT16}; \cref{section:introduction} has been expanded and

### 2016-11-07 paper/paper.tex
A bit more abstract

@@
     this size. We show how to generate ``really hard'' random instances for subgraph isomorphism
-    problems, which are computationally challenging with only a couple of hundred vertices. For the
-    non-induced version of the problem, these instances lie on a satisfiable / unsatisfiable phase
-    transition, whose location we can predict; for the induced variant, much richer behaviour is
-    observed, and constrainedness gives a better measure of difficulty than does proximity to a
-    phase transition. In both cases, we can use this to understand and improve variable and value
-    ordering heuristics.
+    problems, which are computationally challenging with only a couple of hundred vertices in the
+    target, and only twenty to thirty pattern vertices. For the non-induced version of the problem,
+    these instances lie on a satisfiable / unsatisfiable phase transition, whose location we can
+    predict; for the induced variant, much richer behaviour is observed, and constrainedness gives a
+    better measure of difficulty than does proximity to a phase transition. In both cases, we can
+    use this to understand and improve variable and value ordering heuristics.
 
@@
     flawed design based upon a repeated misunderstanding of what makes subgraph isomorphism hard. We
-    explain why this technique would not be beneficial, even with harder instances than those
+    explain why filter / verify cannot be beneficial, even with harder instances than those
     currently used, when paired with any reasonable modern subgraph isomorphism algorithm.

### 2017-07-20 paper/paper.tex
Little tweaks
Bring things in

@@
 
+\definecolor{chromaygb1}{rgb}{1.0, 1.0, 0.878431}
+\definecolor{chromaygb1b}{rgb}{0.8, 0.972549, 0.72549}
+\definecolor{chromaygb2}{rgb}{0.560784,0.929412,0.564706}
+\definecolor{chromaygb3}{rgb}{0.407843,0.776471,0.560784}
+\definecolor{chromaygb4}{rgb}{0.227451,0.631373,0.552941}
+\definecolor{chromaygb5}{rgb}{0.301961,0.431373,0.670588}
+\definecolor{chromaygb6}{rgb}{0.317647,0.254902,0.588235}
+\definecolor{chromaygb7}{rgb}{0.294118,0.0,0.509804}
+
+\crefname{algorithm}{Algorithm}{Algorithms}
+\Crefname{algorithm}{Algorithm}{Algorithms}
+\crefname{figure}{Figure}{Figures}
+\Crefname{figure}{Figure}{Figures}
+\crefname{table}{Table}{Tables}
+\Crefname{table}{Table}{Tables}
+\crefname{section}{Section}{Sections}
+\Crefname{section}{Section}{Sections}
+\crefname{proposition}{Proposition}{Propositions}
+\Crefname{proposition}{Proposition}{Propositions}
+
 \newcommand{\citet}[1]{\citeA{#1}}
@@
 \author{\name Ciaran McCreesh \thanks{This work was supported by the Engineering and Physical Sciences
-           Research Council [grant numbers EP/K503058/1 and EP/M508056/1]} \email c.mccreesh.1@research.gla.ac.uk \\
+           Research Council [grant numbers EP/K503058/1 and EP/M508056/1]} \email
+           ciaran.mccreesh@glasgow.ac.uk \\
        \name Patrick Prosser \email patrick.prosser@glasgow.ac.uk \\
@@
     When labels are introduced, as commonly seen in real-world applications and in graph databases,
-    additional rich behaviour arises.  We show that when only slightly better algorithms are used,
-    all commonly used benchmark suites for evaluating graph database systems become trivial, and
-    that existing notions of difficulty are based almost entirely upon unnecessarily poor
-    performance from overly simplistic subgraph isomorphism algorithms. With this in mind, we argue
-    that the widely researched ``filter / verify'' indexing technique used in graph databases is a
-    flawed design based upon a repeated misunderstanding of what makes subgraph isomorphism hard. We
-    explain why filter / verify cannot be beneficial, even with harder instances than those
-    currently used, when paired with any reasonable modern subgraph isomorphism algorithm.
+    additional rich behaviour arises.  We show that when even the simplest constraint programming
+    algorithms are used, all commonly used benchmark suites for evaluating graph database systems
+    become trivial, and that existing notions of difficulty are based almost entirely upon
+    unnecessarily poor performance from overly simplistic subgraph isomorphism algorithms. With this
+    in mind, we demonstrate that the widely researched ``filter~/ verify'' indexing technique used in
+    graph databases is a flawed design based upon a repeated misunderstanding of what makes subgraph
+    isomorphism hard. We explain why filter~/ verify cannot be beneficial, even with harder
+    instances than those currently used, when paired with any reasonable modern subgraph isomorphism
+    algorithm.
 \end{abstract}
@@
 graph, and up to ten thousand vertices in the target graph
-\cite{Cordella:2004,Solnon:2010,Audemard:2014,McCreesh:2015}, leading to successful applications in
-areas such as computer vision \cite{Damiand:2011,Solnon:2015}, biochemistry
-\cite{Giugno:2013,Carletti:2015}, and pattern recognition \citep{Conte:2004}.  In particular, a
-labelled version of subgraph isomorphism is one of the key components in supporting complex queries
-in graph databases---typically, these systems store a fixed set of target graphs, and for a sequence
-of pattern queries, they must return every target graph which contains that pattern.  One common
-approach to this problem is to combine a subgraph isomorphism algorithm with an indexing system in a
-so-called ``filter / verify'' model, where invariants are used to attempt to pre-exclude
-unsatisfiable instances to avoid the cost of a subgraph isomorphism call. In this context, the terms
-\emph{matching} and \emph{verification} are often used for the subgraph isomorphism step (or a
-slightly broader problem, for example permitting wildcards).
-
-\begin{figure}[t]
+\citep{DBLP:journals/pami/CordellaFSV04,DBLP:journals/ai/Solnon10,DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16},
+leading to successful applications in areas such as computer vision
+\citep{DBLP:journals/cviu/DamiandSHJS11,DBLP:journals/pr/SolnonDHJ15}, biochemistry
+\citep{o:10.1371/journal.pone.0076911,DBLP:conf/gbrpr/CarlettiFV15}, and pattern recognition
+\citep{DBLP:journals/ijprai/ConteFSV04}.  In particular, a labelled version of subgraph isomorphism
+is one of the key components in supporting complex queries in graph databases---typically, these
+systems store a fixed set of target graphs, and for a sequence of pattern queries, they must return
+every target graph which contains that pattern.  One common approach to this problem is to combine a
+subgraph isomorphism algorithm with an indexing system in a so-called ``filter~/ verify'' model,
+where invariants are used to attempt to pre-exclude unsatisfiable instances to avoid the cost of a
+subgraph isomorphism call. In this context, the terms \emph{matching} and \emph{verification} are
+often used for the subgraph isomorphism step (or a slightly broader problem, for example permitting
+wildcards).
+
+However, subgraph isomorphism algorithms cannot handle \emph{arbitrary} instances involving hundreds
+or thousands of vertices. Experimental evaluations of subgraph isomorphism algorithms are usually
+performed using randomly generated graph pairs, sometimes together with a mix of real-world graphs,
+and graphs that encode biochemistry and computer vision problems. Using random instances to evaluate
+algorithm behaviour can be beneficial, because it provides a way of generating many instances
+cheaply, and reduces the risk of over-fitting when tuning design parameters. The random instances
+used typically come from common datasets
+\citep{DBLP:journals/prl/SantoFSV03,DBLP:journals/constraints/ZampelliDS10}, which were generated by
+taking a random subgraph of a random (Erd\H{o}s-R\'enyi, scale-free, bounded degree, or mesh)
+graph and permuting the vertices. Such instances are guaranteed to be
+satisfiable---\citet{o:AntonO09} exploited this property to create large sets of random satisfiable
+boolean satisfiability instances.  This is the most common approach to generating random subgraph
+isomorphism instances, meaning existing benchmark suites contain relatively few non-trivial
+unsatisfiable instances (although a few of the patterns in the instances by
+\citeauthor{DBLP:journals/constraints/ZampelliDS10} have had extra edges added, making them
+unsatisfiable). Also, the satisfiable instances tend to be computationally fairly easy, with most of
+the difficulty being in dealing with the size of the model. This has lead to bias in algorithm
+design, to the extent that some proposed techniques, such as those of
+\citet{DBLP:conf/sls/BattitiM07}, will \emph{only} work on satisfiable instances.
+
+The first contribution of this paper is to present and evaluate new methods for creating random
+pattern~/ target pairs.  The method we introduce generates both satisfiable and unsatisfiable
+instances, and can produce computationally challenging instances with only a few tens of vertices in
+the pattern, and 150 vertices in the target. This is not entirely straightforward---the lack of
+unsatisfiable instances for testing purposes cannot be addressed simply by taking a pattern graph
+from one of the existing random suites with the ``wrong'' target graph, as this tends to give either
+a trivially unsatisfiable instance, or a satisfiable instance. (In particular, it is \emph{not} the
+case that a relatively small random graph is unlikely to appear in a larger random graph.)
+
+\begin{figure}[b]
     \centering
-    \begin{tikzpicture}[scale=0.5,every node/.style={font=\footnotesize}]
-        \begin{scope}[yshift=-2cm]
-            \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (Na) at (1,  0) {1};
-            \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (Nb) at (1, -1.5) {2};
-            \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (Nc) at (0, -3) {3};
-            \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (Nd) at (2, -3) {4};
+    \begin{tikzpicture}[scale=0.3,every node/.style={font=\footnotesize}]
+        \begin{scope}
+            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (Na) at (1,  0) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (Nb) at (1, -2) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (Nc) at (0, -4) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (Nd) at (2, -4) {\phantom{0}};
 
@@
 
-            \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (N1) at (5.5,  0) {1};
-            \node [draw, circle, fill=white, inner sep=1.5pt] (N2) at (7.5,  0) {2};
-            \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (N3) at (5.5, -1.5) {3};
-            \node [draw, circle, fill=white, inner sep=1.5pt] (N4) at (7.5, -1.5) {4};
-            \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (N5) at (5.5, -3) {5};
-            \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (N6) at (7.5, -3) {6};
-
-            \draw [thick, color=uofgsandstone!50] (N1) -- (N2);
+            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (N1) at (7.5,  0) {\phantom{0}};
+            \node[draw, circle, fill=white, inner sep=0.5pt] (N2) at (9.5,  0) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (N3) at (7.5, -2) {\phantom{0}};
+            \node[draw, circle, fill=white, inner sep=0.5pt] (N4) at (9.5, -2) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (N5) at (7.5, -4) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (N6) at (9.5, -4) {\phantom{0}};
+            \node[draw, circle, fill=white, inner sep=0.5pt] (N7) at (5.5,  0) {\phantom{0}};
+            \node[draw, circle, fill=white, inner sep=0.5pt] (N8) at (5.5, -2) {\phantom{0}};
+            \node[draw, circle, fill=white, inner sep=0.5pt] (N9) at (5.5, -4) {\phantom{0}};
+
+            \draw [] (N1) -- (N2);
             \draw [ultra thick] (N1) -- (N3);
-            \draw [thick, color=uofgsandstone!50] (N1) -- (N4);
-            \draw [thick, color=uofgsandstone!50] (N2) -- (N4);
+            \draw [] (N1) -- (N4);
+            \draw [] (N2) -- (N4);
             \draw [ultra thick] (N3) -- (N5);
             \draw [ultra thick] (N3) -- (N6);
-            \draw [thick, color=uofgsandstone!50] (N4) -- (N6);
+            \draw [] (N4) -- (N6);
             \draw [ultra thick] (N5) -- (N6);
-            \draw [thick, color=uofgsandstone!50] (N2) to [in=45, out=315] (N6);
-            \draw [thick, color=uofgsandstone!50] (N1) to [in=135, out=225] (N5);
-
-            \node [draw, circle, fill=white, inner sep=1.5pt] (M1) at (-5.5,  0) {1};
-            \node [draw, circle, fill=uofgcobalt, inner sep=1.5pt] (M2) at (-3.5,  0) {2};
-            \node [draw, circle, fill=uofgthistle, inner sep=1.5pt] (M3) at (-5.5, -1.5) {3};
-            \node [draw, circle, fill=white, inner sep=1.5pt] (M4) at (-3.5, -1.5) {4};
-            \node [draw, circle, fill=uofglawn, inner sep=1.5pt] (M5) at (-5.5, -3) {5};
-            \node [draw, circle, fill=uofgpumpkin, inner sep=1.5pt] (M6) at (-3.5, -3) {6};
-
-            \draw [thick, color=uofgsandstone!50] (M1) -- (M2);
-            \draw [thick, color=uofgsandstone!50] (M1) -- (M3);
-            \draw [thick, color=uofgsandstone!50] (M1) -- (M4);
-            \draw [thick, color=uofgsandstone!50] (M2) -- (M4);
-            \draw [ultra thick] (M3) -- (M5);
-            \draw [ultra thick] (M3) -- (M6);
-            \draw [thick, color=uofgsandstone!50] (M4) -- (M6);
-            \draw [ultra thick] (M5) -- (M6);
-            \draw [ultra thick] (M2) to [in=45, out=315] (M6);
-            \draw [thick, color=uofgsandstone!50] (M1) to [in=135, out=225] (M5);
-
-            \node [anchor=center, font=\Large] (A1) at (-1.5, -1.5) { $\hookleftarrow$ };
-            \node [anchor=center, font=\Large] (A2) at ( 3.5, -1.5) { $\rightarrowtail$ };
+            \draw [] (N2) to [in=45, out=315] (N6);
+            \draw [] (N1) -- (N7);
+            \draw [] (N5) -- (N9);
+            \draw [] (N7) -- (N8);
+            \draw [] (N8) -- (N9);
         \end{scope}
-        \begin{scope}[xshift=13cm]
-            \foreach \n in {1, ..., 6}{
-                \node [draw, circle, fill=uofgcobalt, inner sep=1pt] (N1\n) at (\n, 0) {\n};
-            }
-
-            \draw [decorate, decoration={brace, raise=0.2cm}, very thick] (N11.north west) -- (N16.north
-                east) node [midway, above=0.3cm] { $1 \mapsto$ };
-
-            \foreach \n in {1, ..., 6}{
-                \node [draw, circle, fill=uofgpumpkin, inner sep=1pt] (N2\n) at (0, -\n) {\n};
-            }
-
-            \draw [decorate, decoration={brace, raise=0.2cm, mirror}, very thick] (N21.north west) --
-                (N26.south west) node [midway, left=0.3cm] { $2 \mapsto$ };
-
-            \foreach \n in {1, ..., 6}{
-                \node [draw, circle, fill=uofglawn, inner sep=1pt] (N3\n) at (7, -\n) {\n};
-            }
-
-            \draw [decorate, decoration={brace, raise=0.2cm}, very thick] (N31.north east) -- (N36.south
-                east) node [midway, right=0.3cm] { $3 \mapsto$ };
-
-            \foreach \n in {1, ..., 6}{
-                \node [draw, circle, fill=uofgthistle, inner sep=1pt] (N4\n) at (\n, -7) {\n};
-            }
-
-            \draw [decorate, decoration={brace, raise=0.2cm, mirror}, very thick] (N41.south west) --
-                (N46.south east) node [midway, below=0.3cm] { $4 \mapsto$ };
-
-            \begin{scope}[on background layer]
-                \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N22)!0.8!(N22.east)$);
-                \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N23)!0.8!(N23.east)$);
-                \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N24)!0.8!(N24.east)$);
-                \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N25)!0.8!(N25.east)$);
-                \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N36)!0.8!(N36.west)$);
-                \draw [color=uofgsandstone!50] ($(N11)!0.8!(N11.south)$) -- ($(N46)!0.8!(N46.north)$);
-                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N12)!0.8!(N12.south)$);
-                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N32)!0.8!(N32.west)$);
-                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N42)!0.8!(N42.north)$);
-                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N13)!0.8!(N13.south)$);
-                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N33)!0.8!(N33.west)$);
-                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N43)!0.8!(N43.north)$);
-                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N14)!0.8!(N14.south)$);
-                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N34)!0.8!(N34.west)$);
-                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N44)!0.8!(N44.north)$);
-                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N15)!0.8!(N15.south)$);
-                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N35)!0.8!(N35.west)$);
-                \draw [color=uofgsandstone!50] ($(N21)!0.8!(N21.east)$) -- ($(N45)!0.8!(N45.north)$);
-                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N22)!0.8!(N22.east)$);
-                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N42)!0.8!(N42.north)$);
-                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N23)!0.8!(N23.east)$);
-                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N43)!0.8!(N43.north)$);
-                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N24)!0.8!(N24.east)$);
-                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N44)!0.8!(N44.north)$);
-                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N25)!0.8!(N25.east)$);
-                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N45)!0.8!(N45.north)$);
-                \draw [color=uofgsandstone!50] ($(N31)!0.8!(N31.west)$) -- ($(N16)!0.8!(N16.south)$);
-                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N22)!0.8!(N22.east)$);
-                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N32)!0.8!(N32.west)$);
-                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N23)!0.8!(N23.east)$);
-                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N33)!0.8!(N33.west)$);
-                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N24)!0.8!(N24.east)$);
-                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N34)!0.8!(N34.west)$);
-                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N25)!0.8!(N25.east)$);
-                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N35)!0.8!(N35.west)$);
-                \draw [color=uofgsandstone!50] ($(N41)!0.8!(N41.north)$) -- ($(N16)!0.8!(N16.south)$);
-                \draw [color=uofgsandstone!50] ($(N12)!0.8!(N12.south)$) -- ($(N33)!0.8!(N33.west)$);
-                \draw [color=uofgsandstone!50] ($(N12)!0.8!(N12.south)$) -- ($(N24)!0.8!(N24.east)$);
-                \draw [color=uofgsandstone!50] ($(N12)!0.8!(N12.south)$) -- ($(N45)!0.8!(N45.north)$);
-                \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N14)!0.8!(N14.south)$);
-                \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N34)!0.8!(N34.west)$);
-                \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N44)!0.8!(N44.north)$);
-                \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N16)!0.8!(N16.south)$);
-                \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N36)!0.8!(N36.west)$);
-                \draw [color=uofgsandstone!50] ($(N22)!0.8!(N22.east)$) -- ($(N46)!0.8!(N46.north)$);
-                \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N13)!0.8!(N13.south)$);
-                \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N24)!0.8!(N24.east)$);
-                \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N44)!0.8!(N44.north)$);
-                \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N15)!0.8!(N15.south)$);
-                \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N26)!0.8!(N26.east)$);
-                \draw [color=uofgsandstone!50] ($(N32)!0.8!(N32.west)$) -- ($(N46)!0.8!(N46.north)$);
-                \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N13)!0.8!(N13.south)$);
-                \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N24)!0.8!(N24.east)$);
-                \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N34)!0.8!(N34.west)$);
-                \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N15)!0.8!(N15.south)$);
-                \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N26)!0.8!(N26.east)$);
-                \draw [color=uofgsandstone!50] ($(N42)!0.8!(N42.north)$) -- ($(N36)!0.8!(N36.west)$);
-                \draw [color=uofgsandstone!50] ($(N13)!0.8!(N13.south)$) -- ($(N34)!0.8!(N34.west)$);
-                \draw [color=uofgsandstone!50] ($(N13)!0.8!(N13.south)$) -- ($(N44)!0.8!(N44.north)$);
-                \draw [color=uofgsandstone!50] ($(N13)!0.8!(N13.south)$) -- ($(N25)!0.8!(N25.east)$);
-                \draw [color=uofgsandstone!50] ($(N13)!0.8!(N13.south)$) -- ($(N26)!0.8!(N26.east)$);
-                \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N15)!0.8!(N15.south)$);
-                \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N35)!0.8!(N35.west)$);
-                \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N45)!0.8!(N45.north)$);
-                \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N16)!0.8!(N16.south)$);
-                \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N36)!0.8!(N36.west)$);
-                \draw [color=uofgsandstone!50] ($(N23)!0.8!(N23.east)$) -- ($(N46)!0.8!(N46.north)$);
-                \draw [color=uofgsandstone!50] ($(N33)!0.8!(N33.west)$) -- ($(N14)!0.8!(N14.south)$);
-                \draw [color=uofgsandstone!50] ($(N33)!0.8!(N33.west)$) -- ($(N25)!0.8!(N25.east)$);
-                \draw [color=uofgsandstone!50] ($(N33)!0.8!(N33.west)$) -- ($(N45)!0.8!(N45.north)$);
-                \draw [color=uofgsandstone!50] ($(N33)!0.8!(N33.west)$) -- ($(N26)!0.8!(N26.east)$);
-                \draw [color=uofgsandstone!50] ($(N33)!0.8!(N33.west)$) -- ($(N46)!0.8!(N46.north)$);
-                \draw [color=uofgsandstone!50] ($(N43)!0.8!(N43.north)$) -- ($(N14)!0.8!(N14.south)$);
-                \draw [color=uofgsandstone!50] ($(N43)!0.8!(N43.north)$) -- ($(N25)!0.8!(N25.east)$);
-                \draw [color=uofgsandstone!50] ($(N43)!0.8!(N43.north)$) -- ($(N36)!0.8!(N36.west)$);
-                \draw [color=uofgsandstone!50] ($(N14)!0.8!(N14.south)$) -- ($(N35)!0.8!(N35.west)$);
-                \draw [color=uofgsandstone!50] ($(N14)!0.8!(N14.south)$) -- ($(N45)!0.8!(N45.north)$);
-                \draw [color=uofgsandstone!50] ($(N14)!0.8!(N14.south)$) -- ($(N26)!0.8!(N26.east)$);
-                \draw [color=uofgsandstone!50] ($(N24)!0.8!(N24.east)$) -- ($(N16)!0.8!(N16.south)$);
-                \draw [color=uofgsandstone!50] ($(N24)!0.8!(N24.east)$) -- ($(N36)!0.8!(N36.west)$);
-                \draw [color=uofgsandstone!50] ($(N24)!0.8!(N24.east)$) -- ($(N46)!0.8!(N46.north)$);
-                \draw [color=uofgsandstone!50] ($(N34)!0.8!(N34.west)$) -- ($(N15)!0.8!(N15.south)$);
-                \draw [color=uofgsandstone!50] ($(N34)!0.8!(N34.west)$) -- ($(N26)!0.8!(N26.east)$);
-                \draw [color=uofgsandstone!50] ($(N34)!0.8!(N34.west)$) -- ($(N46)!0.8!(N46.north)$);
-                \draw [color=uofgsandstone!50] ($(N44)!0.8!(N44.north)$) -- ($(N15)!0.8!(N15.south)$);
-                \draw [color=uofgsandstone!50] ($(N44)!0.8!(N44.north)$) -- ($(N26)!0.8!(N26.east)$);
-                \draw [color=uofgsandstone!50] ($(N44)!0.8!(N44.north)$) -- ($(N36)!0.8!(N36.west)$);
-                \draw [color=uofgsandstone!50] ($(N15)!0.8!(N15.south)$) -- ($(N26)!0.8!(N26.east)$);
-                \draw [color=uofgsandstone!50] ($(N25)!0.8!(N25.east)$) -- ($(N16)!0.8!(N16.south)$);
-                \draw [color=uofgsandstone!50] ($(N25)!0.8!(N25.east)$) -- ($(N36)!0.8!(N36.west)$);
-                \draw [color=uofgsandstone!50] ($(N25)!0.8!(N25.east)$) -- ($(N46)!0.8!(N46.north)$);
-                \draw [color=uofgsandstone!50] ($(N35)!0.8!(N35.west)$) -- ($(N46)!0.8!(N46.north)$);
-                \draw [color=uofgsandstone!50] ($(N45)!0.8!(N45.north)$) -- ($(N26)!0.8!(N26.east)$);
-                \draw [color=uofgsandstone!50] ($(N45)!0.8!(N45.north)$) -- ($(N36)!0.8!(N36.west)$);
-                \draw [ultra thick] ($(N12)!0.8!(N12.south)$) -- ($(N43)!0.8!(N43.north)$);
-                \draw [ultra thick] ($(N12)!0.8!(N12.south)$) -- ($(N35)!0.8!(N35.west)$);
-                \draw [ultra thick] ($(N12)!0.8!(N12.south)$) -- ($(N26)!0.8!(N26.east)$);
-                \draw [ultra thick] ($(N43)!0.8!(N43.north)$) -- ($(N35)!0.8!(N35.west)$);
-                \draw [ultra thick] ($(N43)!0.8!(N43.north)$) -- ($(N26)!0.8!(N26.east)$);
-                \draw [ultra thick] ($(N35)!0.8!(N35.west)$) -- ($(N26)!0.8!(N26.east)$);
-            \end{scope}
+        \begin{scope}[xshift=20cm]
+            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (Na) at (1,  0) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (Nb) at (1, -2) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (Nc) at (0, -4) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (Nd) at (2, -4) {\phantom{0}};
+
+            \draw [ultra thick] (Na) -- (Nb);
+            \draw [ultra thick] (Nb) -- (Nc);
+            \draw [ultra thick] (Nc) -- (Nd);
+            \draw [ultra thick] (Nb) -- (Nd);
+
+            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (N1) at (7.5,  0) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (N2) at (9.5,  0) {\phantom{0}};
+            \node[draw, circle, fill=white, inner sep=0.5pt] (N3) at (7.5, -2) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (N4) at (9.5, -2) {\phantom{0}};
+            \node[draw, circle, fill=white, inner sep=0.5pt] (N5) at (7.5, -4) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (N6) at (9.5, -4) {\phantom{0}};
+            \node[draw, circle, fill=white, inner sep=0.5pt] (N7) at (5.5,  0) {\phantom{0}};
+            \node[draw, circle, fill=white, inner sep=0.5pt] (N8) at (5.5, -2) {\phantom{0}};
+            \node[draw, circle, fill=white, inner sep=0.5pt] (N9) at (5.5, -4) {\phantom{0}};
+
+            \draw [ultra thick] (N1) -- (N2);
+            \draw [] (N1) -- (N3);
+            \draw [ultra thick] (N1) -- (N4);
+            \draw [ultra thick] (N2) -- (N4);
+            \draw [] (N3) -- (N5);
+            \draw [] (N3) -- (N6);
+            \draw [ultra thick] (N4) -- (N6);
+            \draw [] (N5) -- (N6);
+            \draw [ultra thick, dashed] (N2) to [in=45, out=315] (N6);
+            \draw [] (N1) -- (N7);
+            \draw [] (N5) -- (N9);
+            \draw [] (N7) -- (N8);
+            \draw [] (N8) -- (N9);
         \end{scope}
@@
 
-    \caption{On the left, there is a non-induced isomorphism from the pattern graph, in the center, to the target
-    on the right, mapping vertex 1 to 1, 2 to 3, 3 to 5 and 4 to 6. This is not an induced
-    isomorphism, since there is an edge between 1 and 5 in the target but not between 1 and 3 in the
-    pattern. The mapping from the pattern to the (same) target on the left, sending 1 to 2, 2 to 6, 3 to
-    5 and 4 to 3, is both non-induced and induced.
-    To the right, the association graph encoding for the induced version: the highlighted clique
-    corresponds to the same mapping.}
+    \caption{On the left, an induced subgraph isomorphism. On the right, a non-induced subgraph
+    isomorphism: the extra dashed edge is not present in the pattern graph.}
     \label{figure:sip}
@@
 
-Despite these successes, subgraph isomorphism algorithms cannot handle \emph{arbitrary} instances
-involving hundreds or thousands of vertices. Experimental evaluations of subgraph isomorphism
-algorithms are usually performed using a mix of real-world graphs, graphs that encode biochemistry
-and computer vision problems, and randomly generated graph pairs. Using random instances to evaluate
-algorithm behaviour can be beneficial, because it provides a way of generating many instances
-cheaply, and reduces the risk of over-fitting when tuning design parameters. The random instances
-used in each case came from common datasets \citep{DeSanto:2003,Zampelli:2010}, which were generated
-by taking a random subgraph of a random (Erd\H{o}s-R\'enyi, scale-free, bounded degree, or mesh)
-graph and permuting the vertices. Such instances are guaranteed to be
-satisfiable---\citet{Anton:2009} exploited this property to create large sets of random satisfiable
-boolean satisfiability instances.  This has been by far the most common approach used so far to
-generate random subgraph isomorphism instances so far, meaning existing benchmark suites contain
-relatively few non-trivial unsatisfiable instances (although some of the patterns in the instances
-by \citeauthor{Zampelli:2010} have had extra edges added, making them unsatisfiable). Also, the
-satisfiable instances tend to be computationally fairly easy, with most of the difficulty being in
-dealing with the size of the model. This has lead to bias in algorithm design, to the extent that
-some proposed techniques will \emph{only} work on satisfiable instances \citep{Battiti:2007}.
-
-Our first contribution is to present and evaluate a new method for creating random pattern/target
-pairs. This method generates both satisfiable and unsatisfiable instances, and can produce
-computationally challenging instances with only a few tens of vertices in the pattern, and 150
-vertices in the target. This is not entirely straightforward---the lack of unsatisfiable instances
-for testing purposes cannot be addressed simply by taking a pattern graph from one of the
-existing random suites with the ``wrong'' target graph, as this tends to give either a trivially
-unsatisfiable instance, or a satisfiable instance. (In particular, it is \emph{not} the case that a
-relatively small random graph is unlikely to appear in a larger random graph.)
-
-?? This needs to be reworded if we include non-ER random graphs.
-
 This work builds upon the phase transition phenomena observed in satisfiability and graph colouring
-problems first described by \citet{DBLP:conf/ijcai/CheesemanKT91} and \citet{Mitchell:1992}.  For
-subgraph isomorphism we identify three relevant control parameters: we can independently alter the
-edge probability of the pattern graph, the edge probability of the target graph, and the relative
-orders (number of vertices) of the pattern and target graphs.  For non-induced isomorphisms, with
-the correct choice of parameters we see results very similar to those observed with boolean
-satisfiability problems: there is a phase transition (whose location we can predict) from
-satisfiable to unsatisfiable, we see a solver-independent complexity peak occur near this phase
-transition, and understanding this behaviour helps us to select variable and value ordering
-heuristics.
+problems first described by \citet{DBLP:conf/ijcai/CheesemanKT91} and
+\citet{DBLP:conf/aaai/MitchellSL92}. For subgraph isomorphism we identify three relevant control
+parameters: we can independently alter the edge probability of the pattern graph, the edge
+probability of the target graph, and the relative orders (number of vertices) of the pattern and
+target graphs.  For non-induced isomorphisms, with the correct choice of parameters we see results
+very similar to those observed with boolean satisfiability problems: there is a phase transition
+(whose location we can predict) from satisfiable to unsatisfiable, and we see a solver-independent
+complexity peak near this phase transition. Additionally, understanding this behaviour
+helps us to select better variable- and value-ordering heuristics---this is the second contribution
+of this paper.
 
@@
 hard when using a pseudo-boolean encoding, and under reduction to the clique problem. Interestingly,
-the constrainedness measure proposed by \citet{Gent:1996:Kappa} \emph{does} predict this difficult
-region---these instances provide evidence in favour of constrainedness, rather than proximity to a
-phase transition, being an accurate predictor of difficulty, and show that constrainedness is not
-simply a refinement of a phase transition prediction.
+the constrainedness measure proposed by \citet{DBLP:conf/aaai/GentMPW96} \emph{does} predict this difficult
+region---the third contribution of this paper is to use these instances to provide evidence in
+favour of constrainedness, rather than proximity to a phase transition, being an accurate predictor
+of difficulty, and to show that constrainedness is not simply a refinement of a phase transition
+prediction.
+
+% What about other random graph models? Graphs where every vertex has the same degree, which are known
+% as $k$-regular, exhibit very similar characteristics.
 
 When labels on vertices are introduced, as is commonly seen in real-world applications and in graph
-database systems, richer behaviour emerges. We look at some of the datasets commonly used to test
-graph database systems, which the literature suggests are extremely hard to solve. Our theory
-predicts that these instances should be trivial, and indeed we see that the entire perceived
-difficulty of every dataset we investigate comes down to the widespread use of a particular subgraph
-isomorphsim algorithm which does not use a theoretically-strong variable ordering heuristic. Our
-second contribution is to show that the use of the right variable ordering heuristic, even without
-the more sophisticated recent advances in subgraph algorithms, makes the entire graph database
-filter / verify paradigm unnecessary.  Finally, we explain why filter / verify cannot be beneficial
-even on more challenging instances except if an extremely poor choice of subgraph isomorphism
-algorithm is made.
-
-\subsection{Definitions}
-
-?? Check whether any of the real-world graphs we look at have loops.
-
-Throughout, our graphs are undirected, and do not have any loops.  The \emph{order} of a graph is
-the cardinality of its vertex set. We write $\operatorname{V}(G)$ for the vertex set of a graph $G$.
-The \emph{complement} of a graph $G$, denoted $\shortoverline{G}$, is the graph with the same vertex
-set as $G$, and with an edge between distinct vertices $v$ and $w$ if and only if $v$ and $w$ are
-not adjacent in $G$. We write $G(n, p)$ for an Erd\H{o}s-R\'enyi random graph with $n$ vertices, and
-an edge between each distinct pair of vertices with independent probability $p$.  A graph may have
-labels on its vertices, in which case we write $\ell(v)$ for the label of a given vertex $v$.
-
-A \emph{non-induced subgraph isomorphism} from a graph $P$ (called the \emph{pattern}) to a graph
-$T$ (called the \emph{target}) is an injective mapping $i$ from $\operatorname{V}(P)$ to
-$\operatorname{V}(T)$ which preserves adjacency---that is, for every adjacent vertices $v$ and $w$
-in $\operatorname{V}(P)$, the vertices $i(v)$ and $i(w)$ are adjacent in $T$. If vertices are
-labelled, the mapping must preserve labels, so $\ell(v) = \ell(i(v)) \forall v \in
-\operatorname{V}(P)$. An \emph{induced subgraph isomorphism} additionally preserves
-non-adjacency---that is, if $v$ and $w$ are not adjacent in $P$, then $i(v)$ and $i(w)$ must not be
-adjacent in $T$. We use the notation $i : P \rightarrowtail T$ for a non-induced subgraph
-isomorphism, and $i : P \hookrightarrow T$ for an induced subgraph isomorphism. Observe that an
-induced subgraph isomorphism $i : P \hookrightarrow T$ is a non-induced subgraph isomorphism $i : P
-\rightarrowtail T$ which is also a non-induced subgraph isomorphism $\shortoverline{i} :
-\shortoverline{P} \rightarrowtail \shortoverline{T}$.
-
-\subsection{Experimental Setup}
-
-Our experiments were performed on systems with Intel Xeon E5-4650 v2 CPUs, running Scientific Linux
-release 6.7. We selected three subgraph isomorphism solvers: the Glasgow solver
-\citep{McCreesh:2015}, LAD \citep{Solnon:2010}, and VF2 \citep{Cordella:2004}; each was compiled
-using GCC 4.9.
-
-The Glasgow and LAD solvers use backtracking search to build up an assignment of target vertices
-(values) to pattern vertices (variables), but differ in terms of inference and ordering heuristics.
-The approach used by VF2 is similar, although the domains of variables are not stored (in the style
-of conventional backtracking, rather than forward-checking), and so domain wipeouts are not detected
-until an assignment is made.
-
-In each case we measure the number of recursive calls (guessed assignments) made, not runtimes. We
-are not aiming to compare absolute performance between solvers; rather, we are looking for
-solver-independent patterns of difficulty. We used a timeout of 1,000 seconds, which was enough for
-the Glasgow solver to solve nearly all our instances (whose orders were selected with this timeout
-in mind), although we may slightly overestimate the proportion of unsatisfiable instances for
-extremely sparse or dense pattern graphs. The LAD and VF2 solvers experienced many more failures
+database systems, richer behaviour emerges, particularly when moving away from a uniform labelling
+scheme, and we see that VF2 behaves much worse than the Glasgow and LAD algorithms in certain cases.
+A close look shows that these cases \emph{should} be easy to solve.
+
+This is not simply a theoretical curiosity. A labelled version of subgraph isomorphism is one of
+the key components in supporting complex queries in graph databases---typically, these systems store
+a fixed set of target graphs, and for a sequence of pattern queries, they must return every target
+graph which contains that pattern.  One common approach to this problem is to combine a subgraph
+isomorphism algorithm with an indexing system in a so-called ``filter~/ verify'' model, where
+invariants are used to attempt to pre-exclude unsatisfiable instances to avoid the cost of a
+subgraph isomorphism call. In this context, the terms \emph{matching} and \emph{verification} are
+often used for the subgraph isomorphism step (or a slightly broader problem, for example permitting
+wildcards).
+
+We look at some of the datasets commonly used to test graph database systems, which the literature
+suggests are extremely hard to solve. Simple experiments show that the entire perceived difficulty
+of each of these datasets comes down to the widespread use of especially poor subgraph isomorphism
+algorithms. Our fourth contribution is to show that the use of forward-checking and the right
+variable ordering heuristic, even without the more sophisticated recent advances in subgraph
+algorithms, makes the entire graph database filter~/ verify paradigm unnecessary.  Finally, we
+explain why filter~/ verify (and other recently proposed techniques) cannot be beneficial even on
+more challenging instances except if an extremely poor choice of subgraph isomorphism algorithm is
+made.
+
+\section{Experimental Setup}
+
+The experiments in this paper are performed on systems with Intel Xeon E5-4650 v2 CPUs, running
+Scientific Linux release 6.7. We will be working with three subgraph isomorphism solvers: the
+Glasgow solver \citet{DBLP:conf/cp/McCreeshP15}; LAD \citep{DBLP:journals/ai/Solnon10}; and VF2
+\citep{DBLP:journals/pami/CordellaFSV04}. Each was compiled using GCC 4.9.
+
+Recall that the Glasgow and LAD solvers use backtracking search to build up an assignment of target
+vertices (values) to pattern vertices (variables), but differ in terms of inference and ordering
+heuristics.  The approach used by VF2 is similar, although the domains of variables are not stored
+(in the style of conventional backtracking, rather than forward-checking), and so domain wipeouts
+are not detected until an assignment is made. Both the Glasgow and LAD solvers use domain sizes to
+guide search, whilst VF2 uses adjacency to previously-assigned vertices.
+
+We measure only the number of recursive calls (guessed assignments) made, not
+runtimes. We are not aiming to compare absolute performance between solvers; rather, we are looking
+for solver-independent patterns of difficulty. All experiments use a timeout of 1,000 seconds, which
+is enough for the Glasgow solver to solve nearly all instances (whose orders were selected with this
+timeout in mind), although we may slightly overestimate the proportion of unsatisfiable instances
+for extremely sparse or dense pattern graphs. The LAD and VF2 solvers experienced many more failures
 with this timeout, so our picture of just how hard the hardest instances are with these solvers is
@@
 
-\begin{figure}[tb]
-    \centering\includegraphics{gen-graph-phase-transition.pdf}
+\begin{figure}[t]
+    \centering
+    \includegraphics*{plots/phase-transition.pdf}
+
     \caption{With a fixed pattern graph order of 20, a target graph order of 150, a target edge
@@
 pattern, we can always find a match) to entirely unsatisfiable instances (a maximum clique in this
-order and edge probability of target graph will usually have between 9 and 12 vertices). The dark
-line in \cref{figure:phase-transition} shows that this is the case. For densities of 0.67 or
-greater, no instance is satisfiable; with densities of 0.44 or less, every instance is satisfiable;
-and with a density of 0.55, roughly half the instances are satisfiable.
-
-The light line plots mean search effort using the Glasgow solver: for sparse patterns, the problem
-is trivial, for dense patterns proving unsatisfiability is not particularly difficult, and we see a
+order and edge probability of target graph will usually have between 9 and 12 vertices). The move
+from green circles (satisfiable) to blue crosses (unsatisfiable) in \cref{figure:phase-transition}
+shows that this is the case. For densities of 0.67 or greater, no instance is satisfiable; with
+densities of 0.44 or less, every instance is satisfiable; and with a density of 0.55, roughly half
+the instances are satisfiable.
+
+The line plots mean search effort using the Glasgow solver: for sparse patterns, the problem is
+trivial, for dense patterns proving unsatisfiability is not particularly difficult, and we see a
 complexity peak around the point where half the instances are satisfiable.  We also plot the search
-cost of individual instances, as points. The behaviour we observe looks remarkably similar to random
-3SAT problems---compare, for example, Figure 1 of \citet{LeytonBrown:2014}. In particular,
-satisfiable instances tend to be easier, but show greater variation than unsatisfiable instances, and
-there are exceptionally hard satisfiable instances \citep{Smith:1997}. (The Glasgow solver supports
-parallel search with a work-stealing strategy explicitly designed to eliminate these. We have not
-enabled this option to avoid dealing with the complexity of search tree measurements under
-parallelism.)
+cost of individual instances, as points. The behaviour we observe looks remarkably similar to the
+clique decision problem and to random 3SAT problems---compare, for example,
+Figure~1 of
+\citet{DBLP:journals/cacm/Leyton-BrownHHX14}. In particular, satisfiable instances tend to be easier, but show greater
+variation than unsatisfiable instances, and there are exceptionally hard satisfiable instances
+\citep{DBLP:conf/cp/SmithG97}. (The Glasgow solver supports
+parallel search with a work-stealing strategy explicitly designed to reduce or eliminate these. We
+are not using parallel search to avoid dealing with the complexity of search tree
+measurements under parallelism.)
 
@@
             \node [anchor=center] {}; &
-            \node [anchor=center] {\scriptsize $G(10, x){\rightarrowtail}G(150, y)$}; &
-            \node [anchor=center] {\scriptsize $G(20, x){\rightarrowtail}G(150, y)$}; &
-            \node [anchor=center] {\scriptsize $G(30, x){\rightarrowtail}G(150, y)$}; &
-            \\
-            \node [anchor=center] [rotate=90] {\scriptsize Satisfiable?}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-satisfiable-10-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-satisfiable-20-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-satisfiable-30-150.pdf}}; &
-            \\
-            \node [anchor=center] [rotate=90] {\scriptsize Glasgow}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-nodes-10-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-nodes-20-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-nodes-30-150.pdf}}; &
-            \\
-            \node [anchor=center] [rotate=90] {\scriptsize LAD}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-lad-non-induced-nodes-10-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-lad-non-induced-nodes-20-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-lad-non-induced-nodes-30-150.pdf}}; &
-            \\
-            \node [anchor=center] [rotate=90] {\scriptsize VF2}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-vf2-non-induced-nodes-10-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-vf2-non-induced-nodes-20-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-vf2-non-induced-nodes-30-150.pdf}}; \\
+            \node [anchor=center] {\scriptsize $G(10, x)$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(20, x)$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(30, x)$\hspace*{3em}}; &
+            \\
+            \node [anchor=center, inner sep=4pt] [rotate=90] {\scriptsize Satisfiable?}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-satisfiable-10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-satisfiable-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-satisfiable-30-150.pdf}}; &
+            \\[0.3cm]
+            \node [anchor=center, inner sep=4pt] [rotate=90] {\scriptsize Glasgow}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-nodes-10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-nodes-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-nodes-30-150.pdf}}; &
+            \\[0.3cm]
+            \node [anchor=center, inner sep=4pt] [rotate=90] {\scriptsize LAD}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-lad-non-induced-nodes-10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-lad-non-induced-nodes-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-lad-non-induced-nodes-30-150.pdf}}; &
+            \\[0.3cm]
+            \node [anchor=center, inner sep=4pt] [rotate=90] {\scriptsize VF2}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-nodes-10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-nodes-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-nodes-30-150.pdf}}; \\
         };
@@
         Along the top row, we show the proportion of instances which are satisfiable; the white
-        bands show the phase transitions, and the black lines are our predictions of where the
-        phase transition will occur. On the final three rows, we show the number of search nodes used by the
+        bands show the phase transitions, and the black lines are the predictions using
+        equation~\eqref{equation:non-induced-prediction} of where the
+        phase transition will occur. On the subsequent three rows, we show the number of search nodes used by the
         Glasgow, LAD and VF2 solvers; the dark regions indicate ``really hard'' instances.}
@@
 of 0.01. For each of these points, we generate ten random instances. The colour denotes the
-proportion of these instances which were found to be satisfiable.  Inside the orange region, at the
+proportion of these instances which were found to be satisfiable.  Inside the red region, at the
 bottom right of each plot, every instance is unsatisfiable---here we are trying to find a dense
-pattern in a sparse target. In the purple region, at the top left, every instance is
+pattern in a sparse target. In the green region, at the top left, every instance is
 satisfiable---we are looking for a sparse pattern in a dense target (which is easy, since we only
@@
 general, satisfiable instances are easy, until very close to the phase transition. As we hit the
-phase transition and move into the unsatisfiable region, we see complexity increase. Finally, as
-we pass through the phase transition and move deeper into the unsatisfiable region, instances become
+phase transition and move into the unsatisfiable region, we see complexity increase. Finally, as we
+pass through the phase transition and move deeper into the unsatisfiable region, instances become
 easier again. This behaviour is largely solver-independent, although VF2 has a larger hard region
-than Glasgow or LAD (and the observant reader may notice that VF2 finds some instances extremely
-hard when the target is empty but the pattern is not---this turns out to be extremely important, and
-we return to it in \cref{section:labelled,section:filterverify}). Thus, although we have moved away
-from a single control parameter, we still observe the easy-hard-easy pattern seen in many other
-\NP-complete problems.
+than Glasgow or LAD. VF2 also finds some instances extremely hard when the target is empty but the
+pattern is not---this turns out to be extremely important, and we return to it in
+\cref{section:labelled,section:filterverify}. Thus, although we have moved away from a single
+control parameter, we still observe the easy-hard-easy pattern seen in many \NP-complete
+problems.
 
@@
 to a target $T = G(t, d_t)$, there are \[ t^{\underline{p}} = t \cdot (t - 1) \cdot \ldots \cdot (t -
-p + 1) \] possible assignments of target vertices to pattern vertices.  We expect the pattern to have
-$d_p \cdot \binom{p}{2}$ edges, so we obtain the probability of all of these edges being mapped to
+p + 1) \] possible assignments of target vertices to pattern vertices.  If we assume the pattern
+has exactly $d_p \cdot \binom{p}{2}$ edges, we obtain the probability of all of these edges being mapped to
 edges in the target by raising $d_t$ to this power, giving an expected number of solutions of
-\begin{equation}
+\begin{equation}\label{equation:non-induced-prediction}
     \langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \textnormal{.}
@@
 similar behaviour is seen with random constraint satisfaction problems
-\citep{Smith:1996}.
+\citep{DBLP:journals/ai/SmithD96}.
 
@@
 heuristics for backtracking search algorithms---one of these is to try to maximise the expected
-number of solutions inside any subproblem considered during search \citep{Gent:1996:EN}.  This is
+number of solutions inside any subproblem considered during search \citep{DBLP:conf/cp/GentMPSW96}.  This is
 usually done by cheaper surrogates, rather than direct calculation. When branching, both LAD and
@@
 
-\begin{figure}[tb]
+\begin{figure}[p]
     \centering
@@
             \node [anchor=center] {}; &
-            \node [anchor=center] {\scriptsize $G(10, x)$}; &
-            \node [anchor=center] {\scriptsize $G(14, x)$}; &
-            \node [anchor=center] {\scriptsize $G(15, x)$}; &
-            \node [anchor=center] {\scriptsize $G(16, x)$}; &
-            \node [anchor=center] {\scriptsize $G(20, x)$}; &
-            \node [anchor=center] {\scriptsize $G(30, x)$}; &
+            \node [anchor=center] {\scriptsize $G(10, x)$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(14, x)$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(15, x)$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(16, x)$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(20, x)$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(30, x)$\hspace*{3em}}; &
             \\
             \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Satisfiable?}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-10-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-14-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-15-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-16-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-20-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-30-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-satisfiable-10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-satisfiable-14-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-satisfiable-15-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-satisfiable-16-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-satisfiable-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-satisfiable-30-150.pdf}}; &
             \\
             \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Glasgow}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-10-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-14-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-15-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-16-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-20-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-30-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-nodes-10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-nodes-14-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-nodes-15-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-nodes-16-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-nodes-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-nodes-30-150.pdf}}; &
             \\
             \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize LAD}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-lad-induced-nodes-10-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-lad-induced-nodes-14-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-lad-induced-nodes-15-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-lad-induced-nodes-16-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-lad-induced-nodes-20-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-lad-induced-nodes-30-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-lad-induced-nodes-10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-lad-induced-nodes-14-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-lad-induced-nodes-15-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-lad-induced-nodes-16-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-lad-induced-nodes-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-lad-induced-nodes-30-150.pdf}}; &
             \\
             \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize VF2}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-vf2-induced-nodes-10-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-vf2-induced-nodes-14-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-vf2-induced-nodes-15-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-vf2-induced-nodes-16-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-vf2-induced-nodes-20-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-vf2-induced-nodes-30-150.pdf}}; \\
+            \node [anchor=center] {\includegraphics*{plots/graph-vf2-induced-nodes-10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-vf2-induced-nodes-14-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-vf2-induced-nodes-15-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-vf2-induced-nodes-16-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-vf2-induced-nodes-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-vf2-induced-nodes-30-150.pdf}}; \\
             \\
             \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Constrainedness}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-kappa-10-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-kappa-14-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-kappa-15-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-kappa-16-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-kappa-20-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-kappa-30-150.pdf}}; \\
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-kappa-10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-kappa-14-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-kappa-15-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-kappa-16-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-kappa-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-kappa-30-150.pdf}}; \\
             \\
             \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Complement?}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-which-10-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-which-14-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-which-15-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-which-16-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-which-20-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-which-30-150.pdf}}; \\
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-which-10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-which-14-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-which-15-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-which-16-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-which-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-which-30-150.pdf}}; \\
         };
@@
     \caption{Behaviour of algorithms on the induced variant, shown in the style
-    of \cref{figure:non-induced}. The fifth row plots constrainedness: the darkest region is where
-    $\kappa = 1$, and the lighter regions show where the problem is either over-
-    or under-constrained. The final row shows when the Glasgow algorithm performs better when given
-    the complements of the pattern and target graphs as inputs---the solid lines show the location
-of the phase transition, and the dotted lines are $t_d=0.5$ and the $p_d=t_d$ diagonal.}\label{figure:induced}
+    of \cref{figure:non-induced}. The fifth row plots constrainedness using
+    equation~\eqref{equation:constrainedness}: the
+    darkest region is where $\kappa = 1$, and the lighter regions show where the problem is either
+    over- or under-constrained. The final row shows when the Glasgow algorithm performs better when
+    given the complements of the pattern and target graphs as inputs---the solid lines show the
+    location of the phase transition, and the dotted lines are $t_d=0.5$ and the $p_d=t_d$ diagonal.}\label{figure:induced}
 \end{figure}
@@
 
+We might expect these complexity plots to be symmetric along the diagonal, since for the induced
+problem, if we replace both inputs with their complements, the solutions remain the same.  For the
+Glasgow solver, this is the case. This should be expected, because this complement property is
+precisely how the Glasgow solver handles the induced variant (although the heuristics may differ
+between the two, which we discuss below). For LAD, some of the very dense patterns are slightly
+harder than their diagonal opposites (LAD reasons about degrees, but not about complement-degrees);
+it is interesting to note that for larger target graphs, VF2 finds \emph{all} dense pattern graphs
+difficult.
+
 \subsection{Predictions and Heuristics}
@@
 To predict the location of the induced phase transition, we repeat the argument for locating the
-non-induced phase transition and additionally considering non-edges, to get an expected number of
-solutions of \begin{equation} \langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot
+non-induced phase transition and additionally consider non-edges, to get an expected number of
+solutions of \begin{equation}\label{equation:induced-prediction} \langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot
 {(1 - d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}} \textnormal{.} \end{equation} We plot this using black lines on
@@
 towards the top left), and if we are in the top left we should use the complement instead. This
-goes against a suggestion by \citet{Walsh:1998} that switching heuristics based upon an estimate of
+goes against a suggestion by \citet{DBLP:conf/aaai/Walsh98} that switching heuristics based upon an estimate of
 the solubility of the problem may offer good performance.
@@
 the data suggests that the actual location of the phase transition may be involved (and perhaps
-\citeS{Walsh:1998} suggestion applies only in these conditions). In any case, $\langle Sol
+\citeauthor{DBLP:conf/aaai/Walsh98}'s suggestion applies only in these conditions). In any case, $\langle Sol
 \rangle$ is insufficient to explain the observed behaviour in these two eighths of the parameter space.
@@
 In practice, this is unlikely to be a problem: most real-world instances are extremely sparse and
-are usually easy, which perhaps explains the continuing popularity of VF2's connectivity-based
-heuristics \citep{Carletti:2015}. In this situation, these experiments justify reusing the
+are usually easy. In this situation, these experiments justify reusing the
 non-induced heuristics on induced problems.
@@
             \node [anchor=center] {}; &
-            \node [anchor=center] {\scriptsize $G(10, x)$}; &
-            \node [anchor=center] {\scriptsize $G(12, x)$}; &
-            \node [anchor=center] {\scriptsize $G(14, x)$}; &
-            \node [anchor=center] {\scriptsize $G(16, x)$}; &
-            \node [anchor=center] {\scriptsize $G(18, x)$}; &
-            \node [anchor=center] {\scriptsize $G(25, x)$}; &
+            \node [anchor=center] {\scriptsize $G(10, x)$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(12, x)$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(14, x)$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(16, x)$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(18, x)$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(25, x)$\hspace*{3em}}; &
             \\
             \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Satisfiable?}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-10-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-12-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-14-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-16-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-18-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-satisfiable-25-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-satisfiable-10-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-satisfiable-12-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-satisfiable-14-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-satisfiable-16-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-satisfiable-18-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-satisfiable-25-75.pdf}}; &
             \\
             \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Glasgow}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-10-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-12-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-14-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-16-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-18-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-induced-nodes-25-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-nodes-10-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-nodes-12-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-nodes-14-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-nodes-16-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-nodes-18-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-induced-nodes-25-75.pdf}}; &
             \\
             \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Clasp (PB)}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-clasp-induced-nodes-10-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-clasp-induced-nodes-12-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-clasp-induced-nodes-14-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-clasp-induced-nodes-16-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-clasp-induced-nodes-18-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-clasp-induced-nodes-25-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-clasp-induced-nodes-10-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-clasp-induced-nodes-12-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-clasp-induced-nodes-14-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-clasp-induced-nodes-16-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-clasp-induced-nodes-18-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-clasp-induced-nodes-25-75.pdf}}; &
             \\
             \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Glucose (SAT)}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-glucose-induced-nodes-10-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-glucose-induced-nodes-12-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-glucose-induced-nodes-14-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-glucose-induced-nodes-16-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-glucose-induced-nodes-18-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-glucose-induced-nodes-25-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-glucose-induced-nodes-10-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-glucose-induced-nodes-12-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-glucose-induced-nodes-14-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-glucose-induced-nodes-16-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-glucose-induced-nodes-18-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-glucose-induced-nodes-25-75.pdf}}; &
+            \\
+            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Gurobi (MIP)}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-gurobi-induced-nodes-10-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-gurobi-induced-nodes-12-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-gurobi-induced-nodes-14-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-gurobi-induced-nodes-16-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-gurobi-induced-nodes-18-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-gurobi-induced-nodes-25-75.pdf}}; &
             \\
             \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize BBMC (Clique)}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-clique-induced-nodes-10-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-clique-induced-nodes-12-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-clique-induced-nodes-14-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-clique-induced-nodes-16-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-clique-induced-nodes-18-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-clique-induced-nodes-25-75.pdf}}; &
-            \\
-            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Gurobi (MIP)}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-gurobi-induced-nodes-10-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-gurobi-induced-nodes-12-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-gurobi-induced-nodes-14-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-gurobi-induced-nodes-16-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-gurobi-induced-nodes-18-75.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-gurobi-induced-nodes-25-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-clique-induced-nodes-10-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-clique-induced-nodes-12-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-clique-induced-nodes-14-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-clique-induced-nodes-16-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-clique-induced-nodes-18-75.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-clique-induced-nodes-25-75.pdf}}; &
             \\
@@
     Glasgow algorithm, the third and fourth rows show the number of decisions made by the pseudo-boolean
-and SAT solvers, the fifth shows the number of search nodes used on the clique encoding, and the
-final row the MIP encoding.}\label{figure:alt}
+and SAT solvers, the fifth shows the number of search nodes used on the MIP encoding, and the
+final row the clique encoding.}\label{figure:alt}
 \end{figure}
@@
 
-The pseudo-boolean (PB) encoding is as follows. For each pattern vertex $v$ and each target vertex
+Our pseudo-boolean (PB) encoding is as follows. For each pattern vertex $v$ and each target vertex
 $w$, we have a binary variable which takes the value 1 if and only if $v$ is mapped to $w$.
@@
 each target vertex is mapped to by at most one pattern vertex, that adjacent vertices are mapped to
-adjacent vertices, and that non-adjacent vertices are mapped to non-adjacent vertices. We used the
-Clasp solver~\citep{gekakaosscsc11a} version 3.1.3 to solve the pseudo-boolean instances.  The
+adjacent vertices, and that non-adjacent vertices are mapped to non-adjacent vertices. We use the
+Clasp solver \citep{DBLP:journals/aicom/GebserKKOSS11} version 3.1.3 to solve the pseudo-boolean instances.  The
 instances that are hard for the Glasgow solver remain hard for the PB solver, including instances
-inside the central region, and the easy satisfiable instances remain easy. Similar results were seen
-with the Glucose SAT solver \cite{glucose} using a direct encoding of the cardinality constraints.
-We also implemented an integer program encoding: the Gurobi solver was only able to solve some of
-the trivial satisfiable instances, and was almost never able to prove unsatisfiability within the
-time limit.
-
-The \emph{association graph encoding} of a subgraph isomorphism problem (illustrated in
-\cref{figure:sip}) is constructed by creating a new graph with a vertex for each pair $(p, t)$ of
-vertices from the pattern and target graphs respectively. There is an edge between vertex $(p_1,
-t_1)$ and vertex $(p_2, t_2)$ if mapping $p_1$ to $t_1$ and $p_2$ to $t_2$ simultaneously is
-permitted, i.e.\ $p_1$ is adjacent to $p_2$ if and only if $t_1$ is adjacent to $t_2$. A clique of
-size equal to the order of the pattern graph exists in the association graph if and only if the
-problem is satisfiable \citep{Levi:1973}. We used this encoding with the BBMC clique algorithm
-\citep{SanSegundo:2011}, which we implemented in C++. Usually BBMC solves the optimisation version
-of the clique problem; we adapted it to solve the decision problem by initialising the incumbent to
-be one less than the decision value, and allowing it to exit as soon as a clique with size equal to
-the decision value is encountered.  Again, our results show that the instances in the central region
-remain hard, and additionally, some of the easy unsatisfiable instances become hard.
+inside the central region, and the easy satisfiable instances remain easy. Similar results are seen
+with the Glucose SAT solver \citep{o:glucose} using a direct encoding of the cardinality constraints.
+We also show an integer program encoding: the Gurobi solver is only able to solve some of the
+trivial satisfiable instances, and was almost never able to prove unsatisfiability within the time
+limit.
+
+The \emph{association graph encoding} of a subgraph isomorphism problem is a reduction to the clique
+decision problem. \citet{DBLP:conf/cp/McCreeshNPS16} describe and study this approach in more
+detail. Briefly, the association graph is constructed by creating a new graph with a vertex for each
+pair $(p, t)$ of vertices from the pattern and target graphs respectively. There is an edge between
+vertex $(p_1, t_1)$ and vertex $(p_2, t_2)$ if mapping $p_1$ to $t_1$ and $p_2$ to $t_2$
+simultaneously is permitted, i.e.\ $p_1$ is adjacent to $p_2$ if and only if $t_1$ is adjacent to
+$t_2$. A clique of size equal to the order of the pattern graph exists in the association graph if
+and only if the problem is satisfiable \citep{o:Levi73}. We used this encoding with an
+implementation of \citet{DBLP:journals/cor/SegundoRJ11}'s bit-parallel maximum clique algorithm,
+modified to solve the decision problem rather than the optimisation problem.  Again, our results
+show that the instances in the central region remain hard, and additionally, some of the easy
+unsatisfiable instances become hard.
 
@@
 isomorphism solvers only find this region hard due to not reasoning simultaneously about adjacency
-and non-adjacency, since the association graph encoding constraints consider compatibility rather
-than adjacency and non-adjacency.
+and non-adjacency, since the constraints in the association graph encoding consider compatibility
+rather than adjacency and non-adjacency.
 
@@
 phase transition concept, and to generalise hardness parameters across different combinatorial
-problems \citep{Gent:1996:Kappa}. A problem with $\kappa < 1$ is said to be underconstrained, and is
+problems \citep{DBLP:conf/aaai/GentMPW96}. A problem with $\kappa < 1$ is said to be underconstrained, and is
 likely to be satisfiable; a problem with $\kappa > 1$ is overconstrained, and is likely to be
@@
 is very small or very large are usually easy. By handling injectivity as a restriction on the size
-of the state space rather than as a constraint, we derive \begin{equation} \kappa = 1 - \frac{\log \left(
+of the state space rather than as a constraint, we derive
+\begin{equation}\label{equation:constrainedness}\kappa = 1 - \frac{\log \left(
 t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 - d_{t})}^{(1 - d_{p}) \cdot
@@
 
-\section{Other Models of Randomness}\label{section:models}
+\section{$k$-Regular Graphs}\label{section:models}
 
@@
             \node [anchor=center] {}; &
-            \node [anchor=center] {\scriptsize $\mathcal{G}(10, x){\rightarrowtail}\mathcal{G}(150, y)$}; &
-            \node [anchor=center] {\scriptsize $\mathcal{G}(20, x){\rightarrowtail}\mathcal{G}(150, y)$}; &
-            \node [anchor=center] {\scriptsize $\mathcal{G}(30, x){\rightarrowtail}\mathcal{G}(150, y)$}; &
-            \\
-            \node [anchor=center] [rotate=90] {\scriptsize Satisfiable?}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-kr-satisfiable-10-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-kr-satisfiable-20-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-kr-satisfiable-30-150.pdf}}; &
-            \\
-            \node [anchor=center] [rotate=90] {\scriptsize Glasgow}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-kr-nodes-10-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-kr-nodes-20-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-kr-nodes-30-150.pdf}}; &
-            \\
-            \node [anchor=center] [rotate=90] {\scriptsize LAD}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-lad-non-induced-kr-nodes-10-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-lad-non-induced-kr-nodes-20-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-lad-non-induced-kr-nodes-30-150.pdf}}; &
-            \\
-            \node [anchor=center] [rotate=90] {\scriptsize VF2}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-vf2-non-induced-kr-nodes-10-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-vf2-non-induced-kr-nodes-20-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-vf2-non-induced-kr-nodes-30-150.pdf}}; &
+            \node [anchor=center] {\scriptsize $R(10, x)$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $R(20, x)$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $R(30, x)$\hspace*{3em}}; &
+            \\
+            \node [anchor=center, inner sep=4pt] [rotate=90] {\scriptsize Satisfiable?}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-kr-satisfiable-10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-kr-satisfiable-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-kr-satisfiable-30-150.pdf}}; &
+            \\[0.3cm]
+            \node [anchor=center, inner sep=4pt] [rotate=90] {\scriptsize Glasgow}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-kr-nodes-10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-kr-nodes-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-kr-nodes-30-150.pdf}}; &
+            \\[0.3cm]
+            \node [anchor=center, inner sep=4pt] [rotate=90] {\scriptsize LAD}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-lad-non-induced-kr-nodes-10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-lad-non-induced-kr-nodes-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-lad-non-induced-kr-nodes-30-150.pdf}}; &
+            \\[0.3cm]
+            \node [anchor=center, inner sep=4pt] [rotate=90] {\scriptsize VF2}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-kr-nodes-10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-kr-nodes-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-kr-nodes-30-150.pdf}}; &
             \\
@@
 
-By $\mathcal{G}(n, k)$ we mean a random graph with $k$ vertices, each of which has degree $k$. In
+What about richer graph structures? We now look at what happens with the non-induced problem if we
+use $k$-regular graphs rather than the Erd\H{o}s-R\'enyi model, and then see the effects of
+introducing vertex labels. Both models still exhibit phase transition behaviour, although with
+sufficiently many labels the ``satisfiable'' region now contains a mix of satisfiable and
+unsatisfiable instances.
+
+?? We then move on to more structured instances.
+
+By $R(n, k)$ we mean a random graph with $k$ vertices, each of which has degree $k$. In
 \cref{figure:kr} we recreate \cref{figure:non-induced}, using the NetworkX implementation
-\citep{NetworkX} of the \citet{DBLP:journals/cpc/StegerW99} algorithm to generate regular graphs.
+\citep{o:NetworkX} of the \citet{DBLP:journals/cpc/StegerW99} algorithm to generate regular graphs.
 The $x$-axes range from degree 0 to degree $n - 1$, and the y-axes range from degree 0 to degree
-149.
+149.  The satisfiable / unsatisfiable plots show very similar results to the non-induced problem on
+Erd\H{o}s-R\'enyi random graphs, with a similarly sharp phase transition, and
+again only instances near the phase transition are difficult.
+
+?? Talk about degree heuristics
 
 \section{Labelled Graphs}\label{section:labelled}
-
-\begin{figure}[t]
-    \begin{tikzpicture}[every node/.style={inner sep=0pt, outer sep=0pt}]
-        \matrix {
-            \node [anchor=center] {}; &
-            \node [anchor=center] {\scriptsize No labels}; &
-            \node [anchor=center] {\scriptsize 1 label}; &
-            \node [anchor=center] {\scriptsize 2 labels}; &
-            \node [anchor=center] {\scriptsize 5 labels}; &
-            \node [anchor=center] {\scriptsize 10 labels}; &
-            \node [anchor=center] {\scriptsize 20 labels}; &
-            \\
-            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Satisfiable?}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-satisfiable-20-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-satisfiable-20-l2-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-satisfiable-20-l3-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-satisfiable-20-l5-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-satisfiable-20-l10-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-satisfiable-20-l20-150.pdf}}; &
-            \\
-            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Glasgow}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-nodes-20-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-nodes-20-l2-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-nodes-20-l3-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-nodes-20-l5-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-nodes-20-l10-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-non-induced-nodes-20-l20-150.pdf}}; &
-            \\
-            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize LAD}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-lad-non-induced-nodes-20-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-lad-non-induced-nodes-20-l2-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-lad-non-induced-nodes-20-l3-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-lad-non-induced-nodes-20-l5-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-lad-non-induced-nodes-20-l10-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-lad-non-induced-nodes-20-l20-150.pdf}}; &
-            \\
-            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize VF2}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-vf2-non-induced-nodes-20-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-vf2-non-induced-nodes-20-l2-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-vf2-non-induced-nodes-20-l3-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-vf2-non-induced-nodes-20-l5-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-vf2-non-induced-nodes-20-l10-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{gen-graph-vf2-non-induced-nodes-20-l20-150.pdf}}; &
-            \\
-        };
-    \end{tikzpicture}
-    \caption{On the top row, predicted and actual location of the phase transition for labelled
-    non-induced random subgraph isomorphism, with a pattern order of 20, a target order of 150,
-    varying pattern (x-axis) and target (y-axis) density, and varying numbers of labels. On
-    subsequent rows, the average number of search nodes needed to solve an instance, for three
-    different solvers.}\label{figure:labels}
-\end{figure}
 
@@
 hydrogen atoms with hydrogen atoms, and so on.  We will look at the non-induced variant, as this
-seems to be more common in the literature, and return to the Erd\H{o}s-R\'enyi model.
+seems to be more common in the literature.
+% , and return to the Erd\H{o}s-R\'enyi model.
+
+\subsection{Predictions and Empirical Hardness}
 
 Suppose our labels are drawn randomly from a set $L = \{ 1, \ldots, k \}$, where $k$ is reasonably
-small compared to $p$. By defining \[ \operatorname{V}(P)\vert_n = \{ v \in \operatorname{V}(P) :
-\ell(v) = n \} \] to be the set of vertices with label $k$, we can partition the pattern
-vertices by label into disjoint sets \[\{ \operatorname{V}(P)\vert_1, \ldots,
-\operatorname{V}(P)\vert_k \}\textnormal{\,,}\] each of which is expected to contain
-$\nicefrac{p}{k}$ vertices. Similarly, we may partition the target vertices into disjoint
-sets $\{ \operatorname{V}(T)\vert_1, \ldots, \operatorname{V}(T)\vert_k \}$.
+small compared to $p$. Let $\ell(v)$ be the label of vertex $v$. By defining \[
+    \operatorname{V}(P)\vert_n = \{ v \in \operatorname{V}(P) : \ell(v) = n \} \] to be the set of
+vertices with label $k$, we can partition the pattern vertices by label into disjoint sets \[\{
+    \operatorname{V}(P)\vert_1, \ldots, \operatorname{V}(P)\vert_k \}\textnormal{\,,}\] each of
+which is expected to contain $\nicefrac{p}{k}$ vertices. Similarly, we may partition the target
+vertices into disjoint sets $\{ \operatorname{V}(T)\vert_1, \ldots, \operatorname{V}(T)\vert_k \}$.
 
@@
 
-As before, we expect the pattern to have $d_p \cdot \binom{p}{2}$ edges, so we obtain the
+As before, we expect the pattern to have $d_p \cdot \binom{p}{2}$ edges, and so if we simplify by
+assuming the pattern will have \emph{exactly} this many edges, we obtain the
 probability of all of these edges being mapped to edges in the target by raising $d_t$ to this
-power, giving an expected number of solutions of \begin{equation} \langle Sol \rangle = \left(
+power, giving an estimate of \begin{equation}\label{equation:non-induced-labelled-prediction} \langle Sol \rangle = \left(
     \frac{\Gamma\left(\nicefrac{t}{k} + 1\right)}{\Gamma\left(\nicefrac{t}{k} - \nicefrac{p}{k} +
@@
 
-So how good are these predictions? We draw some heatmaps in \cref{figure:labels}. For small numbers
+\begin{figure}[t]
+    \centering
+    \begin{tikzpicture}[every node/.style={inner sep=0pt, outer sep=0pt}]
+        \matrix {
+            \node [anchor=center] {}; &
+            \node [anchor=center] {\scriptsize No labels\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize 2 labels\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize 3 labels\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize 5 labels\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize 10 labels\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize 20 labels\hspace*{2em}}; &
+            \\
+            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Satisfiable?}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-satisfiable-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-satisfiable-20-l2-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-satisfiable-20-l3-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-satisfiable-20-l5-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-satisfiable-20-l10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-satisfiable-20-l20-150.pdf}}; &
+            \\
+            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Glasgow}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-nodes-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-nodes-20-l2-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-nodes-20-l3-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-nodes-20-l5-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-nodes-20-l10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-non-induced-nodes-20-l20-150.pdf}}; &
+            \\
+            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize LAD}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-lad-non-induced-nodes-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-lad-non-induced-nodes-20-l2-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-lad-non-induced-nodes-20-l3-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-lad-non-induced-nodes-20-l5-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-lad-non-induced-nodes-20-l10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-lad-non-induced-nodes-20-l20-150.pdf}}; &
+            \\
+            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize VF2}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-nodes-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-nodes-20-l2-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-nodes-20-l3-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-nodes-20-l5-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-nodes-20-l10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-nodes-20-l20-150.pdf}}; &
+            \\
+            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize VF2 Failures}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-l2-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-l3-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-l5-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-l10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-l20-150.pdf}}; &
+            \\
+        };
+    \end{tikzpicture}
+    \caption{On the top row, predicted and actual location of the phase transition for labelled
+    non-induced random subgraph isomorphism, with a pattern order of 20, a target order of 150,
+    varying pattern (x-axis) and target (y-axis) density, and varying numbers of labels. On
+    subsequent rows, the average number of search nodes needed to solve an instance, for three
+    different solvers, and on the final row, the proportion of instances aborted due to a timeout
+    with VF2.}\label{figure:labels}
+\end{figure}
+
+So how good are these predictions? The black lines on the first row of heatmaps in
+\cref{figure:labels} plot where we calculate $\langle Sol \rangle = 1$ will occur. For small numbers
 of labels, our predictions are slightly better than in the unlabelled case: there seems to be less
 of a variance problem for very dense patterns. Even as the numbers of labels becomes relatively
-large, the prediction of the phase transition still occurs in the right place, but we start to see
-sporadic unsatisfiable instances deep inside the satisfiable region. With sufficiently many labels,
-we get a kind of phase transition from ``all unsatisfiable'' to ``mixed satisfiable and
-unsatisfiable''.
-
-\subsection{What's Wrong with VF2?}
-
-?? What goes on here? With labels, there are two regions where VF2 often finds instances really
-hard, but Glasgow and LAD do not. Looking at the data more closely, although the average runtime is
-high, there is a wide distribution of runtimes in these areas, and VF2 does sometimes find these
-instances trivial too.
-
-?? Give examples here: VF2 does not detect simple things like there being two red vertices in the
-pattern but one in the target, or even no red vertices in the target. In contrast, Glasgow and LAD
-do. Explain thrashing.
-
-?? Explain what's going on in the satisfiable region. This is a bit more subtle: is it to do with
-neighbours, rather than a root node property? Would be nice if we could find a representative
-example.
-
-?? We call these \emph{spuriously hard} instances, to distinguish them from \emph{really hard}
-problems. Not the same as \emph{exceptionally hard}.
-
-?? Look more at size of min unsat core in these areas?
-
-\subsection{Spuriously Hard Instances?}
-
-?? Use of VF2 in other areas.
-
-?? Need for better benchmark instances, not running millisecond versus microsecond benchmarks.
-
-?? However, by far the biggest problem is in graph databases. Not just lead to overly pessimistic
-conclusions, but has misdirected the design of systems. Note that these deal with sparse, non-random
-graphs.
+large, the prediction of the phase transition still occurs in the right place, but with ten labels
+we start to see sporadic unsatisfiable instances deep inside the satisfiable region. With twenty
+labels, we instead get a kind of phase transition from ``all unsatisfiable'' to ``mixed satisfiable
+and unsatisfiable''. We can understand this intuitively: with sufficiently many labels, we might
+generate, say, a red vertex adjacent to a blue vertex in the pattern, but not in the target.  With
+twenty labels, we even sometimes generate no vertices with a particular label in the target at all.
+This in many ways resembles ``flaws'' generated by certain random constraint satisfaction problem
+instance generators
+\citep{DBLP:journals/constraints/AchlioptasMKSKK01,DBLP:journals/constraints/GentMPSW01}.
+
+What about empirical hardness? As before, some instances on the phase transition are hard for all
+solvers (although as the number of labels increases, the hardest instances become easier).
+Instances far from the phase transition are easy, except for VF2, which occasionally finds
+some of instances with larger numbers of labels very difficult. This behaviour occurs both on
+satisfiable instances, and on the flawed unsatisfiable instances deep inside the ``satisfiable''
+region.  It is interesting to observe that all such unsatisfiable instances that VF2 finds difficult
+have a very small proof of unsatisfiability, with neither Glasgow nor LAD requiring more than one
+hundred recursive calls.
+
+\subsection{Richer Label Models, and VF2's Deficiencies}\label{subsection:richer-labels}
+
+Why does VF2 find some of these instances so hard? VF2 does not track domains, like other
+algorithms, and does not detect a domain wipeout until it branches on a particular vertex.  It also
+cannot detect small domains, and only uses ``adjacency to an existing assignment'' as a branching
+heuristic. This can make it very hard for VF2 to detect that it is in an obviously failed state, or
+that an instance or subproblem is trivially unsatisfiable.
+
+\begin{figure}[t]
+    \centering
+    \includegraphics*{plots/skewed.pdf}
+
+    \caption{Cumulative number of instances solved over time, using the richer model of randomness
+    described in \cref{subsection:richer-labels}.}
+    \label{figure:skewed-cumulative}
+\end{figure}
+
+To illustrate this point further, we will now look at some instances created using a slightly more
+structured random model. We create a family of one thousand labelled instance pairs, as follows. To
+create a pattern graph, we create ten vertices with label zero, with edges between these vertices
+with probability 0.2. We then add another ten vertices, with labels chosen randomly between one and
+thirty, and add edges between these vertices and the zero-labelled vertices with probability 0.1.
+To generate a target, we follow a similar process: we have fifty vertices with label zero and edge
+probability 0.2, fifty vertices with labels randomly between one and thirty, and edges from the
+first set of vertices to the second set with probability 0.3.  This model was selected and the
+parameters tuned to provide a demonstration of particular behaviours of VF2, not because of any natural
+property (although inspiration came from seeking a very crude approximation of chemical graphs,
+which often contain a lot of carbon in the centre, and other atoms around the outside). From our set
+of one thousand such instances, sixty six are satisfiable.
+
+We plot the cumulative number of instances solved over time for these instances in
+\cref{figure:skewed-cumulative}. Both Glasgow and LAD find all of these instances trivial, but VF2
+finds many of them extremely difficult.  The ``VF2${\leftrightarrow}$'' line shows what happens with
+VF2 if the graphs are permuted, so that the non-zero labelled instances are given lower vertex
+numbers rather than higher vertex numbers. This
+makes VF2 behave better, but still extremely poorly compared to the other two solvers; we return to
+this in reference to work by \citet{DBLP:conf/edbt/KatsarouNT17} in the following section.
+It is also important to note for later that both VF2 variations find some satisfiable and some
+unsatisfiable instances extremely hard.
+
+\begin{figure}[tb]
+    \centering
+    \begin{tikzpicture}[scale=0.5,every node/.style={font=\footnotesize}]
+        \begin{scope}
+            \newcount \myc
+            \newcount \myd
+            \newcount \mye
+            \foreach \n in {1, ..., 9}{
+                \myc=\n \advance\myc by -1 \multiply\myc by -360 \divide\myc by 10 \advance\myc by 18.0
+                \node[draw, circle, fill=chromaygb1, inner sep=0.5pt] (N\n) at (\the\myc:2.4) {\phantom{0}};
+            }
+            \node[draw, circle, fill=chromaygb3, inner sep=0.5pt] (N10) at (54.0:2.4) {\phantom{0}};
+
+            \foreach \n in {1, ..., 8}{
+                \myd=\n
+                \myc=\n \advance\myc by 1
+                \foreach \m in {\the\myc, ..., 9}{
+                    \mye=\m
+                    \draw (N\the\myd) -- (N\the\mye);
+                }
+            }
+        \end{scope}
+        \begin{scope}[xshift=9cm]
+            \newcount \myc
+            \newcount \myd
+            \newcount \mye
+            \foreach \n in {1, ..., 9}{
+                \myc=\n \advance\myc by -1 \multiply\myc by -360 \divide\myc by 10 \advance\myc by 18.0
+                \node[draw, circle, fill=chromaygb1, inner sep=0.5pt] (N\n) at (\the\myc:2.4) {\phantom{0}};
+            }
+            \node[draw, circle, fill=chromaygb7, inner sep=0.5pt] (N10) at (54.0:2.4) {\phantom{0}};
+
+            \foreach \n in {1, ..., 8}{
+                \myd=\n
+                \myc=\n \advance\myc by 1
+                \foreach \m in {\the\myc, ..., 9}{
+                    \mye=\m
+                    \draw (N\the\myd) -- (N\the\mye);
+                }
+            }
+        \end{scope}
+    \end{tikzpicture}
+    \caption{A trivially unsatisfiable subgraph isomorphism instance which VF2 finds exponentially
+    difficult. The pattern (left) and target (right) both consist of a clique, plus one extra vertex
+    which has a different label in each graph.}\label{figure:vf2-stupidity}
+\end{figure}
+
+An even more extreme example of VF2's misbehaviour can be seen in \cref{figure:vf2-stupidity}. Here
+we have a pattern graph and a target graph, both of which are cliques plus one isolated vertex, and
+the isolated vertices have different unique labels. This is trivially unsatisfiable (and Glasgow and
+LAD detect this without search), but unless the labelled vertex is given the lowest vertex number
+(so it is branched on first), VF2 takes nearly a million recursive calls to detect this: VF2 will
+try to map every adjacent vertex in the clique before considering the difficult isolated vertex.
+This example can also be extended slightly to fool any simple static heuristic, or any simple label
+counting mechanism.
+
+These experiments further highlight that VF2 occasionally finds some instances which should be easy hard.
+But does this cause problems in practice? The literature suggests that it does. For example,
+\citet{o:Gromping14} uses VF2 in a package for the R statistics language, and states
+\begin{displayquote}``There are (not so many) instances for which creation of a clear design is
+    prohibitively slow in the current implementation that evaluates subgraph isomorphism with the
+    VF2 algorithm \ldots Recent experiences with a few of these showed that the LAD algorithm was
+    very fast in ruling out impossible matches, where VF2 took a long time.''\end{displayquote}
+Similarly, \citet{o:Murray12} uses VF2 inside a compiler, and observes extremely variable
+(and prohibitively high) compile times in some cases due to the expense of subgraph isomorphism
+calls---given the heavily labelled nature of these graphs, we conjecture that any domain-based
+algorithm would eliminate this cost.
+
+However, by far the biggest problem is in graph databases. The following section shows how
+widespread use of VF2 has not just lead to overly pessimistic conclusions regarding performance, but
+has misdirected the design of larger systems.
 
@@
 
-The general problem is, for a set of target graphs, to process a pattern query and return every
-target graph which is subgraph-isomorphic to that pattern. The set of target graphs is usually seen
-as fixed, or at least rarely-changing, whilst the patterns arrive dynamically. This has lead to the
-development of systems which perform extensive computations on the target graphs beforehand, in the
-hopes of reducing the response times for individual pattern queries. The most popular of these
-strategies is a form of indexing which is often named \emph{filter / verify}.
+A particularly common use of subgraph isomorphism algorithms is inside graph databases.  The
+general problem these systems solve is, for a set of target graphs, to process a pattern query and
+return every target graph which is subgraph-isomorphic to that pattern. The set of target graphs is
+usually seen as fixed, or at least rarely-changing, whilst the patterns arrive dynamically. This has
+lead to the development of systems which perform extensive computations on the target graphs
+beforehand, in the hopes of reducing the response times for individual pattern queries. The most
+popular of these strategies is a form of indexing which is often named \emph{filter~/ verify}.
 
@@
 
-The filter / verify approach has an interesting history, of which we now give a very selective and
-incomplete overview. Our description is biased by a general modern understanding of \NP-completeness
-which was not widely known at least at the time of the earlier papers we discuss. The common theme
-of all of the following papers is that pre-computed information is used to eliminate certain
-unsatisfiable instances from consideration, without performing a subgraph isomorphism test.
-
-?? Give a quick example here of how an index would eliminate the two red vertices in the pattern but
-one red vertex in the target example.
+The filter~/ verify approach has an interesting history, of which we now give a very selective and
+incomplete overview. Our description is biased by a general modern understanding of the empirical
+hardness of \NP-complete problems, which was not widely known at least at the time of the earlier
+papers we discuss. The common theme of all of the following papers is that pre-computed information
+is used to eliminate certain unsatisfiable instances from consideration, without performing a
+subgraph isomorphism test. For example, an index might contain a bit of information expressing
+whether a target graph contains two red vertices. When a pattern graph with two red vertices is used
+as a query, any target whose feature set does not have this bit set would not be considered, and so
+a subgraph isomorphism call would not be made for that pattern / target pair. (In practice, feature
+hashing is often used, which can lead to false positives, but this is not relevant to our
+discussion.)
 
@@
 experiments do not answer whether the indexing is effective. However, the work was influenced by a commercial
-graph database system, whose documentation \cite[section 7.1]{Daylight} states that indexing is
+graph database system, whose documentation \citep{o:Daylight} states that indexing is
 used to minimise disk accesses.
@@
 graph indices in order to help processing graph queries.''\end{displayquote}
-
 With what we now know about the behaviour of modern subgraph isomorphism algorithms, and the nature
@@
 cannot immediately detect unsatisfiability in simple cases like there being two red vertices in the
-pattern but only one in the target\footnote{?? expand on: Although interestingly, this algorithm
-effectively does forward-checking and decent value ordering heuristics before these concepts
-appeared in the constraints literature.}.
+pattern but only one in the target\footnote{Although interestingly, this algorithm effectively does
+forward-checking and has a variable-ordering heuristic, before these concepts appeared in the
+constraints literature.}.
 
@@
 \citet{DBLP:conf/icdcs/CaoYWRL11} propose a privacy-preserving cloud graph database system using
-filter / verify, stating that ``checking subgraph isomorphism is \NP-complete, and therefore it is
+filter~/ verify, stating that ``checking subgraph isomorphism is \NP-complete, and therefore it is
 infeasible to employ such a costly solution'' which simply checks every graph for a match.
@@
 
-The filter / verify paradigm also influences other research. For example, it is used by
+The filter~/ verify paradigm also influences other research. For example, it is used by
 \citet{DBLP:conf/icde/TianP08} in an approximate subgraph searching system: they state that
@@
 are now target graphs). \citet{DBLP:journals/tkde/HongZLY15} look at graph database subgraph
-matching with an additional set similarity constraint, and state that Ullmann's algorithm and VF2
+matching with an additional set-similarity constraint, and state that Ullmann's algorithm and VF2
 are ``usually costly for large graphs'' because they ``do not utilize any index structure''. They
@@
 
-After some early ambiguity, it becomes clear that the intent behind filter / verify systems is to
-reduce the number of subgraph isomorphism calls, and that the cost of loading graphs from disk is
+After some early ambiguity, then, it becomes clear that the intent behind filter~/ verify systems is
+to reduce the number of subgraph isomorphism calls, and that the cost of loading graphs from disk is
 not considered to be problematic. It is worth noting that the entire test datasets from most of
@@
 propagator \citep{DBLP:conf/aaai/Regin94}.  However, as \citet{DBLP:journals/pvldb/KatsarouNT15}
-note, some variation on VF2 \citep{Cordella:2004} is the usual matching algorithm of choice for
-graph database systems, although \citeS{DBLP:journals/jacm/Ullmann76}'s algorithm is sometimes
-chosen. Other approaches have been considered: for example, \citet{DBLP:journals/pvldb/ShangZLY08}
-propose an algorithm which makes use of the frequency of various features to guide search.
+note, some variation on VF2 \citep{DBLP:journals/pami/CordellaFSV04} is the usual matching algorithm of choice for
+graph database systems, although \citeauthor{DBLP:journals/jacm/Ullmann76}'s algorithm is sometimes
+chosen. Other approaches have been considered, albeit not with algorithms strong enough to
+establish all-difference. For example, \citet{DBLP:journals/pvldb/ShangZLY08}
+propose an algorithm which makes use of the frequency of various features to guide search;
 \citet{DBLP:journals/pvldb/LeeHKL12} determine experimentally that this technique tends to be very
-effective, even on families of graphs for which it was not designed.) We believe it would be
-unlikely to cause too much astonishment if we suggested that a better matching algorithm could be
-dropped in as a black box replacement in graph databases systems to improve their performance.
-However, it appears not to have occurred to the graph databases community that a better subgraph
-isomorphism algorithm could invalidate the premise underlying the entire filter / verify approach.
-
-?? Eliminating this filtering step would not merely be a matter of simplicity: indices take ??  hours
-to build.
-
-?? Additionally, this stuff works on satisfiable instances too. Indexing addresses some of the
-obviously-unsatisfiable cases that VF2 struggles with, but does nothing on the satisfiable
-instances.
-
-\subsection{An Algorithm which Requires No Filtering}
-
-?? We build up a simple algorithm. Use smallest domain first, all different, NDS, possibly distance
-2 filtering. Show its effectiveness at eliminating unsatisfiable instances from graph database test
-sets, and give its total performance.  None of the steps in this algorithm are new. Our aim in this
-section is to show which aspects of modern algorithm design are critical, and that an indexless
-sequential scan is entirely plausible with a decent algorithm.
-
-?? To tie this into earlier parts of the paper, we also show what happens if we use the VF2
-heuristic in this setting, rather than smallest domains first.
-
-?? This stuff eliminates all the spuriously hard unsatisfiable instances along the bottom of the
-graph, but also helps on the satisfiable instances that an indexing system cannot touch.
-
-\subsection{What Should the Future of Graph Matching in Databases Be?}
-
-?? Clearly lots of scope for cooperation between AI and graph databases. AI is not just good for
-better algorithms, but the design of entire systems.
-
-?? We are not presenting the ultimate big data subgraph matching algorithm.
-
-?? We need optimisations for trivial cases. Lazy forward checking, cheap alternatives to heuristics
-and all-different. Presolving. Huge domains.  Precalculating NDS, distance 2 etc, not for indexing.
-Edge labels.
-
-?? Degrade to clique. Symmetries.
-
-?? Invariant which are effective on labels used to inform algorithm design and to provide filtering
-during search. Look at minimum unsat core. If we are prepared to spend hours, indexing-like system
-can be used to enhance propagation, for example by pre-computing degree sequences and paths on
-target graphs. This is stronger, because, say, counting red vertices and counting degrees are
-automatically combined.
-
-?? On the other hand, our key message is not just saying use a better algorithm. Need a full
-understanding of every aspect of the system, and cannot treat subgraph matching as a black box.
+effective, even on families of graphs for which it was not designed.
+
+We therefore believe it would be unlikely to cause too much astonishment if we suggested that a
+better matching algorithm could be dropped in as a black box replacement in graph databases systems
+to improve their performance.  This is not our claim. Instead, this paper shows that better
+algorithms invalidate the flawed premise underlying the entire filter~/ verify approach. We will now
+demonstrate that filter~/ verify is simply a poor workaround for the kinds of deficiency in VF2
+demonstrated towards the end of the previous section.
+
+\subsection{Is Filtering Necessary?}
+
+To demonstrate that a pure subgraph isomorphism approach is feasible, with no indexing or supporting
+preprocessing, we look at four datasets commonly used in graph indexing evaluations
+\citep{o:10.1371/journal.pone.0076911}. We do not claim that these are high-quality datasets or
+that the associated queries are sensible, merely that they are widely used.
+
+\begin{itemize}
+    \item The AIDS dataset contains graphs representing 40,000 chemical molecules. These graphs are
+        labelled, and are fairly small (mean 45 vertices) and sparse. Following
+        \citeauthor{o:10.1371/journal.pone.0076911}, the queries are compounds with 8, 16 or 32
+        edges.
+    \item The PDBS dataset \citep{o:HeLCBBSKMR02} contains 30 labelled graphs representing DNA, RNA,
+        and proteins, each duplicated twenty times. These
+        can be relatively large, having up to tens of thousands of vertices, but are extremely
+        sparse. The queries are randomly selected connected subgraphs and do not have a real-world
+        meaning.
+    \item The PCM dataset \citep{o:doi:10.1093/bioinformatics/btr163} contains 50 protein contact
+        maps, each duplicated four times. These graphs have under a thousand vertices and below
+        twenty thousand edges; they are slightly less sparse than the previous two datasets. As for
+        PDBS, the queries are randomly generated and do not have a particular meaning.
+    \item The PPI dataset contains 20 protein interaction networks, with up to ten thousand
+        vertices. The queries have either four or eight vertices.
+\end{itemize}
+
+\begin{figure}[t]
+    \centering
+    \hspace*{1em}
+    \includegraphics*{plots/biiiig-data-aids.pdf}
+    \hfill
+    \includegraphics*{plots/biiiig-data-pcms.pdf}
+    \hspace*{1em}
+
+    \vspace*{1em}
+
+    \centering
+    \hspace*{1em}
+    \includegraphics*{plots/biiiig-data-pdbs.pdf}
+    \hfill
+    \includegraphics*{plots/biiiig-data-ppigo.pdf}
+    \hspace*{1em}
+
+    \caption{Cumulative number of instances solved as a function of search space size, using four
+    graph database datasets and a simple domain-based subgraph isomorphism algorithm. Note that the
+    $x$-axis shows recursive calls, and does \emph{not} use a log scale.}
+    \label{figure:biiiig-data}
+\end{figure}
+
+Rather than the modern techniques discussed used by Glasgow and LAD, we deliberately select a very
+simple starting point: a forward-checking algorithm which uses the smallest domain first heuristic,
+the simple counting all-different propagator, and degree and neighbourhood degree sequence filtering
+at the top of search. Even this simple algorithm finds every instance nearly trivial, and no
+exponential behaviour is observed. As we show in \cref{figure:biiiig-data}, the \emph{hardest}
+instance for the AIDS dataset requires 59 recursive calls, and 146,175 of the instances can be
+solved without search (we measure recursive calls to demonstrate hardness, rather than algorithm
+tuning and good constant factors; $10^4$ to $10^6$ recursive calls per second is feasible).  For
+PDBS, the hardest problem requires 1,776 recursive calls (which occurs 20 times, due to duplicated
+queries), and none of the rest require more than 182 calls. For PCM the hardest problem requires
+3,261 calls, and for PPI, 26. In other words, none of these instances are in any way computationally
+hard even for a simple domain-based algorithm, even before introducing stronger filtering,
+inference, or heuristics. Furthermore it is certainly not the case that a sequential scan is
+infeasible as so many filter~/ verify papers claim.  The exponential behaviour seen by others in
+these instances is purely down to a bad choice of subgraph isomorphism algorithm, and these
+instances are not inherently ``really hard''.
+
+There are at least four benefits beyond simplicity towards abolishing filter~/ verify:
+
+\begin{itemize}
+    \item Domain filtering is useful on both satisfiable and unsatisfiable instances, whilst
+        filtering can only eliminate trivially unsatisfiable instances, and is entirely wasted on
+        satisfiable instances. Domain filtering is also useful even on relatively hard instances,
+        since the information cuts down the search space rather than providing a simple ``yes'' or
+        ``no''.
+
+    \item Domain-based heuristics are much stronger than VF2's adjacency branching rules. Picking
+        from small domains dynamically allows search to focus on the hardest part of the problem.
+
+    \item This approach automatically combines features. For example, a pattern~/ target pair may
+        have matching label features, and matching degree features, but if the only red pattern
+        vertex has degree three whilst no red target vertex has degree more than two, domain
+        filtering will detect this immediately. When combined with all-different propagation and
+        maintained during search, this effect is even stronger.
+
+    \item Finally, as indexing systems get more and more complex in an attempt to filter out a few more
+        instances where VF2 performs poorly, the cost of index construction and maintenance is
+        considerable.  Indices proposed in the literature often take many hours to build, and can
+        consume much more space than the original graphs.
+\end{itemize}
+
+In other words, using domain-based algorithms would not simply be a viable alternative to filter~/
+verify with VF2, but rather would be a much better solution.
+
+\subsection{Rethinking Graph Matching for Database Systems}
+
+The evidence so far clearly shows a need for graph database systems to be redesigned making use of
+knowledge from constraint programming and artificial intelligence, and that this knowledge must inform
+the systems as a whole, and not just parts of them.  Having demonstrated that filter~/ verify is
+both theoretically and practically a flawed approach to designing graph databases based upon a
+repeated misconception that \NP-completeness means long runtimes, we now briefly discuss how future
+graph database systems should be designed, and how such systems can help support strong subgraph
+isomorphism algorithms.
+
+First, though, we briefly consider whether other lessons can be learned from constraint programming:
+unfortunately indexing is not the only incorrect design choice being in these systems. For example,
+recent work by \citet{DBLP:conf/edbt/KatsarouNT17} looks at what they call ``straggler'' queries,
+which are the small subset of queries that they observe taking much longer than others to solve. The
+authors present an approach to address this perceived problem that they describe as novel: ``instead
+of trying to come up with new algorithms for sub-iso testing, we utilize isomorphic query rewritings
+and existing alternative algorithms in parallel''. Their first claim is that permuting graphs before
+running the subgraph isomorphism algorithm can lead to ``wildly different'' execution times. They
+note that VF2 does not define ``a strict order in which the nodes of the query are matched'', and so
+permuting the graphs (for example, by using degree or label frequency information) before search can
+sometimes give orders of magnitude improvements. The connection to variable- and value-ordering
+heuristics is not noted, and no consideration is given to even the simplest dynamic ordering
+heuristics like smallest domain first \citep{DBLP:journals/ai/HaralickE80}. A little thought shows
+that of the orderings proposed, those involving placing rare labels first are effectively poor
+approximations to a static ``smallest domain at top of search first'' ordering, whilst the remainder
+use degree as we did earlier in this paper. However, \citeauthor{DBLP:conf/edbt/KatsarouNT17} do
+not investigate any algorithm which employs domains, let alone strong variable- or value-ordering
+heuristics or all-different propagation, and do not consider that their apparent successes could be
+due to VF2's weaknesses rather than an inherent property of \NP-completeness. We saw in the previous
+section that although permuting input graphs could improve VF2's behaviour on some instances, doing
+so does not make its performance come close to that of domain-based algorithms.
+\citeauthor{DBLP:conf/edbt/KatsarouNT17} say they ``hope that our findings will open up new research
+directions, striving to find appropriate, per-query, isomorphic rewritings, in combination with
+alternate per-query sub-iso algorithms that can yield large improvements''. We believe it is
+important instead to emphasis previous research directions that have already solved most of this
+problem, before any further research effort is wasted.
+
+\citet{DBLP:conf/edbt/KatsarouNT17}'s second claim is that ``different algorithms find different
+queries hard''. To address this, they run many subgraph isomorphism algorithms and input
+permutations in parallel; the extensive literature on parallel portfolios in general
+\citep{DBLP:journals/ai/GomesS01}, and the approach by \citet{DBLP:conf/sls/BattitiM07} for subgraph
+isomorphism in particular, is not noted (and nor are portfolios mentioned when they say that ``using
+machine learning models to predict which version \ldots to employ per query is of high interest'').
+The evidence so far in this paper suggests that we should be wary. It is certainly true that there
+are some instances that all algorithms find hard, and there are good theoretical reasons to believe
+that these instances genuinely are really hard.  Furthermore, algorithm portfolios are also
+well-known to be a successful technique, and are beneficial even with modern subgraph isomorphism
+algorithms \citep{DBLP:conf/lion/KotthoffMS16}.  However, in \cref{section:labelled} we saw that there
+are many instances that VF2 finds hard that should not be hard, and that are not hard for other
+algorithms.  If permuting graphs sometimes addresses the difficulty of some of these instances, then
+it is likely that they are not genuinely hard instances at all, and perhaps it would be better
+simply to start using an algorithm with domain tracking and domain-based ordering heuristics, rather
+than assembling a portfolio of bad algorithms in the hopes that at least one of them will often get
+lucky.
+
+To test this suggestion, we return briefly to the experiments on the datasets discussed earlier in
+this section. What if we had not used a smallest domain first heuristic, and instead used only
+degree-based heuristics explained earlier in this paper? Although we still find all four of these
+datasets easy, and the changes to the runtimes are hard to detect (more time is spent in
+initialisation than in search), the worst case in the AIDS dataset now takes 4,051 recursive calls
+as opposed to 59.  How about if, additionally, we do not detect domain wipeouts until attempting to
+branch on an empty variable?  Suddenly three of the four datasets become extremely hard, with many
+instances now timing out after making hundreds of millions of recursive calls; the PPI dataset now
+requires up to 168,451 calls for some queries. Furthermore, slight changes to the input vertex
+ordering can now make some of these ``hard'' instances easy again.
+
+This backs up our suspicions: the apparent success of \citet{DBLP:conf/edbt/KatsarouNT17}'s
+technique is due to the ease of sometimes getting better results out of a poor algorithm, rather
+than being an inherent hardness property, and it would not help on really hard problem instances. We
+point to \citet{DBLP:journals/heuristics/Gent98}'s admonition on how heuristics are benchmarked as
+being similarly relevant here: \begin{displayquote}``Benchmark problems should be hard. I report on
+    the solution of the five open benchmark problems introduced \ldots for testing bin packing
+    problems. Since the solutions were found either by hand or by using very simple heuristic
+    methods, these problems would appear to be easy. In four cases I give improved packings to
+    refute conjectures that previously reported packings were optimal, and I give a proof that the
+    fifth conjecture was correct. \ldots Future experimenters should be careful to perform tests on
+problems that can reasonably be regarded as hard.''\end{displayquote} Claiming huge improvements
+from an algorithm portfolio consisting only of variations of poor algorithms does not demonstrate a
+genuine improvement over the state of the art.
+
+We have seen, then, that domain tracking and domain-based ordering heuristics are critical, and we
+have seen that these alone are enough to make indexing and other techniques irrelevant. There
+is therefore no point in continued research such as that by
+\citet{DBLP:conf/gbrpr/CarlettiFV15,o:Carletti16}
+into tweaking VF2's connectivity-based heuristics.  However, we do not claim that the ultimate
+``big data'' subgraph matching algorithm already exists: on the contrary, there is likely to be
+plenty of room for future improvement now that we understand the importance of getting algorithm
+design right.  For example, a major disadvantage of using domains is the relatively expensive
+initialisation costs, which quickly add up when dealing with large numbers of trivial instances.
+Employing a presolver is an obvious approach---and VF2 is actually good in this role
+\citep{DBLP:conf/lion/KotthoffMS16}---but there are other possibilities. For example,
+minimal or lazy forward-checking
+\citep{DBLP:conf/ictai/DentM94,DBLP:conf/cp/BacchusG95,o:Dent96,DBLP:conf/ecai/LarrosaM98} avoids
+constructing every domain upfront, although adopting this require alternatives to smallest domain first
+and to all-different. Such an approach may also be beneficial for huge target graphs, where having
+domains range over the entire target is impractical.
+
+There is also scope for precalculating supporting information about target graphs. For example,
+neighbourhood degree sequences and supplemental graphs could both be pre-computed and stored. The aim
+here is to reduce the initialisation costs of a good subgraph isomorphism algorithm, and not to
+provide indexing (although additionally using this information as an index may not hurt, if
+initialisation is still costly).
+
+Our central message, though, is not just to say that graph database systems must use a better
+algorithm.  Instead, this paper shows that such systems need to be designed with a better
+understanding of the empirical hardness of \NP-complete problems, that subgraph isomorphism
+algorithms must not be treated as a black box, and that lessons learned in constraint programming
+and artificial intelligence must not go unheeded in other domains.
 
@@
 
-It is worth noting that this technique does \emph{not} give a way of generating hard instances for
-graph isomorphism problems: the pattern graph must be substantially smaller than the target graph
-for independent pairs of randomly generated instances to give interesting behaviour.
-
-?? This is not purely of theoretical interest. Stuff about graph databases.
-
-\section*{Acknowledgements}
-
-The authors wish to thank Kitty Meeks and Craig Reilly for their comments.
+When labels were introduced, we saw that VF2 finds some instances hard which other algorithms find
+easy. We looked at this kind of instance in more detail, and argued that this was due to flaws in
+VF2's design, rather than an interesting property of \NP-completeness. Although inevitably there
+will be instances that every algorithm finds hard, we see no excuse for an algorithm to exhibit
+exponential behaviour in the case when there are two red vertices in a pattern and only one in a
+target.
+
+Unfortunately, we saw that this aspect of VF2's performance has had practical consequences, most
+notably being the misdesign of graph database systems. By not trying even the simplest constraint
+programming techniques from the literature, and ignoring all that is known about the empirical
+hardness of \NP-complete problems, members of the graph databases community have misled themselves
+into believing that extensive research into supporting techniques is important, whereas really all
+they are doing is working around some but not all of the defects in their choices of subgraph
+isomorphism algorithms. With this new understanding of what does and does not make subgraph
+isomorphism hard, it is time for a radical rethink of how graph database systems work.
 
@@
 \appendix
-
 \section{Reproducing These Experiments}

### 2017-07-22 paper/paper.tex
Figure placement
We have new authors

@@
 \author{\name Ciaran McCreesh \thanks{This work was supported by the Engineering and Physical Sciences
-           Research Council [grant numbers EP/K503058/1 and EP/M508056/1]} \email
+           Research Council [grant numbers EP/K503058/1, EP/M508056/1, and EP/P026842/1]} \email
            ciaran.mccreesh@glasgow.ac.uk \\
-       \name Patrick Prosser \email patrick.prosser@glasgow.ac.uk \\
-       \name James Trimble \samethanks \email j.trimble.1@research.gla.ac.uk \\
+       \addr University of Glasgow, Glasgow, Scotland \\
+       \name Patrick Prosser \samethanks[2] \email patrick.prosser@glasgow.ac.uk \\
+       \addr University of Glasgow, Glasgow, Scotland \\
+       \name Christine Solnon \thanks{This work was supported
+       by the ANR project SoLStiCe (ANR-13-BS02-0002-01)} \email christine.solnon@insa-lyon.fr \\
+       \addr INSA-Lyon, LIRIS, UMR5205, F-69621, France \\
+       \name James Trimble \samethanks[2] \email j.trimble.1@research.gla.ac.uk \\
        \addr University of Glasgow, Glasgow, Scotland}
@@
 
+\begin{figure}[t]
+    \centering
+    \begin{tikzpicture}[scale=0.3,every node/.style={font=\footnotesize}]
+        \begin{scope}
+            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (Na) at (1,  0) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (Nb) at (1, -2) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (Nc) at (0, -4) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (Nd) at (2, -4) {\phantom{0}};
+
+            \draw [ultra thick] (Na) -- (Nb);
+            \draw [ultra thick] (Nb) -- (Nc);
+            \draw [ultra thick] (Nc) -- (Nd);
+            \draw [ultra thick] (Nb) -- (Nd);
+
+            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (N1) at (7.5,  0) {\phantom{0}};
+            \node[draw, circle, fill=white, inner sep=0.5pt] (N2) at (9.5,  0) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (N3) at (7.5, -2) {\phantom{0}};
+            \node[draw, circle, fill=white, inner sep=0.5pt] (N4) at (9.5, -2) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (N5) at (7.5, -4) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (N6) at (9.5, -4) {\phantom{0}};
+            \node[draw, circle, fill=white, inner sep=0.5pt] (N7) at (5.5,  0) {\phantom{0}};
+            \node[draw, circle, fill=white, inner sep=0.5pt] (N8) at (5.5, -2) {\phantom{0}};
+            \node[draw, circle, fill=white, inner sep=0.5pt] (N9) at (5.5, -4) {\phantom{0}};
+
+            \draw [] (N1) -- (N2);
+            \draw [ultra thick] (N1) -- (N3);
+            \draw [] (N1) -- (N4);
+            \draw [] (N2) -- (N4);
+            \draw [ultra thick] (N3) -- (N5);
+            \draw [ultra thick] (N3) -- (N6);
+            \draw [] (N4) -- (N6);
+            \draw [ultra thick] (N5) -- (N6);
+            \draw [] (N2) to [in=45, out=315] (N6);
+            \draw [] (N1) -- (N7);
+            \draw [] (N5) -- (N9);
+            \draw [] (N7) -- (N8);
+            \draw [] (N8) -- (N9);
+        \end{scope}
+        \begin{scope}[xshift=20cm]
+            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (Na) at (1,  0) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (Nb) at (1, -2) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (Nc) at (0, -4) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (Nd) at (2, -4) {\phantom{0}};
+
+            \draw [ultra thick] (Na) -- (Nb);
+            \draw [ultra thick] (Nb) -- (Nc);
+            \draw [ultra thick] (Nc) -- (Nd);
+            \draw [ultra thick] (Nb) -- (Nd);
+
+            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (N1) at (7.5,  0) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (N2) at (9.5,  0) {\phantom{0}};
+            \node[draw, circle, fill=white, inner sep=0.5pt] (N3) at (7.5, -2) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (N4) at (9.5, -2) {\phantom{0}};
+            \node[draw, circle, fill=white, inner sep=0.5pt] (N5) at (7.5, -4) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (N6) at (9.5, -4) {\phantom{0}};
+            \node[draw, circle, fill=white, inner sep=0.5pt] (N7) at (5.5,  0) {\phantom{0}};
+            \node[draw, circle, fill=white, inner sep=0.5pt] (N8) at (5.5, -2) {\phantom{0}};
+            \node[draw, circle, fill=white, inner sep=0.5pt] (N9) at (5.5, -4) {\phantom{0}};
+
+            \draw [ultra thick] (N1) -- (N2);
+            \draw [] (N1) -- (N3);
+            \draw [ultra thick] (N1) -- (N4);
+            \draw [ultra thick] (N2) -- (N4);
+            \draw [] (N3) -- (N5);
+            \draw [] (N3) -- (N6);
+            \draw [ultra thick] (N4) -- (N6);
+            \draw [] (N5) -- (N6);
+            \draw [ultra thick, dashed] (N2) to [in=45, out=315] (N6);
+            \draw [] (N1) -- (N7);
+            \draw [] (N5) -- (N9);
+            \draw [] (N7) -- (N8);
+            \draw [] (N8) -- (N9);
+        \end{scope}
+    \end{tikzpicture}
+
+    \caption{On the left, an induced subgraph isomorphism. On the right, a non-induced subgraph
+    isomorphism: the extra dashed edge is not present in the pattern graph.}
+    \label{figure:sip}
+\end{figure}
+
 However, subgraph isomorphism algorithms cannot handle \emph{arbitrary} instances involving hundreds
@@
 case that a relatively small random graph is unlikely to appear in a larger random graph.)
-
-\begin{figure}[b]
-    \centering
-    \begin{tikzpicture}[scale=0.3,every node/.style={font=\footnotesize}]
-        \begin{scope}
-            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (Na) at (1,  0) {\phantom{0}};
-            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (Nb) at (1, -2) {\phantom{0}};
-            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (Nc) at (0, -4) {\phantom{0}};
-            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (Nd) at (2, -4) {\phantom{0}};
-
-            \draw [ultra thick] (Na) -- (Nb);
-            \draw [ultra thick] (Nb) -- (Nc);
-            \draw [ultra thick] (Nc) -- (Nd);
-            \draw [ultra thick] (Nb) -- (Nd);
-
-            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (N1) at (7.5,  0) {\phantom{0}};
-            \node[draw, circle, fill=white, inner sep=0.5pt] (N2) at (9.5,  0) {\phantom{0}};
-            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (N3) at (7.5, -2) {\phantom{0}};
-            \node[draw, circle, fill=white, inner sep=0.5pt] (N4) at (9.5, -2) {\phantom{0}};
-            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (N5) at (7.5, -4) {\phantom{0}};
-            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (N6) at (9.5, -4) {\phantom{0}};
-            \node[draw, circle, fill=white, inner sep=0.5pt] (N7) at (5.5,  0) {\phantom{0}};
-            \node[draw, circle, fill=white, inner sep=0.5pt] (N8) at (5.5, -2) {\phantom{0}};
-            \node[draw, circle, fill=white, inner sep=0.5pt] (N9) at (5.5, -4) {\phantom{0}};
-
-            \draw [] (N1) -- (N2);
-            \draw [ultra thick] (N1) -- (N3);
-            \draw [] (N1) -- (N4);
-            \draw [] (N2) -- (N4);
-            \draw [ultra thick] (N3) -- (N5);
-            \draw [ultra thick] (N3) -- (N6);
-            \draw [] (N4) -- (N6);
-            \draw [ultra thick] (N5) -- (N6);
-            \draw [] (N2) to [in=45, out=315] (N6);
-            \draw [] (N1) -- (N7);
-            \draw [] (N5) -- (N9);
-            \draw [] (N7) -- (N8);
-            \draw [] (N8) -- (N9);
-        \end{scope}
-        \begin{scope}[xshift=20cm]
-            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (Na) at (1,  0) {\phantom{0}};
-            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (Nb) at (1, -2) {\phantom{0}};
-            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (Nc) at (0, -4) {\phantom{0}};
-            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (Nd) at (2, -4) {\phantom{0}};
-
-            \draw [ultra thick] (Na) -- (Nb);
-            \draw [ultra thick] (Nb) -- (Nc);
-            \draw [ultra thick] (Nc) -- (Nd);
-            \draw [ultra thick] (Nb) -- (Nd);
-
-            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (N1) at (7.5,  0) {\phantom{0}};
-            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (N2) at (9.5,  0) {\phantom{0}};
-            \node[draw, circle, fill=white, inner sep=0.5pt] (N3) at (7.5, -2) {\phantom{0}};
-            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (N4) at (9.5, -2) {\phantom{0}};
-            \node[draw, circle, fill=white, inner sep=0.5pt] (N5) at (7.5, -4) {\phantom{0}};
-            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (N6) at (9.5, -4) {\phantom{0}};
-            \node[draw, circle, fill=white, inner sep=0.5pt] (N7) at (5.5,  0) {\phantom{0}};
-            \node[draw, circle, fill=white, inner sep=0.5pt] (N8) at (5.5, -2) {\phantom{0}};
-            \node[draw, circle, fill=white, inner sep=0.5pt] (N9) at (5.5, -4) {\phantom{0}};
-
-            \draw [ultra thick] (N1) -- (N2);
-            \draw [] (N1) -- (N3);
-            \draw [ultra thick] (N1) -- (N4);
-            \draw [ultra thick] (N2) -- (N4);
-            \draw [] (N3) -- (N5);
-            \draw [] (N3) -- (N6);
-            \draw [ultra thick] (N4) -- (N6);
-            \draw [] (N5) -- (N6);
-            \draw [ultra thick, dashed] (N2) to [in=45, out=315] (N6);
-            \draw [] (N1) -- (N7);
-            \draw [] (N5) -- (N9);
-            \draw [] (N7) -- (N8);
-            \draw [] (N8) -- (N9);
-        \end{scope}
-    \end{tikzpicture}
-
-    \caption{On the left, an induced subgraph isomorphism. On the right, a non-induced subgraph
-    isomorphism: the extra dashed edge is not present in the pattern graph.}
-    \label{figure:sip}
-\end{figure}
 

### 2017-07-22 paper/paper.tex
Christine's comments

@@
 \begin{abstract}
-    Despite subgraph isomorphism being \NP-complete, modern algorithms can operate comfortably on
-    most real-world graphs with thousands of vertices. However, they cannot handle arbitrary instances of
-    this size. We show how to generate ``really hard'' random instances for subgraph isomorphism
-    problems, which are computationally challenging with only a couple of hundred vertices in the
-    target, and only twenty to thirty pattern vertices. For the non-induced version of the problem,
-    these instances lie on a satisfiable / unsatisfiable phase transition, whose location we can
-    predict; for the induced variant, much richer behaviour is observed, and constrainedness gives a
-    better measure of difficulty than does proximity to a phase transition. In both cases, we can
-    use this to understand and improve variable and value ordering heuristics.
+    The subgraph isomorphism problem involves finding a copy of a pattern graph inside a larger
+    target graph. The non-induced version allows extra edges in the target, whilst the induced
+    version does not.  Although both subgraph isomorphism variants are \NP-complete, modern
+    algorithms can operate comfortably on most real-world graphs with thousands of vertices.
+    However, they cannot handle arbitrary instances of this size. We show how to generate ``really
+    hard'' random instances for subgraph isomorphism problems, which are computationally challenging
+    with only a couple of hundred vertices in the target, and only twenty to thirty pattern
+    vertices. For the non-induced version of the problem, these instances lie on a satisfiable /
+    unsatisfiable phase transition, whose location we can predict; for the induced variant, much
+    richer behaviour is observed, and constrainedness gives a better measure of difficulty than does
+    proximity to a phase transition. In both cases, we can use this to understand and improve
+    variable and value ordering heuristics.
 
@@
 algorithm behaviour can be beneficial, because it provides a way of generating many instances
-cheaply, and reduces the risk of over-fitting when tuning design parameters. The random instances
-used typically come from common datasets
+cheaply, and reduces the risk of over-fitting when tuning design parameters. It also allows us to
+control graph parameters such as order, density, or degree distribution, and to study scaling
+properties of algorithms with respect to these parameters.
+
+The random instances used in previous evaluations typically come from common datasets
 \citep{DBLP:journals/prl/SantoFSV03,DBLP:journals/constraints/ZampelliDS10}, which were generated by
-taking a random subgraph of a random (Erd\H{o}s-R\'enyi, scale-free, bounded degree, or mesh)
-graph and permuting the vertices. Such instances are guaranteed to be
-satisfiable---\citet{o:AntonO09} exploited this property to create large sets of random satisfiable
-boolean satisfiability instances.  This is the most common approach to generating random subgraph
-isomorphism instances, meaning existing benchmark suites contain relatively few non-trivial
-unsatisfiable instances (although a few of the patterns in the instances by
-\citeauthor{DBLP:journals/constraints/ZampelliDS10} have had extra edges added, making them
-unsatisfiable). Also, the satisfiable instances tend to be computationally fairly easy, with most of
-the difficulty being in dealing with the size of the model. This has lead to bias in algorithm
-design, to the extent that some proposed techniques, such as those of
-\citet{DBLP:conf/sls/BattitiM07}, will \emph{only} work on satisfiable instances.
+taking a random subgraph of a random (Erd\H{o}s-R\'enyi, scale-free, bounded degree, or mesh) graph
+and permuting the vertices. Such instances are guaranteed to be satisfiable---\citet{o:AntonO09}
+exploited this property to create large sets of random satisfiable boolean satisfiability instances.
+This is the most common approach to generating random subgraph isomorphism instances, meaning
+existing benchmark suites contain relatively few non-trivial unsatisfiable instances (although a few
+of the patterns in the instances by \citeauthor{DBLP:journals/constraints/ZampelliDS10} have had
+extra edges added, making them unsatisfiable). Also, the satisfiable instances tend to be
+computationally fairly easy, with most of the difficulty being in dealing with the size of the
+model. This has lead to bias in algorithm design, to the extent that some proposed techniques, such
+as those of \citet{DBLP:conf/sls/BattitiM07}, will \emph{only} work on satisfiable instances.
 
@@
 instances, and can produce computationally challenging instances with only a few tens of vertices in
-the pattern, and 150 vertices in the target. This is not entirely straightforward---the lack of
+the pattern, and 150 vertices in the target. Note that the lack of
 unsatisfiable instances for testing purposes cannot be addressed simply by taking a pattern graph
@@
 problems first described by \citet{DBLP:conf/ijcai/CheesemanKT91} and
-\citet{DBLP:conf/aaai/MitchellSL92}. For subgraph isomorphism we identify three relevant control
-parameters: we can independently alter the edge probability of the pattern graph, the edge
-probability of the target graph, and the relative orders (number of vertices) of the pattern and
-target graphs.  For non-induced isomorphisms, with the correct choice of parameters we see results
-very similar to those observed with boolean satisfiability problems: there is a phase transition
-(whose location we can predict) from satisfiable to unsatisfiable, and we see a solver-independent
-complexity peak near this phase transition. Additionally, understanding this behaviour
-helps us to select better variable- and value-ordering heuristics---this is the second contribution
-of this paper.
+\citet{DBLP:conf/aaai/MitchellSL92}. For subgraph isomorphism in Erd\H{o}s-R\'enyi random graphs, we
+identify three relevant control parameters: we can independently alter the edge probability of the
+pattern graph, the edge probability of the target graph, and the relative orders (number of
+vertices) of the pattern and target graphs.  For non-induced isomorphisms, with the correct choice
+of parameters we see results very similar to those observed with boolean satisfiability problems:
+there is a phase transition (whose location we can predict) from satisfiable to unsatisfiable, and
+we see a solver-independent complexity peak near this phase transition. Additionally, understanding
+this behaviour helps us to select better variable- and value-ordering heuristics---this is the
+second contribution of this paper.
 
@@
 database systems, richer behaviour emerges, particularly when moving away from a uniform labelling
-scheme, and we see that VF2 behaves much worse than the Glasgow and LAD algorithms in certain cases.
-A close look shows that these cases \emph{should} be easy to solve.
+scheme, and we see that the popular VF2 algorithm behaves much worse than two more recent
+constraint-based algorithms in certain cases.  A close look shows that these cases \emph{should} be
+easy to solve.
 
@@
 We look at some of the datasets commonly used to test graph database systems, which the literature
-suggests are extremely hard to solve. Simple experiments show that the entire perceived difficulty
+suggests are hard to solve. Simple experiments show that the entire perceived difficulty
 of each of these datasets comes down to the widespread use of especially poor subgraph isomorphism
@@
 
-\section{Experimental Setup}
+\subsection{Definitions}
+
+Throughout, our graphs are unlabelled, undirected, and do not have any loops.  The \emph{order} of a
+graph is the cardinality of its vertex set. We write $\operatorname{V}(G)$ for the vertex set of a
+graph $G$. The \emph{complement} of a graph $G$, denoted $\shortoverline{G}$, is the graph with the
+same vertex set as $G$, and with an edge between distinct vertices $v$ and $w$ if and only if $v$
+and $w$ are not adjacent in $G$. We write $G(n, p)$ for an Erd\H{o}s-R\'enyi random graph with $n$
+vertices, and an edge between each distinct pair of vertices with independent probability $p$.
+
+A \emph{non-induced subgraph isomorphism} from a graph $P$ (called the \emph{pattern}) to a graph
+$T$ (called the \emph{target}) is an injective mapping from $\operatorname{V}(P)$ to
+$\operatorname{V}(T)$ which preserves adjacency---that is, for every adjacent $v$ and $w$ in
+$\operatorname{V}(P)$, the vertices $i(v)$ and $i(w)$ are adjacent in $T$. An \emph{induced subgraph
+isomorphism} additionally preserves non-adjacency---that is, if $v$ and $w$ are not adjacent in $P$,
+then $i(v)$ and $i(w)$ must not be adjacent in $T$. We use the notation $i : P \rightarrowtail T$
+for a non-induced isomorphism, and $i : P \hookrightarrow T$ for an induced isomorphism. Observe
+that an induced isomorphism $i : P \hookrightarrow T$ is a non-induced isomorphism $i : P
+\rightarrowtail T$ which is also a non-induced isomorphism $\shortoverline{i} : \shortoverline{P}
+\rightarrowtail \shortoverline{T}$.
+
+\subsection{Experimental Setup}
 
@@
 
-Recall that the Glasgow and LAD solvers use backtracking search to build up an assignment of target
-vertices (values) to pattern vertices (variables), but differ in terms of inference and ordering
-heuristics.  The approach used by VF2 is similar, although the domains of variables are not stored
-(in the style of conventional backtracking, rather than forward-checking), and so domain wipeouts
-are not detected until an assignment is made. Both the Glasgow and LAD solvers use domain sizes to
-guide search, whilst VF2 uses adjacency to previously-assigned vertices.
+The Glasgow and LAD solvers store domains, and use backtracking search to build up an assignment of
+target vertices (values) to pattern vertices (variables), but differ in terms of inference and
+ordering heuristics.  The approach used by VF2 is similar, although the domains of variables are not
+stored (in the style of conventional backtracking, rather than forward-checking), and so domain
+wipeouts are not detected until an assignment is made. Both the Glasgow and LAD solvers use domain
+sizes to guide search, whilst VF2 uses adjacency to previously-assigned vertices.
 
@@
 less detailed.
+
+?? Glasgow not parallel
+
+?? VF3 does not do non-induced
 
@@
 
+?? LAD and Glasgow not really designed for induced
+
 \begin{figure}[p]
@@
 isomorphisms. With a pattern of order 10, we get two independent phase transitions: the bottom right
-half of the plots resemble the non-induced results, and the top left half is close to a mirror
+halves of the plots resemble the non-induced results, and the top left halves are close to a mirror
 image. The central satisfiable region, which is away from either phase transition, is
@@
 Instances far from the phase transition are easy, except for VF2, which occasionally finds
-some of instances with larger numbers of labels very difficult. This behaviour occurs both on
+some of the instances with larger numbers of labels very difficult. This behaviour occurs both on
 satisfiable instances, and on the flawed unsatisfiable instances deep inside the ``satisfiable''
@@
 finds many of them extremely difficult.  The ``VF2${\leftrightarrow}$'' line shows what happens with
-VF2 if the graphs are permuted, so that the non-zero labelled instances are given lower vertex
+VF2 if the graphs are permuted, so that the non-zero labelled vertices are given lower vertex
 numbers rather than higher vertex numbers. This
@@
 algorithm chosen for the verification step \citep{DBLP:journals/jacm/Ullmann76}: as the algorithm
-predates techniques like arc-consistent all-different propagation \citep{DBLP:conf/aaai/Regin94}, it
+predates techniques like generalised arc-consistent all-different propagation \citep{DBLP:conf/aaai/Regin94}, it
 cannot immediately detect unsatisfiability in simple cases like there being two red vertices in the

### 2017-07-23 paper/paper.tex
Little changes

@@
 \jairheading{1}{2016}{0-0}{0/0}{0/0}
-\ShortHeadings{Why Real-World Subgraph Matching is Easy}
+\ShortHeadings{Why Real-World Subgraph Isomorphism is Easy}
 {McCreesh, Prosser \& Trimble}
@@
 
-\title{Why Real-World Subgraph Matching is Easy, and \\ What This Means for Filter / Verify in
+\title{Why Real-World Subgraph Isomorphism is Easy, and \\ What This Means for Filter / Verify in
 Graph Databases \thanks{Part of this paper appeared at IJCAI
@@
 going from satisfiable to unsatisfiable (from either direction), instances go from being trivial to
-really hard to solve. However, each of the three solvers we tested also finds the central
+really hard to solve. However, each of the three solvers we test also finds the central
 unsatisfiable region to be hard, despite it not being near a phase transition. To show that this is
 not a simple weakness of current subgraph isomorphism algorithms, we verify that this region is also
-hard when using a pseudo-boolean encoding, and under reduction to the clique problem. Interestingly,
-the constrainedness measure proposed by \citet{DBLP:conf/aaai/GentMPW96} \emph{does} predict this difficult
-region---the third contribution of this paper is to use these instances to provide evidence in
-favour of constrainedness, rather than proximity to a phase transition, being an accurate predictor
-of difficulty, and to show that constrainedness is not simply a refinement of a phase transition
-prediction.
+hard for boolean satisfiability, pseudo-boolean, and mixed integer solvers, and under reduction to
+the clique problem. Interestingly, the constrainedness measure proposed by
+\citet{DBLP:conf/aaai/GentMPW96} \emph{does} predict this difficult region---the third contribution
+of this paper is to use these instances to provide evidence in favour of constrainedness, rather
+than proximity to a phase transition, being an accurate predictor of difficulty, and to show that
+constrainedness is not simply a refinement of a phase transition prediction.
 
@@
 
+An even more extreme example of VF2's misbehaviour can be seen in \cref{figure:vf2-stupidity}. Here
+we have a pattern graph and a target graph, both of which are cliques plus one isolated vertex, and
+the isolated vertices have different unique labels. This is trivially unsatisfiable (and Glasgow and
+LAD detect this without search), but unless the labelled vertex is given the lowest vertex number
+(so it is branched on first), VF2 takes nearly a million recursive calls to detect this: VF2 will
+try to map every adjacent vertex in the clique before considering the difficult isolated vertex.
+This example can also be extended slightly to fool any simple static heuristic, or any simple label
+counting mechanism.
+
+These experiments further highlight that VF2 occasionally finds some instances which should be easy hard.
+But does this cause problems in practice? The literature suggests that it does. For example,
+\citet{o:Gromping14} uses VF2 in a package for the R statistics language, and states
+\begin{displayquote}``There are (not so many) instances for which creation of a clear design is
+    prohibitively slow in the current implementation that evaluates subgraph isomorphism with the
+    VF2 algorithm \ldots Recent experiences with a few of these showed that the LAD algorithm was
+    very fast in ruling out impossible matches, where VF2 took a long time.''\end{displayquote}
+Similarly, \citet{o:Murray12} uses VF2 inside a compiler, and observes extremely variable
+(and prohibitively high) compile times in some cases due to the expense of subgraph isomorphism
+calls---given the heavily labelled nature of these graphs, we conjecture that any domain-based
+algorithm would eliminate this cost.
+
 \begin{figure}[tb]
@@
 \end{figure}
-
-An even more extreme example of VF2's misbehaviour can be seen in \cref{figure:vf2-stupidity}. Here
-we have a pattern graph and a target graph, both of which are cliques plus one isolated vertex, and
-the isolated vertices have different unique labels. This is trivially unsatisfiable (and Glasgow and
-LAD detect this without search), but unless the labelled vertex is given the lowest vertex number
-(so it is branched on first), VF2 takes nearly a million recursive calls to detect this: VF2 will
-try to map every adjacent vertex in the clique before considering the difficult isolated vertex.
-This example can also be extended slightly to fool any simple static heuristic, or any simple label
-counting mechanism.
-
-These experiments further highlight that VF2 occasionally finds some instances which should be easy hard.
-But does this cause problems in practice? The literature suggests that it does. For example,
-\citet{o:Gromping14} uses VF2 in a package for the R statistics language, and states
-\begin{displayquote}``There are (not so many) instances for which creation of a clear design is
-    prohibitively slow in the current implementation that evaluates subgraph isomorphism with the
-    VF2 algorithm \ldots Recent experiences with a few of these showed that the LAD algorithm was
-    very fast in ruling out impossible matches, where VF2 took a long time.''\end{displayquote}
-Similarly, \citet{o:Murray12} uses VF2 inside a compiler, and observes extremely variable
-(and prohibitively high) compile times in some cases due to the expense of subgraph isomorphism
-calls---given the heavily labelled nature of these graphs, we conjecture that any domain-based
-algorithm would eliminate this cost.
 

### 2017-07-24 paper/paper.tex
Abstract tweak

@@
     in mind, we demonstrate that the widely researched ``filter~/ verify'' indexing technique used in
-    graph databases is a flawed design based upon a repeated misunderstanding of what makes subgraph
-    isomorphism hard. We explain why filter~/ verify cannot be beneficial, even with harder
-    instances than those currently used, when paired with any reasonable modern subgraph isomorphism
-    algorithm.
+    graph databases is a flawed design based upon a repeated misunderstanding of the empirical
+    hardness of \NP-complete problems. We explain why filter~/ verify cannot be beneficial, even
+    with harder instances than those currently used, when paired with any reasonable modern subgraph
+    isomorphism algorithm.
 \end{abstract}

### 2017-07-25 paper/paper.tex
Updated plottery

@@
 \usepackage{csquotes}
+
+\usepackage{showframe}
 
@@
             \\
+            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize VF3}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-vf3-induced-nodes-10-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-vf3-induced-nodes-14-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-vf3-induced-nodes-15-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-vf3-induced-nodes-16-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-vf3-induced-nodes-20-150.pdf}}; &
+            \node [anchor=center] {\includegraphics*{plots/graph-vf3-induced-nodes-30-150.pdf}}; \\
+            \\
             \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Constrainedness}; &
@@
     \caption{Behaviour of algorithms on the induced variant, shown in the style
-    of \cref{figure:non-induced}. The fifth row plots constrainedness using
+    of \cref{figure:non-induced}. The sixth row plots constrainedness using
     equation~\eqref{equation:constrainedness}: the
@@
     given the complements of the pattern and target graphs as inputs---the solid lines show the
-    location of the phase transition, and the dotted lines are $t_d=0.5$ and the $p_d=t_d$ diagonal.}\label{figure:induced}
+    location of the phase transition, and the dotted lines are $t_d=0.5$ and the $p_d=t_d$
+    diagonal.}\label{figure:induced}
 \end{figure}

### 2017-07-25 paper/paper.tex
Title

@@
 \jairheading{1}{2016}{0-0}{0/0}{0/0}
-\ShortHeadings{Why Real-World Subgraph Isomorphism is Easy}
+\ShortHeadings{When Subgraph Isomorphism is Really Hard}
 {McCreesh, Prosser \& Trimble}
@@
 
-\title{Why Real-World Subgraph Isomorphism is Easy, and \\ What This Means for Filter / Verify in
-Graph Databases \thanks{Part of this paper appeared at IJCAI
-\citep{DBLP:conf/ijcai/McCreeshPT16}; \cref{section:introduction} has been expanded and
+\title{When Subgraph Isomorphism is Really Hard, and Why This Matters for Graph
+Databases\thanks{Part of this paper appeared at IJCAI \citep{DBLP:conf/ijcai/McCreeshPT16};
+\cref{section:introduction} has been expanded and
 \cref{section:models,section:labelled,section:filterverify} are new material.}}

### 2017-09-29 analysis/search_nodes_heatmap.cc
Analyse this

@@ -1,0 +1,226 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include <algorithm>
+#include <cstdlib>
+#include <iostream>
+#include <fstream>
+#include <string>
+
+#include <boost/format.hpp>
+
+using std::cerr;
+using std::cout;
+using std::endl;
+using std::getline;
+using std::ifstream;
+using std::max;
+using std::stoi;
+using std::stoull;
+using std::string;
+using std::to_string;
+
+using boost::format;
+using boost::str;
+
+auto add_results(
+        const string & dir, const string & alg,
+        int ps, int ts, int pd, int td, int r,
+        const std::string & pd_p, const std::string & td_p,
+        int & failures, unsigned long long & total_nodes) -> void
+{
+    string pd_s = (pd_p == "pkr" ? to_string(pd) : str(format("%|1|.%|02|") % (pd / 100) % (pd % 100)));
+    string td_s = (td_p == "tkr" ? to_string(td) : str(format("%|1|.%|02|") % (td / 100) % (td % 100)));
+    string r_s = str(format("%|02|") % r);
+
+    string filename{dir + "/ps" + to_string(ps) + "-ts" + to_string(ts) + "/" + pd_p + pd_s + "-" + td_p + td_s + "/" + r_s + "." + alg + ".out"};
+    ifstream infile{filename};
+    if (! infile) {
+        cerr << "missing " << filename << endl;
+        ++failures;
+        return;
+    }
+
+    string success, aborted;
+    unsigned long long this_nodes;
+
+    if (string::npos != alg.find("vf3")) {
+        infile >> success >> this_nodes;
+        if (! infile) {
+            cerr << "reading " << filename << endl;
+            ++failures;
+            return;
+        }
+        if (success == "aborted")
+            aborted = success;
+    }
+    else if (string::npos != alg.find("vf2")) {
+        infile >> success;
+        if (success == "aborted")
+            aborted = success;
+        else
+            infile >> this_nodes;
+
+        if (! infile) {
+            cerr << "reading " << filename << endl;
+            ++failures;
+            return;
+        }
+    }
+    else if (string::npos != alg.find("lad")) {
+        string line;
+        getline(infile, line);
+        if (string::npos == line.find("Run completed") && string::npos == line.find("CPU time exceeded"))
+            getline(infile, line);
+
+        if (! infile) {
+            cerr << "reading " << filename << endl;
+            ++failures;
[…]

### 2017-09-29 paper/paper.tex
Pictures etc

@@
 exponent in the second half of the formula, maximising the overall value. LAD does not apply a value
-ordering heuristic, but Glasgow does: it prefers target vertices of lowest degree.  Again, this was
-determined empirically, but it has the effect of increasing $\langle Sol \rangle$ by increasing the
-remaining target density. The VF2 heuristics, in contrast, are based around preserving connectivity,
-which gives very little discrimination except on the sparsest of inputs.
+ordering heuristic, but Glasgow does: it prefers target vertices of highest degree.  Again, this was
+determined empirically, but it has the effect of increasing $\langle Sol \rangle$ by leaving as many
+vertices as possible available for future use. The VF2 heuristics, in contrast, are based around
+preserving connectivity, which gives very little discrimination except on the sparsest of inputs.
 

### 2017-10-01 paper/paper.tex
kr

@@
 
-    \caption{Behaviour of algorithms on random $k$-regular graphs, as the pattern degree (x-axis)
-    and target degree (y-axis) are varied from 0 to $\left|\operatorname{V}\right| - 1$.}
+    \caption{Behaviour of algorithms on the non-induced problem on random $k$-regular graphs, as the
+    pattern degree (x-axis) and target degree (y-axis) are varied from 0 to
+    $\left|\operatorname{V}\right| - 1$.}
     \label{figure:kr}

### 2017-10-01 analysis/search_nodes_heatmap.cc
Also labels

@@ -25,14 +25,14 @@ using boost::str;
 auto add_results(
         const string & dir, const string & alg,
         int ps, int ts, int pd, int td, int r,
-        const std::string & pd_p, const std::string & td_p,
+        const string & pd_p, const string & td_p, const string & ed,
         int & failures, unsigned long long & total_nodes) -> void
 {
     string pd_s = (pd_p == "pkr" ? to_string(pd) : str(format("%|1|.%|02|") % (pd / 100) % (pd % 100)));
     string td_s = (td_p == "tkr" ? to_string(td) : str(format("%|1|.%|02|") % (td / 100) % (td % 100)));
     string r_s = str(format("%|02|") % r);
 
-    string filename{dir + "/ps" + to_string(ps) + "-ts" + to_string(ts) + "/" + pd_p + pd_s + "-" + td_p + td_s + "/" + r_s + "." + alg + ".out"};
+    string filename{dir + "/ps" + to_string(ps) + "-ts" + to_string(ts) + "/" + pd_p + pd_s + "-" + td_p + td_s + "/" + ed + r_s + "." + alg + ".out"};
     ifstream infile{filename};
     if (! infile) {
         cerr << "missing " << filename << endl;
@@ -185,21 +185,22 @@ auto add_results(
 
 auto main(int argc, char * argv[]) -> int
 {
-    if (argc != 11) {
-        cerr << "Usage: " << argv[0] << " " << "directory algorithm pattern_size target_size pattern_prefix target_prefix density_increment repeats pattern_stop target_stop" << endl;
+    if (argc != 12) {
+        cerr << "Usage: " << argv[0] << " " << "directory extra-directory algorithm pattern_size target_size pattern_prefix target_prefix density_increment repeats pattern_stop target_stop" << endl;
         return EXIT_FAILURE;
     }
 
     string dir = argv[1];
-    string alg = argv[2];
-    int ps = stoi(argv[3]);
-    int ts = stoi(argv[4]);
-    string pd_p = argv[5];
-    string td_p = argv[6];
-    int di = stoi(argv[7]);
-    int nr = stoi(argv[8]);
-    int px = stoi(argv[9]);
-    int tx = stoi(argv[10]);
+    string ed = argv[2];
+    string alg = argv[3];
+    int ps = stoi(argv[4]);
+    int ts = stoi(argv[5]);
+    string pd_p = argv[6];
+    string td_p = argv[7];
+    int di = stoi(argv[8]);
[…]

### 2017-10-15 paper/paper.tex
Plotting
Caption

@@
         phase transition will occur. On the subsequent three rows, we show the number of search nodes used by the
-        Glasgow, LAD and VF2 solvers; the dark regions indicate ``really hard'' instances.}
+        Glasgow, LAD and VF2 solvers; the dark regions indicate ``really hard'' instances, and black
+        points indicate that at least one timeout occurred.}
     \label{figure:non-induced}
@@
 
-\clearpage
-\appendix
-\section{Reproducing These Experiments}
-
 \end{document}

### 2017-10-27 paper/paper.tex
Missing author
More things

@@
 
-\usepackage{showframe}
+% \usepackage{showframe}
 
@@
 
-\jairheading{1}{2016}{0-0}{0/0}{0/0}
+\jairheading{1}{2017}{0-0}{0/0}{0/0}
 \ShortHeadings{When Subgraph Isomorphism is Really Hard}
-{McCreesh, Prosser \& Trimble}
+{McCreesh, Prosser, Solnon \& Trimble}
 \firstpageno{1}
@@
 
-\crefname{algorithm}{Algorithm}{Algorithms}
-\Crefname{algorithm}{Algorithm}{Algorithms}
 \crefname{figure}{Figure}{Figures}
 \Crefname{figure}{Figure}{Figures}
-\crefname{table}{Table}{Tables}
-\Crefname{table}{Table}{Tables}
 \crefname{section}{Section}{Sections}
 \Crefname{section}{Section}{Sections}
-\crefname{proposition}{Proposition}{Propositions}
-\Crefname{proposition}{Proposition}{Propositions}
 
@@
 
-\definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}
-\definecolor{uofglawn}{rgb}{0.517647, 0.741176, 0}
-\definecolor{uofgcobalt}{rgb}{0, 0.615686, 0.92549}
-\definecolor{uofgpumpkin}{rgb}{1.0, 0.72549, 0.282353}
-\definecolor{uofgthistle}{rgb}{0.584314, 0.070588, 0.447059}
-
 \begin{document}
@@
 \title{When Subgraph Isomorphism is Really Hard, and Why This Matters for Graph
-Databases\thanks{Part of this paper appeared at IJCAI \citep{DBLP:conf/ijcai/McCreeshPT16};
-\cref{section:introduction} has been expanded and
-\cref{section:models,section:labelled,section:filterverify} are new material.}}
+Databases}
 
@@
     proximity to a phase transition. In both cases, we can use this to understand and improve
-    variable and value ordering heuristics.
+    variable- and value-ordering heuristics.
 
@@
 \citep{DBLP:journals/ijprai/ConteFSV04}.  In particular, a labelled version of subgraph isomorphism
-is one of the key components in supporting complex queries in graph databases---typically, these
-systems store a fixed set of target graphs, and for a sequence of pattern queries, they must return
-every target graph which contains that pattern.  One common approach to this problem is to combine a
-subgraph isomorphism algorithm with an indexing system in a so-called ``filter~/ verify'' model,
-where invariants are used to attempt to pre-exclude unsatisfiable instances to avoid the cost of a
-subgraph isomorphism call. In this context, the terms \emph{matching} and \emph{verification} are
-often used for the subgraph isomorphism step (or a slightly broader problem, for example permitting
-wildcards).
-
-\begin{figure}[t]
+is one of the key components in supporting complex queries in graph databases, which we discuss in
+detail in \cref{section:filterverify}. Typically, these systems store a fixed set of target graphs,
+and for a sequence of pattern queries, they must return every target graph which contains that
+pattern.  One common approach to this problem is to combine a subgraph isomorphism algorithm with an
+indexing system in a so-called ``filter~/ verify'' model, where invariants are used to attempt to
+pre-exclude unsatisfiable instances to avoid the cost of a subgraph isomorphism call. In this
+context, the terms \emph{matching} and \emph{verification} are often used for the subgraph
+isomorphism step (or a slightly broader problem, for example permitting wildcards).
+
+\begin{figure}[tb]
     \centering
@@
 The first contribution of this paper is to present and evaluate new methods for creating random
-pattern~/ target pairs.  The method we introduce generates both satisfiable and unsatisfiable
-instances, and can produce computationally challenging instances with only a few tens of vertices in
-the pattern, and 150 vertices in the target. Note that the lack of
+pattern~/ target pairs.  The method we introduce in \cref{section:non-induced} generates both
+satisfiable and unsatisfiable instances, and can produce computationally challenging instances with
+only a few tens of vertices in the pattern, and 150 vertices in the target. Note that the lack of
 unsatisfiable instances for testing purposes cannot be addressed simply by taking a pattern graph
@@
 
-For certain choices of parameters for induced isomorphisms, there are two phase transitions, going
-from satisfiable to unsatisfiable, and then from unsatisfiable back to satisfiable. Again, when
-going from satisfiable to unsatisfiable (from either direction), instances go from being trivial to
-really hard to solve. However, each of the three solvers we test also finds the central
-unsatisfiable region to be hard, despite it not being near a phase transition. To show that this is
-not a simple weakness of current subgraph isomorphism algorithms, we verify that this region is also
-hard for boolean satisfiability, pseudo-boolean, and mixed integer solvers, and under reduction to
-the clique problem. Interestingly, the constrainedness measure proposed by
-\citet{DBLP:conf/aaai/GentMPW96} \emph{does} predict this difficult region---the third contribution
-of this paper is to use these instances to provide evidence in favour of constrainedness, rather
-than proximity to a phase transition, being an accurate predictor of difficulty, and to show that
-constrainedness is not simply a refinement of a phase transition prediction.
-
-% What about other random graph models? Graphs where every vertex has the same degree, which are known
-% as $k$-regular, exhibit very similar characteristics.
-
-When labels on vertices are introduced, as is commonly seen in real-world applications and in graph
-database systems, richer behaviour emerges, particularly when moving away from a uniform labelling
-scheme, and we see that the popular VF2 algorithm behaves much worse than two more recent
-constraint-based algorithms in certain cases.  A close look shows that these cases \emph{should} be
-easy to solve.
+In \cref{section:induced} we look at the induced isomorphism problem.  For certain choices of
+parameters, there are two phase transitions, going from satisfiable to unsatisfiable, and then from
+unsatisfiable back to satisfiable. Again, when going from satisfiable to unsatisfiable (from either
+direction), instances go from being trivial to really hard to solve. However, each of the three
+solvers we test also finds the central unsatisfiable region to be hard, despite it not being near a
+phase transition. To show that this is not a simple weakness of current subgraph isomorphism
+algorithms, we verify that this region is also hard for boolean satisfiability, pseudo-boolean, and
+mixed integer solvers, and under reduction to the clique problem. Interestingly, the constrainedness
+measure proposed by \citet{DBLP:conf/aaai/GentMPW96} \emph{does} predict this difficult region---the
+third contribution of this paper is to use these instances to provide evidence in favour of
+constrainedness, rather than proximity to a phase transition, being an accurate predictor of
+difficulty, and to show that constrainedness is not simply a refinement of a phase transition
+prediction.
+
+What about other random graph models? In \cref{section:regular} we see that graphs where every vertex
+has the same degree (known as $k$-regular) exhibit very similar characteristics.  However, when
+labels on vertices are introduced (as is commonly seen in real-world applications and in graph
+database systems), richer behaviour emerges, particularly when moving away from a uniform labelling
+scheme, and the popular VF2 algorithm behaves much worse than two more recent
+constraint-based algorithms in certain cases. In \cref{section:labelled} we take a close look at
+these cases, and argue that they \emph{should} be easy to solve.
 
@@
 
-We look at some of the datasets commonly used to test graph database systems, which the literature
-suggests are hard to solve. Simple experiments show that the entire perceived difficulty
-of each of these datasets comes down to the widespread use of especially poor subgraph isomorphism
-algorithms. Our fourth contribution is to show that the use of forward-checking and the right
-variable ordering heuristic, even without the more sophisticated recent advances in subgraph
-algorithms, makes the entire graph database filter~/ verify paradigm unnecessary.  Finally, we
-explain why filter~/ verify (and other recently proposed techniques) cannot be beneficial even on
+In \cref{section:filterverify} we look at some of the datasets commonly used to test graph database
+systems, which the literature suggests are hard to solve. Simple experiments show that the entire
+perceived difficulty of each of these datasets comes down to the widespread use of especially poor
+subgraph isomorphism algorithms. Our fourth contribution is to show that the use of forward-checking
+and the right variable ordering heuristic, even without the more sophisticated recent advances in
+subgraph algorithms, makes the entire graph database filter~/ verify paradigm unnecessary.  Finally,
+we explain why filter~/ verify (and other recently proposed techniques) cannot be beneficial even on
 more challenging instances except if an extremely poor choice of subgraph isomorphism algorithm is
@@
 
+\bigskip
+
+Part of this paper appeared at IJCAI \citep{DBLP:conf/ijcai/McCreeshPT16};
+\cref{section:introduction} has been expanded and
+\cref{section:regular,section:labelled,section:filterverify} are new material.
+
 \subsection{Definitions}
 
-Throughout, our graphs are unlabelled, undirected, and do not have any loops.  The \emph{order} of a
-graph is the cardinality of its vertex set. We write $\operatorname{V}(G)$ for the vertex set of a
-graph $G$. The \emph{complement} of a graph $G$, denoted $\shortoverline{G}$, is the graph with the
-same vertex set as $G$, and with an edge between distinct vertices $v$ and $w$ if and only if $v$
-and $w$ are not adjacent in $G$. We write $G(n, p)$ for an Erd\H{o}s-R\'enyi random graph with $n$
-vertices, and an edge between each distinct pair of vertices with independent probability $p$.
+Throughout, our graphs are undirected, and do not have any loops. In some cases vertices have
+labels, which are treated as integers; we write $\ell(v)$ for the label of vertex $v$. The
+\emph{order} of a graph is the cardinality of its vertex set. We write $\operatorname{V}(G)$ for the
+vertex set of a graph $G$. The \emph{complement} of a graph $G$, denoted $\shortoverline{G}$, is the
+graph with the same vertex set as $G$, and with an edge between distinct vertices $v$ and $w$ if and
+only if $v$ and $w$ are not adjacent in $G$. We write $G(n, p)$ for an Erd\H{o}s-R\'enyi random
+graph with $n$ vertices, and an edge between each distinct pair of vertices with independent
+probability $p$.
 
@@
 $\operatorname{V}(T)$ which preserves adjacency---that is, for every adjacent $v$ and $w$ in
-$\operatorname{V}(P)$, the vertices $i(v)$ and $i(w)$ are adjacent in $T$. An \emph{induced subgraph
-isomorphism} additionally preserves non-adjacency---that is, if $v$ and $w$ are not adjacent in $P$,
-then $i(v)$ and $i(w)$ must not be adjacent in $T$. We use the notation $i : P \rightarrowtail T$
-for a non-induced isomorphism, and $i : P \hookrightarrow T$ for an induced isomorphism. Observe
-that an induced isomorphism $i : P \hookrightarrow T$ is a non-induced isomorphism $i : P
-\rightarrowtail T$ which is also a non-induced isomorphism $\shortoverline{i} : \shortoverline{P}
-\rightarrowtail \shortoverline{T}$.
+$\operatorname{V}(P)$, the vertices $i(v)$ and $i(w)$ are adjacent in $T$. Where labels are present,
+the mapping must also preserve labels, so $\ell(v) = \ell(i(v))$ for any $v$. An \emph{induced
+subgraph isomorphism} additionally preserves non-adjacency---that is, if $v$ and $w$ are not
+adjacent in $P$, then $i(v)$ and $i(w)$ must not be adjacent in $T$. We use the notation $i : P
+\rightarrowtail T$ for a non-induced isomorphism, and $i : P \hookrightarrow T$ for an induced
+isomorphism. Observe that an induced isomorphism $i : P \hookrightarrow T$ is a non-induced
+isomorphism $i : P \rightarrowtail T$ which is also a non-induced isomorphism $\shortoverline{i} :
+\shortoverline{P} \rightarrowtail \shortoverline{T}$.
 
@@
 The experiments in this paper are performed on systems with Intel Xeon E5-4650 v2 CPUs, running
-Scientific Linux release 6.7. We will be working with three subgraph isomorphism solvers: the
-Glasgow solver \citet{DBLP:conf/cp/McCreeshP15}; LAD \citep{DBLP:journals/ai/Solnon10}; and VF2
-\citep{DBLP:journals/pami/CordellaFSV04}. Each was compiled using GCC 4.9.
+Scientific Linux release 6.7. We will be working with four subgraph isomorphism solvers: the Glasgow
+solver \citet{DBLP:conf/cp/McCreeshP15}; LAD \citep{DBLP:journals/ai/Solnon10}; VF2
+\citep{DBLP:journals/pami/CordellaFSV04}; and VF3 \citep{DBLP:conf/gbrpr/CarlettiFSV17}. Each was
+compiled using GCC 4.9.
 
@@
 target vertices (values) to pattern vertices (variables), but differ in terms of inference and
-ordering heuristics.  The approach used by VF2 is similar, although the domains of variables are not
-stored (in the style of conventional backtracking, rather than forward-checking), and so domain
-wipeouts are not detected until an assignment is made. Both the Glasgow and LAD solvers use domain
-sizes to guide search, whilst VF2 uses adjacency to previously-assigned vertices.
+ordering heuristics.  The approach used by VF2 and VF3 is similar, although the domains of variables
+are not stored (in the style of conventional backtracking, rather than forward-checking), and so
+domain wipeouts are not detected until an assignment is made. Both the Glasgow and LAD solvers use
+domain sizes to guide search, whilst VF2 and VF3 use adjacency to previously-assigned vertices.
 
@@
 timeout in mind), although we may slightly overestimate the proportion of unsatisfiable instances
-for extremely sparse or dense pattern graphs. The LAD and VF2 solvers experienced many more failures
+for extremely sparse or dense pattern graphs. The LAD, VF2, and VF3 solvers experienced many more failures
 with this timeout, so our picture of just how hard the hardest instances are with these solvers is
@@
 
-?? Glasgow not parallel
-
-?? VF3 does not do non-induced
-
-\section{Non-Induced Subgraph Isomorphisms}
+The Glasgow, LAD, and VF2 solvers support both non-induced and induced isomorphisms (although
+neither Glasgow nor LAD were designed with the induced version in mind), whilst VF3 supports only
+the induced problem. Note also that VF3 will sometimes crash on especially sparse graphs: we have
+treated these instances as timeouts.
+
+\section{Non-Induced Subgraph Isomorphisms}\label{section:non-induced}
 
@@
             \node [anchor=center] {}; &
-            \node [anchor=center] {\scriptsize $G(10, x)$\hspace*{1em}}; &
-            \node [anchor=center] {\scriptsize $G(20, x)$\hspace*{1em}}; &
-            \node [anchor=center] {\scriptsize $G(30, x)$\hspace*{3em}}; &
+            \node [anchor=center] {\scriptsize $G(10, x) \rightarrowtail G(150, y)$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(20, x) \rightarrowtail G(150, y)$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(30, x) \rightarrowtail G(150, y)$\hspace*{3em}}; &
             \\
-            \node [anchor=center, inner sep=4pt] [rotate=90] {\scriptsize Satisfiable?}; &
+            \node [anchor=center, inner sep=2pt] [rotate=90] {\scriptsize Satisfiable?}; &
             \node [anchor=center] {\includegraphics*{plots/graph-non-induced-satisfiable-10-150.pdf}}; &
@@
             \\[0.3cm]
-            \node [anchor=center, inner sep=4pt] [rotate=90] {\scriptsize Glasgow}; &
+            \node [anchor=center, inner sep=2pt] [rotate=90] {\scriptsize Glasgow}; &
             \node [anchor=center] {\includegraphics*{plots/graph-non-induced-nodes-10-150.pdf}}; &
@@
             \\[0.3cm]
-            \node [anchor=center, inner sep=4pt] [rotate=90] {\scriptsize LAD}; &
+            \node [anchor=center, inner sep=2pt] [rotate=90] {\scriptsize LAD}; &
             \node [anchor=center] {\includegraphics*{plots/graph-lad-non-induced-nodes-10-150.pdf}}; &
@@
             \\[0.3cm]
-            \node [anchor=center, inner sep=4pt] [rotate=90] {\scriptsize VF2}; &
+            \node [anchor=center, inner sep=2pt] [rotate=90] {\scriptsize VF2}; &
             \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-nodes-10-150.pdf}}; &
@@
 
-    \caption{Behaviour of algorithms on the non-induced variant. For each plot, the x-axis is the
+    \caption{Behaviour of algorithms on the non-induced variant, using target graphs of 150
+        vertices, and pattern graphs of 10, 20, or 30 vertices. For each plot, the x-axis is the
         pattern edge probability and the y-axis is the target edge probability, both from 0 to 1.
@@
 
-On subsequent rows, we show the average number of search nodes used by the different algorithms. In
-general, satisfiable instances are easy, until very close to the phase transition. As we hit the
-phase transition and move into the unsatisfiable region, we see complexity increase. Finally, as we
-pass through the phase transition and move deeper into the unsatisfiable region, instances become
-easier again. This behaviour is largely solver-independent, although VF2 has a larger hard region
-than Glasgow or LAD. VF2 also finds some instances extremely hard when the target is empty but the
-pattern is not---this turns out to be extremely important, and we return to it in
-\cref{section:labelled,section:filterverify}. Thus, although we have moved away from a single
-control parameter, we still observe the easy-hard-easy pattern seen in many \NP-complete
-problems.
+On subsequent rows, we show the average number of search nodes used by the different algorithms.
+Darker regions indicate harder difficulties, and black is used to indicate that at least one timeout
+occurred. In general, satisfiable instances are easy, until very close to the phase transition. As
+we hit the phase transition and move into the unsatisfiable region, we see complexity increase.
+Finally, as we pass through the phase transition and move deeper into the unsatisfiable region,
+instances become easier again. This behaviour is largely solver-independent, although VF2 has a
+larger hard region than Glasgow or LAD. Thus, although we have moved away from a single control
+parameter, we still observe the easy-hard-easy pattern seen in many \NP-complete problems.
+
+Finally, a close look at the very bottom left of the plots shows that VF2 will sometimes give a
+timeout on instances where the target is empty but the pattern is not---this turns out to be
+extremely important, and we return to it in \cref{section:labelled,section:filterverify}.
 
@@
 heuristics for backtracking search algorithms---one of these is to try to maximise the expected
-number of solutions inside any subproblem considered during search \citep{DBLP:conf/cp/GentMPSW96}.  This is
-usually done by cheaper surrogates, rather than direct calculation. When branching, both LAD and
-Glasgow pick a variable with fewest remaining values in its domain: doing this will generally reduce
-the first part of the $\langle Sol \rangle$ equation by as little as possible. When two or more
-domains are of equal size, LAD simply breaks ties lexicographically, whereas Glasgow will pick a
-variable corresponding to a pattern vertex of highest degree. This strategy was determined
-empirically, but could have been derived from the $\langle Sol \rangle$ formula: picking a pattern
-vertex of high degree will make the remaining pattern subgraph sparser, which will decrease the
-exponent in the second half of the formula, maximising the overall value. LAD does not apply a value
-ordering heuristic, but Glasgow does: it prefers target vertices of highest degree.  Again, this was
-determined empirically, but it has the effect of increasing $\langle Sol \rangle$ by leaving as many
-vertices as possible available for future use. The VF2 heuristics, in contrast, are based around
-preserving connectivity, which gives very little discrimination except on the sparsest of inputs.
-
-\section{Induced Subgraph Isomorphisms}
-
-?? LAD and Glasgow not really designed for induced
+number of solutions inside any subproblem considered during search \citep{DBLP:conf/cp/GentMPSW96}.
+This is usually done by cheaper surrogates, rather than direct calculation. When branching, both LAD
+and Glasgow pick a variable with fewest remaining values in its domain: doing this will generally
+reduce the first part of the $\langle Sol \rangle$ equation by as little as possible. When two or
+more domains are of equal size, LAD simply breaks ties lexicographically, whereas Glasgow will pick
+a variable corresponding to a pattern vertex of highest degree (?? check carefully!). This strategy
+was determined empirically, but could have been derived from the $\langle Sol \rangle$ formula:
+picking a pattern vertex of high degree will make the remaining pattern subgraph sparser, which will
+decrease the exponent in the second half of the formula, maximising the overall value. LAD does not
+apply a value ordering heuristic, but Glasgow does: it prefers target vertices of highest degree (??
+check carefully!).  Again, this was determined empirically, but it has the effect of increasing
+$\langle Sol \rangle$ by leaving as many vertices as possible available for future use. The VF2
+heuristics, in contrast, are based around preserving connectivity, which gives very little
+discrimination except on the sparsest of inputs.
+
+\subsection{Instances for Benchmarking}
+
+?? Instances for benchmarking
+
+?? Something about distance-based filtering being no use on these instances
+
+?? Low degree spread
+
+\section{Induced Subgraph Isomorphisms}\label{section:induced}
 
@@
             \node [anchor=center] {}; &
-            \node [anchor=center] {\scriptsize $G(10, x)$\hspace*{1em}}; &
-            \node [anchor=center] {\scriptsize $G(14, x)$\hspace*{1em}}; &
-            \node [anchor=center] {\scriptsize $G(15, x)$\hspace*{1em}}; &
-            \node [anchor=center] {\scriptsize $G(16, x)$\hspace*{1em}}; &
-            \node [anchor=center] {\scriptsize $G(20, x)$\hspace*{1em}}; &
-            \node [anchor=center] {\scriptsize $G(30, x)$\hspace*{3em}}; &
+            \node [anchor=center] {\scriptsize $G(10, x) \hookrightarrow$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(14, x) \hookrightarrow$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(15, x) \hookrightarrow$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(16, x) \hookrightarrow$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(20, x) \hookrightarrow$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(30, x) \hookrightarrow$\hspace*{3em}}; &
             \\
-            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Satisfiable?}; &
+            \node [anchor=center] {}; &
+            \node [anchor=center] {\scriptsize $G(150, y)$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(150, y)$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(150, y)$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(150, y)$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(150, y)$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(150, y)$\hspace*{3em}}; &
+            \\[-0.08cm]
+            \node [anchor=center, inner sep=2pt, rotate=90] {\scriptsize Satisfiable?}; &
             \node [anchor=center] {\includegraphics*{plots/graph-induced-satisfiable-10-150.pdf}}; &
@@
             \node [anchor=center] {\includegraphics*{plots/graph-induced-satisfiable-30-150.pdf}}; &
-            \\
-            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Glasgow}; &
+            \\[-0.08cm]
+            \node [anchor=center, inner sep=2pt, rotate=90] {\scriptsize Glasgow}; &
             \node [anchor=center] {\includegraphics*{plots/graph-induced-nodes-10-150.pdf}}; &
@@
             \node [anchor=center] {\includegraphics*{plots/graph-induced-nodes-30-150.pdf}}; &
-            \\
-            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize LAD}; &
+            \\[-0.08cm]
+            \node [anchor=center, inner sep=2pt, rotate=90] {\scriptsize LAD}; &
             \node [anchor=center] {\includegraphics*{plots/graph-lad-induced-nodes-10-150.pdf}}; &
@@
             \node [anchor=center] {\includegraphics*{plots/graph-lad-induced-nodes-30-150.pdf}}; &
-            \\
-            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize VF2}; &
+            \\[-0.08cm]
+            \node [anchor=center, inner sep=2pt, rotate=90] {\scriptsize VF2}; &
             \node [anchor=center] {\includegraphics*{plots/graph-vf2-induced-nodes-10-150.pdf}}; &
@@
             \node [anchor=center] {\includegraphics*{plots/graph-vf2-induced-nodes-30-150.pdf}}; \\
-            \\
-            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize VF3}; &
+            \\[-0.08cm]
+            \node [anchor=center, inner sep=2pt, rotate=90] {\scriptsize VF3}; &
             \node [anchor=center] {\includegraphics*{plots/graph-vf3-induced-nodes-10-150.pdf}}; &
@@
             \node [anchor=center] {\includegraphics*{plots/graph-vf3-induced-nodes-30-150.pdf}}; \\
-            \\
-            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Constrainedness}; &
+            \\[-0.08cm]
+            \node [anchor=center, inner sep=2pt, rotate=90] {\scriptsize Constrainedness}; &
             \node [anchor=center] {\includegraphics*{plots/graph-induced-kappa-10-150.pdf}}; &
@@
             \node [anchor=center] {\includegraphics*{plots/graph-induced-kappa-30-150.pdf}}; \\
-            \\
-            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Complement?}; &
+            \\[-0.08cm]
+            \node [anchor=center, inner sep=2pt, rotate=90] {\scriptsize Complement?}; &
             \node [anchor=center] {\includegraphics*{plots/graph-induced-which-10-150.pdf}}; &
@@
     \end{tikzpicture}
-    \caption{Behaviour of algorithms on the induced variant, shown in the style
-    of \cref{figure:non-induced}. The sixth row plots constrainedness using
-    equation~\eqref{equation:constrainedness}: the
-    darkest region is where $\kappa = 1$, and the lighter regions show where the problem is either
-    over- or under-constrained. The final row shows when the Glasgow algorithm performs better when
-    given the complements of the pattern and target graphs as inputs---the solid lines show the
-    location of the phase transition, and the dotted lines are $t_d=0.5$ and the $p_d=t_d$
-    diagonal.}\label{figure:induced}
+    \caption{Behaviour of algorithms on the induced variant with target graphs of 150 vertices,
+    shown in the style of \cref{figure:non-induced}. The sixth row plots constrainedness using
+    equation~\eqref{equation:constrainedness}: the darkest region is where $\kappa = 1$, and the
+    lighter regions show where the problem is either over- or under-constrained. The final row shows
+    when the Glasgow algorithm performs better when given the complements of the pattern and target
+    graphs as inputs---the solid lines show the location of the phase transition, and the dotted
+    lines are $t_d=0.5$ and the $p_d=t_d$ diagonal.}\label{figure:induced}
 \end{figure}
 
-In the first four rows of \cref{figure:induced} we repeat our experiments, finding induced
+In the first four rows of \cref{figure:induced} we repeat our experiments, now finding induced
 isomorphisms. With a pattern of order 10, we get two independent phase transitions: the bottom right
@@
 it is interesting to note that for larger target graphs, VF2 finds \emph{all} dense pattern graphs
-difficult.
+difficult. Finally, VF3's performance generally improves upon VF2's on instances that VF2 finds easy
+(except in the cases where VF3 crashes), but worsens on instances that VF2 finds hard.
 
@@
             \node [anchor=center] {}; &
-            \node [anchor=center] {\scriptsize $G(10, x)$\hspace*{1em}}; &
-            \node [anchor=center] {\scriptsize $G(12, x)$\hspace*{1em}}; &
-            \node [anchor=center] {\scriptsize $G(14, x)$\hspace*{1em}}; &
-            \node [anchor=center] {\scriptsize $G(16, x)$\hspace*{1em}}; &
-            \node [anchor=center] {\scriptsize $G(18, x)$\hspace*{1em}}; &
-            \node [anchor=center] {\scriptsize $G(25, x)$\hspace*{3em}}; &
+            \node [anchor=center] {\scriptsize $G(10, x) \hookrightarrow$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(12, x) \hookrightarrow$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(14, x) \hookrightarrow$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(16, x) \hookrightarrow$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(18, x) \hookrightarrow$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(25, x) \hookrightarrow$\hspace*{3em}}; &
             \\
-            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Satisfiable?}; &
+            \node [anchor=center] {}; &
+            \node [anchor=center] {\scriptsize $G(75, y)$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(75, y)$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(75, y)$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(75, y)$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(75, y)$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $G(75, y)$\hspace*{3em}}; &
+            \\
+            \node [anchor=center, inner sep=2pt, rotate=90] {\scriptsize Satisfiable?}; &
             \node [anchor=center] {\includegraphics*{plots/graph-induced-satisfiable-10-75.pdf}}; &
@@
             \node [anchor=center] {\includegraphics*{plots/graph-induced-satisfiable-25-75.pdf}}; &
-            \\
-            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Glasgow}; &
+            \\[0.3cm]
+            \node [anchor=center, inner sep=2pt, rotate=90] {\scriptsize Glasgow}; &
             \node [anchor=center] {\includegraphics*{plots/graph-induced-nodes-10-75.pdf}}; &
@@
             \node [anchor=center] {\includegraphics*{plots/graph-induced-nodes-25-75.pdf}}; &
-            \\
-            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Clasp (PB)}; &
+            \\[0.3cm]
+            \node [anchor=center, inner sep=2pt, rotate=90] {\scriptsize Clasp (PB)}; &
             \node [anchor=center] {\includegraphics*{plots/graph-clasp-induced-nodes-10-75.pdf}}; &
@@
             \node [anchor=center] {\includegraphics*{plots/graph-clasp-induced-nodes-25-75.pdf}}; &
-            \\
-            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Glucose (SAT)}; &
+            \\[0.3cm]
+            \node [anchor=center, inner sep=2pt, rotate=90] {\scriptsize Glucose (SAT)}; &
             \node [anchor=center] {\includegraphics*{plots/graph-glucose-induced-nodes-10-75.pdf}}; &
@@
             \node [anchor=center] {\includegraphics*{plots/graph-glucose-induced-nodes-25-75.pdf}}; &
-            \\
-            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Gurobi (MIP)}; &
+            \\[0.3cm]
+            \node [anchor=center, inner sep=2pt, rotate=90] {\scriptsize Gurobi (MIP)}; &
             \node [anchor=center] {\includegraphics*{plots/graph-gurobi-induced-nodes-10-75.pdf}}; &
@@
             \node [anchor=center] {\includegraphics*{plots/graph-gurobi-induced-nodes-25-75.pdf}}; &
-            \\
-            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize BBMC (Clique)}; &
+            \\[0.3cm]
+            \node [anchor=center, inner sep=2pt, rotate=90] {\scriptsize BBMC (Clique)}; &
             \node [anchor=center] {\includegraphics*{plots/graph-clique-induced-nodes-10-75.pdf}}; &
@@
     \end{tikzpicture}
-    \caption{Behaviour of other solvers on the induced variant on smaller graphs, shown in the style of
-        \cref{figure:non-induced}. The second row shows the number of search nodes used by the
-    Glasgow algorithm, the third and fourth rows show the number of decisions made by the pseudo-boolean
-and SAT solvers, the fifth shows the number of search nodes used on the MIP encoding, and the
+    \caption{Behaviour of other solvers on the induced variant using smaller target graphs with 75
+        vertices, shown in the style of \cref{figure:non-induced}. The second row shows the number
+        of search nodes used by the Glasgow algorithm, the third and fourth rows show the number of
+        decisions made by the pseudo-boolean and SAT solvers, the fifth shows the number of search
+        nodes used on the MIP encoding, and the
     final row the clique encoding.}\label{figure:alt}
@@
 
-\section{$k$-Regular Graphs}\label{section:models}
+\section{$k$-Regular Graphs}\label{section:regular}
 
@@
             \node [anchor=center] {}; &
-            \node [anchor=center] {\scriptsize $R(10, x)$\hspace*{1em}}; &
-            \node [anchor=center] {\scriptsize $R(20, x)$\hspace*{1em}}; &
-            \node [anchor=center] {\scriptsize $R(30, x)$\hspace*{3em}}; &
+            \node [anchor=center] {\scriptsize $R(10, x) \rightarrowtail R(150, y)$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $R(20, x) \rightarrowtail R(150, y)$\hspace*{1em}}; &
+            \node [anchor=center] {\scriptsize $R(30, x) \rightarrowtail R(150, y)$\hspace*{3em}}; &
             \\
-            \node [anchor=center, inner sep=4pt] [rotate=90] {\scriptsize Satisfiable?}; &
+            \node [anchor=center, inner sep=2pt] [rotate=90] {\scriptsize Satisfiable?}; &
             \node [anchor=center] {\includegraphics*{plots/graph-non-induced-kr-satisfiable-10-150.pdf}}; &
@@
             \\[0.3cm]
-            \node [anchor=center, inner sep=4pt] [rotate=90] {\scriptsize Glasgow}; &
+            \node [anchor=center, inner sep=2pt] [rotate=90] {\scriptsize Glasgow}; &
             \node [anchor=center] {\includegraphics*{plots/graph-non-induced-kr-nodes-10-150.pdf}}; &
@@
             \\[0.3cm]
-            \node [anchor=center, inner sep=4pt] [rotate=90] {\scriptsize LAD}; &
+            \node [anchor=center, inner sep=2pt] [rotate=90] {\scriptsize LAD}; &
             \node [anchor=center] {\includegraphics*{plots/graph-lad-non-induced-kr-nodes-10-150.pdf}}; &
@@
             \\[0.3cm]
-            \node [anchor=center, inner sep=4pt] [rotate=90] {\scriptsize VF2}; &
+            \node [anchor=center, inner sep=2pt] [rotate=90] {\scriptsize VF2}; &
             \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-kr-nodes-10-150.pdf}}; &
@@
     pattern degree (x-axis) and target degree (y-axis) are varied from 0 to
-    $\left|\operatorname{V}\right| - 1$.}
+    $\left|\operatorname{V}\right| - 1$, using target graphs of 150 vertices.}
     \label{figure:kr}
@@
 What about richer graph structures? We now look at what happens with the non-induced problem if we
-use $k$-regular graphs rather than the Erd\H{o}s-R\'enyi model, and then see the effects of
-introducing vertex labels. Both models still exhibit phase transition behaviour, although with
-sufficiently many labels the ``satisfiable'' region now contains a mix of satisfiable and
-unsatisfiable instances.
-
-?? We then move on to more structured instances.
+use $k$-regular graphs rather than the Erd\H{o}s-R\'enyi model, and then in the following section we
+see the effects of introducing vertex labels. Both models still exhibit phase transition behaviour,
+although with sufficiently many labels the ``satisfiable'' region now contains a mix of satisfiable
+and unsatisfiable instances.
 
@@
 Suppose our labels are drawn randomly from a set $L = \{ 1, \ldots, k \}$, where $k$ is reasonably
-small compared to $p$. Let $\ell(v)$ be the label of vertex $v$. By defining \[
+small compared to $p$. Recall that $\ell(v)$ is the label of vertex $v$. By defining \[
     \operatorname{V}(P)\vert_n = \{ v \in \operatorname{V}(P) : \ell(v) = n \} \] to be the set of
-vertices with label $k$, we can partition the pattern vertices by label into disjoint sets \[\{
-    \operatorname{V}(P)\vert_1, \ldots, \operatorname{V}(P)\vert_k \}\textnormal{\,,}\] each of
+vertices with label $k$, we can partition the pattern vertices by label into disjoint sets
+$\{\operatorname{V}(P)\vert_1, \ldots, \operatorname{V}(P)\vert_k \}\textnormal{\,,}$ each of
 which is expected to contain $\nicefrac{p}{k}$ vertices. Similarly, we may partition the target
@@
 continuous extension by taking \[ |S| = \left(\frac{\Gamma\left(\nicefrac{t}{k} +
-1\right)}{\Gamma\left(\nicefrac{t}{k} - \nicefrac{p}{k} + 1\right)}\right)^{k} \textnormal{\,,} \]
+1\right)}{\Gamma\left(\nicefrac{t}{k} - \nicefrac{p}{k} + 1\right)}\right)^{k} \textnormal{\,.} \]
 
@@
             \\
-            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Satisfiable?}; &
+            \node [anchor=center, inner sep=2pt, rotate=90] {\scriptsize Satisfiable?}; &
             \node [anchor=center] {\includegraphics*{plots/graph-non-induced-satisfiable-20-150.pdf}}; &
@@
             \\
-            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize Glasgow}; &
+            \node [anchor=center, inner sep=2pt, rotate=90] {\scriptsize Glasgow}; &
             \node [anchor=center] {\includegraphics*{plots/graph-non-induced-nodes-20-150.pdf}}; &
@@
             \\
-            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize LAD}; &
+            \node [anchor=center, inner sep=2pt, rotate=90] {\scriptsize LAD}; &
             \node [anchor=center] {\includegraphics*{plots/graph-lad-non-induced-nodes-20-150.pdf}}; &
@@
             \\
-            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize VF2}; &
+            \node [anchor=center, inner sep=2pt, rotate=90] {\scriptsize VF2}; &
             \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-nodes-20-150.pdf}}; &
@@
             \\
-            \node [anchor=center, inner sep=4pt, rotate=90] {\scriptsize VF2 Failures}; &
+            \node [anchor=center, inner sep=2pt, rotate=90] {\scriptsize VF2 Failures}; &
             \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-150.pdf}}; &
@@
 
-\subsection{Richer Label Models, and VF2's Deficiencies}\label{subsection:richer-labels}
-
-Why does VF2 find some of these instances so hard? VF2 does not track domains, like other
-algorithms, and does not detect a domain wipeout until it branches on a particular vertex.  It also
-cannot detect small domains, and only uses ``adjacency to an existing assignment'' as a branching
-heuristic. This can make it very hard for VF2 to detect that it is in an obviously failed state, or
-that an instance or subproblem is trivially unsatisfiable.
+\subsection{VF2's Struggles on Richer Label Models}\label{subsection:richer-labels}
+
+Why does VF2 find some of these instances so hard? Unlike Glasgow and LAD, VF2 does not track
+domains, and so cannot detect that there are no suitable target vertices available for a given
+pattern vertex until it branches on that vertex.  It also cannot detect small domains, and only uses
+``adjacency to an existing assignment'' as a branching heuristic. This can make it very hard for VF2
+to detect that it is in an obviously failed state, or that an instance or subproblem is trivially
+unsatisfiable.
 
@@
 
-    \caption{Cumulative number of instances solved over time, using the richer model of randomness
-    described in \cref{subsection:richer-labels}.}
+    \caption{The cumulative number of instances solved over time, using the richer model of
+    randomness described in \cref{subsection:richer-labels}.}
     \label{figure:skewed-cumulative}
@@
 We plot the cumulative number of instances solved over time for these instances in
-\cref{figure:skewed-cumulative}. Both Glasgow and LAD find all of these instances trivial, but VF2
-finds many of them extremely difficult.  The ``VF2${\leftrightarrow}$'' line shows what happens with
-VF2 if the graphs are permuted, so that the non-zero labelled vertices are given lower vertex
-numbers rather than higher vertex numbers. This
-makes VF2 behave better, but still extremely poorly compared to the other two solvers; we return to
-this in reference to work by \citet{DBLP:conf/edbt/KatsarouNT17} in the following section.
-It is also important to note for later that both VF2 variations find some satisfiable and some
-unsatisfiable instances extremely hard.
+\cref{figure:skewed-cumulative}: for a given runtime choice along the $x$-axis, the $y$ value for a
+given algorithm shows how many instances, individually, took at most $x$ milliseconds to solve with
+that algorithm. Both Glasgow and LAD find all of these instances trivial, with no instance requiring
+even ten milliseconds to solve. VF2, in contrast, finds many of them extremely difficult, and cannot
+solve 352 of the 1,000 instances with a one thousand second timeout.
+
+The ``VF2${\leftrightarrow}$'' line shows what happens with VF2 if the graphs
+are permuted, so that the non-zero labelled vertices are given lower vertex numbers rather than
+higher vertex numbers.  This makes VF2 behave better, but still extremely poorly compared to the
+other two solvers; the motivation behind this will become clear in the following section.  It is
+also important to note for the following section that both VF2 variations find some satisfiable
+\emph{and} some unsatisfiable instances extremely hard.
+
+?? Also plot CP
 
@@
 However, by far the biggest problem is in graph databases. The following section shows how
-widespread use of VF2 has not just lead to overly pessimistic conclusions regarding performance, but
-has misdirected the design of larger systems.
+widespread use of poor subgraph isomorphism algorithms has not just lead to overly pessimistic
+conclusions regarding performance, but has misdirected the design of larger systems.
 
@@
 build, and they suggest that Ullmann's algorithm and VF2 are state of the art for verification.
-And \citet{DBLP:journals/www/PengZ0LZ16} state that ``obviously, it is impossible to employ some
+\citet{DBLP:journals/www/PengZ0LZ16} state that ``obviously, it is impossible to employ some
 subgraph isomorphism algorithm, such as Ullmann or VF2'', and argue that ``in order to speed up
-query processing'', they need to create indices.
+query processing'', they need to create indices.  And \citet{DBLP:conf/isda/AzaouziR16} argue that
+``since the subgraph isomorphism test is expensive, checking all graphs of a large database can be
+unfeasible'', saying that ``naturally, the verification step is computationally more expensive since
+it requires a subgraph isomorphism''; their experiments look at candidate set reduction sizes, and
+their choice of subgraph matching algorithm is not mentioned.
 
@@
 of space.  There is also research into maintaining indices when the set of target graph changes: for
-example \citet{DBLP:conf/sigmod/YuanMYG15} look at algorithms for updating graph indices. And in
-describing a system for reusing results of queries which are sub- or super-graphs of previous
+example \citet{DBLP:conf/sigmod/YuanMYG15} look at algorithms for updating graph indices.
+And describing a system for reusing results of queries which are sub- or super-graphs of previous
 queries, \citet{DBLP:conf/edbt/WangNT16} state that querying is a ``very costly operation as it
@@
 \end{figure}
+
+?? Redo these using CP.
 
@@
 is therefore no point in continued research such as that by
-\citet{DBLP:conf/gbrpr/CarlettiFV15,o:Carletti16}
-into tweaking VF2's connectivity-based heuristics.  However, we do not claim that the ultimate
-``big data'' subgraph matching algorithm already exists: on the contrary, there is likely to be
-plenty of room for future improvement now that we understand the importance of getting algorithm
-design right.  For example, a major disadvantage of using domains is the relatively expensive
-initialisation costs, which quickly add up when dealing with large numbers of trivial instances.
-Employing a presolver is an obvious approach---and VF2 is actually good in this role
-\citep{DBLP:conf/lion/KotthoffMS16}---but there are other possibilities. For example,
-minimal or lazy forward-checking
+\citet{DBLP:conf/gbrpr/CarlettiFV15} and \citet{o:Carletti16}
+into tweaking VF2's connectivity-based heuristics.
+
+?? In particular, we saw that VF3 \citep{DBLP:conf/gbrpr/CarlettiFSV17} performed \emph{worse} than
+VF2 on harder instances.
+
+However, we do not claim that the ultimate ``big data'' subgraph matching algorithm already exists:
+on the contrary, there is likely to be plenty of room for future improvement now that we understand
+the importance of getting algorithm design right.  For example, a major disadvantage of using
+domains is the relatively expensive initialisation costs, which quickly add up when dealing with
+large numbers of trivial instances.  Employing a presolver is an obvious approach---and VF2 is
+actually good in this role \citep{DBLP:conf/lion/KotthoffMS16}---but there are other possibilities.
+For example, minimal or lazy forward-checking
 \citep{DBLP:conf/ictai/DentM94,DBLP:conf/cp/BacchusG95,o:Dent96,DBLP:conf/ecai/LarrosaM98} avoids
-constructing every domain upfront, although adopting this require alternatives to smallest domain first
-and to all-different. Such an approach may also be beneficial for huge target graphs, where having
-domains range over the entire target is impractical.
+constructing every domain upfront, although adopting this require alternatives to smallest domain
+first and to all-different. Such an approach may also be beneficial for huge target graphs, where
+having domains range over the entire target is impractical.
 

### 2017-10-31 paper/paper.tex
Stuffing

@@
        \addr University of Glasgow, Glasgow, Scotland \\
-       \name Patrick Prosser \samethanks[2] \email patrick.prosser@glasgow.ac.uk \\
+       \name Patrick Prosser \samethanks[1] \email patrick.prosser@glasgow.ac.uk \\
        \addr University of Glasgow, Glasgow, Scotland \\
@@
        \addr INSA-Lyon, LIRIS, UMR5205, F-69621, France \\
-       \name James Trimble \samethanks[2] \email j.trimble.1@research.gla.ac.uk \\
+       \name James Trimble \samethanks[1] \email j.trimble.1@research.gla.ac.uk \\
        \addr University of Glasgow, Glasgow, Scotland}
@@
     target graph. The non-induced version allows extra edges in the target, whilst the induced
-    version does not.  Although both subgraph isomorphism variants are \NP-complete, modern
-    algorithms can operate comfortably on most real-world graphs with thousands of vertices.
-    However, they cannot handle arbitrary instances of this size. We show how to generate ``really
-    hard'' random instances for subgraph isomorphism problems, which are computationally challenging
-    with only a couple of hundred vertices in the target, and only twenty to thirty pattern
-    vertices. For the non-induced version of the problem, these instances lie on a satisfiable /
-    unsatisfiable phase transition, whose location we can predict; for the induced variant, much
-    richer behaviour is observed, and constrainedness gives a better measure of difficulty than does
-    proximity to a phase transition. In both cases, we can use this to understand and improve
-    variable- and value-ordering heuristics.
-
-    When labels are introduced, as commonly seen in real-world applications and in graph databases,
-    additional rich behaviour arises.  We show that when even the simplest constraint programming
-    algorithms are used, all commonly used benchmark suites for evaluating graph database systems
-    become trivial, and that existing notions of difficulty are based almost entirely upon
-    unnecessarily poor performance from overly simplistic subgraph isomorphism algorithms. With this
-    in mind, we demonstrate that the widely researched ``filter~/ verify'' indexing technique used in
-    graph databases is a flawed design based upon a repeated misunderstanding of the empirical
-    hardness of \NP-complete problems. We explain why filter~/ verify cannot be beneficial, even
-    with harder instances than those currently used, when paired with any reasonable modern subgraph
-    isomorphism algorithm.
+    version does not.  Although both variants are \NP-complete, algorithms inspired by constraint programming
+    can operate comfortably on most real-world graphs with thousands of vertices.  However, they
+    cannot handle arbitrary instances of this size. We show how to generate ``really hard'' random
+    instances for subgraph isomorphism problems, which are computationally challenging with a couple
+    of hundred vertices in the target, and only twenty pattern vertices. For the non-induced version
+    of the problem, these instances lie on a satisfiable / unsatisfiable phase transition, whose
+    location we can predict; for the induced variant, much richer behaviour is observed, and
+    constrainedness gives a better measure of difficulty than does proximity to a phase transition.
+    These results have practical consequences: we explain why the widely researched ``filter~/
+    verify'' indexing technique used in graph databases is a flawed design based upon a repeated
+    misunderstanding of the empirical hardness of \NP-complete problems, that cannot be beneficial
+    when paired with any reasonable subgraph isomorphism algorithm.
 \end{abstract}
@@
 so there are no ``extra edges'' in the copy of the pattern that we find. We illustrate both variants
-in \cref{figure:sip}.  Although these problems are \NP-complete, modern practical subgraph
-isomorphism algorithms can handle problem instances with many hundreds of vertices in the pattern
-graph, and up to ten thousand vertices in the target graph
-\citep{DBLP:journals/pami/CordellaFSV04,DBLP:journals/ai/Solnon10,DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16},
-leading to successful applications in areas such as computer vision
+in \cref{figure:sip}.  Although these problems are \NP-complete, modern subgraph
+isomorphism algorithms based upon constraint programming techniques can handle problem instances
+with many hundreds of vertices in the pattern graph, and up to ten thousand vertices in the target
+graph
+\citep{DBLP:journals/ai/Solnon10,DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16},
+and subgraph isomorphism is used successfully in application areas including computer vision
 \citep{DBLP:journals/cviu/DamiandSHJS11,DBLP:journals/pr/SolnonDHJ15}, biochemistry
 \citep{o:10.1371/journal.pone.0076911,DBLP:conf/gbrpr/CarlettiFV15}, and pattern recognition
-\citep{DBLP:journals/ijprai/ConteFSV04}.  In particular, a labelled version of subgraph isomorphism
-is one of the key components in supporting complex queries in graph databases, which we discuss in
-detail in \cref{section:filterverify}. Typically, these systems store a fixed set of target graphs,
-and for a sequence of pattern queries, they must return every target graph which contains that
-pattern.  One common approach to this problem is to combine a subgraph isomorphism algorithm with an
-indexing system in a so-called ``filter~/ verify'' model, where invariants are used to attempt to
-pre-exclude unsatisfiable instances to avoid the cost of a subgraph isomorphism call. In this
-context, the terms \emph{matching} and \emph{verification} are often used for the subgraph
-isomorphism step (or a slightly broader problem, for example permitting wildcards).
+\citep{DBLP:journals/ijprai/ConteFSV04}.  A labelled version of subgraph isomorphism
+is also one of the key components in supporting complex queries in graph databases, which we discuss in
+detail in \cref{section:filterverify}.
 
@@
 algorithm behaviour can be beneficial, because it provides a way of generating many instances
-cheaply, and reduces the risk of over-fitting when tuning design parameters. It also allows us to
-control graph parameters such as order, density, or degree distribution, and to study scaling
-properties of algorithms with respect to these parameters.
-
-The random instances used in previous evaluations typically come from common datasets
+cheaply, and reduces the risk of over-fitting when tuning design parameters. It also allows
+experimenters to control graph parameters such as order, density, or degree distribution, and to
+study scaling properties of algorithms with respect to these parameters.  The random instances used
+in previous evaluations come from common datasets
 \citep{DBLP:journals/prl/SantoFSV03,DBLP:journals/constraints/ZampelliDS10}, which were generated by
@@
 subgraph isomorphism algorithms. Our fourth contribution is to show that the use of forward-checking
-and the right variable ordering heuristic, even without the more sophisticated recent advances in
+and the right variable-ordering heuristic, even without the more sophisticated recent advances in
 subgraph algorithms, makes the entire graph database filter~/ verify paradigm unnecessary.  Finally,
-we explain why filter~/ verify (and other recently proposed techniques) cannot be beneficial even on
-more challenging instances except if an extremely poor choice of subgraph isomorphism algorithm is
-made.
+we explain why filter~/ verify (and other recently proposed techniques, which we discuss in
+\cref{section:angryranting}) cannot be beneficial even on more challenging instances except if an
+extremely poor choice of subgraph isomorphism algorithm is made.
 
@@
 
-Part of this paper appeared at IJCAI \citep{DBLP:conf/ijcai/McCreeshPT16};
-\cref{section:introduction} has been expanded and
-\cref{section:regular,section:labelled,section:filterverify} are new material.
+Part of this paper were presented at IJCAI \citep{DBLP:conf/ijcai/McCreeshPT16}; the first three
+sections have been expanded, whilst
+\cref{section:regular,section:labelled,section:filterverify,section:angryranting} are new material.
 
@@
 only if $v$ and $w$ are not adjacent in $G$. We write $G(n, p)$ for an Erd\H{o}s-R\'enyi random
-graph with $n$ vertices, and an edge between each distinct pair of vertices with independent
-probability $p$.
+graph with $n$ vertices, and an edge between each distinct unordered pair of vertices with
+independent probability $p$.
 
 A \emph{non-induced subgraph isomorphism} from a graph $P$ (called the \emph{pattern}) to a graph
-$T$ (called the \emph{target}) is an injective mapping from $\operatorname{V}(P)$ to
+$T$ (called the \emph{target}) is an injective mapping $i$ from $\operatorname{V}(P)$ to
 $\operatorname{V}(T)$ which preserves adjacency---that is, for every adjacent $v$ and $w$ in
@@
 Scientific Linux release 6.7. We will be working with four subgraph isomorphism solvers: the Glasgow
-solver \citet{DBLP:conf/cp/McCreeshP15}; LAD \citep{DBLP:journals/ai/Solnon10}; VF2
-\citep{DBLP:journals/pami/CordellaFSV04}; and VF3 \citep{DBLP:conf/gbrpr/CarlettiFSV17}. Each was
-compiled using GCC 4.9.
-
-The Glasgow and LAD solvers store domains, and use backtracking search to build up an assignment of
-target vertices (values) to pattern vertices (variables), but differ in terms of inference and
-ordering heuristics.  The approach used by VF2 and VF3 is similar, although the domains of variables
-are not stored (in the style of conventional backtracking, rather than forward-checking), and so
-domain wipeouts are not detected until an assignment is made. Both the Glasgow and LAD solvers use
-domain sizes to guide search, whilst VF2 and VF3 use adjacency to previously-assigned vertices.
-
-We measure only the number of recursive calls (guessed assignments) made, not
-runtimes. We are not aiming to compare absolute performance between solvers; rather, we are looking
-for solver-independent patterns of difficulty. All experiments use a timeout of 1,000 seconds, which
-is enough for the Glasgow solver to solve nearly all instances (whose orders were selected with this
+solver \citep{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16}; LAD
+\citep{DBLP:journals/ai/Solnon10}; VF2 \citep{DBLP:journals/pami/CordellaFSV04}; and VF3
+\citep{DBLP:conf/gbrpr/CarlettiFSV17}. Each was compiled using GCC 4.9.
+
+The Glasgow and LAD solvers are constraint programming style algorithms (although they use dedicated
+data structures, and do not use a constraint programming toolkit). For each vertex in the pattern
+graph, they create a \emph{variable} whose \emph{domain} contains one \emph{value} for each vertex
+in the target graph. The algorithms perform backtracking search, with the goal of assigning to each
+variable a value from its domain so as to form a subgraph isomorphism. The algorithms use
+\emph{inference} to process \emph{constraints} which eliminate values from domains, ensuring that
+the injectivity (an \emph{all-different} constraint) and adjacency rules are respected. If at any
+point during search a \emph{domain wipeout} occurs (that is, if every value is removed from any
+domain), the algorithms will backtrack immediately. The search is guided by \emph{variable-ordering
+heuristics}, which select which variable (i.e.\ pattern vertex) to branch on next, and
+\emph{value-ordering} heuristics which select which value (i.e.\ target vertex) to try as part of a
+partial mapping. The two algorithms differ in terms of additional inference performed, their
+choices of underlying data structures and algorithms, and in the details of their variable- and
+value-ordering heuristics.
+
+The approach used by VF2 and VF3 is a little different: domains of variables are not stored (so the
+algorithm operates in the style of conventional backtracking), and wipeouts are not detected until
+an assignment is made. Instead, an assignment is built up by repeatedly branching only on vertices
+which are adjacent to a previously-assigned vertex.
+
+We measure only the number of recursive calls (guessed assignments) made, not runtimes. We are not
+aiming to compare absolute performance between solvers; rather, we are looking for
+solver-independent patterns of difficulty. All experiments use a timeout of 1,000 seconds, which is
+enough for the Glasgow solver to solve nearly all instances (whose orders were selected with this
 timeout in mind), although we may slightly overestimate the proportion of unsatisfiable instances
-for extremely sparse or dense pattern graphs. The LAD, VF2, and VF3 solvers experienced many more failures
-with this timeout, so our picture of just how hard the hardest instances are with these solvers is
-less detailed.
+for extremely sparse or dense pattern graphs. The LAD, VF2, and VF3 solvers experienced many more
+failures with this timeout, so our picture of just how hard the hardest instances are with these
+solvers is less detailed.
 
@@
         complexity peak with the Glasgow solver in the non-induced variant. Each point represents
-        one instance. The lines show mean search effort and mean proportion satisfiable.}
+        one instance, and the line shows mean search effort.}
     \label{figure:phase-transition}
@@
 complexity peak around the point where half the instances are satisfiable.  We also plot the search
-cost of individual instances, as points. The behaviour we observe looks remarkably similar to the
-clique decision problem and to random 3SAT problems---compare, for example,
-Figure~1 of
-\citet{DBLP:journals/cacm/Leyton-BrownHHX14}. In particular, satisfiable instances tend to be easier, but show greater
-variation than unsatisfiable instances, and there are exceptionally hard satisfiable instances
-\citep{DBLP:conf/cp/SmithG97}. (The Glasgow solver supports
-parallel search with a work-stealing strategy explicitly designed to reduce or eliminate these. We
-are not using parallel search to avoid dealing with the complexity of search tree
-measurements under parallelism.)
+cost of individual instances, as points. The behaviour we observe looks remarkably similar to random
+3SAT problems---compare, for example, Figure~1 of \citet{DBLP:journals/cacm/Leyton-BrownHHX14}. In
+particular, satisfiable instances tend to be easier, but show greater variation than unsatisfiable
+instances, and there are exceptionally hard satisfiable instances \citep{DBLP:conf/cp/SmithG97}.
 
@@
 
-\subsection{Variable and Value Ordering Heuristics}
-
-Various general principles have been considered when designing variable and value ordering
+\subsection{Variable- and Value-Ordering Heuristics}
+
+Various general principles have been considered when designing variable- and value-ordering
 heuristics for backtracking search algorithms---one of these is to try to maximise the expected
@@
 more domains are of equal size, LAD simply breaks ties lexicographically, whereas Glasgow will pick
-a variable corresponding to a pattern vertex of highest degree (?? check carefully!). This strategy
-was determined empirically, but could have been derived from the $\langle Sol \rangle$ formula:
-picking a pattern vertex of high degree will make the remaining pattern subgraph sparser, which will
-decrease the exponent in the second half of the formula, maximising the overall value. LAD does not
-apply a value ordering heuristic, but Glasgow does: it prefers target vertices of highest degree (??
-check carefully!).  Again, this was determined empirically, but it has the effect of increasing
-$\langle Sol \rangle$ by leaving as many vertices as possible available for future use. The VF2
-heuristics, in contrast, are based around preserving connectivity, which gives very little
-discrimination except on the sparsest of inputs.
-
-\subsection{Instances for Benchmarking}
-
-?? Instances for benchmarking
-
-?? Something about distance-based filtering being no use on these instances
-
-?? Low degree spread
+a variable corresponding to a pattern vertex of highest degree. This strategy was determined
+empirically, but could have been derived from the $\langle Sol \rangle$ formula: picking a pattern
+vertex of high degree will make the remaining pattern subgraph sparser, which will decrease the
+exponent in the second half of the formula, maximising the overall value. LAD does not apply a
+value-ordering heuristic, but Glasgow does: it prefers target vertices of highest degree.  Again, this was
+determined empirically, but it has the effect of increasing $\langle Sol \rangle$ by leaving as many
+vertices as possible available for future use. The search strategy used by VF2, in contrast, is
+based around preserving connectivity, which gives very little discrimination except on the sparsest
+of inputs.
 
@@
 
-In the first four rows of \cref{figure:induced} we repeat our experiments, now finding induced
-isomorphisms. With a pattern of order 10, we get two independent phase transitions: the bottom right
-halves of the plots resemble the non-induced results, and the top left halves are close to a mirror
-image. The central satisfiable region, which is away from either phase transition, is
+So far we have looked at the non-induced problem, where extra edges in the target graph are
+permitted.  In the first five rows of \cref{figure:induced} we repeat our experiments, now finding
+induced isomorphisms. With a pattern of order 10, we get two independent phase transitions: the
+bottom right halves of the plots resemble the non-induced results, and the top left halves are close
+to a mirror image. The central satisfiable region, which is away from either phase transition, is
 computationally easy, but instances near the phase transition are hard.
@@
 not being near either phase transition, instances in the centre remain computationally challenging.
-We also plot patterns of orders 14, 15 and 16, to show the transition between the two behaviours.
+We also plot patterns of orders 14, 15 and 16, to show the change between the two behaviours.
 
@@
 problem, if we replace both inputs with their complements, the solutions remain the same.  For the
-Glasgow solver, this is the case. This should be expected, because this complement property is
+Glasgow solver, this is almost the case. This should be expected, because this complement property is
 precisely how the Glasgow solver handles the induced variant (although the heuristics may differ
@@
 We might guess that degree-based heuristics would just not work for the induced problem: for any
-claim about the degree, the opposite will hold for the complement constraints. However, empirically,
-this is not the case: on the final row of \cref{figure:induced}, we show whether it is better to use
-the original pattern and target as the input to the Glasgow algorithm, or to take the complements.
-(The only steps performed by the Glagsow algorithm which differ under taking the complements are
-the degree-based heuristics.  LAD and VF2 are not symmetric in this way: LAD performs a filtering
-step using degree information, but does not consider the complement degree, and VF2 uses
-connectivity in the pattern graph.)
+claim about the degree, the opposite will hold for the complement constraints. Should we therefore
+resort to just using ``smallest domain first'', abandoning degree-based tiebreaking and the
+value-ordering heuristic? Empirically, this is not the case: on the final row of
+\cref{figure:induced}, we show whether it is better to use the original pattern and target as the
+input to the Glasgow algorithm, or to take the complements.  (The only steps performed by the
+Glagsow algorithm which differ under taking the complements are the degree-based heuristics.  LAD,
+VF2, and VF3 are not symmetric in this way: LAD performs a filtering step using degree information,
+but does not consider the complement degree, and VF2 and VF3 use connectivity.)
 
@@
 
-In practice, this is unlikely to be a problem: most real-world instances are extremely sparse and
-are usually easy. In this situation, these experiments justify reusing the
-non-induced heuristics on induced problems.
+In practice, this is unlikely to be a problem: most real-world instances are relatively sparse. In
+this situation, these experiments justify reusing the non-induced heuristics on induced problems.
 
@@
 
-Our pseudo-boolean (PB) encoding is as follows. For each pattern vertex $v$ and each target vertex
+The pseudo-boolean (PB) encoding is as follows. For each pattern vertex $v$ and each target vertex
 $w$, we have a binary variable which takes the value 1 if and only if $v$ is mapped to $w$.
@@
 inside the central region, and the easy satisfiable instances remain easy. Similar results are seen
-with the Glucose SAT solver \citep{o:glucose} using a direct encoding of the cardinality constraints.
-We also show an integer program encoding: the Gurobi solver is only able to solve some of the
-trivial satisfiable instances, and was almost never able to prove unsatisfiability within the time
-limit.
+with the Glucose SAT solver \citep{o:glucose} using a direct encoding.  We also show an integer
+program encoding: the Gurobi solver is only able to solve some of the trivial satisfiable instances,
+and was almost never able to prove unsatisfiability within the time limit.
 
@@
 \binom{p}{2}} \right)}{\log t^{\underline{p}}} \end{equation} for induced isomorphisms, which we plot on the
-fifth row of \cref{figure:induced}. We see that constrainedness predicts that the central region
+sixth row of \cref{figure:induced}. We see that constrainedness predicts that the central region
 will still be relatively difficult for larger pattern graphs: although the problem is
@@
 
-What about richer graph structures? We now look at what happens with the non-induced problem if we
-use $k$-regular graphs rather than the Erd\H{o}s-R\'enyi model, and then in the following section we
-see the effects of introducing vertex labels. Both models still exhibit phase transition behaviour,
-although with sufficiently many labels the ``satisfiable'' region now contains a mix of satisfiable
-and unsatisfiable instances.
+What about richer graph structures? We briefly look at what happens with the non-induced problem if
+we use $k$-regular graphs rather than the Erd\H{o}s-R\'enyi model, and then in the following section
+we see the effects of introducing vertex labels. Both models still exhibit phase transition
+behaviour, although with sufficiently many labels the ``satisfiable'' region now contains a mix of
+satisfiable and unsatisfiable instances.
 
@@
 
-?? Talk about degree heuristics
+Regularity means that degree-based heuristics have no information to work with. Additionally, the
+degree-based filtering techniques used by Glasgow and LAD have no effect on these graphs, and so at
+the top of search, every domain is identical. However, once a guessed assignment has been made,
+selecting from small domains first remains an effective strategy to guide the remainder of the
+search.
 
@@
     \operatorname{V}(P)\vert_n = \{ v \in \operatorname{V}(P) : \ell(v) = n \} \] to be the set of
-vertices with label $k$, we can partition the pattern vertices by label into disjoint sets
+vertices with label $n$, we can partition the pattern vertices by label into disjoint sets
 $\{\operatorname{V}(P)\vert_1, \ldots, \operatorname{V}(P)\vert_k \}\textnormal{\,,}$ each of
@@
             \\
-            \node [anchor=center, inner sep=2pt, rotate=90] {\scriptsize VF2 Failures}; &
-            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-l2-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-l3-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-l5-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-l10-150.pdf}}; &
-            \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-l20-150.pdf}}; &
-            \\
+            % \node [anchor=center, inner sep=2pt, rotate=90] {\scriptsize VF2 Failures}; &
+            % \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-150.pdf}}; &
+            % \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-l2-150.pdf}}; &
+            % \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-l3-150.pdf}}; &
+            % \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-l5-150.pdf}}; &
+            % \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-l10-150.pdf}}; &
+            % \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-l20-150.pdf}}; &
+            % \\
         };
@@
     subsequent rows, the average number of search nodes needed to solve an instance, for three
-    different solvers, and on the final row, the proportion of instances aborted due to a timeout
-    with VF2.}\label{figure:labels}
+    different solvers}\label{figure:labels}
 \end{figure}
@@
 
-\subsection{VF2's Struggles on Richer Label Models}\label{subsection:richer-labels}
+\subsection{Richer Label Models}\label{subsection:richer-labels}
 
@@
 
-The ``VF2${\leftrightarrow}$'' line shows what happens with VF2 if the graphs
-are permuted, so that the non-zero labelled vertices are given lower vertex numbers rather than
-higher vertex numbers.  This makes VF2 behave better, but still extremely poorly compared to the
-other two solvers; the motivation behind this will become clear in the following section.  It is
-also important to note for the following section that both VF2 variations find some satisfiable
-\emph{and} some unsatisfiable instances extremely hard.
+The ``VF2${\leftrightarrow}$'' line shows what happens with VF2 if the graphs are permuted, so that
+the non-zero labelled vertices are given lower vertex numbers rather than higher vertex numbers.
+This makes VF2 behave better, but still extremely poorly compared to the other two solvers; the
+motivation behind this modification will become clear later in this paper.  It will also be
+important to remember that both VF2 variations find some satisfiable \emph{and} some unsatisfiable
+instances extremely hard.
 
@@
 try to map every adjacent vertex in the clique before considering the difficult isolated vertex.
-This example can also be extended slightly to fool any simple static heuristic, or any simple label
-counting mechanism.
+This example can be extended slightly to fool any simple static heuristic, or any simple label
+counting mechanism (for example, by attaching an additional vertex with a unique label to the
+clique).
 
@@
 must have a very small proof of unsatisfiability. These instances \emph{should} be trivial with any
-decent matching algorithm. Thus, all filtering \emph{should} be doing is eliminating the startup
+decent subgraph isomorphism algorithm. Thus, all filtering \emph{should} be doing is eliminating the startup
 costs of a trivial subgraph isomorphism call. The fact that filtering was successful empirically
 should make us wonder whether the subgraph isomorphism algorithm being used was excessively
-primitive. Indeed, the matching algorithm used is described only as ``the simplest approach'' in the
+primitive. Indeed, the subgraph isomorphism algorithm used is described only as ``the simplest approach'' in the
 paper.  Additionally, the experiments focus on reducing the size of the candidate set, without
@@
 $T_{\mathit{iso\_test}}$ does not change much is not justified experimentally, and no consideration
-is given as to whether this would hold true for other subgraph matching algorithms.
+is given as to whether this would hold true for other subgraph isomorphism algorithms.
 
@@
 the key to reducing query response time is to minimize the size of the candidate answer
-set''.\end{displayquote} This claim becomes understandable when one examines the subgraph matching
-algorithm chosen for the verification step \citep{DBLP:journals/jacm/Ullmann76}: as the algorithm
-predates techniques like generalised arc-consistent all-different propagation \citep{DBLP:conf/aaai/Regin94}, it
-cannot immediately detect unsatisfiability in simple cases like there being two red vertices in the
-pattern but only one in the target\footnote{Although interestingly, this algorithm effectively does
-forward-checking and has a variable-ordering heuristic, before these concepts appeared in the
-constraints literature.}.
+set''.\end{displayquote} This claim becomes understandable when one examines the subgraph
+isomorphism algorithm chosen for the verification step \citep{DBLP:journals/jacm/Ullmann76}: as the
+algorithm predates techniques like generalised arc-consistent all-different propagation
+\citep{DBLP:conf/aaai/Regin94}, it cannot immediately detect unsatisfiability in simple cases like
+there being two red vertices in the pattern but only one in the target\footnote{Although
+interestingly, this algorithm effectively does forward-checking and has a variable-ordering
+heuristic, before these concepts appeared in the constraints literature.}.
 
@@
 it requires a subgraph isomorphism''; their experiments look at candidate set reduction sizes, and
-their choice of subgraph matching algorithm is not mentioned.
+their choice of subgraph isomorphism algorithm is not mentioned.
 
@@
 are now target graphs). \citet{DBLP:journals/tkde/HongZLY15} look at graph database subgraph
-matching with an additional set-similarity constraint, and state that Ullmann's algorithm and VF2
+isomorphism with an additional set-similarity constraint, and state that Ullmann's algorithm and VF2
 are ``usually costly for large graphs'' because they ``do not utilize any index structure''. They
@@
 
-These beliefs are not entirely unfounded: none of the matching algorithms considered in these papers
-will immediately detect if a pattern contains two red vertices, whilst the target graph contains
-only one. This kind of flaw \emph{should} be picked up at the top of search by an all-different
-propagator \citep{DBLP:conf/aaai/Regin94}.  However, as \citet{DBLP:journals/pvldb/KatsarouNT15}
-note, some variation on VF2 \citep{DBLP:journals/pami/CordellaFSV04} is the usual matching algorithm of choice for
+These beliefs are not entirely unfounded: none of the isomorphism algorithms considered in these
+papers will immediately detect if a pattern contains two red vertices, whilst the target graph
+contains only one. This kind of flaw \emph{should} be picked up at the top of search by an
+all-different propagator \citep{DBLP:conf/aaai/Regin94}.  However, as
+\citet{DBLP:journals/pvldb/KatsarouNT15} note, some variation on VF2
+\citep{DBLP:journals/pami/CordellaFSV04} is the usual subgraph isomorphism algorithm of choice for
 graph database systems, although \citeauthor{DBLP:journals/jacm/Ullmann76}'s algorithm is sometimes
-chosen. Other approaches have been considered, albeit not with algorithms strong enough to
-establish all-difference. For example, \citet{DBLP:journals/pvldb/ShangZLY08}
-propose an algorithm which makes use of the frequency of various features to guide search;
-\citet{DBLP:journals/pvldb/LeeHKL12} determine experimentally that this technique tends to be very
-effective, even on families of graphs for which it was not designed.
+chosen. Other approaches have been considered, albeit not with algorithms strong enough to establish
+all-difference. For example, \citet{DBLP:journals/pvldb/ShangZLY08} propose an algorithm which makes
+use of the frequency of various features to guide search; \citet{DBLP:journals/pvldb/LeeHKL12}
+determine experimentally that this technique tends to be very effective, even on families of graphs
+for which it was not designed.
 
 We therefore believe it would be unlikely to cause too much astonishment if we suggested that a
-better matching algorithm could be dropped in as a black box replacement in graph databases systems
-to improve their performance.  This is not our claim. Instead, this paper shows that better
-algorithms invalidate the flawed premise underlying the entire filter~/ verify approach. We will now
-demonstrate that filter~/ verify is simply a poor workaround for the kinds of deficiency in VF2
-demonstrated towards the end of the previous section.
+better subgraph isomorphism algorithm could be dropped in as a black box replacement in graph
+databases systems to improve their performance.  This is not our claim. Instead, this paper shows
+that better algorithms invalidate the flawed premise underlying the entire filter~/ verify approach.
+We will now demonstrate that filter~/ verify is simply a poor workaround for the kinds of deficiency
+in VF2 demonstrated towards the end of the previous section.
 
@@
 Rather than the modern techniques discussed used by Glasgow and LAD, we deliberately select a very
-simple starting point: a forward-checking algorithm which uses the smallest domain first heuristic,
-the simple counting all-different propagator, and degree and neighbourhood degree sequence filtering
-at the top of search. Even this simple algorithm finds every instance nearly trivial, and no
-exponential behaviour is observed. As we show in \cref{figure:biiiig-data}, the \emph{hardest}
-instance for the AIDS dataset requires 59 recursive calls, and 146,175 of the instances can be
-solved without search (we measure recursive calls to demonstrate hardness, rather than algorithm
-tuning and good constant factors; $10^4$ to $10^6$ recursive calls per second is feasible).  For
-PDBS, the hardest problem requires 1,776 recursive calls (which occurs 20 times, due to duplicated
-queries), and none of the rest require more than 182 calls. For PCM the hardest problem requires
-3,261 calls, and for PPI, 26. In other words, none of these instances are in any way computationally
-hard even for a simple domain-based algorithm, even before introducing stronger filtering,
-inference, or heuristics. Furthermore it is certainly not the case that a sequential scan is
-infeasible as so many filter~/ verify papers claim.  The exponential behaviour seen by others in
-these instances is purely down to a bad choice of subgraph isomorphism algorithm, and these
-instances are not inherently ``really hard''.
+simple starting point: a forward-checking algorithm which uses the ``smallest domain first''
+heuristic, the simple counting all-different propagator, and degree and neighbourhood degree
+sequence filtering at the top of search. Even this simple algorithm finds every instance nearly
+trivial, and no exponential behaviour is observed. As we show in \cref{figure:biiiig-data}, the
+\emph{hardest} instance for the AIDS dataset requires 59 recursive calls, and 146,175 of the
+instances can be solved without search.  For PDBS, the hardest problem requires 1,776 recursive
+calls (which occurs 20 times, due to duplicated queries), and none of the rest require more than 182
+calls. For PCM the hardest problem requires 3,261 calls, and for PPI, 26. In other words, none of
+these instances are in any way computationally hard even for a simple domain-based algorithm, even
+before introducing stronger filtering, inference, or heuristics. Furthermore it is certainly not the
+case that a sequential scan is infeasible as so many filter~/ verify papers claim.  The exponential
+behaviour seen by others in these instances is purely down to a bad choice of subgraph isomorphism
+algorithm, and these instances are not inherently ``really hard''.
 
@@
 
-\subsection{Rethinking Graph Matching for Database Systems}
-
-The evidence so far clearly shows a need for graph database systems to be redesigned making use of
-knowledge from constraint programming and artificial intelligence, and that this knowledge must inform
-the systems as a whole, and not just parts of them.  Having demonstrated that filter~/ verify is
-both theoretically and practically a flawed approach to designing graph databases based upon a
-repeated misconception that \NP-completeness means long runtimes, we now briefly discuss how future
-graph database systems should be designed, and how such systems can help support strong subgraph
-isomorphism algorithms.
-
-First, though, we briefly consider whether other lessons can be learned from constraint programming:
-unfortunately indexing is not the only incorrect design choice being in these systems. For example,
-recent work by \citet{DBLP:conf/edbt/KatsarouNT17} looks at what they call ``straggler'' queries,
-which are the small subset of queries that they observe taking much longer than others to solve. The
+\section{Further Lessons}\label{section:angryranting}
+
+Unfortunately, indexing is not the only incorrect design choice being made in systems built around
+subgraph isomorphism algorithms, and lessons from constraint programming can be applied to other
+subgraph-related research. We illustrate this by looking at several recent papers on subgraph
+isomorphism, showing how misleading conclusions are being reached due to domain-based algorithms not
+being used or understood.
+
+\subsection{Straggler Queries?}
+
+\citet{DBLP:conf/edbt/KatsarouNT17} look at what they call ``straggler'' queries, which they define
+to be the small subset of queries that they observe taking much longer than others to solve. The
 authors present an approach to address this perceived problem that they describe as novel: ``instead
@@
 heuristics is not noted, and no consideration is given to even the simplest dynamic ordering
-heuristics like smallest domain first \citep{DBLP:journals/ai/HaralickE80}. A little thought shows
-that of the orderings proposed, those involving placing rare labels first are effectively poor
+heuristics like ``smallest domain first'' \citep{DBLP:journals/ai/HaralickE80}. A little thought
+shows that of the orderings proposed, those involving placing rare labels first are effectively poor
 approximations to a static ``smallest domain at top of search first'' ordering, whilst the remainder
-use degree as we did earlier in this paper. However, \citeauthor{DBLP:conf/edbt/KatsarouNT17} do
-not investigate any algorithm which employs domains, let alone strong variable- or value-ordering
+use degree as we did earlier in this paper. However, \citeauthor{DBLP:conf/edbt/KatsarouNT17} do not
+investigate any algorithm which employs domains, let alone strong variable- or value-ordering
 heuristics or all-different propagation, and do not consider that their apparent successes could be
-due to VF2's weaknesses rather than an inherent property of \NP-completeness. We saw in the previous
-section that although permuting input graphs could improve VF2's behaviour on some instances, doing
-so does not make its performance come close to that of domain-based algorithms.
-\citeauthor{DBLP:conf/edbt/KatsarouNT17} say they ``hope that our findings will open up new research
-directions, striving to find appropriate, per-query, isomorphic rewritings, in combination with
-alternate per-query sub-iso algorithms that can yield large improvements''. We believe it is
-important instead to emphasis previous research directions that have already solved most of this
-problem, before any further research effort is wasted.
+due to VF2's weaknesses rather than an inherent property of \NP-completeness.
+
+We saw in the previous sections that although permuting input graphs could improve VF2's behaviour
+on some instances, doing so does not make its performance come close to that of domain-based
+algorithms.  \citeauthor{DBLP:conf/edbt/KatsarouNT17} say they ``hope that our findings will open up
+new research directions, striving to find appropriate, per-query, isomorphic rewritings, in
+combination with alternate per-query sub-iso algorithms that can yield large improvements''. We
+believe it is important instead to emphasis previous research directions that have already solved
+most of this problem, and to help ensure that these techniques become more widely known.
+
+\subsection{Portfolios}
 
@@
 To test this suggestion, we return briefly to the experiments on the datasets discussed earlier in
-this section. What if we had not used a smallest domain first heuristic, and instead used only
+this section. What if we had not used a ``smallest domain first'' heuristic, and instead used only
 degree-based heuristics explained earlier in this paper? Although we still find all four of these
@@
 technique is due to the ease of sometimes getting better results out of a poor algorithm, rather
-than being an inherent hardness property, and it would not help on really hard problem instances. We
-point to \citet{DBLP:journals/heuristics/Gent98}'s admonition on how heuristics are benchmarked as
-being similarly relevant here: \begin{displayquote}``Benchmark problems should be hard. I report on
-    the solution of the five open benchmark problems introduced \ldots for testing bin packing
-    problems. Since the solutions were found either by hand or by using very simple heuristic
-    methods, these problems would appear to be easy. In four cases I give improved packings to
-    refute conjectures that previously reported packings were optimal, and I give a proof that the
-    fifth conjecture was correct. \ldots Future experimenters should be careful to perform tests on
-problems that can reasonably be regarded as hard.''\end{displayquote} Claiming huge improvements
-from an algorithm portfolio consisting only of variations of poor algorithms does not demonstrate a
-genuine improvement over the state of the art.
+than being an inherent hardness property, and it would not help on really hard problem instances.
+
+\subsection{Search Ordering}
 
@@
 
+?? NVMe \citep{DBLP:conf/wise/ShenZ17}
+
 ?? In particular, we saw that VF3 \citep{DBLP:conf/gbrpr/CarlettiFSV17} performed \emph{worse} than
@@
 
-However, we do not claim that the ultimate ``big data'' subgraph matching algorithm already exists:
-on the contrary, there is likely to be plenty of room for future improvement now that we understand
-the importance of getting algorithm design right.  For example, a major disadvantage of using
-domains is the relatively expensive initialisation costs, which quickly add up when dealing with
-large numbers of trivial instances.  Employing a presolver is an obvious approach---and VF2 is
-actually good in this role \citep{DBLP:conf/lion/KotthoffMS16}---but there are other possibilities.
-For example, minimal or lazy forward-checking
-\citep{DBLP:conf/ictai/DentM94,DBLP:conf/cp/BacchusG95,o:Dent96,DBLP:conf/ecai/LarrosaM98} avoids
-constructing every domain upfront, although adopting this require alternatives to smallest domain
-first and to all-different. Such an approach may also be beneficial for huge target graphs, where
-having domains range over the entire target is impractical.
-
-There is also scope for precalculating supporting information about target graphs. For example,
-neighbourhood degree sequences and supplemental graphs could both be pre-computed and stored. The aim
-here is to reduce the initialisation costs of a good subgraph isomorphism algorithm, and not to
-provide indexing (although additionally using this information as an index may not hurt, if
-initialisation is still costly).
-
-Our central message, though, is not just to say that graph database systems must use a better
-algorithm.  Instead, this paper shows that such systems need to be designed with a better
-understanding of the empirical hardness of \NP-complete problems, that subgraph isomorphism
-algorithms must not be treated as a black box, and that lessons learned in constraint programming
-and artificial intelligence must not go unheeded in other domains.
+\subsection{Benchmarking on Hard Instances}
+
+?? VF3 milliseconds
+
+?? \citet{DBLP:journals/tcbb/BonniciG17} compare various non-CP approaches to LAD. Benchmark sets
+whose difficulties are measured between microseconds and milliseconds
+
+\citet{DBLP:journals/heuristics/Gent98} describes the importance of hard benchmark instances:
+
+\begin{displayquote}
+    ``Benchmark problems should be hard. I report on the solution of the five open benchmark
+    problems introduced \ldots for testing bin packing problems. Since the solutions were found
+    either by hand or by using very simple heuristic methods, these problems would appear to be
+    easy. In four cases I give improved packings to refute conjectures that previously reported
+    packings were optimal, and I give a proof that the fifth conjecture was correct. \ldots Future
+    experimenters should be careful to perform tests on problems that can reasonably be regarded as
+    hard.''
+\end{displayquote}
+
+?? Claiming huge improvements from an algorithm portfolio consisting only of variations of poor
+algorithms does not demonstrate a genuine improvement over the state of the art.
+
+?? Need for more hard instances
+
+?? Something about distance-based filtering being no use on our instances, low degree spread
 
@@
 
+We do not claim that the ultimate ``big data'' subgraph matching algorithm already exists:
+on the contrary, there is likely to be plenty of room for future improvement now that we understand
+the importance of getting algorithm design right.  For example, a major disadvantage of using
+domains is the relatively expensive initialisation costs, which quickly add up when dealing with
+large numbers of trivial instances. Employing a presolver is an obvious approach---and VF2 is
+actually good in this role \citep{DBLP:conf/lion/KotthoffMS16}---but there are other possibilities.
+For example, minimal or lazy forward-checking
+\citep{DBLP:conf/ictai/DentM94,DBLP:conf/cp/BacchusG95,o:Dent96,DBLP:conf/ecai/LarrosaM98} avoids
+constructing every domain upfront, although adopting this require alternatives to ``smallest domain
+first'' and to all-different. Such an approach may also be beneficial for huge target graphs, where
+having domains range over the entire target is impractical.
+
+There is also scope for precalculating supporting information about target graphs. For example,
+neighbourhood degree sequences and supplemental graphs could both be pre-computed and stored. The aim
+here is to reduce the initialisation costs of a good subgraph isomorphism algorithm, and not to
+provide indexing (although additionally using this information as an index may not hurt, if
+initialisation is still costly).
+
+We must stress that we are not simply concluding that graph database systems should use a better
+subgraph isomorphism algorithm.  Instead, this paper has shown that such systems need to be designed
+respecting our understanding of the empirical hardness of \NP-complete problems, that subgraph
+isomorphism algorithms must not be treated as a black box, and that lessons learned in constraint
+programming and artificial intelligence must not go unheeded in other domains.
+
+\acks{The authors would like to thank Kitty Meeks and Craig Reilly for their comments. The heatmap
+schemes used in this paper were created with the assistance of the ``Chroma.js Color Scale Helper''
+by Gregor Aisch.}
+
 \bibliography{paper}

### 2017-10-31 biiiiiig-data/gecode/sip.cc
Biiiiig data

@@ -1,0 +1,198 @@
+#include <gecode/int.hh>
+#include <gecode/search.hh>
+
+#include <algorithm>
+#include <cstdlib>
+#include <fstream>
+#include <iostream>
+#include <map>
+#include <memory>
+#include <numeric>
+#include <tuple>
+#include <utility>
+#include <vector>
+
+using namespace Gecode;
+
+using std::accumulate;
+using std::cerr;
+using std::cout;
+using std::endl;
+using std::flush;
+using std::iota;
+using std::ifstream;
+using std::make_tuple;
+using std::make_unique;
+using std::map;
+using std::sort;
+using std::string;
+using std::tuple;
+using std::unique_ptr;
+using std::vector;
+
+struct Graph
+{
+    string name;
+    int size;
+    vector<int> label;
+    vector<vector<int> > adj;
+};
+
+auto load(const string & filename, vector<Graph> & graphs, map<string, int> & labels_map) -> bool
+{
+    ifstream file{ filename };
+    if (! file) {
+        cerr << "Error opening " << filename << endl;
+        return false;
+    }
+
+    string name;
+    while (file >> name) {
+        if (0 != name.compare(0, 1, "#", 0, 1)) {
+            cerr << "Bad name '" << name << "'" << endl;
+            return false;
+        }
+        name.erase(0, 1);
+
+        int size;
+        file >> size;
+
+        graphs.push_back(Graph{ });
+        auto & g = graphs.back();
+        g.name = name;
+        g.size = size;
+        g.label.resize(size);
+        g.adj.resize(size, vector<int>(size, 0));
+
+        for (int l = 0 ; l < size ; ++l) {
+            string label;
+            file >> label;
+            if (! labels_map.count(label))
+                labels_map.emplace(label, labels_map.size() + 1);
+
+            g.label[l] = labels_map.find(label)->second;
+        }
+
+        int edge_count;
+        file >> edge_count;
+        for (int e = 0 ; e < edge_count ; ++e) {
+            int f, t;
+            file >> f >> t;
+            g.adj[f][t] = g.adj[t][f] = 1;
+        }
+    }
+
+    cerr << filename << " " << graphs.size() << endl;
+
+    return true;
+}
+
+class SIP : public Space
+{
+    public:
+        IntVarArray mapping;
+
+        SIP(const Graph & pattern, const Graph & target, int maximum_label) :
+            mapping(*this, pattern.size, 0, target.size - 1)
+        {
+            // injectivity
+            distinct(*this, mapping);
+
[…]

### 2017-11-01 paper/paper.tex
More of the things

@@
     \caption{Cumulative number of instances solved as a function of search space size, using four
-    graph database datasets and a simple domain-based subgraph isomorphism algorithm. Note that the
-    $x$-axis shows recursive calls, and does \emph{not} use a log scale.}
+    graph database datasets and a simple constraint programming subgraph isomorphism algorithm. Note
+    that the $x$-axis shows recursive calls, and does \emph{not} use a log scale.}
     \label{figure:biiiig-data}
@@
 
-?? Redo these using CP.
-
-Rather than the modern techniques discussed used by Glasgow and LAD, we deliberately select a very
-simple starting point: a forward-checking algorithm which uses the ``smallest domain first''
-heuristic, the simple counting all-different propagator, and degree and neighbourhood degree
-sequence filtering at the top of search. Even this simple algorithm finds every instance nearly
-trivial, and no exponential behaviour is observed. As we show in \cref{figure:biiiig-data}, the
-\emph{hardest} instance for the AIDS dataset requires 59 recursive calls, and 146,175 of the
-instances can be solved without search.  For PDBS, the hardest problem requires 1,776 recursive
-calls (which occurs 20 times, due to duplicated queries), and none of the rest require more than 182
-calls. For PCM the hardest problem requires 3,261 calls, and for PPI, 26. In other words, none of
-these instances are in any way computationally hard even for a simple domain-based algorithm, even
-before introducing stronger filtering, inference, or heuristics. Furthermore it is certainly not the
-case that a sequential scan is infeasible as so many filter~/ verify papers claim.  The exponential
-behaviour seen by others in these instances is purely down to a bad choice of subgraph isomorphism
-algorithm, and these instances are not inherently ``really hard''.
+Rather than the modern techniques used by Glasgow and LAD, we deliberately select a very simple
+starting point: a constraint programming model implemented in Gecode 5.1.0 using only
+toolkit-provided constraints and heuristics. The model uses a ``smallest domain first'' variable
+ordering heuristic with tie-breaking on degree, an all-different constraint, extensional constraints
+for adjacency, and labelled-degree filtering at the top of search. Even this simple approach finds
+nearly every instance trivial, and no exponential behaviour is observed. As we show in
+\cref{figure:biiiig-data}, the \emph{hardest} instance for the AIDS dataset requires 287 recursive
+calls, and 149,268 of the instances can be solved without search.  For PDBS, the hardest problem
+requires 3,161 recursive calls (which occurs 20 times, due to duplicated queries). For both PCM and
+PPI, the hardest problem requires 23 recursive calls.  In other words, none of these instances are
+in any way computationally hard even for a simple domain-based algorithm, even before introducing
+stronger filtering, inference, or heuristics.  Furthermore it is certainly not the case that a
+sequential scan is infeasible as so many filter~/ verify papers claim.  The exponential behaviour
+seen by others in these instances is purely down to a bad choice of subgraph isomorphism algorithm,
+and these instances are not inherently ``really hard''.
 
@@
 To test this suggestion, we return briefly to the experiments on the datasets discussed earlier in
-this section. What if we had not used a ``smallest domain first'' heuristic, and instead used only
-degree-based heuristics explained earlier in this paper? Although we still find all four of these
-datasets easy, and the changes to the runtimes are hard to detect (more time is spent in
-initialisation than in search), the worst case in the AIDS dataset now takes 4,051 recursive calls
-as opposed to 59.  How about if, additionally, we do not detect domain wipeouts until attempting to
-branch on an empty variable?  Suddenly three of the four datasets become extremely hard, with many
-instances now timing out after making hundreds of millions of recursive calls; the PPI dataset now
-requires up to 168,451 calls for some queries. Furthermore, slight changes to the input vertex
-ordering can now make some of these ``hard'' instances easy again.
-
-This backs up our suspicions: the apparent success of \citet{DBLP:conf/edbt/KatsarouNT17}'s
-technique is due to the ease of sometimes getting better results out of a poor algorithm, rather
-than being an inherent hardness property, and it would not help on really hard problem instances.
+this section. What if we modify our solver so that it does not use ``smallest domain first'', and
+does not detect domain wipeouts until attempting to branch on an empty variable?  Suddenly three of
+the four datasets become extremely hard, with many instances now timing out after making hundreds of
+millions of recursive calls, and the PPI dataset now requires hundreds of thousands of calls for
+some queries.  Furthermore, slight changes to the input vertex ordering can now make some of these
+``hard'' instances easy again.  This backs up our suspicions: the apparent success of
+\citet{DBLP:conf/edbt/KatsarouNT17}'s technique is due to the ease of sometimes getting better
+results out of a poor algorithm, rather than being an inherent hardness property, and it would not
+help on really hard problem instances.
 

### 2017-11-01 paper/paper.tex
Bits of things

@@
 perceived difficulty of each of these datasets comes down to the widespread use of especially poor
-subgraph isomorphism algorithms. Our fourth contribution is to show that the use of forward-checking
-and the right variable-ordering heuristic, even without the more sophisticated recent advances in
-subgraph algorithms, makes the entire graph database filter~/ verify paradigm unnecessary.  Finally,
-we explain why filter~/ verify (and other recently proposed techniques, which we discuss in
-\cref{section:angryranting}) cannot be beneficial even on more challenging instances except if an
-extremely poor choice of subgraph isomorphism algorithm is made.
+subgraph isomorphism algorithms. Our fourth contribution is to show that the simplest constraint
+programming approach, even without the more sophisticated recent advances in subgraph algorithms,
+makes the entire graph database filter~/ verify paradigm unnecessary.  Finally, we explain why
+filter~/ verify and other recently proposed techniques cannot be beneficial even on more challenging
+instances except if an extremely poor choice of subgraph isomorphism algorithm is made.
 
@@
 sections have been expanded, whilst
-\cref{section:regular,section:labelled,section:filterverify,section:angryranting} are new material.
+\cref{section:regular,section:labelled,section:filterverify} are new material.
 
@@
 
-The experiments in this paper are performed on systems with Intel Xeon E5-4650 v2 CPUs, running
-Scientific Linux release 6.7. We will be working with four subgraph isomorphism solvers: the Glasgow
-solver \citep{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16}; LAD
+The experiments in this paper are performed on systems with Intel Xeon E5-4650 v2 CPUs and 768GBytes
+RAM, running Scientific Linux release 6.7. We will be working with four subgraph isomorphism
+solvers: the Glasgow solver \citep{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16}; LAD
 \citep{DBLP:journals/ai/Solnon10}; VF2 \citep{DBLP:journals/pami/CordellaFSV04}; and VF3
@@
 
-?? Also plot CP
-
 An even more extreme example of VF2's misbehaviour can be seen in \cref{figure:vf2-stupidity}. Here
@@
 that better algorithms invalidate the flawed premise underlying the entire filter~/ verify approach.
-We will now demonstrate that filter~/ verify is simply a poor workaround for the kinds of deficiency
-in VF2 demonstrated towards the end of the previous section.
+We will now demonstrate empirically that filter~/ verify is simply a poor workaround for the kinds
+of deficiency in VF2 demonstrated towards the end of the previous section.
 
@@
 
-To demonstrate that a pure subgraph isomorphism approach is feasible, with no indexing or supporting
-preprocessing, we look at four datasets commonly used in graph indexing evaluations
-\citep{o:10.1371/journal.pone.0076911}. We do not claim that these are high-quality datasets or
-that the associated queries are sensible, merely that they are widely used.
+To show that a pure subgraph isomorphism approach is feasible, with no indexing or supporting
+preprocessing, we look at four datasets commonly used in graph indexing evaluations. We do not claim
+that these are high-quality datasets or that the associated queries are sensible, merely that
+following the example of \citet{o:10.1371/journal.pone.0076911}, they are widely used.
 
@@
 for adjacency, and labelled-degree filtering at the top of search. Even this simple approach finds
-nearly every instance trivial, and no exponential behaviour is observed. As we show in
-\cref{figure:biiiig-data}, the \emph{hardest} instance for the AIDS dataset requires 287 recursive
-calls, and 149,268 of the instances can be solved without search.  For PDBS, the hardest problem
-requires 3,161 recursive calls (which occurs 20 times, due to duplicated queries). For both PCM and
-PPI, the hardest problem requires 23 recursive calls.  In other words, none of these instances are
-in any way computationally hard even for a simple domain-based algorithm, even before introducing
-stronger filtering, inference, or heuristics.  Furthermore it is certainly not the case that a
-sequential scan is infeasible as so many filter~/ verify papers claim.  The exponential behaviour
-seen by others in these instances is purely down to a bad choice of subgraph isomorphism algorithm,
-and these instances are not inherently ``really hard''.
-
-There are at least four benefits beyond simplicity towards abolishing filter~/ verify:
+nearly every instance in each of the four datasets trivial, and no exponential behaviour is
+observed. As we show in \cref{figure:biiiig-data}, the \emph{hardest} instance for the AIDS dataset
+requires 287 recursive calls, and 149,268 of the instances can be solved without search.  For PDBS,
+the hardest problem requires 3,161 recursive calls (which occurs 20 times, due to duplicated
+queries). For both PCM and PPI, the hardest problem requires 23 recursive calls.  In other words,
+none of these instances are in any way computationally hard even for a simple domain-based
+algorithm, even before introducing stronger filtering, inference, or heuristics.  Furthermore it is
+certainly not the case that a sequential scan is infeasible as so many filter~/ verify papers claim.
+The exponential behaviour seen by others in these instances is purely down to a bad choice of
+subgraph isomorphism algorithm, and these instances are not inherently ``really hard''.
+
+\subsection{Benefits of Constraint Programming over Filter / Verify}
+
+There are at least four benefits beyond simplicity towards abolishing filter~/ verify in favour of a
+purely constraint programming style approach:
 
@@
 
-\section{Further Lessons}\label{section:angryranting}
+\subsection{Other Implications}
 
@@
 subgraph-related research. We illustrate this by looking at several recent papers on subgraph
-isomorphism, showing how misleading conclusions are being reached due to domain-based algorithms not
-being used or understood.
-
-\subsection{Straggler Queries?}
+isomorphism, discussing how a deeper understanding of the empirical hardness of subgraph isomorphism
+could lead the reader to different conclusions than the ones presented.
 
@@
 
-\subsection{Portfolios}
-
-\citet{DBLP:conf/edbt/KatsarouNT17}'s second claim is that ``different algorithms find different
-queries hard''. To address this, they run many subgraph isomorphism algorithms and input
-permutations in parallel; the extensive literature on parallel portfolios in general
-\citep{DBLP:journals/ai/GomesS01}, and the approach by \citet{DBLP:conf/sls/BattitiM07} for subgraph
-isomorphism in particular, is not noted (and nor are portfolios mentioned when they say that ``using
-machine learning models to predict which version \ldots to employ per query is of high interest'').
-The evidence so far in this paper suggests that we should be wary. It is certainly true that there
-are some instances that all algorithms find hard, and there are good theoretical reasons to believe
-that these instances genuinely are really hard.  Furthermore, algorithm portfolios are also
-well-known to be a successful technique, and are beneficial even with modern subgraph isomorphism
-algorithms \citep{DBLP:conf/lion/KotthoffMS16}.  However, in \cref{section:labelled} we saw that there
-are many instances that VF2 finds hard that should not be hard, and that are not hard for other
-algorithms.  If permuting graphs sometimes addresses the difficulty of some of these instances, then
-it is likely that they are not genuinely hard instances at all, and perhaps it would be better
-simply to start using an algorithm with domain tracking and domain-based ordering heuristics, rather
-than assembling a portfolio of bad algorithms in the hopes that at least one of them will often get
-lucky.
+\citet{DBLP:conf/edbt/KatsarouNT17}'s second claim is that ``different
+algorithms find different queries hard''. To address this, they run many subgraph isomorphism
+algorithms and input permutations in parallel; the extensive literature on parallel portfolios in
+general \citep{DBLP:journals/ai/GomesS01}, and the approach by \citet{DBLP:conf/sls/BattitiM07} for
+subgraph isomorphism in particular, is not noted (and nor are portfolios mentioned when they say
+that ``using machine learning models to predict which version \ldots to employ per query is of high
+interest'').  The evidence so far in this paper suggests that we should be wary. It is certainly
+true that there are some instances that all algorithms find hard, and there are good theoretical
+reasons to believe that these instances genuinely are really hard.  Furthermore, algorithm
+portfolios are also well-known to be a successful technique, and are beneficial even with modern
+subgraph isomorphism algorithms \citep{DBLP:conf/lion/KotthoffMS16}.  However, in
+\cref{section:labelled} we saw that there are many instances that VF2 finds hard that should not be
+hard, and that are not hard for other algorithms.  If permuting graphs sometimes addresses the
+difficulty of some of these instances, then it is likely that they are not genuinely hard instances
+at all, and perhaps it would be better simply to start using an algorithm with domain tracking and
+domain-based ordering heuristics, rather than assembling a portfolio of bad algorithms in the hopes
+that at least one of them will often get lucky.
 
@@
 \citet{DBLP:conf/edbt/KatsarouNT17}'s technique is due to the ease of sometimes getting better
-results out of a poor algorithm, rather than being an inherent hardness property, and it would not
-help on really hard problem instances.
-
-\subsection{Search Ordering}
-
-We have seen, then, that domain tracking and domain-based ordering heuristics are critical, and we
-have seen that these alone are enough to make indexing and other techniques irrelevant. There
-is therefore no point in continued research such as that by
-\citet{DBLP:conf/gbrpr/CarlettiFV15} and \citet{o:Carletti16}
-into tweaking VF2's connectivity-based heuristics.
-
-?? NVMe \citep{DBLP:conf/wise/ShenZ17}
-
-?? In particular, we saw that VF3 \citep{DBLP:conf/gbrpr/CarlettiFSV17} performed \emph{worse} than
-VF2 on harder instances.
-
-\subsection{Benchmarking on Hard Instances}
-
-?? VF3 milliseconds
-
-?? \citet{DBLP:journals/tcbb/BonniciG17} compare various non-CP approaches to LAD. Benchmark sets
-whose difficulties are measured between microseconds and milliseconds
-
-\citet{DBLP:journals/heuristics/Gent98} describes the importance of hard benchmark instances:
+results out of a poor algorithm. Claiming huge improvements from an algorithm portfolio consisting
+only of variations of poor algorithms does not demonstrate a genuine improvement over the state of
+the art.
+
+Improved search orderings for VF2 also appear in recent work by
+\citet{DBLP:conf/gbrpr/CarlettiFV15} and \citet{o:Carletti16}, cumulating in VF3
+\citep{DBLP:conf/gbrpr/CarlettiFSV17}, and by \citet{DBLP:conf/wise/ShenZ17}. We have seen in
+\cref{section:labelled} that such an approach could give improvements, but will not bring an
+algorithm that does not use domains close to the performance of one that does on harder instances.
+Indeed, we saw in \cref{figure:induced} that VF3 performs \emph{worse} than VF2 on harder instances.
+The importance of using hard benchmark instances was stressed by
+\citet{DBLP:journals/heuristics/Gent98}:
 
@@
 
-?? Claiming huge improvements from an algorithm portfolio consisting only of variations of poor
-algorithms does not demonstrate a genuine improvement over the state of the art.
-
-?? Need for more hard instances
-
-?? Something about distance-based filtering being no use on our instances, low degree spread
+Unfortunately, this lesson has not sunk in. For example, \citet{DBLP:journals/tcbb/BonniciG17}
+recently compared various non-constraint-based approaches to LAD using benchmark sets whose
+difficulties are measured between microseconds and milliseconds. Their results suggest that LAD is
+an order of magnitude slower than other approaches. This should not surprise us: the initialisation
+costs of LAD are non-trivial due to it constructing domains, calculating neighbourhood degree
+sequences, and performing all-different filtering. However, the risks of occasionally encountering
+exponential performance from weaker algorithms on easy instances should now be clear.
+
+Of course, hard random instances of the kind we describe in
+\cref{section:non-induced,section:induced} are also atypical, and should not form the sole basis for
+benchmarking. In particular, distance-based filtering
+\citep{DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16} has little
+to no effect on these instances, but is extremely beneficial on other kinds of graph, including
+those from real-world applications.  These graphs also have a low degree spread and very little
+structure to exploit, which could mislead experimenters as to the value of certain kinds of
+inference. \citet{o:GentW95} give an example of an exam timetabling graph which contains an
+unexpected ten-vertex clique, which would be extremely unlikely in a randomly generated graph with a
+similar order and density.
 

### 2017-11-03 paper/paper.tex
Tweaky things

@@
 we see a solver-independent complexity peak near this phase transition. Additionally, understanding
-this behaviour helps us to select better variable- and value-ordering heuristics---this is the
+this behaviour helps us to design better variable- and value-ordering heuristics---this is the
 second contribution of this paper.
@@
 
-Part of this paper were presented at IJCAI \citep{DBLP:conf/ijcai/McCreeshPT16}; the first three
+Parts of this paper were presented at IJCAI \citep{DBLP:conf/ijcai/McCreeshPT16}; the first three
 sections have been expanded, whilst
@@
 
-The Glasgow and LAD solvers are constraint programming style algorithms (although they use dedicated
+The Glasgow and LAD solvers are constraint programming inspired algorithms (although they use dedicated
 data structures, and do not use a constraint programming toolkit). For each vertex in the pattern
@@
 reduce the first part of the $\langle Sol \rangle$ equation by as little as possible. When two or
-more domains are of equal size, LAD simply breaks ties lexicographically, whereas Glasgow will pick
+more domains are of equal size, LAD breaks ties lexicographically, whereas Glasgow will pick
 a variable corresponding to a pattern vertex of highest degree. This strategy was determined
@@
 
-In practice, this is unlikely to be a problem: most real-world instances are relatively sparse. In
-this situation, these experiments justify reusing the non-induced heuristics on induced problems.
+In practice, this is unlikely to be a problem: most of the real-world instances we have seen tend to
+be relatively sparse. In this situation, these experiments justify reusing the non-induced
+heuristics on induced problems.
 
@@
 The region in the parameter space where both pattern and target have medium density is far from a
-phase transition, but nevertheless contains instances that are hard for all three solvers. We would
+phase transition, but nevertheless contains instances that are hard for all four solvers. We would
 like to know whether this is due to a weakness in current solvers (perhaps our solvers cannot reason
@@
 and only if the problem is satisfiable \citep{o:Levi73}. We used this encoding with an
-implementation of \citet{DBLP:journals/cor/SegundoRJ11}'s bit-parallel maximum clique algorithm,
-modified to solve the decision problem rather than the optimisation problem.  Again, our results
-show that the instances in the central region remain hard, and additionally, some of the easy
-unsatisfiable instances become hard.
+implementation of \citet{DBLP:journals/cor/SegundoRJ11}'s bit-parallel maximum clique algorithm
+BBMC, modified to solve the decision problem rather than the optimisation problem.  Again, our
+results show that the instances in the central region remain hard, and additionally, some of the
+easy unsatisfiable instances become hard.
 
@@
 phase transition concept, and to generalise hardness parameters across different combinatorial
-problems \citep{DBLP:conf/aaai/GentMPW96}. A problem with $\kappa < 1$ is said to be underconstrained, and is
-likely to be satisfiable; a problem with $\kappa > 1$ is overconstrained, and is likely to be
-unsatisfiable. Empirically, problems with $\kappa$ close to 1 are hard, and problems where $\kappa$
-is very small or very large are usually easy. By handling injectivity as a restriction on the size
-of the state space rather than as a constraint, we derive
-\begin{equation}\label{equation:constrainedness}\kappa = 1 - \frac{\log \left(
-t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 - d_{t})}^{(1 - d_{p}) \cdot
-\binom{p}{2}} \right)}{\log t^{\underline{p}}} \end{equation} for induced isomorphisms, which we plot on the
-sixth row of \cref{figure:induced}. We see that constrainedness predicts that the central region
-will still be relatively difficult for larger pattern graphs: although the problem is
-overconstrained, it is less overconstrained than in the regions the Glasgow and LAD solvers found
-easy.  Thus it seems that rather than just being a unification of existing generalised heuristic
-techniques, constrainedness also gives a better predictor of difficulty than proximity to a phase
+problems \citep{DBLP:conf/aaai/GentMPW96}. A problem with $\kappa < 1$ is said to be
+underconstrained, and is likely to be satisfiable; a problem with $\kappa > 1$ is overconstrained,
+and is likely to be unsatisfiable. Empirically, problems with $\kappa$ close to 1 are hard, and
+problems where $\kappa$ is very small or very large are usually easy. By handling injectivity as a
+restriction on the size of the state space rather than as a constraint, we derive
+\begin{equation}\label{equation:constrainedness}\kappa = 1 - \frac{\log \left( t^{\underline{p}}
+\cdot {d_t}^{d_p \cdot \binom{p}{2}} \cdot {(1 - d_{t})}^{(1 - d_{p}) \cdot \binom{p}{2}}
+\right)}{\log t^{\underline{p}}} \end{equation} for induced isomorphisms, which we plot on the sixth
+row of \cref{figure:induced}. We see that constrainedness predicts that the central region will
+still be relatively difficult for larger pattern graphs: although the problem is overconstrained, it
+is less overconstrained than in the regions the Glasgow and LAD solvers found easy.  Thus it seems
+that rather than just being a unification of existing generalised heuristic techniques,
+constrainedness also gives a better predictor of difficulty than proximity to a phase
 transition---our method generates instances where constrainedness and ``close to a phase
-transition'' give very different predictions, and constrainedness gives the better prediction.
-
-Unfortunately, constrainedness does not help us with heuristics: minimising constrainedness gives
-the same predictions as maximising the expected number of solutions.
+transition'' give very different predictions, and constrainedness closely matches the empirical
+results.
+
+Unfortunately, constrainedness does not give additional heuristic information: minimising
+constrainedness gives the same predictions as maximising the expected number of solutions.
 
@@
 
-What about richer graph structures? We briefly look at what happens with the non-induced problem if
-we use $k$-regular graphs rather than the Erd\H{o}s-R\'enyi model, and then in the following section
-we see the effects of introducing vertex labels. Both models still exhibit phase transition
-behaviour, although with sufficiently many labels the ``satisfiable'' region now contains a mix of
-satisfiable and unsatisfiable instances.
+What about richer graph structures? In this section we briefly look at what happens with the
+non-induced problem if we use $k$-regular graphs rather than the Erd\H{o}s-R\'enyi model, and then
+in the following section we see the effects of introducing vertex labels. Both models still exhibit
+phase transition behaviour, although with sufficiently many labels the ``satisfiable'' region now
+contains a mix of satisfiable and unsatisfiable instances.
 
@@
 
-To illustrate this point further, we will now look at some instances created using a slightly more
+To illustrate this point further, we now look at some instances created using a slightly more
 structured random model. We create a family of one thousand labelled instance pairs, as follows. To
@@
 counting mechanism (for example, by attaching an additional vertex with a unique label to the
-clique).
+clique). Note also the resemblance to the cases seen in the bottom left of the VF2 plots in
+\cref{figure:non-induced}, where a pattern graph which has a few edges combined with a target graph
+with no edges gives a timeout.
 
@@
 incomplete overview. Our description is biased by a general modern understanding of the empirical
-hardness of \NP-complete problems, which was not widely known at least at the time of the earlier
+hardness of \NP-complete problems, which was not widely known at the time of the earlier
 papers we discuss. The common theme of all of the following papers is that pre-computed information
@@
 will be very long due to the NP-hard complexity.''\end{displayquote} They suggest that indexing is
-a way of avoiding this.
+a way of avoiding this. Similarly, \citet{DBLP:journals/pvldb/ZhaoH10} argue that ``the graph query
+problem is hard in that \ldots it requires subgraph isomorphism checking \ldots which has proven to
+be NP-complete'' and that working with large networks is hard or impossible ``due to the lack of
+scalable graph indexing mechanisms''.
 
@@
 querying is costly because it is \NP-complete, and that indices can improve the performance of graph
-database queries. Again, new indexing techniques are introduced. Subsequently, \citet{DBLP:journals/pvldb/KatsarouNT15} perform a comprehensive comparison of graph database
+database queries. Again, new indexing techniques are introduced. Subsequently,
+\citet{DBLP:journals/pvldb/KatsarouNT15} perform a comprehensive comparison of graph database
 filtering techniques. They state that performing a query against each graph in the dataset
@@
     \item The PDBS dataset \citep{o:HeLCBBSKMR02} contains 30 labelled graphs representing DNA, RNA,
-        and proteins, each duplicated twenty times. These
+        and proteins, each duplicated twenty times to give a dataset of 600 graphs. These
         can be relatively large, having up to tens of thousands of vertices, but are extremely
@@
     \item The PCM dataset \citep{o:doi:10.1093/bioinformatics/btr163} contains 50 protein contact
-        maps, each duplicated four times. These graphs have under a thousand vertices and below
-        twenty thousand edges; they are slightly less sparse than the previous two datasets. As for
-        PDBS, the queries are randomly generated and do not have a particular meaning.
+        maps, each duplicated four times to give a dataset of 200 graphs. These graphs have under a
+        thousand vertices and below twenty thousand edges; they are slightly less sparse than the
+        previous two datasets. As for PDBS, the queries are randomly generated and do not have a
+        particular meaning.
     \item The PPI dataset contains 20 protein interaction networks, with up to ten thousand
@@
 There are at least four benefits beyond simplicity towards abolishing filter~/ verify in favour of a
-purely constraint programming style approach:
+purely constraint programming inspired approach:
 
@@
 
-Improved search orderings for VF2 also appear in recent work by
+Improved search orderings for VF2-style algorithms also appear in recent work by
 \citet{DBLP:conf/gbrpr/CarlettiFV15} and \citet{o:Carletti16}, cumulating in VF3
-\citep{DBLP:conf/gbrpr/CarlettiFSV17}, and by \citet{DBLP:conf/wise/ShenZ17}. We have seen in
-\cref{section:labelled} that such an approach could give improvements, but will not bring an
-algorithm that does not use domains close to the performance of one that does on harder instances.
-Indeed, we saw in \cref{figure:induced} that VF3 performs \emph{worse} than VF2 on harder instances.
+\citep{DBLP:conf/gbrpr/CarlettiFSV17,o:CarlettiFSV17}, and by \citet{DBLP:conf/wise/ShenZ17}. We
+have seen in \cref{section:labelled} that such an approach could give improvements, but will not
+bring an algorithm that does not use domains close to the performance of one that does on harder
+instances.
+
 The importance of using hard benchmark instances was stressed by
@@
 exponential performance from weaker algorithms on easy instances should now be clear.
+
+\citet{o:CarlettiFSV17} correctly observe that ``a performance improvement on [hard instances] has a
+far greater practical impact than it would have on easier graphs, since the saved time can be of
+several hours or even days'', but mistakenly believe that being larger and slightly denser is what
+makes an instance harder. To evaluate VF3, they introduce a new dataset consisting entirely of
+satisfiable instances created by taking a connected subgraph of a randomly generated graph and
+permuting the nodes. They conclude that VF3 is the best algorithm based upon its performance solving
+the enumeration problem on these new instances.  We urge a more cautious interpretation of these
+results: as we saw in \cref{section:induced}, being genuinely hard (rather than simply larger) is a
+subtle property, and more care must be taken to avoid accidentally measuring which algorithm scales
+best to large, easy instances due to it using the simplest inference routines. Indeed, we saw in
+\cref{figure:induced} that VF3 performs \emph{worse} than VF2 on genuinely harder instances, and an
+induced analogue of \cref{figure:labels} confirms that VF3 still struggles with many obviously
+unsatisfiable instances.
 
@@
 respecting our understanding of the empirical hardness of \NP-complete problems, that subgraph
-isomorphism algorithms must not be treated as a black box, and that lessons learned in constraint
-programming and artificial intelligence must not go unheeded in other domains.
+isomorphism algorithms should not be treated as a black box, and that lessons learned in constraint
+programming and artificial intelligence should not go unheeded in other domains.
 

### 2017-11-03 paper/paper.tex
VF3 a bit differently

@@
 neither Glasgow nor LAD were designed with the induced version in mind), whilst VF3 supports only
-the induced problem. Note also that VF3 will sometimes crash on especially sparse graphs: we have
-treated these instances as timeouts.
+the induced problem.
 
@@
 For larger patterns of order 20 and 30, we have a large unsatisfiable region in the middle. Despite
-not being near either phase transition, instances in the centre remain computationally challenging.
-We also plot patterns of orders 14, 15 and 16, to show the change between the two behaviours.
+not being near either phase transition, instances in the centre remain computationally challenging
+for every solver.  We also plot patterns of orders 14, 15 and 16, to show the change between the two
+behaviours.
 
@@
 between the two, which we discuss below). For LAD, some of the very dense patterns are slightly
-harder than their diagonal opposites (LAD reasons about degrees, but not about complement-degrees);
-it is interesting to note that for larger target graphs, VF2 finds \emph{all} dense pattern graphs
-difficult. Finally, VF3's performance generally improves upon VF2's on instances that VF2 finds easy
-(except in the cases where VF3 crashes), but worsens on instances that VF2 finds hard.
+harder than their diagonal opposites (LAD reasons about degrees, but not about complement-degrees).
+
+It is interesting to note that for larger target graphs, VF2 finds \emph{all} dense pattern graphs
+difficult. Meanwhile, VF3 behaves better than VF2 in some regions, but worse than VF2 in others, and
+has a larger hard region than either domain-based algorithm.\footnote{The version of
+VF3 supplied by its inventors will crash if given an empty pattern graph. We have chosen to treat
+these results as successes taking zero search, even though empty pattern graphs are not trivial for
+the induced problem. If the reader prefers to interpret a crash as a timeout, the VF3 plots would
+contain an additional area of black points down the left-hand side.} VF3 also occasionally finds
+some instances with sparse target graphs (along the bottom of the plot) extremely difficult, whilst
+other solvers do not.
 
@@
 results: as we saw in \cref{section:induced}, being genuinely hard (rather than simply larger) is a
-subtle property, and more care must be taken to avoid accidentally measuring which algorithm scales
-best to large, easy instances due to it using the simplest inference routines. Indeed, we saw in
-\cref{figure:induced} that VF3 performs \emph{worse} than VF2 on genuinely harder instances, and an
-induced analogue of \cref{figure:labels} confirms that VF3 still struggles with many obviously
-unsatisfiable instances.
+subtle property, and it is not safe to say that an instance is hard simply because several
+minor variations of a particular kind of algorithm find it hard. More care must thus be taken to
+avoid accidentally measuring which algorithm scales best to large, easy instances due to it using
+the simplest inference routines. Indeed, we saw in \cref{figure:induced} that VF3 performs
+\emph{worse} than VF2 on genuinely harder instances, and additionally behaves erratically in parts
+of the parameter space that every other solver finds easy; finally, an induced analogue of
+\cref{figure:labels} confirms that VF3 also still struggles with many obviously unsatisfiable
+instances.
 

### 2017-11-07 biiiiiig-data/gecode/sip.cc
More experimenty things

@@ -92,7 +92,7 @@ class SIP : public Space
     public:
         IntVarArray mapping;
 
-        SIP(const Graph & pattern, const Graph & target, int maximum_label) :
+        SIP(const Graph & pattern, const Graph & target, int maximum_label, bool sdf) :
             mapping(*this, pattern.size, 0, target.size - 1)
         {
             // injectivity
@@ -154,7 +154,7 @@ class SIP : public Space
             for (int v = 0 ; v < pattern.size ; ++v)
                 mapping_permuted[v] = mapping[mapping_permutation[v]];
 
-            branch(*this, mapping_permuted, INT_VAR_SIZE_MIN(), INT_VAL_MIN());
+            branch(*this, mapping_permuted, sdf ? INT_VAR_SIZE_MIN() : INT_VAR_NONE(), INT_VAL_MIN());
         }
 
         SIP(bool share, SIP & s) :
@@ -171,8 +171,8 @@ class SIP : public Space
 
 int main(int argc, char * argv[])
 {
-    if (3 != argc) {
-        cerr << "Usage: " << argv[0] << " pattern-file target-file" << endl;
+    if (4 != argc) {
+        cerr << "Usage: " << argv[0] << " pattern-file target-file sdf" << endl;
         return EXIT_FAILURE;
     }
 
@@ -185,7 +185,7 @@ int main(int argc, char * argv[])
         for (const auto & target : targets) {
             cout << pattern.name << " -> " << target.name << " " << flush;
 
-            DFS<SIP> e(make_unique<SIP>(pattern, target, labels_map.size() + 1).get());
+            DFS<SIP> e(make_unique<SIP>(pattern, target, labels_map.size() + 1, argv[3] == string("true")).get());
             if (unique_ptr<SIP> s{ e.next() })
                 cout << e.statistics().node << " sat " << s->mapping << endl;
             else

### 2017-11-07 biiiiiig-data/vflib/solve_vf.cc

@@ -1,0 +1,178 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "argraph.h"
+#include "argedit.h"
+#include "argloader.h"
+#include "match.h"
+#include "vf2_mono_state.h"
+
+#include <fstream>
+#include <iostream>
+#include <thread>
+#include <condition_variable>
+#include <mutex>
+#include <atomic>
+#include <cstdlib>
+#include <chrono>
+#include <vector>
+
+#include <boost/regex.hpp>
+
+using std::chrono::milliseconds;
+using std::chrono::steady_clock;
+using std::chrono::duration_cast;
+
+struct GraphInfo
+{
+    string name;
+    int size;
+    vector<int> label;
+    vector<vector<int> > adj;
+};
+
+auto load(const string & filename, vector<GraphInfo> & graphs, map<string, int> & labels_map) -> bool
+{
+    ifstream file{ filename };
+    if (! file) {
+        cerr << "Error opening " << filename << endl;
+        return false;
+    }
+
+    string name;
+    while (file >> name) {
+        if (0 != name.compare(0, 1, "#", 0, 1)) {
+            cerr << "Bad name '" << name << "'" << endl;
+            return false;
+        }
+        name.erase(0, 1);
+
+        int size;
+        file >> size;
+
+        graphs.push_back(GraphInfo{ });
+        auto & g = graphs.back();
+        g.name = name;
+        g.size = size;
+        g.label.resize(size);
+        g.adj.resize(size, vector<int>(size, 0));
+
+        for (int l = 0 ; l < size ; ++l) {
+            string label;
+            file >> label;
+            if (! labels_map.count(label))
+                labels_map.emplace(label, labels_map.size() + 1);
+
+            g.label[l] = labels_map.find(label)->second;
+        }
+
+        int edge_count;
+        file >> edge_count;
+        for (int e = 0 ; e < edge_count ; ++e) {
+            int f, t;
+            file >> f >> t;
+            g.adj[f][t] = g.adj[t][f] = 1;
+        }
+    }
+
+    cerr << filename << " " << graphs.size() << endl;
+
+    return true;
+}
+
+namespace
+{
+    unsigned long long hacky_global_nodes = 0;
+
+    struct OurComparator : AttrComparator
+    {
+        virtual bool compatible(void *attr1, void *attr2)
+        {
+            return *static_cast<int *>(attr1) == *static_cast<int *>(attr2);
+        }
[…]

### 2017-11-07 analysis/search_nodes_heatmap.cc
Handle VF3 crashes

@@ -26,7 +26,7 @@ auto add_results(
         const string & dir, const string & alg,
         int ps, int ts, int pd, int td, int r,
         const string & pd_p, const string & td_p, const string & ed,
-        int & failures, unsigned long long & total_nodes) -> void
+        int & failures, unsigned long long & total_nodes) -> bool
 {
     string pd_s = (pd_p == "pkr" ? to_string(pd) : str(format("%|1|.%|02|") % (pd / 100) % (pd % 100)));
     string td_s = (td_p == "tkr" ? to_string(td) : str(format("%|1|.%|02|") % (td / 100) % (td % 100)));
@@ -37,21 +37,23 @@ auto add_results(
     if (! infile) {
         cerr << "missing " << filename << endl;
         ++failures;
-        return;
+        return false;
     }
 
     string success, aborted;
     unsigned long long this_nodes;
 
     if (string::npos != alg.find("vf3")) {
-        infile >> success >> this_nodes;
+        infile >> success;
         if (! infile) {
             cerr << "reading " << filename << endl;
             ++failures;
-            return;
+            return false;
         }
         if (success == "aborted")
             aborted = success;
+        else
+            infile >> this_nodes;
     }
     else if (string::npos != alg.find("vf2")) {
         infile >> success;
@@ -63,7 +65,7 @@ auto add_results(
         if (! infile) {
             cerr << "reading " << filename << endl;
             ++failures;
-            return;
+            return false;
         }
     }
     else if (string::npos != alg.find("lad")) {
@@ -75,7 +77,7 @@ auto add_results(
         if (! infile) {
             cerr << "reading " << filename << endl;
             ++failures;
-            return;
+            return false;
         }
 
         if (string::npos != line.find("CPU time exceeded"))
@@ -109,7 +111,7 @@ auto add_results(
         if (! (found_status && found_nodes)) {
             cerr << "reading " << filename << endl;
             ++failures;
-            return;
+            return false;
         }
     }
     else if (string::npos != alg.find("glucose")) {
@@ -134,7 +136,7 @@ auto add_results(
         if (! (found_status && found_nodes)) {
             cerr << "reading " << filename << endl;
             ++failures;
-            return;
+            return false;
         }
     }
     else if (string::npos != alg.find("gurobi")) {
@@ -165,7 +167,7 @@ auto add_results(
         if (! (found_status && found_nodes)) {
[…]

### 2017-11-07 paper/paper.tex
Iva
Some of Christine's comments

@@
     These results have practical consequences: we explain why the widely researched ``filter~/
-    verify'' indexing technique used in graph databases is a flawed design based upon a repeated
-    misunderstanding of the empirical hardness of \NP-complete problems, that cannot be beneficial
+    verify'' indexing technique used in graph databases is a flawed design (based upon a repeated
+    misunderstanding of the empirical hardness of \NP-complete problems) that cannot be beneficial
     when paired with any reasonable subgraph isomorphism algorithm.
@@
 performed using randomly generated graph pairs, sometimes together with a mix of real-world graphs,
-and graphs that encode biochemistry and computer vision problems. Using random instances to evaluate
+for example biochemistry and computer vision problems. Using random instances to evaluate
 algorithm behaviour can be beneficial, because it provides a way of generating many instances
@@
 Throughout, our graphs are undirected, and do not have any loops. In some cases vertices have
-labels, which are treated as integers; we write $\ell(v)$ for the label of vertex $v$. The
-\emph{order} of a graph is the cardinality of its vertex set. We write $\operatorname{V}(G)$ for the
-vertex set of a graph $G$. The \emph{complement} of a graph $G$, denoted $\shortoverline{G}$, is the
-graph with the same vertex set as $G$, and with an edge between distinct vertices $v$ and $w$ if and
-only if $v$ and $w$ are not adjacent in $G$. We write $G(n, p)$ for an Erd\H{o}s-R\'enyi random
-graph with $n$ vertices, and an edge between each distinct unordered pair of vertices with
-independent probability $p$.
+labels, which are treated as integers; we write $\ell(v)$ for the label of vertex $v$. Two vertices
+are \emph{adjacent} if there is an edge between them. The \emph{order} of a graph is the cardinality
+of its vertex set. We write $\operatorname{V}(G)$ for the vertex set of a graph $G$. The
+\emph{complement} of a graph $G$, denoted $\shortoverline{G}$, is the graph with the same vertex set
+as $G$, and with an edge between distinct vertices $v$ and $w$ if and only if $v$ and $w$ are not
+adjacent in $G$. We write $G(n, p)$ for an Erd\H{o}s-R\'enyi random graph with $n$ vertices, and an
+edge between each distinct unordered pair of vertices with independent probability $p$.
 
@@
 $T$ (called the \emph{target}) is an injective mapping $i$ from $\operatorname{V}(P)$ to
-$\operatorname{V}(T)$ which preserves adjacency---that is, for every adjacent $v$ and $w$ in
+$\operatorname{V}(T)$ which preserves adjacency---that is, for every adjacent vertices $v$ and $w$ in
 $\operatorname{V}(P)$, the vertices $i(v)$ and $i(w)$ are adjacent in $T$. Where labels are present,
@@
 \citep{DBLP:journals/ai/Solnon10}; VF2 \citep{DBLP:journals/pami/CordellaFSV04}; and VF3
-\citep{DBLP:conf/gbrpr/CarlettiFSV17}. Each was compiled using GCC 4.9.
+\citep{DBLP:conf/gbrpr/CarlettiFSV17}. Each was compiled using GCC 4.9 with the ``-O3'' option.
 
@@
 
+??
+
+\subsection{A Phase Transitions when Varying Pattern Edge Density}
+
 \begin{figure}[t]
@@
     probability of 0.40, and varying pattern edge probability, we observe a phase transition and
-        complexity peak with the Glasgow solver in the non-induced variant. Each point represents
-        one instance, and the line shows mean search effort.}
+    complexity peak with the Glasgow solver in the non-induced variant. Each point $(x, y)$
+    represents one instance $i$, where $x$ is the pattern edge probability used to generate $i$, $y$
+    is the number of search nodes needed by the Glasgow solver to solve $i$, and the point is drawn
+    as a green circle if $i$ is satisfiable, and a blue cross otherwise. The black line plots the
+    evolution of the mean number of search nodes when increasing the pattern edge probability from
+    $0$ to $1$.}
     \label{figure:phase-transition}
@@
 instances, and there are exceptionally hard satisfiable instances \citep{DBLP:conf/cp/SmithG97}.
+
+\subsection{Phase Transitions when Varying Pattern and Target Edge Densities}
 
@@
         bands show the phase transitions, and the black lines are the predictions using
-        equation~\eqref{equation:non-induced-prediction} of where the
-        phase transition will occur. On the subsequent three rows, we show the number of search nodes used by the
-        Glasgow, LAD and VF2 solvers; the dark regions indicate ``really hard'' instances, and black
+        equation~\eqref{equation:non-induced-prediction} of where the phase transition will occur.
+        On the subsequent three rows, we show the average number of search nodes used by the
+        Glasgow, LAD and VF2 solvers over ten instances; the dark regions indicate ``really hard''
+        instances, and black
         points indicate that at least one timeout occurred.}
@@
 parameters. Since we are trying to find an \emph{injective} mapping from a pattern $P = G(p, d_p)$
-to a target $T = G(t, d_t)$, there are \[ t^{\underline{p}} = t \cdot (t - 1) \cdot \ldots \cdot (t -
-p + 1) \] possible assignments of target vertices to pattern vertices.  If we assume the pattern
-has exactly $d_p \cdot \binom{p}{2}$ edges, we obtain the probability of all of these edges being mapped to
-edges in the target by raising $d_t$ to this power, giving an expected number of solutions of
-\begin{equation}\label{equation:non-induced-prediction}
-    \langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \textnormal{.}
-\end{equation} This
-formula predicts a very sharp phase transition from $\langle Sol \rangle \ll 1$ to $\langle Sol
-\rangle \gg 1$, which may easily be located numerically. We plot where this occurs using black lines
-in the first row of \cref{figure:non-induced}.
+to a target $T = G(t, d_t)$, there are \[ t^{\underline{p}} = t \cdot (t - 1) \cdot \ldots \cdot (t
+- p + 1) \] possible injective assignments of target vertices to pattern vertices.  If we assume the
+pattern has exactly $d_p \cdot \binom{p}{2}$ edges, we obtain the probability of all of these edges
+being mapped to edges in the target by raising $d_t$ to this power, giving an expected number of
+solutions of \begin{equation}\label{equation:non-induced-prediction} \langle Sol \rangle =
+t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}} \textnormal{.} \end{equation} This formula
+predicts a very sharp phase transition from $\langle Sol \rangle \ll 1$ to $\langle Sol \rangle \gg
+1$, which may easily be located numerically. We plot where this occurs using black lines in the
+first row of \cref{figure:non-induced}.
 
@@
 and Glasgow pick a variable with fewest remaining values in its domain: doing this will generally
-reduce the first part of the $\langle Sol \rangle$ equation by as little as possible. When two or
-more domains are of equal size, LAD breaks ties lexicographically, whereas Glasgow will pick
-a variable corresponding to a pattern vertex of highest degree. This strategy was determined
+reduce the first part of the $\langle Sol \rangle$ equation (i.e.\ $t^p$) by as little as possible.
+When two or more domains are of equal size, LAD breaks ties lexicographically, whereas Glasgow will
+pick a variable corresponding to a pattern vertex of highest degree. This strategy was determined
 empirically, but could have been derived from the $\langle Sol \rangle$ formula: picking a pattern
@@
 exponent in the second half of the formula, maximising the overall value. LAD does not apply a
-value-ordering heuristic, but Glasgow does: it prefers target vertices of highest degree.  Again, this was
-determined empirically, but it has the effect of increasing $\langle Sol \rangle$ by leaving as many
-vertices as possible available for future use. The search strategy used by VF2, in contrast, is
-based around preserving connectivity, which gives very little discrimination except on the sparsest
-of inputs.
+value-ordering heuristic, but Glasgow does: it prefers target vertices of highest degree.  Again,
+this was determined empirically, but it has the effect of increasing $\langle Sol \rangle$ by
+leaving as many vertices as possible available for future use. The search strategy used by VF2, in
+contrast, is based around preserving connectivity, which gives very little discrimination except on
+the sparsest of inputs.
 
@@
     when the Glasgow algorithm performs better when given the complements of the pattern and target
-    graphs as inputs---the solid lines show the location of the phase transition, and the dotted
-    lines are $t_d=0.5$ and the $p_d=t_d$ diagonal.}\label{figure:induced}
+    graphs as inputs---the solid lines show the empirical location of the phase transition, and the
+    dotted lines are $t_d=0.5$ and the $p_d=t_d$ diagonal.}\label{figure:induced}
 \end{figure}
@@
 bottom right halves of the plots resemble the non-induced results, and the top left halves are close
-to a mirror image. The central satisfiable region, which is away from either phase transition, is
-computationally easy, but instances near the phase transition are hard.
+to a mirror image. This second phase transition comes from the fact that an empty pattern is not an
+induced subgraph of a complete target: an empty pattern of order $k$ is an induced subgraph only if
+there exist $k$ vertices in the target that are not pairwise connected.  The central satisfiable
+region, which is away from either phase transition, is computationally easy, but instances near the
+phase transition are hard.
 
@@
     pattern degree (x-axis) and target degree (y-axis) are varied from 0 to
-    $\left|\operatorname{V}\right| - 1$, using target graphs of 150 vertices.}
+    $\left|\operatorname{V}(G)\right| - 1$, using target graphs of 150 vertices, and pattern graphs
+    of 10, 20, and 30 vertices.}
     \label{figure:kr}
@@
 
-By $R(n, k)$ we mean a random graph with $k$ vertices, each of which has degree $k$. In
+By $R(n, k)$ we mean a random graph with $n$ vertices, each of which has degree $k$. In
 \cref{figure:kr} we recreate \cref{figure:non-induced}, using the NetworkX implementation
 \citep{o:NetworkX} of the \citet{DBLP:journals/cpc/StegerW99} algorithm to generate regular graphs.
-The $x$-axes range from degree 0 to degree $n - 1$, and the y-axes range from degree 0 to degree
-149.  The satisfiable / unsatisfiable plots show very similar results to the non-induced problem on
-Erd\H{o}s-R\'enyi random graphs, with a similarly sharp phase transition, and
-again only instances near the phase transition are difficult.
+The $x$-axes range from degree 0 to degree $n - 1$ where $n \in \{ 10, 20, 30 \}$ is the number of
+pattern vertices, and the y-axes range from degree 0 to degree
+149 (one less than the number of target vertices).  The satisfiable / unsatisfiable plots show very
+similar results to the non-induced problem on Erd\H{o}s-R\'enyi random graphs, with a similarly
+sharp phase transition, and again only instances near the phase transition are difficult.
 
 Regularity means that degree-based heuristics have no information to work with. Additionally, the
-degree-based filtering techniques used by Glasgow and LAD have no effect on these graphs, and so at
-the top of search, every domain is identical. However, once a guessed assignment has been made,
-selecting from small domains first remains an effective strategy to guide the remainder of the
+degree-based filtering techniques used by Glasgow and LAD have no effect on these graphs at the top
+of search, and so initially every domain is identical. However, once a guessed assignment has been
+made, selecting from small domains first remains an effective strategy to guide the remainder of the
 search.
@@
 Suppose our labels are drawn randomly from a set $L = \{ 1, \ldots, k \}$, where $k$ is reasonably
-small compared to $p$. Recall that $\ell(v)$ is the label of vertex $v$. By defining \[
-    \operatorname{V}(P)\vert_n = \{ v \in \operatorname{V}(P) : \ell(v) = n \} \] to be the set of
-vertices with label $n$, we can partition the pattern vertices by label into disjoint sets
-$\{\operatorname{V}(P)\vert_1, \ldots, \operatorname{V}(P)\vert_k \}\textnormal{\,,}$ each of
+small compared to the number of pattern vertices $p$. Recall that $\ell(v)$ is the label of vertex
+$v$. By defining \[ \operatorname{V}(P)\vert_n = \{ v \in \operatorname{V}(P) : \ell(v) = x \} \] to
+be the set of vertices with label $x$, we can partition the pattern vertices by label into disjoint
+sets $\{\operatorname{V}(P)\vert_1, \ldots, \operatorname{V}(P)\vert_k \}\textnormal{\,,}$ each of
 which is expected to contain $\nicefrac{p}{k}$ vertices. Similarly, we may partition the target
@@
 \textnormal{,} \] but to do this we must state what $a^{\underline{b}}$ means when $b$ is
-fractional. The gamma function $\Gamma(n)$ is equal to $(n - 1)!$ for integers $n \ge 1$, but is also
-defined for positive real numbers, obeying the identify $\Gamma(x + 1) = x\Gamma(x)$. By noting that
-$t^{\underline{p}} = \nicefrac{t!}{(t - p)!}$, we may obtain a reasonable
-continuous extension by taking \[ |S| = \left(\frac{\Gamma\left(\nicefrac{t}{k} +
+fractional. The gamma function $\Gamma(n)$ is equal to $(n - 1)!$ for integers $n \ge 1$
+\citep{Davis59}, but is also defined for positive real numbers, obeying the identify $\Gamma(x + 1)
+= x\Gamma(x)$. By noting that $t^{\underline{p}} = \nicefrac{t!}{(t - p)!}$, we may obtain a
+reasonable continuous extension by taking \[ |S| = \left(\frac{\Gamma\left(\nicefrac{t}{k} +
 1\right)}{\Gamma\left(\nicefrac{t}{k} - \nicefrac{p}{k} + 1\right)}\right)^{k} \textnormal{\,.} \]
@@
 of labels, our predictions are slightly better than in the unlabelled case: there seems to be less
-of a variance problem for very dense patterns. Even as the numbers of labels becomes relatively
+of a variance problem for very dense patterns. Even as the number of labels becomes relatively
 large, the prediction of the phase transition still occurs in the right place, but with ten labels
@@
 the non-zero labelled vertices are given lower vertex numbers rather than higher vertex numbers.
+In other words, we permute pattern vertices in such a way that the first ten vertices have non-zero
+labels, and the ten last vertices have zero labels (and similarly for the target graphs).
 This makes VF2 behave better, but still extremely poorly compared to the other two solvers; the
@@
 (so it is branched on first), VF2 takes nearly a million recursive calls to detect this: VF2 will
-try to map every adjacent vertex in the clique before considering the difficult isolated vertex.
+try to map every adjacent vertex in the clique before considering the unmatchable isolated vertex.
 This example can be extended slightly to fool any simple static heuristic, or any simple label
@@
 subgraph isomorphism test. For example, an index might contain a bit of information expressing
-whether a target graph contains two red vertices. When a pattern graph with two red vertices is used
-as a query, any target whose feature set does not have this bit set would not be considered, and so
-a subgraph isomorphism call would not be made for that pattern / target pair. (In practice, feature
-hashing is often used, which can lead to false positives, but this is not relevant to our
-discussion.)
+whether a target graph contains at least two red vertices. When a pattern graph with two red
+vertices is used as a query, any target whose feature set does not have this bit set would not be
+considered, and so a subgraph isomorphism call would not be made for that pattern / target pair. (In
+practice, feature hashing is often used, which can lead to false positives, but this is not relevant
+to our discussion.)
 
@@
 reason that since isomorphism testing is \NP-complete, by making $\left|C_q\right|$ as small as
-possible, the query response time will be reduced.  (Importantly, it is \emph{not} time taken to
+possible, the query response time will be reduced.  (Importantly, it is \emph{not} the time taken to
 load graphs from disk which contributes to the per-candidate cost, but rather the time to perform a
@@
 all-different propagator \citep{DBLP:conf/aaai/Regin94}.  However, as
-\citet{DBLP:journals/pvldb/KatsarouNT15} note, some variation on VF2
-\citep{DBLP:journals/pami/CordellaFSV04} is the usual subgraph isomorphism algorithm of choice for
-graph database systems, although \citeauthor{DBLP:journals/jacm/Ullmann76}'s algorithm is sometimes
-chosen. Other approaches have been considered, albeit not with algorithms strong enough to establish
-all-difference. For example, \citet{DBLP:journals/pvldb/ShangZLY08} propose an algorithm which makes
-use of the frequency of various features to guide search; \citet{DBLP:journals/pvldb/LeeHKL12}
-determine experimentally that this technique tends to be very effective, even on families of graphs
-for which it was not designed.
+\citet{DBLP:journals/pvldb/KatsarouNT15} note, VF2 \citep{DBLP:journals/pami/CordellaFSV04} or a
+similar algorithm is the usual subgraph isomorphism algorithm of choice for graph database systems,
+although \citeauthor{DBLP:journals/jacm/Ullmann76}'s algorithm is sometimes chosen. Other approaches
+have been considered, albeit not with algorithms strong enough to establish all-difference. For
+example, \citet{DBLP:journals/pvldb/ShangZLY08} propose an algorithm which makes use of the
+frequency of various features to guide search; \citet{DBLP:journals/pvldb/LeeHKL12} determine
+experimentally that this technique tends to be very effective, even on families of graphs for which
+it was not designed.
 
@@
 
-\acks{The authors would like to thank Kitty Meeks and Craig Reilly for their comments. The heatmap
-schemes used in this paper were created with the assistance of the ``Chroma.js Color Scale Helper''
-by Gregor Aisch.}
+\acks{The authors would like to thank Kitty Meeks and Craig Reilly for their comments, and Iva
+Babukova for her help with the datasets. The heatmap schemes used in this paper were created with
+the assistance of the ``Chroma.js Color Scale Helper'' by Gregor Aisch.}
 

### 2017-11-07 biiiiiig-data/vflib/solve_vf.cc
Allow VF2 to fail harder

@@ -100,7 +100,7 @@ namespace
         using VF2MonoState::VF2MonoState;
 
         virtual void BackTrack() {
-            if (++hacky_global_nodes > 1000000)
+            if (++hacky_global_nodes > 10000000)
                 throw TimeForAnAbortion();
 
             VF2MonoState::BackTrack();

### 2017-11-07 paper/paper.tex
More big data

@@
 that these are high-quality datasets or that the associated queries are sensible, merely that
-following the example of \citet{o:10.1371/journal.pone.0076911}, they are widely used.
+following the example of \citet{o:10.1371/journal.pone.0076911}, they are widely used. Each dataset
+has labels on vertices (but not on edges), and we will look at the non-induced problem.
 
@@
     \caption{Cumulative number of instances solved as a function of search space size, using four
-    graph database datasets and a simple constraint programming subgraph isomorphism algorithm. Note
-    that the $x$-axis shows recursive calls, and does \emph{not} use a log scale.}
+    graph database datasets, and either VF2 or a simple constraint programming subgraph isomorphism
+    algorithm implemented in Gecode. The $x$-axis shows recursive calls, using a log scale, and each
+    line ends at the $x$ value where every instance has been solved.}
     \label{figure:biiiig-data}
@@
 certainly not the case that a sequential scan is infeasible as so many filter~/ verify papers claim.
-The exponential behaviour seen by others in these instances is purely down to a bad choice of
-subgraph isomorphism algorithm, and these instances are not inherently ``really hard''.
+In contrast, with VF2, the hardest instance in PDBS requires 31,189 recursive calls; in AIDS,
+168,569 calls; in PCMS, 6,476,072 calls; and in PPI, 2,913. In other words, any exponential
+behaviour seen in these instances is purely down to a bad choice of subgraph isomorphism algorithm,
+and these instances are not inherently ``really hard''.
 

### 2017-11-08 paper/paper.tex
Text
Unleaded

@@
 computationally fairly easy, with most of the difficulty being in dealing with the size of the
-model. This has lead to bias in algorithm design, to the extent that some proposed techniques, such
+model. This has led to bias in algorithm design, to the extent that some proposed techniques, such
 as those of \citet{DBLP:conf/sls/BattitiM07}, will \emph{only} work on satisfiable instances.
@@
 
-??
+We begin by introducing a way of randomly generating a mix of satisfiable and unsatisfiable
+instances for the non-induced problem, and by measuring the empirical difficulty of such instances.
 
@@
 However, by far the biggest problem is in graph databases. The following section shows how
-widespread use of poor subgraph isomorphism algorithms has not just lead to overly pessimistic
+widespread use of poor subgraph isomorphism algorithms has not just led to overly pessimistic
 conclusions regarding performance, but has misdirected the design of larger systems.
@@
 usually seen as fixed, or at least rarely-changing, whilst the patterns arrive dynamically. This has
-lead to the development of systems which perform extensive computations on the target graphs
+led to the development of systems which perform extensive computations on the target graphs
 beforehand, in the hopes of reducing the response times for individual pattern queries. The most

### 2017-11-08 paper/paper.tex
More changes from Christine

@@
 labels, which are treated as integers; we write $\ell(v)$ for the label of vertex $v$. Two vertices
-are \emph{adjacent} if there is an edge between them. The \emph{order} of a graph is the cardinality
-of its vertex set. We write $\operatorname{V}(G)$ for the vertex set of a graph $G$. The
-\emph{complement} of a graph $G$, denoted $\shortoverline{G}$, is the graph with the same vertex set
-as $G$, and with an edge between distinct vertices $v$ and $w$ if and only if $v$ and $w$ are not
-adjacent in $G$. We write $G(n, p)$ for an Erd\H{o}s-R\'enyi random graph with $n$ vertices, and an
-edge between each distinct unordered pair of vertices with independent probability $p$.
+are \emph{adjacent} if there is an edge between them. The \emph{neighbourhood} of a vertex is the
+set of vertices to which it is adjacent, and the \emph{degree} of a vertex is the cardinality of its
+neighbourhood. The \emph{order} of a graph is the cardinality of its vertex set. We write
+$\operatorname{V}(G)$ for the vertex set of a graph $G$. The \emph{complement} of a graph $G$,
+denoted $\shortoverline{G}$, is the graph with the same vertex set as $G$, and with an edge between
+distinct vertices $v$ and $w$ if and only if $v$ and $w$ are not adjacent in $G$. We write $G(n, p)$
+for an Erd\H{o}s-R\'enyi random graph with $n$ vertices, and an edge between each distinct unordered
+pair of vertices with independent probability $p$.
 
@@
 \citep{DBLP:conf/gbrpr/CarlettiFSV17}. Each was compiled using GCC 4.9 with the ``-O3'' option.
-
-The Glasgow and LAD solvers are constraint programming inspired algorithms (although they use dedicated
-data structures, and do not use a constraint programming toolkit). For each vertex in the pattern
-graph, they create a \emph{variable} whose \emph{domain} contains one \emph{value} for each vertex
-in the target graph. The algorithms perform backtracking search, with the goal of assigning to each
-variable a value from its domain so as to form a subgraph isomorphism. The algorithms use
-\emph{inference} to process \emph{constraints} which eliminate values from domains, ensuring that
-the injectivity (an \emph{all-different} constraint) and adjacency rules are respected. If at any
-point during search a \emph{domain wipeout} occurs (that is, if every value is removed from any
-domain), the algorithms will backtrack immediately. The search is guided by \emph{variable-ordering
-heuristics}, which select which variable (i.e.\ pattern vertex) to branch on next, and
-\emph{value-ordering} heuristics which select which value (i.e.\ target vertex) to try as part of a
-partial mapping. The two algorithms differ in terms of additional inference performed, their
-choices of underlying data structures and algorithms, and in the details of their variable- and
-value-ordering heuristics.
-
-The approach used by VF2 and VF3 is a little different: domains of variables are not stored (so the
-algorithm operates in the style of conventional backtracking), and wipeouts are not detected until
-an assignment is made. Instead, an assignment is built up by repeatedly branching only on vertices
-which are adjacent to a previously-assigned vertex.
-
-We measure only the number of recursive calls (guessed assignments) made, not runtimes. We are not
-aiming to compare absolute performance between solvers; rather, we are looking for
-solver-independent patterns of difficulty. All experiments use a timeout of 1,000 seconds, which is
-enough for the Glasgow solver to solve nearly all instances (whose orders were selected with this
-timeout in mind), although we may slightly overestimate the proportion of unsatisfiable instances
-for extremely sparse or dense pattern graphs. The LAD, VF2, and VF3 solvers experienced many more
-failures with this timeout, so our picture of just how hard the hardest instances are with these
-solvers is less detailed.
-
 The Glasgow, LAD, and VF2 solvers support both non-induced and induced isomorphisms (although
@@
 the induced problem.
+
+\subsection{Subgraph Isomorphism Solvers}
+
+The Glasgow and LAD solvers are constraint programming inspired algorithms (although they use
+dedicated data structures, and do not use a constraint programming toolkit). Constraint programming
+is a field concerned with solving \emph{constraint satisfaction} problems. In such a problem, we
+have a set of \emph{variables}, each of which has a \emph{domain} of \emph{values}. We also have a
+set of \emph{constraints}. The objective is to assign each variable a value from its domain, whilst
+respecting every constraint. This is done using a combination of \emph{inference}, which eliminates
+infeasible values from domains, and backtracking search. This search is driven by
+\emph{variable-ordering} heuristics, which select which unassigned variable to branch on, and
+\emph{value-ordering} heuristics, which select which value to try giving that variable. If at any
+point during search a \emph{domain wipeout} occurs (that is, if every value is removed from any
+domain), the algorithms will backtrack immediately.
+
+In the case of subgraph isomorphism, the basic constraint model is to have a domain for each pattern
+vertex, with the domains being each target vertex. An \emph{all-different} constraint ensures
+injectivity, whilst the adjacency rules can be expressed as extensional constraints by listing
+allowed tuples.  With this model, the backtracking search builds up partial assignments of pattern
+vertices to target vertices, until either a complete assignment is reached (and the instance is
+\emph{satisfiable}), or search establishes that no such assignment can exist (and it is
+\emph{unsatisfiable}).
+
+This model can be extended with additional constraints. These allow more inference to be performed,
+and so a solution may be found (or the lack of solution may be proven) with less search.  The
+simplest example is that a pattern vertex of degree $d$ may only be mapped to a target vertex of
+degree at least $d$, and so constraints can be posted which eliminate values corresponding to target
+vertices of low degree from domains corresponding to pattern vertices of high degree. Richer
+constraints involving neighbourhood degree sequences \citep{DBLP:journals/constraints/ZampelliDS10},
+neighbourhood difference constraints \citep{DBLP:journals/ai/Solnon10}, and distances and paths
+\citep{DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16} are also
+possible.
+
+The Glasgow and LAD solvers both use extensions of this basic model, but differ in terms of the
+additional constraints used (and hence the inference which is performed), their choices of
+underlying data structures and algorithms, and in the details of their variable- and value-ordering
+heuristics.
+
+The approach used by VF2 and VF3 is different: domains of variables are not stored (so the algorithm
+operates in the style of conventional backtracking), and wipeouts are not detected until an
+assignment is made. Instead, an assignment is built up by repeatedly branching only on vertices
+which are adjacent to a previously-assigned vertex.
+
+Our experiments will primarily measure the number of recursive calls (guessed assignments) made, not
+runtimes. We are not aiming to give a simple comparison of absolute performance between solvers;
+rather, we are looking for solver-independent patterns of difficulty. All experiments use a timeout
+of 1,000 seconds, which is enough for the Glasgow solver to solve nearly all instances (whose orders
+were selected with this timeout in mind), although we may slightly overestimate the proportion of
+unsatisfiable instances for extremely sparse or dense pattern graphs. The LAD, VF2, and VF3 solvers
+experienced many more failures with this timeout, so our picture of just how hard the hardest
+instances are with these solvers is less detailed.
 
@@
 
-    \item Domain-based heuristics are much stronger than VF2's adjacency branching rules. Picking
-        from small domains dynamically allows search to focus on the hardest part of the problem.
-
-    \item This approach automatically combines features. For example, a pattern~/ target pair may
-        have matching label features, and matching degree features, but if the only red pattern
-        vertex has degree three whilst no red target vertex has degree more than two, domain
+    \item Domain-based heuristics are widely studied and well-understood
+        \citep{DBLP:conf/ecai/SmithG98}, and unlike VF2's adjacency branching rules, provide
+        discrimination even on dense and low-diameter graphs. Picking from small domains dynamically
+        allows search to focus on the hardest part of the problem, and can exploit conditional
+        structure which is not visible at the top of search.
+
+    \item Domain-based algorithms automatically combines features. For example, a pattern~/ target
+        pair may have matching label features, and matching degree features, but if the only red
+        pattern vertex has degree three whilst no red target vertex has degree more than two, domain
         filtering will detect this immediately. When combined with all-different propagation and
@@
         considerable.  Indices proposed in the literature often take many hours to build, and can
-        consume much more space than the original graphs.
+        consume much more space than the original graphs \citep{DBLP:journals/tkde/HongZLY15}.
 \end{itemize}
@@
 
-\acks{The authors would like to thank Kitty Meeks and Craig Reilly for their comments, and Iva
-Babukova for her help with the datasets. The heatmap schemes used in this paper were created with
-the assistance of the ``Chroma.js Color Scale Helper'' by Gregor Aisch.}
+\acks{The authors would like to thank Kitty Meeks and Craig Reilly for their comments. We are
+grateful to Iva Babukova for her help with the datasets and for sharing the results of her
+experiments. The heatmap schemes used in this paper were created with the assistance of the
+``Chroma.js Color Scale Helper'' by Gregor Aisch.}
 

### 2017-11-10 paper/paper.tex
Notation
emphasis -> emphasise

@@
 the complements. This behaviour can be predicted by taking the partial derivatives of $\langle Sol
-\rangle$ in the $-p_d$ and $t_d$ directions.  However, when inside the remaining two eighths of the
+\rangle$ in the $-p_d$ and $d_t$ directions.  However, when inside the remaining two eighths of the
 parameter space, the partial derivatives of $\langle Sol \rangle$ disagree on which heuristic to
@@
 combination with alternate per-query sub-iso algorithms that can yield large improvements''. We
-believe it is important instead to emphasis previous research directions that have already solved
+believe it is important instead to emphasise previous research directions that have already solved
 most of this problem, and to help ensure that these techniques become more widely known.

### 2017-11-10 paper/paper.tex
Some re-wording

@@
 databases systems to improve their performance.  This is not our claim. Instead, this paper shows
-that better algorithms invalidate the flawed premise underlying the entire filter~/ verify approach.
+that better algorithms invalidate the premise underlying the entire filter~/ verify approach.
 We will now demonstrate empirically that filter~/ verify is simply a poor workaround for the kinds
@@
 
-There are at least four benefits beyond simplicity towards abolishing filter~/ verify in favour of a
+There are at least four benefits beyond simplicity towards abandoning filter~/ verify in favour of a
 purely constraint programming inspired approach:
@@
 
-Unfortunately, indexing is not the only incorrect design choice being made in systems built around
+Indexing is not the only incorrect design choice being made in systems built around
 subgraph isomorphism algorithms, and lessons from constraint programming can be applied to other
@@
 
-\citet{DBLP:conf/edbt/KatsarouNT17} look at what they call ``straggler'' queries, which they define
+\citet{DBLP:conf/edbt/KatsarouNT17} look at ``straggler'' queries, which they define
 to be the small subset of queries that they observe taking much longer than others to solve. The
@@
 sometimes give orders of magnitude improvements. The connection to variable- and value-ordering
-heuristics is not noted, and no consideration is given to even the simplest dynamic ordering
-heuristics like ``smallest domain first'' \citep{DBLP:journals/ai/HaralickE80}. A little thought
-shows that of the orderings proposed, those involving placing rare labels first are effectively poor
+heuristics is not noted, and no consideration is given to dynamic ordering
+heuristics like ``smallest domain first'' \citep{DBLP:journals/ai/HaralickE80}.
+Of the orderings proposed, those involving placing rare labels first are effectively poor
 approximations to a static ``smallest domain at top of search first'' ordering, whilst the remainder
@@
 at all, and perhaps it would be better simply to start using an algorithm with domain tracking and
-domain-based ordering heuristics, rather than assembling a portfolio of bad algorithms in the hopes
+domain-based ordering heuristics, rather than assembling a portfolio of weak algorithms in the hopes
 that at least one of them will often get lucky.
@@
 
-Unfortunately, this lesson has not sunk in. For example, \citet{DBLP:journals/tcbb/BonniciG17}
+Unfortunately, this lesson has not been well heeded. For example, \citet{DBLP:journals/tcbb/BonniciG17}
 recently compared various non-constraint-based approaches to LAD using benchmark sets whose
@@
 programming techniques from the literature, and ignoring all that is known about the empirical
-hardness of \NP-complete problems, members of the graph databases community have misled themselves
-into believing that extensive research into supporting techniques is important, whereas really all
+hardness of \NP-complete problems, members of the graph databases community have reached the mistaken
+conclusion that extensive research into supporting techniques is important, whereas really all
 they are doing is working around some but not all of the defects in their choices of subgraph
@@
 subgraph isomorphism algorithm.  Instead, this paper has shown that such systems need to be designed
-respecting our understanding of the empirical hardness of \NP-complete problems, that subgraph
+in light of a modern understanding of the empirical hardness of \NP-complete problems, that subgraph
 isomorphism algorithms should not be treated as a black box, and that lessons learned in constraint

### 2017-11-10 paper/paper.tex
Typso
Better word

@@
 point during search a \emph{domain wipeout} occurs (that is, if every value is removed from any
-domain), the algorithms will backtrack immediately.
+domain), the search will backtrack immediately.
 
@@
 
-\subsection{A Phase Transitions when Varying Pattern Edge Density}
+\subsection{A Phase Transition when Varying Pattern Edge Density}
 

### 2017-11-10 paper/paper.tex
Define search nodes
Better punctuation

@@
 
-Our experiments will primarily measure the number of recursive calls (guessed assignments) made, not
-runtimes. We are not aiming to give a simple comparison of absolute performance between solvers;
-rather, we are looking for solver-independent patterns of difficulty. All experiments use a timeout
-of 1,000 seconds, which is enough for the Glasgow solver to solve nearly all instances (whose orders
-were selected with this timeout in mind), although we may slightly overestimate the proportion of
-unsatisfiable instances for extremely sparse or dense pattern graphs. The LAD, VF2, and VF3 solvers
-experienced many more failures with this timeout, so our picture of just how hard the hardest
-instances are with these solvers is less detailed.
+Our experiments will primarily measure the number of \emph{search nodes} (that is, the number of
+recursive calls made, or the number of guessed assignments), not runtimes. We are not aiming to give
+a simple comparison of absolute performance between solvers; rather, we are looking for
+solver-independent patterns of difficulty. All experiments use a timeout of 1,000 seconds, which is
+enough for the Glasgow solver to solve nearly all instances (whose orders were selected with this
+timeout in mind), although we may slightly overestimate the proportion of unsatisfiable instances
+for extremely sparse or dense pattern graphs. The LAD, VF2, and VF3 solvers experienced many more
+failures with this timeout, so our picture of just how hard the hardest instances are with these
+solvers is less detailed.
 
@@
 The line plots mean search effort using the Glasgow solver: for sparse patterns, the problem is
-trivial, for dense patterns proving unsatisfiability is not particularly difficult, and we see a
+trivial; for dense patterns, proving unsatisfiability is not particularly difficult; and we see a
 complexity peak around the point where half the instances are satisfiable.  We also plot the search

### 2017-11-10 paper/paper.tex
Density -> Probability
Underline

@@
 
-\subsection{A Phase Transition when Varying Pattern Edge Density}
+\subsection{A Phase Transition when Varying Pattern Edge Probability}
 
@@
 
-\subsection{Phase Transitions when Varying Pattern and Target Edge Densities}
+\subsection{Phase Transitions when Varying Pattern and Target Edge Probabilities}
 
@@
 and Glasgow pick a variable with fewest remaining values in its domain: doing this will generally
-reduce the first part of the $\langle Sol \rangle$ equation (i.e.\ $t^p$) by as little as possible.
-When two or more domains are of equal size, LAD breaks ties lexicographically, whereas Glasgow will
-pick a variable corresponding to a pattern vertex of highest degree. This strategy was determined
-empirically, but could have been derived from the $\langle Sol \rangle$ formula: picking a pattern
-vertex of high degree will make the remaining pattern subgraph sparser, which will decrease the
-exponent in the second half of the formula, maximising the overall value. LAD does not apply a
-value-ordering heuristic, but Glasgow does: it prefers target vertices of highest degree.  Again,
-this was determined empirically, but it has the effect of increasing $\langle Sol \rangle$ by
+reduce the first part of the $\langle Sol \rangle$ equation (i.e.\ $t^{\underline{p}}$) by as little
+as possible.  When two or more domains are of equal size, LAD breaks ties lexicographically, whereas
+Glasgow will pick a variable corresponding to a pattern vertex of highest degree. This strategy was
+determined empirically, but could have been derived from the $\langle Sol \rangle$ formula: picking
+a pattern vertex of high degree will make the remaining pattern subgraph sparser, which will
+decrease the exponent in the second half of the formula, maximising the overall value. LAD does not
+apply a value-ordering heuristic, but Glasgow does: it prefers target vertices of highest degree.
+Again, this was determined empirically, but it has the effect of increasing $\langle Sol \rangle$ by
 leaving as many vertices as possible available for future use. The search strategy used by VF2, in

### 2017-11-10 paper/paper.tex
Adjacent is probably better

@@
 induced subgraph of a complete target: an empty pattern of order $k$ is an induced subgraph only if
-there exist $k$ vertices in the target that are not pairwise connected.  The central satisfiable
+there exist $k$ vertices in the target that are not pairwise adjacent.  The central satisfiable
 region, which is away from either phase transition, is computationally easy, but instances near the

### 2017-11-10 paper/paper.tex
Consistency. Properly this time
Shave off a squillimetre
Word deletion because the hyphenation looks awful

@@
             \node [anchor=center] {\scriptsize $G(150, y)$\hspace*{3em}}; &
-            \\[-0.08cm]
+            \\[-0.09cm]
             \node [anchor=center, inner sep=2pt, rotate=90] {\scriptsize Satisfiable?}; &
@@
             \node [anchor=center] {\includegraphics*{plots/graph-induced-satisfiable-30-150.pdf}}; &
-            \\[-0.08cm]
+            \\[-0.09cm]
             \node [anchor=center, inner sep=2pt, rotate=90] {\scriptsize Glasgow}; &
@@
             \node [anchor=center] {\includegraphics*{plots/graph-induced-nodes-30-150.pdf}}; &
-            \\[-0.08cm]
+            \\[-0.09cm]
             \node [anchor=center, inner sep=2pt, rotate=90] {\scriptsize LAD}; &
@@
             \node [anchor=center] {\includegraphics*{plots/graph-lad-induced-nodes-30-150.pdf}}; &
-            \\[-0.08cm]
+            \\[-0.09cm]
             \node [anchor=center, inner sep=2pt, rotate=90] {\scriptsize VF2}; &
@@
             \node [anchor=center] {\includegraphics*{plots/graph-vf2-induced-nodes-30-150.pdf}}; \\
-            \\[-0.08cm]
+            \\[-0.09cm]
             \node [anchor=center, inner sep=2pt, rotate=90] {\scriptsize VF3}; &
@@
             \node [anchor=center] {\includegraphics*{plots/graph-vf3-induced-nodes-30-150.pdf}}; \\
-            \\[-0.08cm]
+            \\[-0.09cm]
             \node [anchor=center, inner sep=2pt, rotate=90] {\scriptsize Constrainedness}; &
@@
             \node [anchor=center] {\includegraphics*{plots/graph-induced-kappa-30-150.pdf}}; \\
-            \\[-0.08cm]
+            \\[-0.09cm]
             \node [anchor=center, inner sep=2pt, rotate=90] {\scriptsize Complement?}; &
@@
     graphs as inputs---the solid lines show the empirical location of the phase transition, and the
-    dotted lines are $t_d=0.5$ and the $p_d=t_d$ diagonal.}\label{figure:induced}
+    dotted lines are $d_t=0.5$ and the $d_p=d_t$ diagonal.}\label{figure:induced}
 \end{figure}
@@
 the complements. This behaviour can be predicted by taking the partial derivatives of $\langle Sol
-\rangle$ in the $-p_d$ and $d_t$ directions.  However, when inside the remaining two eighths of the
+\rangle$ in the $-d_p$ and $d_t$ directions.  However, when inside the remaining two eighths of the
 parameter space, the partial derivatives of $\langle Sol \rangle$ disagree on which heuristic to
@@
 about adjacency and non-adjacency simultaneously?), or whether instances in this region are
-inherently difficult to solve.  Thus we repeat the induced experiments on smaller pattern and target
+inherently difficult to solve.  Thus we repeat the experiments on smaller pattern and target
 graphs, using different solving techniques.  Although these techniques are not competitive in

### 2017-11-10 paper/paper.tex
More notation

@@
 small compared to the number of pattern vertices $p$. Recall that $\ell(v)$ is the label of vertex
-$v$. By defining \[ \operatorname{V}(P)\vert_n = \{ v \in \operatorname{V}(P) : \ell(v) = x \} \] to
+$v$. By defining \[ \operatorname{V}(P)\vert_x = \{ v \in \operatorname{V}(P) : \ell(v) = x \} \] to
 be the set of vertices with label $x$, we can partition the pattern vertices by label into disjoint

### 2017-11-10 paper/paper.tex
Say that it's NP-complete
Footnote positioning

@@
 isomorphism $i : P \rightarrowtail T$ which is also a non-induced isomorphism $\shortoverline{i} :
-\shortoverline{P} \rightarrowtail \shortoverline{T}$.
+\shortoverline{P} \rightarrowtail \shortoverline{T}$. Finding either form of isomorphism between two
+given graphs is \NP-complete.
 
@@
 \citep{DBLP:conf/aaai/Regin94}, it cannot immediately detect unsatisfiability in simple cases like
-there being two red vertices in the pattern but only one in the target\footnote{Although
+there being two red vertices in the pattern but only one in the target.\footnote{Although
 interestingly, this algorithm effectively does forward-checking and has a variable-ordering
-heuristic, before these concepts appeared in the constraints literature.}.
+heuristic, before these concepts appeared in the constraints literature.}
 

### 2018-02-12 paper/paper.tex
A bit less wordy about KatsarouNT17's heuristics

@@
 
-\citet{DBLP:conf/edbt/KatsarouNT17} look at ``straggler'' queries, which they define
-to be the small subset of queries that they observe taking much longer than others to solve. The
-authors present an approach to address this perceived problem that they describe as novel: ``instead
-of trying to come up with new algorithms for sub-iso testing, we utilize isomorphic query rewritings
-and existing alternative algorithms in parallel''. Their first claim is that permuting graphs before
-running the subgraph isomorphism algorithm can lead to ``wildly different'' execution times. They
-note that VF2 does not define ``a strict order in which the nodes of the query are matched'', and so
-permuting the graphs (for example, by using degree or label frequency information) before search can
-sometimes give orders of magnitude improvements. The connection to variable- and value-ordering
-heuristics is not noted, and no consideration is given to dynamic ordering
-heuristics like ``smallest domain first'' \citep{DBLP:journals/ai/HaralickE80}.
-Of the orderings proposed, those involving placing rare labels first are effectively poor
-approximations to a static ``smallest domain at top of search first'' ordering, whilst the remainder
-use degree as we did earlier in this paper. However, \citeauthor{DBLP:conf/edbt/KatsarouNT17} do not
-investigate any algorithm which employs domains, let alone strong variable- or value-ordering
-heuristics or all-different propagation, and do not consider that their apparent successes could be
-due to VF2's weaknesses rather than an inherent property of \NP-completeness.
+\citet{DBLP:conf/edbt/KatsarouNT17} present approaches to improve the run times for what they call
+``straggler queries''---the small subset of queries that they observe taking much longer than others
+to solve.  They observe that permuting graphs (for example, by using degree or label frequency
+information) before running the subgraph isomorphism algorithm can improve run times by orders of
+magnitude. However, the connection between this approach and variable- and value-ordering heuristics
+is not noted, and no consideration is given to dynamic ordering heuristics such as ``smallest domain
+first'' \citep{DBLP:journals/ai/HaralickE80}.  Of the orderings proposed, those involving placing
+rare labels first are effectively approximations to a static ``smallest domain at top of search
+first'' ordering, whilst the remainder use degree as we did earlier in this paper.
+\citeauthor{DBLP:conf/edbt/KatsarouNT17} do not investigate any algorithm which employs strong
+variable- or value-ordering heuristics, all-different propagation, or even domains, and do not
+consider that the apparent successes of their approach could be due to VF2's weaknesses rather than
+an inherent property of \NP-completeness.
 

### 2018-02-12 paper/paper.tex
Removed portfolios bit

@@
 
-We saw in the previous sections that although permuting input graphs could improve VF2's behaviour
-on some instances, doing so does not make its performance come close to that of domain-based
-algorithms.  \citeauthor{DBLP:conf/edbt/KatsarouNT17} say they ``hope that our findings will open up
-new research directions, striving to find appropriate, per-query, isomorphic rewritings, in
-combination with alternate per-query sub-iso algorithms that can yield large improvements''. We
-believe it is important instead to emphasise previous research directions that have already solved
-most of this problem, and to help ensure that these techniques become more widely known.
-
-\citet{DBLP:conf/edbt/KatsarouNT17}'s second claim is that ``different
-algorithms find different queries hard''. To address this, they run many subgraph isomorphism
-algorithms and input permutations in parallel; the extensive literature on parallel portfolios in
-general \citep{DBLP:journals/ai/GomesS01}, and the approach by \citet{DBLP:conf/sls/BattitiM07} for
-subgraph isomorphism in particular, is not noted (and nor are portfolios mentioned when they say
-that ``using machine learning models to predict which version \ldots to employ per query is of high
-interest'').  The evidence so far in this paper suggests that we should be wary. It is certainly
-true that there are some instances that all algorithms find hard, and there are good theoretical
-reasons to believe that these instances genuinely are really hard.  Furthermore, algorithm
-portfolios are also well-known to be a successful technique, and are beneficial even with modern
-subgraph isomorphism algorithms \citep{DBLP:conf/lion/KotthoffMS16}.  However, in
-\cref{section:labelled} we saw that there are many instances that VF2 finds hard that should not be
-hard, and that are not hard for other algorithms.  If permuting graphs sometimes addresses the
-difficulty of some of these instances, then it is likely that they are not genuinely hard instances
-at all, and perhaps it would be better simply to start using an algorithm with domain tracking and
-domain-based ordering heuristics, rather than assembling a portfolio of weak algorithms in the hopes
-that at least one of them will often get lucky.
-
-To test this suggestion, we return briefly to the experiments on the datasets discussed earlier in
-this section. What if we modify our solver so that it does not use ``smallest domain first'', and
-does not detect domain wipeouts until attempting to branch on an empty variable?  Suddenly three of
-the four datasets become extremely hard, with many instances now timing out after making hundreds of
-millions of recursive calls, and the PPI dataset now requires hundreds of thousands of calls for
-some queries.  Furthermore, slight changes to the input vertex ordering can now make some of these
-``hard'' instances easy again.  This backs up our suspicions: the apparent success of
-\citet{DBLP:conf/edbt/KatsarouNT17}'s technique is due to the ease of sometimes getting better
-results out of a poor algorithm. Claiming huge improvements from an algorithm portfolio consisting
-only of variations of poor algorithms does not demonstrate a genuine improvement over the state of
-the art.
-
 Improved search orderings for VF2-style algorithms also appear in recent work by
@@
 bring an algorithm that does not use domains close to the performance of one that does on harder
-instances.
-
-The importance of using hard benchmark instances was stressed by
+instances.  We thus believe that a significant amount of effort in designing algorithms for subgraph
+isomorphism has been misdirected due to a lack of familiarity with the best existing algorithms for
+subgraph isomorphism on one hand, and with broadly-applicable lessons from the constraint
+programming literature on the other.
+
+We now move our focus from the importance of using state-of-the-art techniques for algorithm design,
+to the pitfalls of choosing unrealistically easy instances to benchmark our algorithms.
+This was stressed by
 \citet{DBLP:journals/heuristics/Gent98}:

### 2018-02-12 paper/paper.tex
Make NP font consistent

@@
 verification time dominates the Query Response Time [s]ince the computational complexity of
-$T_{\mathit{iso\_test}}$ is NP-Complete''. They note that \begin{displayquote}``Approximately, the
+$T_{\mathit{iso\_test}}$ is \NP-Complete''. They note that \begin{displayquote}``Approximately, the
     value of $T_{\mathit{iso\_test}}$ does not change too much with the difference of query. Thus,
@@
 Without using the equation, \citet{DBLP:conf/sigmod/ChengKNL07} state that since subgraph
-isomorphism is NP-complete, processing by a sequential scan is infeasible. They introduce new
+isomorphism is \NP-complete, processing by a sequential scan is infeasible. They introduce new
 filtering techniques to try to avoid the subgraph isomorphism step. A similar claim is made by
@@
     graph is small, it can be easily fit in the main memory.  However, the query (matching) time
-will be very long due to the NP-hard complexity.''\end{displayquote} They suggest that indexing is
+will be very long due to the \NP-hard complexity.''\end{displayquote} They suggest that indexing is
 a way of avoiding this. Similarly, \citet{DBLP:journals/pvldb/ZhaoH10} argue that ``the graph query
 problem is hard in that \ldots it requires subgraph isomorphism checking \ldots which has proven to
-be NP-complete'' and that working with large networks is hard or impossible ``due to the lack of
+be \NP-complete'' and that working with large networks is hard or impossible ``due to the lack of
 scalable graph indexing mechanisms''.
@@
 returns a special subset of graphs which match a query, \citet{DBLP:journals/tkde/ZhengLZHZ16}
-suggest that it is ``NP-hard to check the graph isomorphism'' (meaning subgraph isomorphism), and
+suggest that it is ``\NP-hard to check the graph isomorphism'' (meaning subgraph isomorphism), and
 ``in order to improve the time efficiency'' they use an indexing system to ``avoid as many costly

### 2018-02-12 paper/paper.tex
Tone down language

@@
     These results have practical consequences: we explain why the widely researched ``filter~/
-    verify'' indexing technique used in graph databases is a flawed design (based upon a repeated
-    misunderstanding of the empirical hardness of \NP-complete problems) that cannot be beneficial
+    verify'' indexing technique used in graph databases is founded upon a
+    misunderstanding of the empirical hardness of \NP-complete problems, and cannot be beneficial
     when paired with any reasonable subgraph isomorphism algorithm.
@@
 filter~/ verify and other recently proposed techniques cannot be beneficial even on more challenging
-instances except if an extremely poor choice of subgraph isomorphism algorithm is made.
+instances, unless the chosen subgraph isomorphism algorithm exhibits unnecessarily poor behavior
+on certain classes of instance.
 
@@
 labels, and the ten last vertices have zero labels (and similarly for the target graphs).
-This makes VF2 behave better, but still extremely poorly compared to the other two solvers; the
+This makes VF2 behave better, but still very poorly compared to the other two solvers; the
 motivation behind this modification will become clear later in this paper.  It will also be
@@
 
-The filter~/ verify approach has an interesting history, of which we now give a very selective and
-incomplete overview. Our description is biased by a general modern understanding of the empirical
+The filter~/ verify approach has an interesting history, of which we now give an abridged
+overview. Our description is biased by a general modern understanding of the empirical
 hardness of \NP-complete problems, which was not widely known at the time of the earlier
@@
 graph indices in order to help processing graph queries.''\end{displayquote}
-With what we now know about the behaviour of modern subgraph isomorphism algorithms, and the nature
-of solving \NP-complete problems in general, we should immediately be suspicious of these claims. We
-do not expect $T_{\mathit{iso\_test}}$ to be anything like a constant, even if the orders of the
+We reassess these claims in light of
+what we now know about the behaviour of modern subgraph isomorphism algorithms and the nature
+of solving \NP-complete problems. Contrary to the authors' claims,
+we do not expect $T_{\mathit{iso\_test}}$ to be anything like a constant, even if the orders of the
 input graphs are similar. In particular, any instance which can be excluded based upon filtering
 must have a very small proof of unsatisfiability. These instances \emph{should} be trivial with any
-decent subgraph isomorphism algorithm. Thus, all filtering \emph{should} be doing is eliminating the startup
+reasonably good subgraph isomorphism algorithm. Thus, all filtering \emph{should} be doing is eliminating the startup
 costs of a trivial subgraph isomorphism call. The fact that filtering was successful empirically
-should make us wonder whether the subgraph isomorphism algorithm being used was excessively
-primitive. Indeed, the subgraph isomorphism algorithm used is described only as ``the simplest approach'' in the
+should make us wonder whether the subgraph isomorphism algorithm used was substantially
+weaker than the state of the art.
+Indeed, the algorithm is described only as ``the simplest approach'' in the
 paper.  Additionally, the experiments focus on reducing the size of the candidate set, without
@@
 168,569 calls; in PCMS, 6,476,072 calls; and in PPI, 2,913. In other words, any exponential
-behaviour seen in these instances is purely down to a bad choice of subgraph isomorphism algorithm,
+behaviour seen in these instances is purely down to the choice of subgraph isomorphism algorithm,
 and these instances are not inherently ``really hard''.
@@
 
-Unfortunately, this lesson has not been well heeded. For example, \citet{DBLP:journals/tcbb/BonniciG17}
+This lesson has not been well heeded. For example, \citet{DBLP:journals/tcbb/BonniciG17}
 recently compared various non-constraint-based approaches to LAD using benchmark sets whose
@@
 When labels were introduced, we saw that VF2 finds some instances hard which other algorithms find
-easy. We looked at this kind of instance in more detail, and argued that this was due to flaws in
+easy. We looked at this kind of instance in more detail, and argued that this was due to shortcomings in
 VF2's design, rather than an interesting property of \NP-completeness. Although inevitably there
-will be instances that every algorithm finds hard, we see no excuse for an algorithm to exhibit
-exponential behaviour in the case when there are two red vertices in a pattern and only one in a
+will be instances that every algorithm finds hard, we see it as a serious weakness for an algorithm to exhibit
+exponential behaviour in such trivial cases as when there are two red vertices in the pattern but only one in the
 target.
 
-Unfortunately, we saw that this aspect of VF2's performance has had practical consequences, most
-notably being the misdesign of graph database systems. By not trying even the simplest constraint
-programming techniques from the literature, and ignoring all that is known about the empirical
+Unfortunately, we saw that this aspect of VF2's performance has had practical consequences, with the
+most notable being the misdesign of graph database systems. By not trying even the simplest constraint
+programming techniques from the literature, and disregarding all that is known about the empirical
 hardness of \NP-complete problems, members of the graph databases community have reached the mistaken
-conclusion that extensive research into supporting techniques is important, whereas really all
-they are doing is working around some but not all of the defects in their choices of subgraph
-isomorphism algorithms. With this new understanding of what does and does not make subgraph
+conclusion that extensive research into supporting techniques is important, whereas really what
+they are doing is working around some but not all of the defects in their chosen algorithms for subgraph
+isomorphism. With this new understanding of what does and does not make subgraph
 isomorphism hard, it is time for a radical rethink of how graph database systems work.
@@
 large numbers of trivial instances. Employing a presolver is an obvious approach---and VF2 is
-actually good in this role \citep{DBLP:conf/lion/KotthoffMS16}---but there are other possibilities.
+in fact good in this role \citep{DBLP:conf/lion/KotthoffMS16}---but there are other possibilities.
 For example, minimal or lazy forward-checking

### 2018-02-12 paper/paper.tex
Correct spacing after et al.

@@
 existing benchmark suites contain relatively few non-trivial unsatisfiable instances (although a few
-of the patterns in the instances by \citeauthor{DBLP:journals/constraints/ZampelliDS10} have had
+of the patterns in the instances by \citeauthor{DBLP:journals/constraints/ZampelliDS10}\ have had
 extra edges added, making them unsatisfiable). Also, the satisfiable instances tend to be
@@
 first'' ordering, whilst the remainder use degree as we did earlier in this paper.
-\citeauthor{DBLP:conf/edbt/KatsarouNT17} do not investigate any algorithm which employs strong
+\citeauthor{DBLP:conf/edbt/KatsarouNT17}\ do not investigate any algorithm which employs strong
 variable- or value-ordering heuristics, all-different propagation, or even domains, and do not

### 2018-02-12 paper/paper.tex
Re-word what's in domains

@@
 In the case of subgraph isomorphism, the basic constraint model is to have a domain for each pattern
-vertex, with the domains being each target vertex. An \emph{all-different} constraint ensures
+vertex, with each domain being the set of target vertices. An \emph{all-different} constraint ensures
 injectivity, whilst the adjacency rules can be expressed as extensional constraints by listing

### 2018-02-14 paper/paper.tex
Minor tweakery

@@
 bring an algorithm that does not use domains close to the performance of one that does on harder
-instances.  We thus believe that a significant amount of effort in designing algorithms for subgraph
-isomorphism has been misdirected due to a lack of familiarity with the best existing algorithms for
-subgraph isomorphism on one hand, and with broadly-applicable lessons from the constraint
-programming literature on the other.
-
-We now move our focus from the importance of using state-of-the-art techniques for algorithm design,
-to the pitfalls of choosing unrealistically easy instances to benchmark our algorithms.
-This was stressed by
-\citet{DBLP:journals/heuristics/Gent98}:
+instances.  We thus believe that a significant amount of effort in designing algorithms for and
+systems involving subgraph isomorphism has been misdirected due to a lack of familiarity with the
+best existing algorithms for subgraph isomorphism, and with more broadly-applicable lessons on
+algorithm design from the constraint programming literature.
+
+Another such lesson is the importance of choosing a good set of benchmark instances. As
+\citet{DBLP:journals/heuristics/Gent98} observes:
 
@@
 
-This lesson has not been well heeded. For example, \citet{DBLP:journals/tcbb/BonniciG17}
-recently compared various non-constraint-based approaches to LAD using benchmark sets whose
+We hope that this paper provides some much-needed help in this direction, as some recent
+evaluations work only with extremely easy instances. For example, \citet{DBLP:journals/tcbb/BonniciG17}
+compared various non-constraint-based approaches to LAD using benchmark sets whose
 difficulties are measured between microseconds and milliseconds. Their results suggest that LAD is
@@
 conclusion that extensive research into supporting techniques is important, whereas really what
-they are doing is working around some but not all of the defects in their chosen algorithms for subgraph
+they are doing is working around some but not all of the weaknesses in their chosen algorithms for subgraph
 isomorphism. With this new understanding of what does and does not make subgraph

### 2018-02-14 paper/paper.tex
Nicerify

@@
 databases systems to improve their performance.  This is not our claim. Instead, this paper shows
-that better algorithms invalidate the premise underlying the entire filter~/ verify approach.
-We will now demonstrate empirically that filter~/ verify is simply a poor workaround for the kinds
-of deficiency in VF2 demonstrated towards the end of the previous section.
+that better algorithms invalidate the premise underlying the entire filter~/ verify approach.  We
+will now demonstrate empirically that filter~/ verify is simply a limited workaround for a subset of
+the weaknesses in non-constraint-based subgraph isomorphism algorithms that we demonstrated towards
+the end of the previous section.
 
@@
 
+We saw in the previous sections that although permuting input graphs could improve VF2's behaviour
+on some instances, doing so does not make its performance come close to that of domain-based
+algorithms.  \citeauthor{DBLP:conf/edbt/KatsarouNT17} say they ``hope that our findings will open up
+new research directions, striving to find appropriate, per-query, isomorphic rewritings, in
+combination with alternate per-query sub-iso algorithms that can yield large improvements''. We
+believe it is important instead to emphasise previous research directions that have already solved
+most of this problem, and to help ensure that these techniques become more widely known.
+
+\citet{DBLP:conf/edbt/KatsarouNT17}'s second claim is that ``different
+algorithms find different queries hard''. To address this, they run many subgraph isomorphism
+algorithms and input permutations in parallel; the extensive literature on parallel portfolios in
+general \citep{DBLP:journals/ai/GomesS01}, and the approach by \citet{DBLP:conf/sls/BattitiM07} for
+subgraph isomorphism in particular, is not noted (and nor are portfolios mentioned when they say
+that ``using machine learning models to predict which version \ldots to employ per query is of high
+interest'').  The evidence so far in this paper suggests that we should be wary. It is certainly
+true that there are some instances that all algorithms find hard, and there are good theoretical
+reasons to believe that these instances genuinely are really hard.  Furthermore, algorithm
+portfolios are also well-known to be a successful technique, and are beneficial even with modern
+subgraph isomorphism algorithms \citep{DBLP:conf/lion/KotthoffMS16}.  However, in
+\cref{section:labelled} we saw that there are many instances that VF2 finds hard that should not be
+hard, and that are not hard for other algorithms.  If permuting graphs sometimes addresses the
+difficulty of some of these instances, then it is likely that they are not genuinely hard instances
+at all, and perhaps it would be better simply to start using an algorithm with domain tracking and
+domain-based ordering heuristics, rather than assembling a portfolio of weak algorithms in the hopes
+that at least one of them will often get lucky.
+
+To test this suggestion, we return briefly to the experiments on the datasets discussed earlier in
+this section. What if we modify our solver so that it does not use ``smallest domain first'', and
+does not detect domain wipeouts until attempting to branch on an empty variable?  Suddenly three of
+the four datasets become extremely hard, with many instances now timing out after making hundreds of
+millions of recursive calls, and the PPI dataset now requires hundreds of thousands of calls for
+some queries.  Furthermore, slight changes to the input vertex ordering can now make some of these
+``hard'' instances easy again.  This backs up our suspicions: the apparent success of
+\citet{DBLP:conf/edbt/KatsarouNT17}'s technique is due to the ease of sometimes getting better
+results out of a poor algorithm. Claiming huge improvements from an algorithm portfolio consisting
+only of variations of poor algorithms does not demonstrate a genuine improvement over the state of
+the art.
+
 Improved search orderings for VF2-style algorithms also appear in recent work by

### 2018-02-14 paper/paper.tex
Final sentence
Explain T_io (review 1)

@@
 This equation is repeated and expanded upon by \citet{DBLP:journals/tods/YanYH05} to explicitly
-separate I/O and isomorphism testing costs, obtaining a query response time of \[
+separate I/O and isomorphism testing costs into $T_{\mathit{io}}$ and $T_{\mathit{iso\_test}}$
+respectively, obtaining a query response time of \[
     T_{\mathit{search}} + \left|C_q\right| \cdot (T_{\mathit{io}} + T_{\mathit{iso\_test}})
@@
 unexpected ten-vertex clique, which would be extremely unlikely in a randomly generated graph with a
-similar order and density.
+similar order and density. We therefore emphasise the importance of working with a diverse set of
+benchmark instances, and of performing a careful family-by-family analysis of the results.
 

### 2018-02-14 paper/paper.tex
Remove commented out stuff
Ref NP-complete (review 2)

@@
 \usepackage{csquotes}
-
-% \usepackage{showframe}
 
@@
 so there are no ``extra edges'' in the copy of the pattern that we find. We illustrate both variants
-in \cref{figure:sip}.  Although these problems are \NP-complete, modern subgraph
+in \cref{figure:sip}.  Although these problems are \NP-complete \citep{DBLP:books/fm/GareyJ79}, modern subgraph
 isomorphism algorithms based upon constraint programming techniques can handle problem instances
@@
 seems to be more common in the literature.
-% , and return to the Erd\H{o}s-R\'enyi model.
 
@@
             \\
-            % \node [anchor=center, inner sep=2pt, rotate=90] {\scriptsize VF2 Failures}; &
-            % \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-150.pdf}}; &
-            % \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-l2-150.pdf}}; &
-            % \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-l3-150.pdf}}; &
-            % \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-l5-150.pdf}}; &
-            % \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-l10-150.pdf}}; &
-            % \node [anchor=center] {\includegraphics*{plots/graph-vf2-non-induced-abortions-20-l20-150.pdf}}; &
-            % \\
         };

### 2018-02-14 paper/paper.tex
Pairs (review 2)

@@
 or thousands of vertices. Experimental evaluations of subgraph isomorphism algorithms are usually
-performed using randomly generated graph pairs, sometimes together with a mix of real-world graphs,
+performed using randomly generated pairs of pattern and target graphs, sometimes together with a mix of real-world graphs,
 for example biochemistry and computer vision problems. Using random instances to evaluate

### 2018-02-14 paper/paper.tex
Poor (review 2)

@@
 systems, which the literature suggests are hard to solve. Simple experiments show that the entire
-perceived difficulty of each of these datasets comes down to the widespread use of especially poor
-subgraph isomorphism algorithms. Our fourth contribution is to show that the simplest constraint
-programming approach, even without the more sophisticated recent advances in subgraph algorithms,
-makes the entire graph database filter~/ verify paradigm unnecessary.  Finally, we explain why
-filter~/ verify and other recently proposed techniques cannot be beneficial even on more challenging
-instances, unless the chosen subgraph isomorphism algorithm exhibits unnecessarily poor behavior
-on certain classes of instance.
+perceived difficulty of each of these datasets comes down to the widespread use of particular
+subgraph isomorphism algorithms, rather than inherent hardness. Our fourth contribution is to show
+that the simplest constraint programming approach, even without the more sophisticated recent
+advances in subgraph algorithms, makes the entire graph database filter~/ verify paradigm
+unnecessary.  Finally, we explain why filter~/ verify and other recently proposed techniques cannot
+be beneficial even on more challenging instances, unless the chosen subgraph isomorphism algorithm
+exhibits unnecessarily poor behavior on certain classes of instance.
 
@@
 However, by far the biggest problem is in graph databases. The following section shows how
-widespread use of poor subgraph isomorphism algorithms has not just led to overly pessimistic
+widespread use of weak subgraph isomorphism algorithms has not just led to overly pessimistic
 conclusions regarding performance, but has misdirected the design of larger systems.

### 2018-02-14 paper/paper.tex
Most (review 2)

@@
     version does not.  Although both variants are \NP-complete, algorithms inspired by constraint programming
-    can operate comfortably on most real-world graphs with thousands of vertices.  However, they
-    cannot handle arbitrary instances of this size. We show how to generate ``really hard'' random
-    instances for subgraph isomorphism problems, which are computationally challenging with a couple
-    of hundred vertices in the target, and only twenty pattern vertices. For the non-induced version
-    of the problem, these instances lie on a satisfiable / unsatisfiable phase transition, whose
-    location we can predict; for the induced variant, much richer behaviour is observed, and
-    constrainedness gives a better measure of difficulty than does proximity to a phase transition.
-    These results have practical consequences: we explain why the widely researched ``filter~/
-    verify'' indexing technique used in graph databases is founded upon a
+    can operate comfortably on many real-world problem instances with thousands of vertices.
+    However, they cannot handle arbitrary instances of this size. We show how to generate ``really
+    hard'' random instances for subgraph isomorphism problems, which are computationally challenging
+    with a couple of hundred vertices in the target, and only twenty pattern vertices. For the
+    non-induced version of the problem, these instances lie on a satisfiable / unsatisfiable phase
+    transition, whose location we can predict; for the induced variant, much richer behaviour is
+    observed, and constrainedness gives a better measure of difficulty than does proximity to a
+    phase transition.  These results have practical consequences: we explain why the widely
+    researched ``filter~/ verify'' indexing technique used in graph databases is founded upon a
     misunderstanding of the empirical hardness of \NP-complete problems, and cannot be beneficial

### 2018-02-14 paper/paper.tex
Kappa is an ensemble (review 3)

@@
 phase transition concept, and to generalise hardness parameters across different combinatorial
-problems \citep{DBLP:conf/aaai/GentMPW96}. A problem with $\kappa < 1$ is said to be
-underconstrained, and is likely to be satisfiable; a problem with $\kappa > 1$ is overconstrained,
-and is likely to be unsatisfiable. Empirically, problems with $\kappa$ close to 1 are hard, and
-problems where $\kappa$ is very small or very large are usually easy. By handling injectivity as a
+problems \citep{DBLP:conf/aaai/GentMPW96}. An ensemble of problem instances with $\kappa < 1$ is
+said to be underconstrained, and is likely to be made up of satisfiable instances; an ensemble with
+$\kappa > 1$ is overconstrained, and is likely to be made of unsatisfiable instances. Empirically,
+problem instances from an ensemble with $\kappa$ close to 1 are hard, and problems drawn from an
+ensemble where $\kappa$ is very small or very large are usually easy.
+
+By handling injectivity as a
 restriction on the size of the state space rather than as a constraint, we derive
@@
 row of \cref{figure:induced}. We see that constrainedness predicts that the central region will
-still be relatively difficult for larger pattern graphs: although the problem is overconstrained, it
-is less overconstrained than in the regions the Glasgow and LAD solvers found easy.  Thus it seems
-that rather than just being a unification of existing generalised heuristic techniques,
-constrainedness also gives a better predictor of difficulty than proximity to a phase
-transition---our method generates instances where constrainedness and ``close to a phase
-transition'' give very different predictions, and constrainedness closely matches the empirical
-results.
+still be relatively difficult for larger pattern graphs: although a family of instances generated
+with these parameters is overconstrained, it is less overconstrained than in the regions the Glasgow
+and LAD solvers found easy.  Thus it seems that rather than just being a unification of existing
+generalised heuristic techniques, constrainedness also gives a better predictor of difficulty than
+proximity to a phase transition---our method generates collections of instances where
+constrainedness and ``close to a phase transition'' give very different predictions, and
+constrainedness closely matches the empirical results.
 

### 2018-02-14 paper/paper.tex
Cumulative (review 1)

@@
 
-    \caption{Cumulative number of instances solved as a function of search space size, using four
+    \caption{The cumulative number of instances solved as a function of search space size, using four
     graph database datasets, and either VF2 or a simple constraint programming subgraph isomorphism

### 2018-02-14 paper/paper.tex
Clarify heuristics (review 1)

@@
 we see a solver-independent complexity peak near this phase transition. Additionally, understanding
-this behaviour helps us to design better variable- and value-ordering heuristics---this is the
-second contribution of this paper.
+this behaviour helps us to design better search strategies, by improving variable- and
+value-ordering heuristics---this is the second contribution of this paper.
 
@@
 \emph{variable-ordering} heuristics, which select which unassigned variable to branch on, and
-\emph{value-ordering} heuristics, which select which value to try giving that variable. If at any
-point during search a \emph{domain wipeout} occurs (that is, if every value is removed from any
-domain), the search will backtrack immediately.
+\emph{value-ordering} heuristics, which select which value to try giving that variable. That is,
+variable-ordering selects a pattern vertex, whilst value-ordering tells us the order in which to try
+different candidate target vertices for that pattern vertex. If at any point during search a
+\emph{domain wipeout} occurs (that is, if every value is removed from any domain), the search will
+backtrack immediately.
 

### 2018-02-14 paper/paper.tex
Ref VF2 when it first appears (review 1)

@@
 database systems), richer behaviour emerges, particularly when moving away from a uniform labelling
-scheme, and the popular VF2 algorithm behaves much worse than two more recent
-constraint-based algorithms in certain cases. In \cref{section:labelled} we take a close look at
-these cases, and argue that they \emph{should} be easy to solve.
+scheme, and the popular VF2 algorithm \citep{DBLP:journals/pami/CordellaFSV04} behaves much worse
+than two more recent constraint-based algorithms in certain cases. In \cref{section:labelled} we
+take a close look at these cases, and argue that they \emph{should} be easy to solve.
 

### 2018-02-19 paper/paper.tex
Cite thesis (review 1)
Sample size (review 1)

@@
 sections have been expanded, whilst
-\cref{section:regular,section:labelled,section:filterverify} are new material.
+\cref{section:regular,section:labelled,section:filterverify} are new material. Most of this work
+also appeared in a PhD thesis \citep{o:McCreesh17}.
 
@@
 
-\begin{figure}[t]
-    \centering
-    \includegraphics*{plots/phase-transition.pdf}
-
-    \caption{With a fixed pattern graph order of 20, a target graph order of 150, a target edge
-    probability of 0.40, and varying pattern edge probability, we observe a phase transition and
-    complexity peak with the Glasgow solver in the non-induced variant. Each point $(x, y)$
-    represents one instance $i$, where $x$ is the pattern edge probability used to generate $i$, $y$
-    is the number of search nodes needed by the Glasgow solver to solve $i$, and the point is drawn
-    as a green circle if $i$ is satisfiable, and a blue cross otherwise. The black line plots the
-    evolution of the mean number of search nodes when increasing the pattern edge probability from
-    $0$ to $1$.}
-    \label{figure:phase-transition}
-\end{figure}
-
 Suppose we arbitrarily decide upon a pattern graph order of 20, a target graph order of 150, a fixed
@@
 the instances are satisfiable.
+
+\begin{figure}[t]
+    \centering
+    \includegraphics*{plots/phase-transition.pdf}
+
+    \caption{With a fixed pattern graph order of 20, a target graph order of 150, a target edge
+    probability of 0.40, and varying pattern edge probability, we observe a phase transition and
+    complexity peak with the Glasgow solver in the non-induced variant. Each point $(x, y)$
+    represents one instance $i$, where $x$ is the pattern edge probability used to generate $i$, $y$
+    is the number of search nodes needed by the Glasgow solver to solve $i$, and the point is drawn
+    as a green circle if $i$ is satisfiable, and a blue cross otherwise. The black line plots the
+    evolution of the mean number of search nodes when increasing the pattern edge probability from
+    $0$ to $1$ in steps of $0.01$, using a larger sample size of $1,000$.}
+    \label{figure:phase-transition}
+\end{figure}
 

### 2018-02-19 paper/paper.tex
Avoid bad hyphenation
Are real world instances hard (review 3)

@@
 
-In \cref{section:induced} we look at the induced isomorphism problem.  For certain choices of
+In \cref{section:induced} we look at the induced subgraph isomorphism problem.  For certain choices of
 parameters, there are two phase transitions, going from satisfiable to unsatisfiable, and then from
@@
 
-When labels were introduced, we saw that VF2 finds some instances hard which other algorithms find
-easy. We looked at this kind of instance in more detail, and argued that this was due to shortcomings in
-VF2's design, rather than an interesting property of \NP-completeness. Although inevitably there
-will be instances that every algorithm finds hard, we see it as a serious weakness for an algorithm to exhibit
-exponential behaviour in such trivial cases as when there are two red vertices in the pattern but only one in the
-target.
+We note that these ``really hard'' instances have a very particular structure, which is unlikely to
+arise naturally. Indeed, the successes seen with constraint programming approaches on the instances
+in \cref{section:filterverify}, and in the wider range of application instances used by
+\citet{DBLP:conf/lion/KotthoffMS16}, suggest that the real-world instances encountered so far are
+generally fairly easy for a good algorithm. Nonetheless, carefully designed synthetic instances
+provide a good starting point for scientific experiments, and can help improve the behaviour of
+algorithms on real-world instances too.
+
+For example, when labels were introduced, we saw that VF2 found some instances hard which other
+algorithms found easy. We looked at this kind of instance in more detail, and argued that this was
+due to shortcomings in VF2's design, rather than an interesting property of \NP-completeness.
+Although inevitably there will be instances that every algorithm finds hard, we see it as a serious
+weakness for an algorithm to exhibit exponential behaviour in such trivial cases as when there are
+two red vertices in the pattern but only one in the target.
 

### 2018-02-19 paper/paper.tex
More cites

@@
 scheme, and the popular VF2 algorithm \citep{DBLP:journals/pami/CordellaFSV04} behaves much worse
-than two more recent constraint-based algorithms in certain cases. In \cref{section:labelled} we
-take a close look at these cases, and argue that they \emph{should} be easy to solve.
+than two more recent constraint-based algorithms
+\citep{DBLP:journals/ai/Solnon10,DBLP:conf/cp/McCreeshP15} in certain cases. In
+\cref{section:labelled} we take a close look at these cases, and argue that they \emph{should} be
+easy to solve.
 

### 2018-02-19 paper/paper.tex
Cite ER

@@
 denoted $\shortoverline{G}$, is the graph with the same vertex set as $G$, and with an edge between
-distinct vertices $v$ and $w$ if and only if $v$ and $w$ are not adjacent in $G$. We write $G(n, p)$
-for an Erd\H{o}s-R\'enyi random graph with $n$ vertices, and an edge between each distinct unordered
-pair of vertices with independent probability $p$.
+distinct vertices $v$ and $w$ if and only if $v$ and $w$ are not adjacent in $G$.
+
+An \emph{Erd\H{o}s-R\'enyi random graph}, written $G(n, p)$, is a graph with $n$ vertices, and an
+edge between each distinct unordered pair of vertices with independent probability $p$
+\citep{o:Gilbert59,o:ErdosR59}.\footnote{Note that elsewhere, this same name sometimes instead
+refers to a graph $G(n, E)$ chosen uniformly from among all graphs with $n$ vertices and exactly $E$
+edges.}
 

### 2018-02-19 paper/paper.tex
Arithmetic mean (review 2)

@@
     as a green circle if $i$ is satisfiable, and a blue cross otherwise. The black line plots the
-    evolution of the mean number of search nodes when increasing the pattern edge probability from
+    evolution of the arithmetic mean number of search nodes when increasing the pattern edge probability from
     $0$ to $1$ in steps of $0.01$, using a larger sample size of $1,000$.}
@@
 
-The line plots mean search effort using the Glasgow solver: for sparse patterns, the problem is
+The line plots arithmetic mean search effort using the Glasgow solver: for sparse patterns, the problem is
 trivial; for dense patterns, proving unsatisfiability is not particularly difficult; and we see a

### 2018-02-21 paper/paper.tex
Cite count (review 2)

@@
 
-Subsequently, in a widely cited paper, \citet{DBLP:conf/sigmod/YanYH04} introduce an indexing system
+Subsequently, in a widely cited\footnote{685 citations according to Google Scholar in February
+2018.} paper, \citet{DBLP:conf/sigmod/YanYH04} introduce an indexing system
 called gIndex. Again, this system handles queries by first producing a set of candidates by

### 2018-02-23 paper/paper.tex
Calls per second
Bit of a reminder for non-CP people

@@
 a simple comparison of absolute performance between solvers; rather, we are looking for
-solver-independent patterns of difficulty. All experiments use a timeout of 1,000 seconds, which is
-enough for the Glasgow solver to solve nearly all instances (whose orders were selected with this
-timeout in mind), although we may slightly overestimate the proportion of unsatisfiable instances
-for extremely sparse or dense pattern graphs. The LAD, VF2, and VF3 solvers experienced many more
-failures with this timeout, so our picture of just how hard the hardest instances are with these
-solvers is less detailed.
+solver-independent patterns of difficulty. However, roughly speaking, on the small graphs used in
+the following sections, Glasgow manages around 300,000 recursive calls per second on our hardware,
+whilst LAD and VF2 perform around 1,000 and 70,000 recursive calls per second respectively.
+All experiments use a timeout of 1,000 seconds, which is enough for the Glasgow solver to solve
+nearly all instances (whose orders were selected with this timeout in mind), although we may
+slightly overestimate the proportion of unsatisfiable instances for extremely sparse or dense
+pattern graphs. The LAD, VF2, and VF3 solvers experienced many more failures with this timeout, so
+our picture of just how hard the hardest instances are with these solvers is less detailed.
 
@@
 
+The Glasgow and LAD solvers are driven by ordering heuristics: variable-ordering tells us which
+pattern vertex to branch on, whilst value-ordering tells the order in which we should try target
+vertices for that pattern vertex.
 Various general principles have been considered when designing variable- and value-ordering

### 2018-02-23 paper/paper.tex
Times

@@
 observed. As we show in \cref{figure:biiiig-data}, the \emph{hardest} instance for the AIDS dataset
-requires 287 recursive calls, and 149,268 of the instances can be solved without search.  For PDBS,
-the hardest problem requires 3,161 recursive calls (which occurs 20 times, due to duplicated
-queries). For both PCM and PPI, the hardest problem requires 23 recursive calls.  In other words,
-none of these instances are in any way computationally hard even for a simple domain-based
-algorithm, even before introducing stronger filtering, inference, or heuristics.  Furthermore it is
-certainly not the case that a sequential scan is infeasible as so many filter~/ verify papers claim.
-In contrast, with VF2, the hardest instance in PDBS requires 31,189 recursive calls; in AIDS,
-168,569 calls; in PCMS, 6,476,072 calls; and in PPI, 2,913. In other words, any exponential
-behaviour seen in these instances is purely down to the choice of subgraph isomorphism algorithm,
-and these instances are not inherently ``really hard''.
+requires 287 recursive calls, and 149,268 of the instances can be solved without search;
+furthermore, no instance individually requires more than 12ms. For both PCM and PPI, the hardest
+problem requires 23 recursive calls, and no instance takes more than 12ms and 282ms respectively.
+For PDBS, the hardest problem requires 3,161 recursive calls (which occurs 20 times, due to
+duplicated queries); however, Gecode struggles on these instances due to its general purpose data
+structures not scaling well to the large target graph sizes, and some of these instances can take as
+long as 30s to solve.
+
+In other words, none of these instances are in any way computationally hard even for a simple
+domain-based algorithm, even before introducing stronger filtering, inference, or heuristics, and
+the only difficulties arise from Gecode requiring the use of extensional constraints rather than the
+specialised adjacency data structures used by Glasgow and LAD.  Furthermore it is certainly not the
+case that a sequential scan is infeasible as so many filter~/ verify papers claim.  In contrast,
+with VF2, the hardest instance in PDBS requires 31,189 recursive calls; in AIDS, 168,569 calls; in
+PCMS, 6,476,072 calls; and in PPI, 2,913. In other words, any exponential behaviour seen in these
+instances is purely down to the choice of subgraph isomorphism algorithm, and these instances are
+not inherently ``really hard''.
 
@@
 first'' and to all-different. Such an approach may also be beneficial for huge target graphs, where
-having domains range over the entire target is impractical.
+having domains range over the entire target is impractical. Indeed, we saw that a raw constraint
+programming approach using Gecode suffered from relatively long runtimes on the very large but
+sparse PDBS instances, despite finding the instances very easy in terms of number of recursive
+calls. An algorithm which is capable of performing most of the reasoning of the Glasgow and LAD
+solvers, but without needing to store domains and adjacency matrices, could represent a major
+breakthrough.
 

### 2018-02-23 paper/paper.tex
Example

@@
 
-\usetikzlibrary{decorations, decorations.pathreplacing, calc, backgrounds}
+\usetikzlibrary{calc, positioning}
 
@@
 
+For example, in \cref{figure:cpsipexample}, we would have three variables $A$, $B$ and $C$, each of
+which has a domain $\{1, 2, 3, 4, 5, 6\}$. Suppose initially we guessed the assignment $A = 1$:
+inference would tell us that vertices adjacent to $A$ must be mapped to vertices adjacent to $1$,
+and so the value $6$ would be eliminated from the domains of both $B$ and $C$, whilst injectivity
+would eliminate $1$ from $B$ and $C$. Now suppose we then guessed $B = 2$: this would leave no
+suitable values for $C$, and so we would backtrack and try a new assignment for $B$. Guessing $B =
+3$ will similarly give a wipeout on $C$, but guessing $B = 4$ will leave the value $5$ in the domain
+of $C$, and we could then guess $C = 5$ to obtain a complete assignment.
+
+\begin{figure}[b]
+    \centering
+    \begin{tikzpicture}[scale=0.3,every node/.style={font=\footnotesize}]
+        \begin{scope}
+            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (Na) at (0, 0) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (Nb) at (-120:3) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb2, inner sep=0.5pt] (Nc) at (-60:3) {\phantom{0}};
+
+            \node [above = 0 of Na] {A};
+            \node [below = 0 of Nb] {B};
+            \node [below = 0 of Nc] {C};
+
+            \draw [ultra thick] (Na) -- (Nb);
+            \draw [ultra thick] (Nb) -- (Nc);
+            \draw [ultra thick] (Nc) -- (Na);
+        \end{scope}
+
+        \begin{scope}[xshift=9cm]
+            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (N1) at (0, 0) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (N2) at (180:3) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (N3) at (0:3) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (N4) at (-120:3) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (N5) at (-60:3) {\phantom{0}};
+            \node[draw, circle, fill=chromaygb5, inner sep=0.5pt] (N6) at ($(-60:3)+(3,0)$) {\phantom{0}};
+
+            \draw [ultra thick] (N1) -- (N2);
+            \draw [ultra thick] (N1) -- (N3);
+            \draw [ultra thick] (N1) -- (N4);
+            \draw [ultra thick] (N1) -- (N5);
+            \draw [ultra thick] (N4) -- (N5);
+            \draw [ultra thick] (N5) -- (N6);
+
+            \node [above = 0 of N1] {1};
+            \node [above = 0 of N2] {2};
+            \node [above = 0 of N3] {3};
+            \node [below = 0 of N4] {4};
+            \node [below = 0 of N5] {5};
+            \node [below = 0 of N6] {6};
+
+        \end{scope}
+    \end{tikzpicture}
+
+    \caption{An example used to illustrate the simple constraint programming model for the problem:
+    the pattern graph is on the left, and the target graph on the right.}
+    \label{figure:cpsipexample}
+\end{figure}
+
 This model can be extended with additional constraints. These allow more inference to be performed,
@@
 degree at least $d$, and so constraints can be posted which eliminate values corresponding to target
-vertices of low degree from domains corresponding to pattern vertices of high degree. Richer
-constraints involving neighbourhood degree sequences \citep{DBLP:journals/constraints/ZampelliDS10},
-neighbourhood difference constraints \citep{DBLP:journals/ai/Solnon10}, and distances and paths
+vertices of low degree from domains corresponding to pattern vertices of high degree. In the example
+in \cref{figure:cpsipexample}, this rule would eliminate the values $2$, $3$ and $6$ from every domain at
+the top of search. Richer constraints involving neighbourhood degree sequences
+\citep{DBLP:journals/constraints/ZampelliDS10}, neighbourhood difference constraints
+\citep{DBLP:journals/ai/Solnon10}, and distances and paths
 \citep{DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16} are also

### 2018-02-28 paper/paper.tex
Don't cite that
Ennicenify

@@
 this behaviour helps us to design better search strategies, by improving variable- and
-value-ordering heuristics---this is the second contribution of this paper.
+value-ordering heuristics (that is, by improving the order in which pattern and target vertices are
+selected for guessed assignments). This is the second contribution of this paper.
 
@@
 \emph{variable-ordering} heuristics, which select which unassigned variable to branch on, and
-\emph{value-ordering} heuristics, which select which value to try giving that variable. That is,
-variable-ordering selects a pattern vertex, whilst value-ordering tells us the order in which to try
-different candidate target vertices for that pattern vertex. If at any point during search a
-\emph{domain wipeout} occurs (that is, if every value is removed from any domain), the search will
-backtrack immediately.
-
-In the case of subgraph isomorphism, the basic constraint model is to have a domain for each pattern
-vertex, with each domain being the set of target vertices. An \emph{all-different} constraint ensures
-injectivity, whilst the adjacency rules can be expressed as extensional constraints by listing
-allowed tuples.  With this model, the backtracking search builds up partial assignments of pattern
-vertices to target vertices, until either a complete assignment is reached (and the instance is
-\emph{satisfiable}), or search establishes that no such assignment can exist (and it is
-\emph{unsatisfiable}).
+\emph{value-ordering} heuristics, which select which value to try giving that variable.
+
+In the case of subgraph isomorphism, the basic constraint model is to have a variable for each
+pattern vertex, with each variable's domain being the set of target vertices. An
+\emph{all-different} constraint ensures injectivity, whilst the adjacency rules can be expressed as
+extensional constraints by listing allowed tuples.  With this model, the backtracking search builds
+up partial assignments of pattern vertices to target vertices, until either a complete assignment is
+reached (and the instance is \emph{satisfiable}), or search establishes that no such assignment can
+exist (and it is \emph{unsatisfiable}).  Hence, the role of the variable-ordering heuristic for
+subgraph isomorphism is to select a pattern vertex, whilst value-ordering tells us the order in
+which to try different candidate target vertices for that pattern vertex. If at any point during
+search a \emph{domain wipeout} occurs (that is, if every value is removed from any domain), the
+search will backtrack immediately.
 
@@
 So far we have looked at the non-induced problem, where extra edges in the target graph are
-permitted.  In the first five rows of \cref{figure:induced} we repeat our experiments, now finding
-induced isomorphisms. With a pattern of order 10, we get two independent phase transitions: the
+permitted.  In the first five rows of \cref{figure:induced} we repeat our experiments, now searching
+for induced isomorphisms. With a pattern of order 10, we get two independent phase transitions: the
 bottom right halves of the plots resemble the non-induced results, and the top left halves are close
@@
 $\kappa > 1$ is overconstrained, and is likely to be made of unsatisfiable instances. Empirically,
-problem instances from an ensemble with $\kappa$ close to 1 are hard, and problems drawn from an
-ensemble where $\kappa$ is very small or very large are usually easy.
+problem instances from an ensemble with $\kappa$ close to 1 are hard, and problem instances drawn
+from an ensemble where $\kappa$ is very small or very large are usually easy.
 
@@
 However, by far the biggest problem is in graph databases. The following section shows how
-widespread use of weak subgraph isomorphism algorithms has not just led to overly pessimistic
-conclusions regarding performance, but has misdirected the design of larger systems.
+widespread use of certain non-constraint-based subgraph isomorphism algorithms has not just led to
+overly pessimistic conclusions regarding performance, but has misdirected the design of larger
+systems.
 
@@
 led to the development of systems which perform extensive computations on the target graphs
-beforehand, in the hopes of reducing the response times for individual pattern queries. The most
+beforehand, in the hope of reducing the response times for individual pattern queries. The most
 popular of these strategies is a form of indexing which is often named \emph{filter~/ verify}.
@@
 $\left|C_q\right|$, and $T_{\mathit{iso\_test}}$ is the cost of a subgraph isomorphism test. They
-reason that since isomorphism testing is \NP-complete, by making $\left|C_q\right|$ as small as
+reason that since subisomorphism testing is \NP-complete, by making $\left|C_q\right|$ as small as
 possible, the query response time will be reduced.  (Importantly, it is \emph{not} the time taken to
@@
 This equation is repeated and expanded upon by \citet{DBLP:journals/tods/YanYH05} to explicitly
-separate I/O and isomorphism testing costs into $T_{\mathit{io}}$ and $T_{\mathit{iso\_test}}$
+separate I/O and subisomorphism testing costs into $T_{\mathit{io}}$ and $T_{\mathit{iso\_test}}$
 respectively, obtaining a query response time of \[
@@
 must have a very small proof of unsatisfiability. These instances \emph{should} be trivial with any
-reasonably good subgraph isomorphism algorithm. Thus, all filtering \emph{should} be doing is eliminating the startup
+domain-based subgraph isomorphism algorithm. Thus, all filtering \emph{should} be doing is eliminating the startup
 costs of a trivial subgraph isomorphism call. The fact that filtering was successful empirically
@@
 \citet{DBLP:conf/vldb/ZhaoYY07}. Again, the work has a focus on reducing the candidate set size
-through indexing. The authors appear to believe that the cost of the isomorphism test is not a major
+through indexing. The authors appear to believe that the cost of the subisomorphism test is not a major
 factor in influencing the result, and focus on the remaining terms in the equation. They use the
@@
 there being two red vertices in the pattern but only one in the target.\footnote{Although
-interestingly, this algorithm effectively does forward-checking and has a variable-ordering
-heuristic, before these concepts appeared in the constraints literature.}
+interestingly, this algorithm effectively maintains (binary) arc consistency, and has a
+variable-ordering heuristic, before these concepts appeared in the constraints literature.}
 
@@
 
-These beliefs are not entirely unfounded: none of the isomorphism algorithms considered in these
+These beliefs are not entirely unfounded: none of the subisomorphism algorithms considered in these
 papers will immediately detect if a pattern contains two red vertices, whilst the target graph
@@
     \item Domain-based heuristics are widely studied and well-understood
-        \citep{DBLP:conf/ecai/SmithG98}, and unlike VF2's adjacency branching rules, provide
+        \citep{DBLP:conf/cp/GentMPSW96}, and unlike VF2's adjacency branching rules, provide
         discrimination even on dense and low-diameter graphs. Picking from small domains dynamically
@@
 
-    \item Domain-based algorithms automatically combines features. For example, a pattern~/ target
+    \item Domain-based algorithms automatically combine features. For example, a pattern~/ target
         pair may have matching label features, and matching degree features, but if the only red
@@
 general \citep{DBLP:journals/ai/GomesS01}, and the approach by \citet{DBLP:conf/sls/BattitiM07} for
-subgraph isomorphism in particular, is not noted (and nor are portfolios mentioned when they say
-that ``using machine learning models to predict which version \ldots to employ per query is of high
-interest'').  The evidence so far in this paper suggests that we should be wary. It is certainly
-true that there are some instances that all algorithms find hard, and there are good theoretical
-reasons to believe that these instances genuinely are really hard.  Furthermore, algorithm
-portfolios are also well-known to be a successful technique, and are beneficial even with modern
-subgraph isomorphism algorithms \citep{DBLP:conf/lion/KotthoffMS16}.  However, in
-\cref{section:labelled} we saw that there are many instances that VF2 finds hard that should not be
-hard, and that are not hard for other algorithms.  If permuting graphs sometimes addresses the
-difficulty of some of these instances, then it is likely that they are not genuinely hard instances
-at all, and perhaps it would be better simply to start using an algorithm with domain tracking and
-domain-based ordering heuristics, rather than assembling a portfolio of weak algorithms in the hopes
-that at least one of them will often get lucky.
+subgraph isomorphism in particular, is not noted. (Nor is the portfolios literature referenced when
+they say that ``using machine learning models to predict which version \ldots to employ per query is
+of high interest'').  The evidence so far in this paper suggests that we should carefully examine
+the reasons behind these speedups. It is certainly true that there are some instances that all
+algorithms find hard, and there are good theoretical reasons to believe that these instances
+genuinely are really hard.  Furthermore, algorithm portfolios are also well-known to be a successful
+technique, and are beneficial even with modern subgraph isomorphism algorithms
+\citep{DBLP:conf/lion/KotthoffMS16}.  However, in \cref{section:labelled} we saw that there are many
+instances that VF2 finds hard that should not be hard, and that are not hard for other algorithms.
+If permuting graphs sometimes addresses the difficulty of some of these instances, then it is likely
+that they are not genuinely hard instances at all. Perhaps then, it would be better simply to use an
+algorithm with domain tracking and domain-based ordering heuristics.
 
@@
 this section. What if we modify our solver so that it does not use ``smallest domain first'', and
-does not detect domain wipeouts until attempting to branch on an empty variable?  Suddenly three of
+does not detect domain wipeouts until attempting to branch on an empty variable?  In this case, three of
 the four datasets become extremely hard, with many instances now timing out after making hundreds of
@@
 some queries.  Furthermore, slight changes to the input vertex ordering can now make some of these
-``hard'' instances easy again.  This backs up our suspicions: the apparent success of
+``hard'' instances easy again.  This verifies our intuition: the apparent success of
 \citet{DBLP:conf/edbt/KatsarouNT17}'s technique is due to the ease of sometimes getting better
-results out of a poor algorithm. Claiming huge improvements from an algorithm portfolio consisting
-only of variations of poor algorithms does not demonstrate a genuine improvement over the state of
-the art.
+results out of a poor algorithm. Obtaining huge improvements from an algorithm portfolio consisting
+only of variations of poor algorithms is not sufficient to demonstrate a genuine improvement over
+the state of the art.
 
@@
 systems involving subgraph isomorphism has been misdirected due to a lack of familiarity with the
-best existing algorithms for subgraph isomorphism, and with more broadly-applicable lessons on
-algorithm design from the constraint programming literature.
+constraint programming literature.
 
@@
 far greater practical impact than it would have on easier graphs, since the saved time can be of
-several hours or even days'', but mistakenly believe that being larger and slightly denser is what
-makes an instance harder. To evaluate VF3, they introduce a new dataset consisting entirely of
+several hours or even days''. To evaluate VF3, they introduce a new dataset consisting entirely of
 satisfiable instances created by taking a connected subgraph of a randomly generated graph and
@@
 the enumeration problem on these new instances.  We urge a more cautious interpretation of these
-results: as we saw in \cref{section:induced}, being genuinely hard (rather than simply larger) is a
-subtle property, and it is not safe to say that an instance is hard simply because several
-minor variations of a particular kind of algorithm find it hard. More care must thus be taken to
-avoid accidentally measuring which algorithm scales best to large, easy instances due to it using
-the simplest inference routines. Indeed, we saw in \cref{figure:induced} that VF3 performs
-\emph{worse} than VF2 on genuinely harder instances, and additionally behaves erratically in parts
-of the parameter space that every other solver finds easy; finally, an induced analogue of
-\cref{figure:labels} confirms that VF3 also still struggles with many obviously unsatisfiable
-instances.
+results: as we saw in \cref{section:induced}, being genuinely hard (rather than simply larger or
+denser) is a subtle property, and care must be taken when empirically evaluating algorithms. In
+particular, if one only considers large, easy instances, then the simpler the algorithm, the better
+it scales. Indeed, we saw in \cref{figure:induced} that VF3 performs \emph{worse} than VF2 on
+genuinely harder instances, and additionally behaves erratically in parts of the parameter space
+that every other solver finds easy; finally, an induced analogue of \cref{figure:labels} confirms
+that VF3 also still struggles with many obviously unsatisfiable instances.
 
@@
 hardness of \NP-complete problems, members of the graph databases community have reached the mistaken
-conclusion that extensive research into supporting techniques is important, whereas really what
-they are doing is working around some but not all of the weaknesses in their chosen algorithms for subgraph
-isomorphism. With this new understanding of what does and does not make subgraph
-isomorphism hard, it is time for a radical rethink of how graph database systems work.
+conclusion that extensive research into supporting techniques is important. With this new
+understanding of what does and does not make subgraph isomorphism hard, it is time for a radical
+rethink of how graph database systems work.
 
@@
 \citep{DBLP:conf/ictai/DentM94,DBLP:conf/cp/BacchusG95,o:Dent96,DBLP:conf/ecai/LarrosaM98} avoids
-constructing every domain upfront, although adopting this require alternatives to ``smallest domain
+constructing every domain upfront, although adopting this may require alternatives to ``smallest domain
 first'' and to all-different. Such an approach may also be beneficial for huge target graphs, where

### 2018-03-01 paper/paper.tex
Remove unnecessary commands

@@
 \Crefname{section}{Section}{Sections}
-
-\newcommand{\citet}[1]{\citeA{#1}}
-\newcommand{\citep}[1]{\cite{#1}}
 
@@
 so there are no ``extra edges'' in the copy of the pattern that we find. We illustrate both variants
-in \cref{figure:sip}.  Although these problems are \NP-complete \citep{DBLP:books/fm/GareyJ79}, modern subgraph
+in \cref{figure:sip}.  Although these problems are \NP-complete \cite{DBLP:books/fm/GareyJ79}, modern subgraph
 isomorphism algorithms based upon constraint programming techniques can handle problem instances
@@
 graph
-\citep{DBLP:journals/ai/Solnon10,DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16},
+\cite{DBLP:journals/ai/Solnon10,DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16},
 and subgraph isomorphism is used successfully in application areas including computer vision
-\citep{DBLP:journals/cviu/DamiandSHJS11,DBLP:journals/pr/SolnonDHJ15}, biochemistry
-\citep{o:10.1371/journal.pone.0076911,DBLP:conf/gbrpr/CarlettiFV15}, and pattern recognition
-\citep{DBLP:journals/ijprai/ConteFSV04}.  A labelled version of subgraph isomorphism
+\cite{DBLP:journals/cviu/DamiandSHJS11,DBLP:journals/pr/SolnonDHJ15}, biochemistry
+\cite{o:10.1371/journal.pone.0076911,DBLP:conf/gbrpr/CarlettiFV15}, and pattern recognition
+\cite{DBLP:journals/ijprai/ConteFSV04}.  A labelled version of subgraph isomorphism
 is also one of the key components in supporting complex queries in graph databases, which we discuss in
@@
 in previous evaluations come from common datasets
-\citep{DBLP:journals/prl/SantoFSV03,DBLP:journals/constraints/ZampelliDS10}, which were generated by
+\cite{DBLP:journals/prl/SantoFSV03,DBLP:journals/constraints/ZampelliDS10}, which were generated by
 taking a random subgraph of a random (Erd\H{o}s-R\'enyi, scale-free, bounded degree, or mesh) graph
-and permuting the vertices. Such instances are guaranteed to be satisfiable---\citet{o:AntonO09}
+and permuting the vertices. Such instances are guaranteed to be satisfiable---\citeA{o:AntonO09}
 exploited this property to create large sets of random satisfiable boolean satisfiability instances.
@@
 model. This has led to bias in algorithm design, to the extent that some proposed techniques, such
-as those of \citet{DBLP:conf/sls/BattitiM07}, will \emph{only} work on satisfiable instances.
+as those of \citeA{DBLP:conf/sls/BattitiM07}, will \emph{only} work on satisfiable instances.
 
@@
 This work builds upon the phase transition phenomena observed in satisfiability and graph colouring
-problems first described by \citet{DBLP:conf/ijcai/CheesemanKT91} and
-\citet{DBLP:conf/aaai/MitchellSL92}. For subgraph isomorphism in Erd\H{o}s-R\'enyi random graphs, we
+problems first described by \citeA{DBLP:conf/ijcai/CheesemanKT91} and
+\citeA{DBLP:conf/aaai/MitchellSL92}. For subgraph isomorphism in Erd\H{o}s-R\'enyi random graphs, we
 identify three relevant control parameters: we can independently alter the edge probability of the
@@
 mixed integer solvers, and under reduction to the clique problem. Interestingly, the constrainedness
-measure proposed by \citet{DBLP:conf/aaai/GentMPW96} \emph{does} predict this difficult region---the
+measure proposed by \citeA{DBLP:conf/aaai/GentMPW96} \emph{does} predict this difficult region---the
 third contribution of this paper is to use these instances to provide evidence in favour of
@@
 database systems), richer behaviour emerges, particularly when moving away from a uniform labelling
-scheme, and the popular VF2 algorithm \citep{DBLP:journals/pami/CordellaFSV04} behaves much worse
+scheme, and the popular VF2 algorithm \cite{DBLP:journals/pami/CordellaFSV04} behaves much worse
 than two more recent constraint-based algorithms
-\citep{DBLP:journals/ai/Solnon10,DBLP:conf/cp/McCreeshP15} in certain cases. In
+\cite{DBLP:journals/ai/Solnon10,DBLP:conf/cp/McCreeshP15} in certain cases. In
 \cref{section:labelled} we take a close look at these cases, and argue that they \emph{should} be
@@
 
-Parts of this paper were presented at IJCAI \citep{DBLP:conf/ijcai/McCreeshPT16}; the first three
+Parts of this paper were presented at IJCAI \cite{DBLP:conf/ijcai/McCreeshPT16}; the first three
 sections have been expanded, whilst
 \cref{section:regular,section:labelled,section:filterverify} are new material. Most of this work
-also appeared in a PhD thesis \citep{o:McCreesh17}.
+also appeared in a PhD thesis \cite{o:McCreesh17}.
 
@@
 edge between each distinct unordered pair of vertices with independent probability $p$
-\citep{o:Gilbert59,o:ErdosR59}.\footnote{Note that elsewhere, this same name sometimes instead
+\cite{o:Gilbert59,o:ErdosR59}.\footnote{Note that elsewhere, this same name sometimes instead
 refers to a graph $G(n, E)$ chosen uniformly from among all graphs with $n$ vertices and exactly $E$
@@
 RAM, running Scientific Linux release 6.7. We will be working with four subgraph isomorphism
-solvers: the Glasgow solver \citep{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16}; LAD
-\citep{DBLP:journals/ai/Solnon10}; VF2 \citep{DBLP:journals/pami/CordellaFSV04}; and VF3
-\citep{DBLP:conf/gbrpr/CarlettiFSV17}. Each was compiled using GCC 4.9 with the ``-O3'' option.
+solvers: the Glasgow solver \cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16}; LAD
+\cite{DBLP:journals/ai/Solnon10}; VF2 \cite{DBLP:journals/pami/CordellaFSV04}; and VF3
+\cite{DBLP:conf/gbrpr/CarlettiFSV17}. Each was compiled using GCC 4.9 with the ``-O3'' option.
 The Glasgow, LAD, and VF2 solvers support both non-induced and induced isomorphisms (although
@@
 the top of search. Richer constraints involving neighbourhood degree sequences
-\citep{DBLP:journals/constraints/ZampelliDS10}, neighbourhood difference constraints
-\citep{DBLP:journals/ai/Solnon10}, and distances and paths
-\citep{DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16} are also
+\cite{DBLP:journals/constraints/ZampelliDS10}, neighbourhood difference constraints
+\cite{DBLP:journals/ai/Solnon10}, and distances and paths
+\cite{DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16} are also
 possible.
@@
 cost of individual instances, as points. The behaviour we observe looks remarkably similar to random
-3SAT problems---compare, for example, Figure~1 of \citet{DBLP:journals/cacm/Leyton-BrownHHX14}. In
+3SAT problems---compare, for example, Figure~1 of \citeA{DBLP:journals/cacm/Leyton-BrownHHX14}. In
 particular, satisfiable instances tend to be easier, but show greater variation than unsatisfiable
-instances, and there are exceptionally hard satisfiable instances \citep{DBLP:conf/cp/SmithG97}.
+instances, and there are exceptionally hard satisfiable instances \cite{DBLP:conf/cp/SmithG97}.
 
@@
 similar behaviour is seen with random constraint satisfaction problems
-\citep{DBLP:journals/ai/SmithD96}.
+\cite{DBLP:journals/ai/SmithD96}.
 
@@
 heuristics for backtracking search algorithms---one of these is to try to maximise the expected
-number of solutions inside any subproblem considered during search \citep{DBLP:conf/cp/GentMPSW96}.
+number of solutions inside any subproblem considered during search \cite{DBLP:conf/cp/GentMPSW96}.
 This is usually done by cheaper surrogates, rather than direct calculation. When branching, both LAD
@@
 towards the top left), and if we are in the top left we should use the complement instead. This
-goes against a suggestion by \citet{DBLP:conf/aaai/Walsh98} that switching heuristics based upon an estimate of
+goes against a suggestion by \citeA{DBLP:conf/aaai/Walsh98} that switching heuristics based upon an estimate of
 the solubility of the problem may offer good performance.
@@
 adjacent vertices, and that non-adjacent vertices are mapped to non-adjacent vertices. We use the
-Clasp solver \citep{DBLP:journals/aicom/GebserKKOSS11} version 3.1.3 to solve the pseudo-boolean instances.  The
+Clasp solver \cite{DBLP:journals/aicom/GebserKKOSS11} version 3.1.3 to solve the pseudo-boolean instances.  The
 instances that are hard for the Glasgow solver remain hard for the PB solver, including instances
 inside the central region, and the easy satisfiable instances remain easy. Similar results are seen
-with the Glucose SAT solver \citep{o:glucose} using a direct encoding.  We also show an integer
+with the Glucose SAT solver \cite{o:glucose} using a direct encoding.  We also show an integer
 program encoding: the Gurobi solver is only able to solve some of the trivial satisfiable instances,
@@
 The \emph{association graph encoding} of a subgraph isomorphism problem is a reduction to the clique
-decision problem. \citet{DBLP:conf/cp/McCreeshNPS16} describe and study this approach in more
+decision problem. \citeA{DBLP:conf/cp/McCreeshNPS16} describe and study this approach in more
 detail. Briefly, the association graph is constructed by creating a new graph with a vertex for each
@@
 $t_2$. A clique of size equal to the order of the pattern graph exists in the association graph if
-and only if the problem is satisfiable \citep{o:Levi73}. We used this encoding with an
-implementation of \citet{DBLP:journals/cor/SegundoRJ11}'s bit-parallel maximum clique algorithm
+and only if the problem is satisfiable \cite{o:Levi73}. We used this encoding with an
+implementation of \citeA{DBLP:journals/cor/SegundoRJ11}'s bit-parallel maximum clique algorithm
 BBMC, modified to solve the decision problem rather than the optimisation problem.  Again, our
@@
 phase transition concept, and to generalise hardness parameters across different combinatorial
-problems \citep{DBLP:conf/aaai/GentMPW96}. An ensemble of problem instances with $\kappa < 1$ is
+problems \cite{DBLP:conf/aaai/GentMPW96}. An ensemble of problem instances with $\kappa < 1$ is
 said to be underconstrained, and is likely to be made up of satisfiable instances; an ensemble with
@@
 \cref{figure:kr} we recreate \cref{figure:non-induced}, using the NetworkX implementation
-\citep{o:NetworkX} of the \citet{DBLP:journals/cpc/StegerW99} algorithm to generate regular graphs.
+\cite{o:NetworkX} of the \citeA{DBLP:journals/cpc/StegerW99} algorithm to generate regular graphs.
 The $x$-axes range from degree 0 to degree $n - 1$ where $n \in \{ 10, 20, 30 \}$ is the number of
@@
 fractional. The gamma function $\Gamma(n)$ is equal to $(n - 1)!$ for integers $n \ge 1$
-\citep{Davis59}, but is also defined for positive real numbers, obeying the identify $\Gamma(x + 1)
+\cite{Davis59}, but is also defined for positive real numbers, obeying the identify $\Gamma(x + 1)
 = x\Gamma(x)$. By noting that $t^{\underline{p}} = \nicefrac{t!}{(t - p)!}$, we may obtain a
@@
 instance generators
-\citep{DBLP:journals/constraints/AchlioptasMKSKK01,DBLP:journals/constraints/GentMPSW01}.
+\cite{DBLP:journals/constraints/AchlioptasMKSKK01,DBLP:journals/constraints/GentMPSW01}.
 
@@
 But does this cause problems in practice? The literature suggests that it does. For example,
-\citet{o:Gromping14} uses VF2 in a package for the R statistics language, and states
+\citeA{o:Gromping14} uses VF2 in a package for the R statistics language, and states
 \begin{displayquote}``There are (not so many) instances for which creation of a clear design is
@@
     very fast in ruling out impossible matches, where VF2 took a long time.''\end{displayquote}
-Similarly, \citet{o:Murray12} uses VF2 inside a compiler, and observes extremely variable
+Similarly, \citeA{o:Murray12} uses VF2 inside a compiler, and observes extremely variable
 (and prohibitively high) compile times in some cases due to the expense of subgraph isomorphism
@@
 
-An early graph database system by \citet{DBLP:conf/pods/ShashaWG02} uses a filtering heuristic to
+An early graph database system by \citeA{DBLP:conf/pods/ShashaWG02} uses a filtering heuristic to
 eliminate unsatisfiable instances based upon simple structural elements. It is not clear whether the
@@
 experiments do not answer whether the indexing is effective. However, the work was influenced by a commercial
-graph database system, whose documentation \citep{o:Daylight} states that indexing is
+graph database system, whose documentation \cite{o:Daylight} states that indexing is
 used to minimise disk accesses.
@@
 Subsequently, in a widely cited\footnote{685 citations according to Google Scholar in February
-2018.} paper, \citet{DBLP:conf/sigmod/YanYH04} introduce an indexing system
+2018.} paper, \citeA{DBLP:conf/sigmod/YanYH04} introduce an indexing system
 called gIndex. Again, this system handles queries by first producing a set of candidates by
@@
 
-This equation is repeated and expanded upon by \citet{DBLP:journals/tods/YanYH05} to explicitly
+This equation is repeated and expanded upon by \citeA{DBLP:journals/tods/YanYH05} to explicitly
 separate I/O and subisomorphism testing costs into $T_{\mathit{io}}$ and $T_{\mathit{iso\_test}}$
@@
 Moving forwards, the (simpler form of the) query response time equation is repeated by
-\citet{DBLP:conf/vldb/ZhaoYY07}. Again, the work has a focus on reducing the candidate set size
+\citeA{DBLP:conf/vldb/ZhaoYY07}. Again, the work has a focus on reducing the candidate set size
 through indexing. The authors appear to believe that the cost of the subisomorphism test is not a major
@@
 
-The equation is also used by \citet{DBLP:conf/icde/JiangWYZ07}, who claim that ``usually the
+The equation is also used by \citeA{DBLP:conf/icde/JiangWYZ07}, who claim that ``usually the
 verification time dominates the Query Response Time [s]ince the computational complexity of
@@
 set''.\end{displayquote} This claim becomes understandable when one examines the subgraph
-isomorphism algorithm chosen for the verification step \citep{DBLP:journals/jacm/Ullmann76}: as the
+isomorphism algorithm chosen for the verification step \cite{DBLP:journals/jacm/Ullmann76}: as the
 algorithm predates techniques like generalised arc-consistent all-different propagation
-\citep{DBLP:conf/aaai/Regin94}, it cannot immediately detect unsatisfiability in simple cases like
+\cite{DBLP:conf/aaai/Regin94}, it cannot immediately detect unsatisfiability in simple cases like
 there being two red vertices in the pattern but only one in the target.\footnote{Although
@@
 
-Without using the equation, \citet{DBLP:conf/sigmod/ChengKNL07} state that since subgraph
+Without using the equation, \citeA{DBLP:conf/sigmod/ChengKNL07} state that since subgraph
 isomorphism is \NP-complete, processing by a sequential scan is infeasible. They introduce new
 filtering techniques to try to avoid the subgraph isomorphism step. A similar claim is made by
-\citet{DBLP:conf/icde/ZhangHY07} in an introduction of another indexing technique: ``obviously it is
+\citeA{DBLP:conf/icde/ZhangHY07} in an introduction of another indexing technique: ``obviously it is
 inefficient to perform a sequential scan on every graph in the database, because the subgraph
@@
 
-Muddying the waters slightly, a survey by \citet{DBLP:journals/datamine/HanCXY07} states that
+Muddying the waters slightly, a survey by \citeA{DBLP:journals/datamine/HanCXY07} states that
 ``large volumes of data'' (not \NP-completeness) is the reason for using indexing in these systems.
 However, in a description of a system tailored to biological networks,
-\citet{DBLP:conf/edbt/ZhangLY09} state that \begin{displayquote}``Since the size of the raw database
+\citeA{DBLP:conf/edbt/ZhangLY09} state that \begin{displayquote}``Since the size of the raw database
     graph is small, it can be easily fit in the main memory.  However, the query (matching) time
 will be very long due to the \NP-hard complexity.''\end{displayquote} They suggest that indexing is
-a way of avoiding this. Similarly, \citet{DBLP:journals/pvldb/ZhaoH10} argue that ``the graph query
+a way of avoiding this. Similarly, \citeA{DBLP:journals/pvldb/ZhaoH10} argue that ``the graph query
 problem is hard in that \ldots it requires subgraph isomorphism checking \ldots which has proven to
@@
 
-\citet{DBLP:conf/icdcs/CaoYWRL11} propose a privacy-preserving cloud graph database system using
+\citeA{DBLP:conf/icdcs/CaoYWRL11} propose a privacy-preserving cloud graph database system using
 filter~/ verify, stating that ``checking subgraph isomorphism is \NP-complete, and therefore it is
 infeasible to employ such a costly solution'' which simply checks every graph for a match.
-\citet{DBLP:journals/tkde/WangWYY12} look at indexing large sparse graphs. They state that ``because
+\citeA{DBLP:journals/tkde/WangWYY12} look at indexing large sparse graphs. They state that ``because
 subgraph isomorphism is an \NP-complete problem, a filter-and-verification method is usually
 employed to speed up the search efficiency of graph similarity matching over a graph set''.
-Similarly, after reviewing the literature, \citet{DBLP:journals/vldb/YuanM13} argue that subgraph
+Similarly, after reviewing the literature, \citeA{DBLP:journals/vldb/YuanM13} argue that subgraph
 querying is costly because it is \NP-complete, and that indices can improve the performance of graph
 database queries. Again, new indexing techniques are introduced. Subsequently,
-\citet{DBLP:journals/pvldb/KatsarouNT15} perform a comprehensive comparison of graph database
+\citeA{DBLP:journals/pvldb/KatsarouNT15} perform a comprehensive comparison of graph database
 filtering techniques. They state that performing a query against each graph in the dataset
 ``obviously does not scale, as subgraph isomorphism is \NP-complete''. In describing a system which
-returns a special subset of graphs which match a query, \citet{DBLP:journals/tkde/ZhengLZHZ16}
+returns a special subset of graphs which match a query, \citeA{DBLP:journals/tkde/ZhengLZHZ16}
 suggest that it is ``\NP-hard to check the graph isomorphism'' (meaning subgraph isomorphism), and
@@
 build, and they suggest that Ullmann's algorithm and VF2 are state of the art for verification.
-\citet{DBLP:journals/www/PengZ0LZ16} state that ``obviously, it is impossible to employ some
+\citeA{DBLP:journals/www/PengZ0LZ16} state that ``obviously, it is impossible to employ some
 subgraph isomorphism algorithm, such as Ullmann or VF2'', and argue that ``in order to speed up
-query processing'', they need to create indices.  And \citet{DBLP:conf/isda/AzaouziR16} argue that
+query processing'', they need to create indices.  And \citeA{DBLP:conf/isda/AzaouziR16} argue that
 ``since the subgraph isomorphism test is expensive, checking all graphs of a large database can be
@@
 The filter~/ verify paradigm also influences other research. For example, it is used by
-\citet{DBLP:conf/icde/TianP08} in an approximate subgraph searching system: they state that
+\citeA{DBLP:conf/icde/TianP08} in an approximate subgraph searching system: they state that
 Ullmann's algorithm ``is prohibitively expensive for querying against [a] database with a large
 number of graphs'', and that indices are used ``to filter out graphs that do not match the query''.
-More recently, \citet{DBLP:journals/pvldb/YuanMG13} continue a line of supergraph search work, again
+More recently, \citeA{DBLP:journals/pvldb/YuanMG13} continue a line of supergraph search work, again
 using a filtering step to avoid subgraph isomorphism calls (in the opposite direction, so queries
-are now target graphs). \citet{DBLP:journals/tkde/HongZLY15} look at graph database subgraph
+are now target graphs). \citeA{DBLP:journals/tkde/HongZLY15} look at graph database subgraph
 isomorphism with an additional set-similarity constraint, and state that Ullmann's algorithm and VF2
@@
 of space.  There is also research into maintaining indices when the set of target graph changes: for
-example \citet{DBLP:conf/sigmod/YuanMYG15} look at algorithms for updating graph indices.
+example \citeA{DBLP:conf/sigmod/YuanMYG15} look at algorithms for updating graph indices.
 And describing a system for reusing results of queries which are sub- or super-graphs of previous
-queries, \citet{DBLP:conf/edbt/WangNT16} state that querying is a ``very costly operation as it
+queries, \citeA{DBLP:conf/edbt/WangNT16} state that querying is a ``very costly operation as it
 entails the \NP-complete problem of subgraph isomorphism'', and place ``an emphasis on the number of
@@
 contains only one. This kind of flaw \emph{should} be picked up at the top of search by an
-all-different propagator \citep{DBLP:conf/aaai/Regin94}.  However, as
-\citet{DBLP:journals/pvldb/KatsarouNT15} note, VF2 \citep{DBLP:journals/pami/CordellaFSV04} or a
+all-different propagator \cite{DBLP:conf/aaai/Regin94}.  However, as
+\citeA{DBLP:journals/pvldb/KatsarouNT15} note, VF2 \cite{DBLP:journals/pami/CordellaFSV04} or a
 similar algorithm is the usual subgraph isomorphism algorithm of choice for graph database systems,
@@
 have been considered, albeit not with algorithms strong enough to establish all-difference. For
-example, \citet{DBLP:journals/pvldb/ShangZLY08} propose an algorithm which makes use of the
-frequency of various features to guide search; \citet{DBLP:journals/pvldb/LeeHKL12} determine
+example, \citeA{DBLP:journals/pvldb/ShangZLY08} propose an algorithm which makes use of the
+frequency of various features to guide search; \citeA{DBLP:journals/pvldb/LeeHKL12} determine
 experimentally that this technique tends to be very effective, even on families of graphs for which
@@
 that these are high-quality datasets or that the associated queries are sensible, merely that
-following the example of \citet{o:10.1371/journal.pone.0076911}, they are widely used. Each dataset
+following the example of \citeA{o:10.1371/journal.pone.0076911}, they are widely used. Each dataset
 has labels on vertices (but not on edges), and we will look at the non-induced problem.
@@
         edges.
-    \item The PDBS dataset \citep{o:HeLCBBSKMR02} contains 30 labelled graphs representing DNA, RNA,
+    \item The PDBS dataset \cite{o:HeLCBBSKMR02} contains 30 labelled graphs representing DNA, RNA,
         and proteins, each duplicated twenty times to give a dataset of 600 graphs. These
@@
         meaning.
-    \item The PCM dataset \citep{o:doi:10.1093/bioinformatics/btr163} contains 50 protein contact
+    \item The PCM dataset \cite{o:doi:10.1093/bioinformatics/btr163} contains 50 protein contact
         maps, each duplicated four times to give a dataset of 200 graphs. These graphs have under a
@@
     \item Domain-based heuristics are widely studied and well-understood
-        \citep{DBLP:conf/cp/GentMPSW96}, and unlike VF2's adjacency branching rules, provide
+        \cite{DBLP:conf/cp/GentMPSW96}, and unlike VF2's adjacency branching rules, provide
         discrimination even on dense and low-diameter graphs. Picking from small domains dynamically
@@
         considerable.  Indices proposed in the literature often take many hours to build, and can
-        consume much more space than the original graphs \citep{DBLP:journals/tkde/HongZLY15}.
+        consume much more space than the original graphs \cite{DBLP:journals/tkde/HongZLY15}.
 \end{itemize}
@@
 
-\citet{DBLP:conf/edbt/KatsarouNT17} present approaches to improve the run times for what they call
+\citeA{DBLP:conf/edbt/KatsarouNT17} present approaches to improve the run times for what they call
 ``straggler queries''---the small subset of queries that they observe taking much longer than others
@@
 is not noted, and no consideration is given to dynamic ordering heuristics such as ``smallest domain
-first'' \citep{DBLP:journals/ai/HaralickE80}.  Of the orderings proposed, those involving placing
+first'' \cite{DBLP:journals/ai/HaralickE80}.  Of the orderings proposed, those involving placing
 rare labels first are effectively approximations to a static ``smallest domain at top of search
@@
 
-\citet{DBLP:conf/edbt/KatsarouNT17}'s second claim is that ``different
+\citeA{DBLP:conf/edbt/KatsarouNT17}'s second claim is that ``different
 algorithms find different queries hard''. To address this, they run many subgraph isomorphism
 algorithms and input permutations in parallel; the extensive literature on parallel portfolios in
-general \citep{DBLP:journals/ai/GomesS01}, and the approach by \citet{DBLP:conf/sls/BattitiM07} for
+general \cite{DBLP:journals/ai/GomesS01}, and the approach by \citeA{DBLP:conf/sls/BattitiM07} for
 subgraph isomorphism in particular, is not noted. (Nor is the portfolios literature referenced when
@@
 technique, and are beneficial even with modern subgraph isomorphism algorithms
-\citep{DBLP:conf/lion/KotthoffMS16}.  However, in \cref{section:labelled} we saw that there are many
+\cite{DBLP:conf/lion/KotthoffMS16}.  However, in \cref{section:labelled} we saw that there are many
 instances that VF2 finds hard that should not be hard, and that are not hard for other algorithms.
@@
 ``hard'' instances easy again.  This verifies our intuition: the apparent success of
-\citet{DBLP:conf/edbt/KatsarouNT17}'s technique is due to the ease of sometimes getting better
+\citeA{DBLP:conf/edbt/KatsarouNT17}'s technique is due to the ease of sometimes getting better
 results out of a poor algorithm. Obtaining huge improvements from an algorithm portfolio consisting
@@
 Improved search orderings for VF2-style algorithms also appear in recent work by
-\citet{DBLP:conf/gbrpr/CarlettiFV15} and \citet{o:Carletti16}, cumulating in VF3
-\citep{DBLP:conf/gbrpr/CarlettiFSV17,o:CarlettiFSV17}, and by \citet{DBLP:conf/wise/ShenZ17}. We
+\citeA{DBLP:conf/gbrpr/CarlettiFV15} and \citeA{o:Carletti16}, cumulating in VF3
+\cite{DBLP:conf/gbrpr/CarlettiFSV17,o:CarlettiFSV17}, and by \citeA{DBLP:conf/wise/ShenZ17}. We
 have seen in \cref{section:labelled} that such an approach could give improvements, but will not
@@
 Another such lesson is the importance of choosing a good set of benchmark instances. As
-\citet{DBLP:journals/heuristics/Gent98} observes:
+\citeA{DBLP:journals/heuristics/Gent98} observes:
 
@@
 We hope that this paper provides some much-needed help in this direction, as some recent
-evaluations work only with extremely easy instances. For example, \citet{DBLP:journals/tcbb/BonniciG17}
+evaluations work only with extremely easy instances. For example, \citeA{DBLP:journals/tcbb/BonniciG17}
 compared various non-constraint-based approaches to LAD using benchmark sets whose
@@
 
-\citet{o:CarlettiFSV17} correctly observe that ``a performance improvement on [hard instances] has a
+\citeA{o:CarlettiFSV17} correctly observe that ``a performance improvement on [hard instances] has a
 far greater practical impact than it would have on easier graphs, since the saved time can be of
@@
 benchmarking. In particular, distance-based filtering
-\citep{DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16} has little
+\cite{DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16} has little
 to no effect on these instances, but is extremely beneficial on other kinds of graph, including
@@
 structure to exploit, which could mislead experimenters as to the value of certain kinds of
-inference. \citet{o:GentW95} give an example of an exam timetabling graph which contains an
+inference. \citeA{o:GentW95} give an example of an exam timetabling graph which contains an
 unexpected ten-vertex clique, which would be extremely unlikely in a randomly generated graph with a
@@
 in \cref{section:filterverify}, and in the wider range of application instances used by
-\citet{DBLP:conf/lion/KotthoffMS16}, suggest that the real-world instances encountered so far are
+\citeA{DBLP:conf/lion/KotthoffMS16}, suggest that the real-world instances encountered so far are
 generally fairly easy for a good algorithm. Nonetheless, carefully designed synthetic instances
@@
 large numbers of trivial instances. Employing a presolver is an obvious approach---and VF2 is
-in fact good in this role \citep{DBLP:conf/lion/KotthoffMS16}---but there are other possibilities.
+in fact good in this role \cite{DBLP:conf/lion/KotthoffMS16}---but there are other possibilities.
 For example, minimal or lazy forward-checking
-\citep{DBLP:conf/ictai/DentM94,DBLP:conf/cp/BacchusG95,o:Dent96,DBLP:conf/ecai/LarrosaM98} avoids
+\cite{DBLP:conf/ictai/DentM94,DBLP:conf/cp/BacchusG95,o:Dent96,DBLP:conf/ecai/LarrosaM98} avoids
 constructing every domain upfront, although adopting this may require alternatives to ``smallest domain

### 2018-03-01 paper/paper.tex
Enn Pee
More Enn Pee

@@
 \begin{abstract}
-    The subgraph isomorphism problem involves finding a copy of a pattern graph inside a larger
+    The subgraph isomorphism problem involves deciding whether a copy of a pattern graph inside a larger
     target graph. The non-induced version allows extra edges in the target, whilst the induced
@@
 isomorphism $i : P \rightarrowtail T$ which is also a non-induced isomorphism $\shortoverline{i} :
-\shortoverline{P} \rightarrowtail \shortoverline{T}$. Finding either form of isomorphism between two
-given graphs is \NP-complete.
+\shortoverline{P} \rightarrowtail \shortoverline{T}$. Deciding whether a subisomorphism of either
+form exists between two given graphs exists is \NP-complete.
 

### 2018-03-07 paper/paper.tex
Formatting comments

@@
 
-\jairheading{1}{2017}{0-0}{0/0}{0/0}
+\jairheading{61}{2018}{0-0}{11/17}{03/18}
 \ShortHeadings{When Subgraph Isomorphism is Really Hard}
@@
 
-\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}
-
 \begin{document}
@@
 
-\author{\name Ciaran McCreesh \thanks{This work was supported by the Engineering and Physical Sciences
-           Research Council [grant numbers EP/K503058/1, EP/M508056/1, and EP/P026842/1]} \email
-           ciaran.mccreesh@glasgow.ac.uk \\
-       \addr University of Glasgow, Glasgow, Scotland \\
-       \name Patrick Prosser \samethanks[1] \email patrick.prosser@glasgow.ac.uk \\
-       \addr University of Glasgow, Glasgow, Scotland \\
-       \name Christine Solnon \thanks{This work was supported
-       by the ANR project SoLStiCe (ANR-13-BS02-0002-01)} \email christine.solnon@insa-lyon.fr \\
-       \addr INSA-Lyon, LIRIS, UMR5205, F-69621, France \\
-       \name James Trimble \samethanks[1] \email j.trimble.1@research.gla.ac.uk \\
+\author{\name Ciaran McCreesh \email ciaran.mccreesh@glasgow.ac.uk \\
+       \name Patrick Prosser \email patrick.prosser@glasgow.ac.uk \\
+       \addr University of Glasgow, Glasgow, Scotland
+       \AND
+       \name Christine Solnon \email christine.solnon@insa-lyon.fr \\
+       \addr INSA-Lyon, LIRIS, UMR5205, F-69621, France
+       \AND
+       \name James Trimble \email j.trimble.1@research.gla.ac.uk \\
        \addr University of Glasgow, Glasgow, Scotland}
@@
 
-\bigskip
-
-Parts of this paper were presented at IJCAI \cite{DBLP:conf/ijcai/McCreeshPT16}; the first three
-sections have been expanded, whilst
-\cref{section:regular,section:labelled,section:filterverify} are new material. Most of this work
-also appeared in a PhD thesis \cite{o:McCreesh17}.
-
 \subsection{Definitions}
@@
 
-\section{Non-Induced Subgraph Isomorphisms}\label{section:non-induced}
+\section{Non-induced Subgraph Isomorphisms}\label{section:non-induced}
 
@@
 cost of individual instances, as points. The behaviour we observe looks remarkably similar to random
-3SAT problems---compare, for example, Figure~1 of \citeA{DBLP:journals/cacm/Leyton-BrownHHX14}. In
-particular, satisfiable instances tend to be easier, but show greater variation than unsatisfiable
-instances, and there are exceptionally hard satisfiable instances \cite{DBLP:conf/cp/SmithG97}.
+3SAT problems---compare, for example, Figure~1 of the work of
+\citeA{DBLP:journals/cacm/Leyton-BrownHHX14}. In particular, satisfiable instances tend to be
+easier, but show greater variation than unsatisfiable instances, and there are exceptionally hard
+satisfiable instances \cite{DBLP:conf/cp/SmithG97}.
 
@@
 use, and using directional derivatives is not enough to resolve the problem. A close observation of
-the data suggests that the actual location of the phase transition may be involved (and perhaps
-\citeauthor{DBLP:conf/aaai/Walsh98}'s suggestion applies only in these conditions). In any case, $\langle Sol
+the data suggests that the actual location of the phase transition may be involved---and perhaps
+\citeS{DBLP:conf/aaai/Walsh98} suggestion applies only in these conditions. In any case, $\langle Sol
 \rangle$ is insufficient to explain the observed behaviour in these two eighths of the parameter space.
@@
 and only if the problem is satisfiable \cite{o:Levi73}. We used this encoding with an
-implementation of \citeA{DBLP:journals/cor/SegundoRJ11}'s bit-parallel maximum clique algorithm
+implementation of \citeS{DBLP:journals/cor/SegundoRJ11} bit-parallel maximum clique algorithm
 BBMC, modified to solve the decision problem rather than the optimisation problem.  Again, our
@@
     VF2 algorithm \ldots Recent experiences with a few of these showed that the LAD algorithm was
-    very fast in ruling out impossible matches, where VF2 took a long time.''\end{displayquote}
+very fast in ruling out impossible matches, where VF2 took a long time.'' (p.\ 44)\end{displayquote}
 Similarly, \citeA{o:Murray12} uses VF2 inside a compiler, and observes extremely variable
@@
 subgraph isomorphism call.) They conclude that ``graph indexing plays a critical role at efficient
-query processing in graph databases''.
+query processing in graph databases'' (p.\ 345).
 
@@
     \textnormal{.} \] The authors explicitly state that ``the value of $T_{\mathit{iso\_test}}$ does
-not change much for a given query''. The argument presented is that
+not change much for a given query'' (p.\ 996). The argument presented is that
 \begin{displayquote}``Sequential scan is very costly
@@
 is known that subgraph isomorphism is an \NP-complete problem. Clearly, it is necessary to build
-graph indices in order to help processing graph queries.''\end{displayquote}
+graph indices in order to help processing graph queries.'' (p.\ 961)\end{displayquote}
 We reassess these claims in light of
@@
 the key to reducing query response time is to minimize the size of the candidate answer
-set''.\end{displayquote} This claim becomes understandable when one examines the subgraph
+set''.  (p.\ 568)\end{displayquote} This claim becomes understandable when one examines the subgraph
 isomorphism algorithm chosen for the verification step \cite{DBLP:journals/jacm/Ullmann76}: as the
@@
 inefficient to perform a sequential scan on every graph in the database, because the subgraph
-isomorphism test is expensive''.
+isomorphism test is expensive'' (p.\ 966).
 
@@
     graph is small, it can be easily fit in the main memory.  However, the query (matching) time
-will be very long due to the \NP-hard complexity.''\end{displayquote} They suggest that indexing is
-a way of avoiding this. Similarly, \citeA{DBLP:journals/pvldb/ZhaoH10} argue that ``the graph query
-problem is hard in that \ldots it requires subgraph isomorphism checking \ldots which has proven to
-be \NP-complete'' and that working with large networks is hard or impossible ``due to the lack of
-scalable graph indexing mechanisms''.
+will be very long due to the \NP-hard complexity.'' (p.\ 193)\end{displayquote} They suggest that
+indexing is a way of avoiding this. Similarly, \citeA{DBLP:journals/pvldb/ZhaoH10} argue that ``the
+graph query problem is hard in that \ldots it requires subgraph isomorphism checking \ldots which
+has proven to be \NP-complete'' and that working with large networks is hard or impossible ``due to
+the lack of scalable graph indexing mechanisms'' (p.\ 340).
 
@@
 filter~/ verify, stating that ``checking subgraph isomorphism is \NP-complete, and therefore it is
-infeasible to employ such a costly solution'' which simply checks every graph for a match.
+infeasible to employ such a costly solution'' (p.\ 393) which simply checks every graph for a match.
 \citeA{DBLP:journals/tkde/WangWYY12} look at indexing large sparse graphs. They state that ``because
 subgraph isomorphism is an \NP-complete problem, a filter-and-verification method is usually
-employed to speed up the search efficiency of graph similarity matching over a graph set''.
+employed to speed up the search efficiency of graph similarity matching over a graph set'' (p.\ 441).
 Similarly, after reviewing the literature, \citeA{DBLP:journals/vldb/YuanM13} argue that subgraph
@@
 filtering techniques. They state that performing a query against each graph in the dataset
-``obviously does not scale, as subgraph isomorphism is \NP-complete''. In describing a system which
-returns a special subset of graphs which match a query, \citeA{DBLP:journals/tkde/ZhengLZHZ16}
-suggest that it is ``\NP-hard to check the graph isomorphism'' (meaning subgraph isomorphism), and
-``in order to improve the time efficiency'' they use an indexing system to ``avoid as many costly
-subgraph isomorphism checkings as possible''. Their index takes tens of thousands of seconds to
-build, and they suggest that Ullmann's algorithm and VF2 are state of the art for verification.
-\citeA{DBLP:journals/www/PengZ0LZ16} state that ``obviously, it is impossible to employ some
-subgraph isomorphism algorithm, such as Ullmann or VF2'', and argue that ``in order to speed up
-query processing'', they need to create indices.  And \citeA{DBLP:conf/isda/AzaouziR16} argue that
-``since the subgraph isomorphism test is expensive, checking all graphs of a large database can be
-unfeasible'', saying that ``naturally, the verification step is computationally more expensive since
-it requires a subgraph isomorphism''; their experiments look at candidate set reduction sizes, and
-their choice of subgraph isomorphism algorithm is not mentioned.
+``obviously does not scale, as subgraph isomorphism is \NP-complete'' (p.\ 1566). In describing a
+system which returns a special subset of graphs which match a query,
+\citeA{DBLP:journals/tkde/ZhengLZHZ16} suggest that it is ``\NP-hard to check the graph
+isomorphism'' (meaning subgraph isomorphism), and ``in order to improve the time efficiency'' they
+use an indexing system to ``avoid as many costly subgraph isomorphism checkings as possible'' (p.\
+1806). Their index takes tens of thousands of seconds to build, and they suggest that Ullmann's
+algorithm and VF2 are state of the art for verification.  \citeA{DBLP:journals/www/PengZ0LZ16} state
+that ``obviously, it is impossible to employ some subgraph isomorphism algorithm, such as Ullmann or
+VF2'', and argue that ``in order to speed up query processing'', they need to create indices (p.\
+418).  And \citeA{DBLP:conf/isda/AzaouziR16} argue that ``since the subgraph isomorphism test is expensive,
+checking all graphs of a large database can be unfeasible'', saying that ``naturally, the
+verification step is computationally more expensive since it requires a subgraph isomorphism'' (p.\
+251); their experiments look at candidate set reduction sizes, and their choice of subgraph isomorphism
+algorithm is not mentioned.
 
@@
 Ullmann's algorithm ``is prohibitively expensive for querying against [a] database with a large
-number of graphs'', and that indices are used ``to filter out graphs that do not match the query''.
-More recently, \citeA{DBLP:journals/pvldb/YuanMG13} continue a line of supergraph search work, again
+number of graphs'', and that indices are used ``to filter out graphs that do not match the query''
+(p.\ 964).  More recently, \citeA{DBLP:journals/pvldb/YuanMG13} continue a line of supergraph search work, again
 using a filtering step to avoid subgraph isomorphism calls (in the opposite direction, so queries
@@
 isomorphism with an additional set-similarity constraint, and state that Ullmann's algorithm and VF2
-are ``usually costly for large graphs'' because they ``do not utilize any index structure''. They
-propose an indexing structure, which takes over 2,000 seconds to construct, and uses nearly 2GBytes
+are ``usually costly for large graphs'' because they ``do not utilize any index structure'' (p.\
+2509). They propose an indexing structure, which takes over 2,000 seconds to construct, and uses nearly 2GBytes
 of space.  There is also research into maintaining indices when the set of target graph changes: for
@@
 queries, \citeA{DBLP:conf/edbt/WangNT16} state that querying is a ``very costly operation as it
-entails the \NP-complete problem of subgraph isomorphism'', and place ``an emphasis on the number of
-unnecessary subgraph isomorphism tests''.
+entails the \NP-complete problem of subgraph isomorphism'' (p.\ 41), and place ``an emphasis on the
+number of unnecessary subgraph isomorphism tests'' (p.\ 42).
 
@@
 on some instances, doing so does not make its performance come close to that of domain-based
-algorithms.  \citeauthor{DBLP:conf/edbt/KatsarouNT17} say they ``hope that our findings will open up
+algorithms.  \citeA{DBLP:conf/edbt/KatsarouNT17} say they ``hope that our findings will open up
 new research directions, striving to find appropriate, per-query, isomorphic rewritings, in
-combination with alternate per-query sub-iso algorithms that can yield large improvements''. We
-believe it is important instead to emphasise previous research directions that have already solved
-most of this problem, and to help ensure that these techniques become more widely known.
-
-\citeA{DBLP:conf/edbt/KatsarouNT17}'s second claim is that ``different
-algorithms find different queries hard''. To address this, they run many subgraph isomorphism
+combination with alternate per-query sub-iso algorithms that can yield large improvements''
+(p.\ 36). We believe it is important instead to emphasise previous research directions that have already
+solved most of this problem, and to help ensure that these techniques become more widely known.
+
+\citeS{DBLP:conf/edbt/KatsarouNT17} second claim is that ``different
+algorithms find different queries hard'' (p.\ 30). To address this, they run many subgraph isomorphism
 algorithms and input permutations in parallel; the extensive literature on parallel portfolios in
@@
 they say that ``using machine learning models to predict which version \ldots to employ per query is
-of high interest'').  The evidence so far in this paper suggests that we should carefully examine
+of high interest'', p.\ 36).  The evidence so far in this paper suggests that we should carefully examine
 the reasons behind these speedups. It is certainly true that there are some instances that all
@@
 ``hard'' instances easy again.  This verifies our intuition: the apparent success of
-\citeA{DBLP:conf/edbt/KatsarouNT17}'s technique is due to the ease of sometimes getting better
+\citeS{DBLP:conf/edbt/KatsarouNT17} technique is due to the ease of sometimes getting better
 results out of a poor algorithm. Obtaining huge improvements from an algorithm portfolio consisting
@@
     experimenters should be careful to perform tests on problems that can reasonably be regarded as
-    hard.''
+    hard.'' (p.\ 299)
 \end{displayquote}
@@
 far greater practical impact than it would have on easier graphs, since the saved time can be of
-several hours or even days''. To evaluate VF3, they introduce a new dataset consisting entirely of
-satisfiable instances created by taking a connected subgraph of a randomly generated graph and
-permuting the nodes. They conclude that VF3 is the best algorithm based upon its performance solving
-the enumeration problem on these new instances.  We urge a more cautious interpretation of these
-results: as we saw in \cref{section:induced}, being genuinely hard (rather than simply larger or
-denser) is a subtle property, and care must be taken when empirically evaluating algorithms. In
-particular, if one only considers large, easy instances, then the simpler the algorithm, the better
-it scales. Indeed, we saw in \cref{figure:induced} that VF3 performs \emph{worse} than VF2 on
-genuinely harder instances, and additionally behaves erratically in parts of the parameter space
-that every other solver finds easy; finally, an induced analogue of \cref{figure:labels} confirms
-that VF3 also still struggles with many obviously unsatisfiable instances.
+several hours or even days'' (p.\ 805). To evaluate VF3, they introduce a new dataset consisting
+entirely of satisfiable instances created by taking a connected subgraph of a randomly generated
+graph and permuting the nodes. They conclude that VF3 is the best algorithm based upon its
+performance solving the enumeration problem on these new instances.  We urge a more cautious
+interpretation of these results: as we saw in \cref{section:induced}, being genuinely hard (rather
+than simply larger or denser) is a subtle property, and care must be taken when empirically
+evaluating algorithms. In particular, if one only considers large, easy instances, then the simpler
+the algorithm, the better it scales. Indeed, we saw in \cref{figure:induced} that VF3 performs
+\emph{worse} than VF2 on genuinely harder instances, and additionally behaves erratically in parts
+of the parameter space that every other solver finds easy; finally, an induced analogue of
+\cref{figure:labels} confirms that VF3 also still struggles with many obviously unsatisfiable
+instances.
 
@@
 
-\acks{The authors would like to thank Kitty Meeks and Craig Reilly for their comments. We are
-grateful to Iva Babukova for her help with the datasets and for sharing the results of her
-experiments. The heatmap schemes used in this paper were created with the assistance of the
-``Chroma.js Color Scale Helper'' by Gregor Aisch.}
+\acks{
+The authors would like to thank Kitty Meeks and Craig Reilly for their comments. We are grateful to
+Iva Babukova for her help with the datasets and for sharing the results of her experiments.  The
+heatmap schemes used in this paper were created with the assistance of the ``Chroma.js Color Scale
+Helper'' by Gregor Aisch.  Parts of this paper were presented at IJCAI
+\cite{DBLP:conf/ijcai/McCreeshPT16}; the first three sections have been expanded, whilst
+\cref{section:regular,section:labelled,section:filterverify} are new material. Most of this work
+also appeared in a PhD thesis \cite{o:McCreesh17}.
+
+This work was supported by the Engineering and Physical Sciences Research Council [grant numbers
+EP/K503058/1, EP/M508056/1, and EP/P026842/1], and by the ANR project SoLStiCe
+(ANR-13-BS02-0002-01).}
 
